### 2014-09-16 Proposal/esoform-95A/usersmanual.tex
Proposal docs
Former-commit-id: 88af7443ff87045ef5e4f54e7253e44bad1257cf

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%.IDENTIFICATION $Id: usersmanual.tex.src,v 1.25 2007/09/04 14:53:32 fsogni Exp $
%.LANGUAGE       TeX, LaTeX
%.ENVIRONMENT    ESOFORM
%.PURPOSE        ESOFORM User manual
%.AUTHOR         This file has been updated by the Observing Programmes Office (OPO)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass{article}
\topmargin=-1cm
\textheight=23.7cm
\textwidth=17cm
\oddsidemargin=-0.5cm
\itemsep=-5pt

\usepackage{color}
\definecolor{webgreen}{rgb}{0,.5,0}
\definecolor{webbrown}{rgb}{.6,0,0}
\definecolor{webyellow}{rgb}{0.98,0.92,0.73}
\definecolor{webgray}{rgb}{.753,.753,.753}
\definecolor{webred}{rgb}{.85,0,0}
\definecolor{webblue}{rgb}{0,0,.8}

\usepackage[pdftex, bookmarksnumbered=true, breaklinks=true, 
  colorlinks=true, pdfstartview=FitV, linkcolor=webred, citecolor=webred,
  urlcolor=webblue]{hyperref}

% Astro definitions
\def\ang{\AA} 
\def\degpoint{\mbox{$^\circ\mskip-7.0mu.\,$}}
\def\halpha{\mbox{H$\alpha$}}
\def\hbeta{\mbox{H$\beta$}}
\def\hgamma{\mbox{H$\gamma$}}
\def\kms{\,km~s$^{-1}$}      % note leading thinspace
\def\lya{\mbox{Ly$\alpha$}}
\def\lyb{\mbox{Ly$\beta$}}
\def\minpoint{\mbox{$'\mskip-4.7mu.\mskip0.8mu$}}
\def\mv{\mbox{$m_{_V}$}}
\def\Mv{\mbox{$M_{_V}$}}
\def\peryr{\mbox{$\>\rm yr^{-1}$}}
\def\secpoint{\mbox{$''\mskip-7.6mu.\,$}}
\def\sqdeg{\mbox{${\rm deg}^2$}}
\def\squig{\sim\!\!}
\def\subsun{\mbox{$_{\normalsize\odot}$}}
\def\deg{\hbox{$^\circ$}}
\def\sun{\hbox{$\odot$}}
\def\earth{\hbox{$\oplus$}}
\def\lesssim{\mathrel{\hbox{\rlap{\hbox{\lower4pt\hbox{$\sim$}}}\hbox{$<$}}}}
\def\gtrsim{\mathrel{\hbox{\rlap{\hbox{\lower4pt\hbox{$\sim$}}}\hbox{$>$}}}}
\def\la{\mathrel{\hbox{\rlap{\hbox{\lower4pt\hbox{$\sim$}}}\hbox{$<$}}}}
\def\ga{\mathrel{\hbox{\rlap{\hbox{\lower4pt\hbox{$\sim$}}}\hbox{$>$}}}}
\def\sq{\hbox{\rlap{$\sqcap$}$\sqcup$}}
\def\arcmin{\hbox{$^\prime$}}
\def\arcsec{\hbox{$^{\prime\prime}$}}
\def\fd{\hbox{$.\!\!^{\rm d}$}}
\def\fh{\hbox{$.\!\!^{\rm h}$}}
\def\fm{\hbox{$.\!\!^{\rm m}$}}
\def\fs{\hbox{$.\!\!^{\rm s}$}}
\def\fdg{\hbox{$.\!\!^\circ$}}
\def\farcm{\hbox{$.\mkern-4mu^\prime$}}
\def\farcs{\hbox{$.\!\!^{\prime\prime}$}}
\def\fp{\hbox{$.\!\!^{\scriptscriptstyle\rm p}$}}
\def\micron{\hbox{$\mu$m}}
\def\case#1#2{\hbox{$\frac{#1}{#2}$}}
\def\slantfrac#1#2{\hbox{$\,^#1\!/_#2$}}
\def\onehalf{\slantfrac{1}{2}}
\def\onethird{\slantfrac{1}{3}}
\def\twothirds{\slantfrac{2}{3}}
\def\onequarter{\slantfrac{1}{4}}
\def\threequarters{\slantfrac{3}{4}}
\def\ubvr{\hbox{$U\!BV\!R$}}            % UBVR system
\def\ub{\hbox{$U\!-\!B$}}               % U-B
\def\bv{\hbox{$B\!-\!V$}}               % B-V
\def\vr{\hbox{$V\!-\!R$}}               % V-R
\def\ur{\hbox{$U\!-\!R$}}               % U-R
\def\jhk{\hbox{$J\!H\!K$}}              % JHK system
\def\jh{\hbox{$J\!-\!H$}}               % J-H
\def\hk{\hbox{$H\!-\!K$}}               % H-K
\def\jk{\hbox{$J\!-\!K$}}               % J-K
\def\nodata{\multicolumn{1}{c}{$\cdots$}}
\makeatletter
\def\ion#1#2{#1$\;${\small\rm\@Roman{#2}}\relax}
\makeatother

\begin{document}

\centerline{{\Large{\bf ESOFORM PDF\LaTeX\ MACROS}}}
\bigskip
\centerline{{\Large{\bf USER MANUAL FOR PHASE~1 PROPOSALS }}}
\bigskip
\centerline{{\bf European Southern Observatory }}
\vspace{3cm}

\def\period{95}

\section*{CONTENTS}

\begin{enumerate}  
\item Introduction 
\item New features for Period 95
\item How to fill a Normal Programme template 
\item How to fill a Large  Programme template 
\item Submission of the application 
\end{enumerate}

\vspace{3.0truecm}

\centerline{\large ESO DEADLINE FOR PROPOSAL SUBMISSION }
\medskip
\centerline{The ESO deadline for Period 95 proposal submission is:}
\medskip
\centerline{\huge \bf 01 October 2014, 12:00 CEST}

\vfill

\noindent This User Manual and the whole ESOFORM Package is
maintained by the Observing Programmes Office (OPO),
while the background software
is provided by the User Support System (USS) Department.

\break
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{INTRODUCTION}

The ESOFORM package has been designed to enable a fully
electronic preparation and submission of applications for observing
time on all telescopes located at the European Southern Observatory's
La Silla Paranal observatory.

{\bf Getting help.} Should you need assistance from ESO to prepare
your proposal, please send emails to the address
\href{mailto:esoform@eso.org}{\tt esoform@eso.org} for questions
related to the ESOFORM package as well as for more general questions
about Observatory policy, etc...
You should receive automatic acknowledgment from the system within 
4 hours. If you do not receive any acknowledgment of your query, 
please email 
\href{mailto:opo@eso.org}{\tt opo@eso.org} directly.

For instrument specific questions,
please email \href{mailto:usd-help@eso.org}{\tt usd-help@eso.org}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{How to Obtain the {\bf New ESOFORM} Proposal Package}

The ESOFORM Proposal Package may be obtained by logging into the ESO
User Portal at the following address:

\begin{center}
  \href{http://www.eso.org/UserPortal}{\bf
    \underline{http://www.eso.org/UserPortal}}. 
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Description of the Content of the ESOFORM Proposal
  Package}

\noindent The ESOFORM package consists of:
\begin{itemize}  
\item two \LaTeX\ class files ({\tt esoform.cls} and {\tt
    esoformlarge.cls}) that, together with the style files ({\tt
    common2e.sty}
 and {\tt config.sty}), define
  all the macros required to generate the application forms for
  Normal and Large Programmes, respectively;
\item two template proposals ({\tt  template.tex} for Normal
  Programme applications, and  {\tt templatelarge.tex}  for  Large  Programme 
    applications), which the  users   may  edit directly in  order  to 
    create a new proposal;
\item this Users' Manual ({\tt usersmanual.tex}), which contains all the 
    information  required to fill the templates, as well as
    instructions on the electronic submission of proposals (via the
    Web-based WASP interface in the ESO User Portal);

\item a short README file.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{General Features}

The present manual describes the use of the ESOFORM templates,
which are composed of macros that are defined in the ESOFORM class and
style files. The macros allow the 
computer controlled typesetting of applications for observing time at
ESO.  If you are already familiar with \TeX\ or \LaTeX, you will
probably have no difficulty using the macros provided.  You should
follow the instructions given below and keep in mind that all your
input must conform to the standard \LaTeX\ rules.

The ESOFORM package has been prepared with the following
version of pdf\LaTeX: pdf\TeX, Version 3.141592 (Web2C 7.5.5).  If you
encounter any serious pdf\TeX\ or pdf\LaTeX\ problem, please send an
email to the address \href{mailto:esoformeso.org}{\tt esoform@eso.org},
describing the problem and indicating which version of pdf\LaTeX\ you
are using.  For ease of use, we have adopted (and already included in
the class files {\tt esoform.cls},  and {\tt esoformlarge.cls}) 
a number of \LaTeX\ definitions of commonly used
astronomical symbols (see list in Table~\ref{tab:symbols}).

\begin{table}[t]
\caption{Astronomical \LaTeX\ Symbols}
\label{tab:symbols}
\medskip
\begin{tabular*}{\hsize}{@{\extracolsep{0pt}}r@{\extracolsep{20pt}}l@{\extracolsep{\fill}}r@{\extracolsep{20pt}}ll@{\extracolsep{0pt}}}
\hline
\hline
& & & &\\
\verb"\ang"      & \ang      & \verb"90\deg"        & 90\deg        & \\
\verb"\halpha"   & \halpha   & \verb"16\sqdeg"      & 16\sqdeg      & \\
\verb"\hbeta"    & \hbeta    & \verb"28\arcmin"     & 28\arcmin     & \\
\verb"\hgamma"   & \hgamma   & \verb"11\arcsec"     & 11\arcsec     & \\
\verb"\lya"      & \lya      & \verb"5\fd4"         & 5\fd4         & \\
\verb"\lyb"      & \lyb      & \verb"8\fh2"         & 8\fh2         & \\
\verb"\mv"       & \mv       & \verb"2\fm56"        & 2\fm56        & \\
\verb"\Mv"       & \Mv       & \verb"10\fs08"       & 10\fs08       & \\
\verb"\ubvr"     & \ubvr     & \verb"23\fdg12"      & 23\fdg12      & \\
\verb"\ub"       & \ub       & \verb"3\farcm6"      & 3\farcm6      & \\
\verb"\bv"       & \bv       & \verb"0\farcs27"     & 0\farcs27     & \\
\verb"\vr"       & \vr       & \verb"0\fp4"         & 0\fp4         & \\
\verb"\ur"       & \ur       & \verb"\onehalf"      & \onehalf      & \\
\verb"\jhk"      & \jhk      & \verb"\onethird"     & \onethird     & \\
\verb"\jh"       & \jh       & \verb"\twothirds"    & \twothirds    & \\
\verb"\hk"       & \hk       & \verb"\onequarter"   & \onequarter   & \\
\verb"\jk"       & \jk       & \verb"\threequarters"& \threequarters& \\
\verb"\ion{C}{4}"& \ion{C}{4}& \verb"\slantfrac{{22}}{7}" & \slantfrac{{22}}{7} & (braces unless one character) \\
\verb"3.6\micron"& 3.6\micron& \verb"$\squig$" & $\squig$ & (math mode only) \\
\verb"25\kms"    & 25\kms    & \verb"$\lesssim$" & $\lesssim$ & (math mode only) \\
\verb"\peryr"    & \peryr    & \verb"$\gtrsim$"  & $\gtrsim$ & (math mode only) \\
\verb"M\subsun"  & M\subsun  & \verb"$\la$"      & $\la$ & (math mode only) \\
\verb"\sun"      & \sun      & \verb"$\ga$"      & $\ga$ & (math mode only) \\
\verb"\earth"    & \earth    & \verb"\nodata"  & \nodata & (tables only) \\
\verb"\sq"       & \sq       & \verb"" & & \\[4pt]
\hline
\end{tabular*}
\end{table}

For every observing period, the layout of the instruments will be
updated according to the anticipated availability of instruments at
the La Silla Paranal Observatory. The ESO Call for Proposals for the 
corresponding period has all the latest information on instrument availability
and any significant changes in performance. {\it Please note that {\bf only}
  proposals prepared using the {\bf latest} version of
  ESOFORM will be valid and accepted by ESO.}

\begin{itemize}

\item{\bf ESO proposal submission deadline: }
Please note that the ESO deadline will be strictly enforced: users should plan accordingly.
It is the PI's responsibility to resolve any problems related to LaTeX, figure uploads or configuration problems and verify the LaTeX proposal form well before the deadline as ESO cannot provide support beyond 11:00 CET on the day of the deadline.
The online receiver will switch off at 12:00 CET and no submissions or amendments to submitted proposals can be accepted after this time.
\end{itemize}

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{IMPORTANT REMINDERS AND NEW FEATURES FOR PERIOD 95}
\label{sec:new}

\begin{itemize}

\item{\bf Normal and Large Programme forms:}
The Large Programme template ({\tt templatelarge.tex}) must be used to
prepare Large Programme proposals. The total amount of time requested in a programme determines which proposal template should be used.
Any programme requesting a total amount of time of 100 hours or more
must use the Large Programme template.
The Normal Programme template form {\bf must} be used for all other programmes,
including the Normal, GTO, ToO, Monitoring and Calibration proposal types.

\item{\bf Large Programme Data Products: }
A key requirement of Large Programmes is the delivery of data products into the ESO Archive. Starting from Period 95, PIs of Large Programmes are required to provide all data products within a reasonable timeframe. Further details are available in the Large Programmes Section of the Call for Proposals document.
A detailed delivery plan along with the relevant experience of proposing teams
should be outlined in Sections 6 and 7 of the Large Programme proposal template
(also see Section~\ref{sec:large}).

\item{\bf Monitoring programme proposals:}  
Proposers can apply for Monitoring Programmes to monitor targets over several periods.  
Monitoring Programme proposals should be prepared using the normal template form \verb|template.tex|. 
The Monitoring Programme type should be specified in this form as follows: \verb|\ProgrammeType{MONITORING}|, also see Table~\ref{tab:insmonitoring}.
More details on the definition of Monitoring Programmes and instruments available for this type of programme are given in Section~4.2 of the Call for Proposals.

\item{\bf Definition of Service Mode and Visitor Mode runs: }
%Remove in P95
An observing programme, as described in a single proposal, may consist of one or more runs.
Multiple runs should only be requested for observations with different instruments, for different
observing modes (e.g., service mode, visitor mode or pre-imaging runs).
Proposers should split Visitor Mode observations at different epochs (e.g., due to different target RAs) into
separate runs. Conversely, Service Mode runs should not be split
according to time-critical windows, or used to group targets according to their Right Ascensions.

\item{\bf Observing conditions:} The definitions of the observing conditions for Phase 1 and Phase 2
can be found on the  \href{http://www.eso.org/sci/observing/phase2/ObsConditions.html}{\bf \underline{Observing Conditions}} webpage.
%Remove in Period 95
Please note in particular the change of the minimum moon distance for grey conditions since Period 92.

\item{\bf Scheduling requirements in visitor mode: }
Any time constraints related to scheduling visitor mode runs should be flagged
using the \verb|\HasTimingConstraints| macro.
Any time constraints specified using the \verb|\TimeCritical| macro 
will be subject to a strict verification  at proposal submission. Please take
particular care to ensure
that the start and end dates of suitable observation windows comply
with the applicable time specification convention, and that the
amount of requested time (in the \verb|\ObservingRun| macro) fits
within the length of the slots indicated in the
\verb|\TimeCritical| macro.  \\
Several time intervals might be acceptable for each run 
(e.g., transits of extrasolar planets
across their host star); if this is the case the \verb|\TimeCritical| macro should be 
repeated as many   times as there are adequate alternative observation dates.
Further details and examples of the
correct usage are shown in Section~\ref{sec:timecrit}.

\item \textbf{Precipitable water vapour (PWV):} Users of
%add VISIR & CRIRES here when they are available again
APEX instruments in Service Mode must specify an upper limit for PWV
  as an observing constraint during their Phase 1 and Phase 2
  preparation. Examples of the correct usage are shown in the ESOFORM package 
  template files.

 \item{\bf VLTI-ATs:} The baselines offered in Period 95 are listed on the
\href{http://www.eso.org/sci/facilities/paranal/telescopes/vlti/configuration/index.html}
{\bf\underline{VLTI baseline}} page.

\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{HOW TO FILL A NORMAL PROGRAMME TEMPLATE}
\label{sec:normal}

As mentioned in the Introduction, you should fill in 
the template file ({\tt template.tex}) with your favourite editor. 
Instructions for Large Programmes
({\tt templatelarge.tex}) are given in Sect.~\ref{sec:large}. 
The easiest way to write a proposal is to modify the
template file {\tt template.tex} by following the examples therein and the
detailed instructions given in the present manual. Input in the
template is allowed {\bf only within the arguments of the provided
  ESOFORM macros}. The presence of text {\bf outside} the macro
arguments will lead to {\bf rejection} of the proposal by the
automatic proposal reception system (see Sect.~\ref{sec:submission}).

Please note that {\bf it is the responsibility of the 
  applicants to stay 
  within the current box limits} and to eliminate potential
  overfill/overwrite problems. A careful visual check of the
  generated pdf file is mandatory.

\subsection{The Cycle, the Title, the Subcategory Code, the GTO,
  ToO, RRM, and XMM Flags: {\bf BOX~1}}

The first macros to check in the {\tt template.tex} files are as follows:
\begin{itemize}
\item \verb|\Cycle| contains the Period ID for this Call for
  Proposals, and should NOT be modified by the users.
\item \verb|\Title| must contain the title of the application (up to
  two lines).
\item \verb|\SubCategoryCode| must contain only one subcategory code,
  corresponding to the keyword (see Table~\ref{tab:categories}) best
  summarizing the aim of your proposal. 
  For example, a study of
  high-redshift clusters of galaxies will have the code A5. 
  Please note: CALIBRATION proposals MUST have a \verb|SubCategoryCode| of L0.
\item \verb|\ProgrammeType| should be {\tt NORMAL} for Normal Programmes, 
 {\tt CALIBRATION} for Calibration, {\tt GTO} for Guaranteed Time Observation 
 (GTO), {\tt MONITORING} for Monitoring Programmes, or {\tt TOO} for Target of 
 Opportunity (ToO) programmes.  
 {\tt LARGE} types are also defined for Large Programmes, but they can only be 
 used within the corresponding specific template, {\tt templatelarge.tex} 
 (see Sect.~\ref{sec:large}).\\
  A CALIBRATION proposal MUST have a SubCategoryCode of L0.\\
  Note that a GTO proposal should ask only for GTO time; it is
  compulsory to fill another, non-GTO proposal if you need more non-GTO
  time, even if it is for exactly the same project. If you submit a
  TOO proposal, your proposal must include a completed ToO page
  (see Sect.~\ref{sec:toopage}). If you submit a GTO proposal, you
  must additionally uncomment and fill in the macro
  \verb|\GTOcontract|, which is described below.
\item \verb|\GTOcontract| {\em must\/} be uncommented if the proposal
  type is GTO. This macro takes one mandatory argument: 
  the keyword corresponding to the contract or agreement
  governing the allocation of Guaranteed Time under which the present
  proposal is submitted. The applicable keyword for each contract or
  agreement has been communicated by ESO to the coordinator of the
  respective GTO Team or to the designated contact person. Every
  single GTO proposal is allowed to request time within the framework
  of only one GTO contract or agreement. If your project involves
  Guaranteed Time corresponding to several different agreements, you
  must submit separate proposals for the time to be charged to each
  agreement. 
\item \verb|\ObservationInRRM| must be uncommented if your proposal is
  a Target of Opportunity proposal applying for Rapid Response Mode
  (RRM) observations.
\end{itemize}

\ifodd\period
 \begin{itemize}
  \item \verb|\ObservationWithXMM| must be uncommented if your proposal
   is applying for time under the VLT-XMM agreement (see the Call for
   Proposals for details on this agreement).  Note that in this case
   users can apply for both Periods 95 and 96.
 \end{itemize}
\else
 \begin{itemize}
 \item \verb|\ObservationWithXMM| should be left commented, as for
   Period 95 it is not possible to apply for time under the
   VLT-XMM agreement.
 \end{itemize}
\fi

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{table}[p]
\caption{ESO OPC categories and subcategories}
\label{tab:categories}
\medskip
{\small\sf
\begin{center}
\begin{tabular}{llcl} 
\hline\hline
& & & \\
Panels & Categories & Code & Subcategories \\
& & & \\
\hline 
& & & \\[-6pt]
{\bf A}& Cosmology & A1 & Surveys of AGNs and high-z galaxies; \\
       &           & A2 & Identification studies of extragalactic surveys; \\
       &           & A3 & Large scale structure and evolution; \\
       &           & A4 & Distance scale; \\  
       &           & A5 & Groups and clusters of galaxies; \\  
       &           & A6 & Gravitational lensing; \\  
       &           & A7 & Intervening absorption line systems; \\  
       &           & A8 & High-redshift galaxies (star formation and ISM). \\[4pt]
%& & & \\
\hline
& & & \\[-6pt]
{\bf B}& Galaxies  & B1 & Morphology and galactic structure; \\  
       & and       & B2 & Unresolved and resolved stellar populations; \\  
       & galactic nuclei & B3 & Chemical evolution; \\  
       &           & B4 & Galaxy dynamics; \\  
       &           & B5 & Peculiar/interacting galaxies; \\   
       &           & B6 & Non-thermal processes in galactic nuclei (incl. \\
       &           &    &  QSRs, QSOs, blazars, Seyfert galaxies, BALs, \\ 
       &           &    &  radio galaxies, and LINERS); \\  
       &           & B7 & Thermal processes in galactic nuclei and starburst\\
       &           &    &  galaxies (incl. ultraluminous \\  
       &           &    &  IR galaxies, outflows, emission lines, and \\
       &           &    &  spectral energy distributions); \\  
       &           & B8 & Central supermassive objects; \\
       &           & B9 & AGN host galaxies. \\[4pt]
%& & & \\
\hline
& & & \\[-6pt]
{\bf C}& ISM, & C1 & Gas and dust, giant molecular clouds, cool and hot gas,\\
       & star formation &   &  diffuse and translucent clouds; \\ 
       & and            &C2& Chemical processes in the interstellar medium; \\
       & planetary systems &C3& Star forming regions, globules, protostars, \\
       &      &    &  HII regions; \\
       &      & C4 & Pre-main-sequence stars (massive PMS stars, \\
       &      &    &  Herbig Ae/Be stars and T Tauri stars); \\  
       &      & C5 & Outflows, stellar jets, HH objects; \\
       &      & C6 & Main-sequence stars with circumstellar matter, \\
       &      &    &  early evolution;   \\
       &      & C7 & Young binaries, brown dwarfs, exosolar planet searches;\\ 
       &      & C8 & Solar system (planets, comets, small bodies). \\[4pt]  
%& & & \\
\hline
& & & \\[-6pt]
{\bf D}& Stellar   & D1 & Main-sequence stars; \\
       & evolution & D2 & Post-main-sequence stars, giants, supergiants, \\
       &           &    &  AGB stars, post-AGB stars; \\
       &           & D3 & Pulsating stars and stellar activity; \\
       &           & D4 & Mass loss and winds; \\
       &           & D5 & Supernovae, pulsars; \\
       &           & D6 & Planetary nebulae, nova remnants and \\
       &           &    &  supernova remnants; \\
       &           & D7 & Pre-white dwarfs and white dwarfs, neutron stars; \\ 
       &           & D8 & Evolved binaries, black-hole candidates, novae, \\ 
       &           &    &  X-ray binaries, CVs; \\
       &           & D9 & Gamma-ray and X-ray bursters; \\
       &           & D10& OB associations, open and globular clusters, \\
       &           &    &  extragalactic star clusters; \\
       &           & D11& Individual stars in external galaxies, resolved stellar populations;\\
       &           & D12& Distance scale -- stars.\\[4pt]
%& & & \\
\hline
%& & & \\
\end{tabular}
\end{center}
}
\end{table}

Your first sequence will then have the following general format:
\begin{verbatim}
        \Cycle{95A}
        \Title{AGN host galaxies}
        \SubCategoryCode{B9}
        \ProgrammeType{GTO}
        \GTOcontract{<your-contract-keyword>)
\end{verbatim}
\noindent which means that you would like to study some AGN host
galaxies, with subcategory code B9, and this would be Guaranteed Time
Observations within the framework of the contract or agreement
referred to by keyword \verb|<your-contract-keyword>|.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{The Abstract/Total Time Requested:  {\bf BOX 2}}

The ESOFORM includes a string called the
``Total Time Requested'' in Box 2. 
This field remains blank when compiling locally; this 
is normal and proposers should not be concerned about it.
The total time requested is only computed in the version of the proposal 
that is reviewed by the OPC.

This macro (\verb|\Abstract|) contains the  abstract of the  proposal,
i.e., a brief summary, in up to nine lines, of your scientific aim.
\begin{verbatim}
          \Abstract{ .
                     .
                     .
                     The text of your summary which will usually be 
                     several lines long. Line breaking will  
                     automatically be taken care of by LaTeX.
                     .
                     .
                     .              }  <-- Do not forget the
                                           closing brace !
\end{verbatim}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Information about the Different Runs: {\bf BOX 3}}
\label{sec:obsrun}

The next macro (\verb|\ObservingRun|) allows the description of the
different parameters characterising your observing run(s) and is
necessary for the scheduling and completion of your programme (see
examples below). This macro takes ten arguments.
These must be specified between the ten pairs of curly braces \verb|{}|, 
which are related to the parameters described below.

\medskip

{\bf 1. RUN ID.} Your programme may involve several observing
runs, e.g. for complementary use of different telescopes or
different instruments. Each observing run (up to 26) required by a
proposal should be identified by a different letter, following the
sequence A, B, C, \dots, Z as needed.  Provide, in the first pair of
curly braces, the run identification(s).  For example,
\begin{verbatim}
    \ObservingRun{A}{}{}{}{}{}{}{}{}{}
    \ObservingRun{B}{}{}{}{}{}{}{}{}{}
\end{verbatim}
A Normal Programme may have up to 26 runs. Since the space for the
run description in Box 3 is limited to 10 lines, a new box containing
the observing runs beyond this limit will be created at the end of the
proposal form if needed.

\medskip

{\bf 2. PERIOD KEYWORD.} Provide the period number in the second pair
of curly braces.  For normal proposals this must always be 95,
while for  Monitoring programmes this can span up to 4 periods.

\ifodd\period
 Uncomment for odd periods
 However, if users are applying under the VLT-XMM agreement,
 a period of 96 is also valid.
\fi

\medskip

{\bf 3. INSTRUMENT.} Provide the keyword of the instrument required
for each observing run. The complete list of keywords of all
instruments offered in Period 95 for Normal Programmes is
given in Table~\ref{tab:insnormal}.

\begin{table}[h]
\caption{Keywords of Available Instruments for Normal Programmes}
\label{tab:insnormal}
\medskip
\begin{center}
\begin{tabular}{@{\extracolsep{0pt}}l@{\extracolsep{40pt}}l@{\extracolsep{0pt}}}
\hline
\hline \\[-6pt]
Telescope&Instrument keywords\\[4pt]
\hline \\[-6pt]
UT1  &NACO$^1$, FORS2, KMOS\\
UT2  &FLAMES, XSHOOTER, UVES\\
UT3  &SPHERE$^2$, VIMOS, VISIR$^3$\\
UT4  &HAWKI, MUSE, SINFONI\\
VISTA&VIRCAM\\
VST  &OMEGACAM$^4$\\
APEX &CHAMPP$^5$, FLASH$^5$, ARTEMIS, LABOCA, SHFI, SUPERCAM, SpecialAPEX\\
NTT$^6$  &SOFI, EFOSC2, SOFOSC$^7$, SpecialNTT\\
3.6$^6$  &HARPS, Special3.6\\
\hline
\end{tabular}
\end{center}
$^1$ NACO availability is subject to  successful re-commissioning in September 2014;\\
$^2$ SPHERE availability is subject to successful commissioning during Period~94;\\
$^3$ VISIR availability is subject to successful re-commissioning in November 2014;\\
$^4$ OMEGACAM Normal Programmes are restricted to poor weather conditions in Period \period.\\
$^5$ CHAMPP and FLASH are APEX PI instruments, in order to propose the use of these instruments the instrument PI must be contacted at least two weeks prior to submitting the proposal (see the Call for Proposals);\\
$^6$Only Visitor Mode observations are allowed on telescopes at La Silla.\\
$^7$ SOFOSC is the instrument configuration to be specified for combined EFOSC-2/SOFI runs (Section~\ref{sec:sofiefosc}).\\
\end{table}

\begin{table}[h]
\caption{Keywords of Available Instruments for Monitoring  Programmes}
\label{tab:insmonitoring}
\medskip
\begin{center}
\begin{tabular}{@{\extracolsep{0pt}}l@{\extracolsep{40pt}}l@{\extracolsep{0pt}}}
\hline
\hline \\[-6pt]
Telescope&Instrument keywords\\[4pt]
\hline \\[-6pt]
UT1       &NACO$^1$, FORS2, KMOS          \\
UT2       &FLAMES, UVES         \\
UT3       &SPHERE$^2$, VIMOS$^3$\\
UT4$^4$   &HAWK-I, SINFONI$^{5}$, MUSE\\
APEX$^6$  &LABOCA, SHFI         \\
\hline
\end{tabular}
\end{center}
$^1$ NACO availability is subject to  successful re-commissioning in September 2014;\\
$^2$ SPHERE availability is subject to successful commissioning during Period~94;\\
$^3$ VIMOS Monitoring Programmes are subject to the
restrictions outlined in Section 1.1 of the Call for Proposals; \\
$^4$ Monitoring programmes for UT4 instruments must be compatible with
constraints imposed by the AOF installation activities described in
Section 1.3 of the Call for Proposals (taking into account that the schedule
described is provisional). \\
$^5$ Monitoring programmes using SINFONI must be compatible with the AOF installation activities described above
as well as the likely changes to SINFONI in Period~96 as described in Section 1.3 of the Call for Proposals;\\
$^6$ APEX observations for approved MPs can only be carried out in  the ESO time-slots. \\
\end{table}

Provide, in the third pair of curly braces, the instrument(s) required
for each observing run.

For example, 
\begin{verbatim}
    \ObservingRun{A}{95}{FORS2}{}{}{}{}{}{}{}
\end{verbatim}

For visitors bringing their own Visitor Instrument, please use
the corresponding ``Special'' keyword, e.g., ``Special3.6'' for 
an instrument to be mounted on the 3.6, and ``SpecialNTT'' for the NTT.
It is then compulsory to fill in the Visitor Instrument information
page (see Sect.~\ref{sec:visins}).

\medskip

{\bf 4.  REQUESTED TIME.}  In order to allow for the automated
scheduling of proposals, you must specify the amount of time 
that you are requesting (hours in Service Mode and nights in Visitor
Mode).

\smallskip

{\bf For Service Mode (SM)}, at this stage only provide the total
number of hours requested, followed by the letter \verb|h| for hours.
This should include also the time related to any special calibrations
required in addition to the standard calibrations provided by ESO.
Any more detailed information about possible particular scheduling
features will be provided during the preparation of the Phase~2
Service Mode proposal.

\smallskip

{\bf For Visitor Mode (VM) proposals}, the runs are scheduled
according to the information supplied in the Phase~1 proposal
form.  Consequently, you are required to provide detailed information 
about any particular scheduling requirements for the successful 
completion of your programme in this form.  You
should code the requested time (in nights), starting with the total
time requested, followed by the letter \verb|n| for nights, according
to the examples below.
If runs are to be split they should be specified using the following
convention:

%\begin{center}
\begin{tabular}{ll}
& \\
4n          & for 4 consecutive nights, \\
4n=4x1      & for 4 times 1 night, with intervals in between,\\
6n=3x2      & for 3 times 2 consecutive nights, \\
7n=3x2+1 \,\,\,\, & for 3 times 2 consecutive nights followed by 1 night. \\
\end{tabular}
%\end{center}

\bigskip

If the programme needs run durations of a half-night, you can code the
requested time (in nights and half-nights), starting with the total
time requested, following the few examples below (with {\bf H1} =
first half of the night, and {\bf H2} = second half of the night):

%\begin{center}
\begin{tabular}{ll}
& \\
2n=4H1     & for 4 consecutive first halves of a night, \\
2n=4x1H2   & for 4 times the second half of a night, with intervals in between,\\   
3n=3x2H1  & for 3 times 2 consecutive first halves of a night, \\   
8n=3x2+4H2& for 3 times 2 consecutive nights followed by 4 consecutive second halves of a night, \\
8n=3x2+4x1H1& for 3 times 2 consecutive nights followed by 4 non-consecutive first halves of a night. \\
\end{tabular}
%\end{center}

\bigskip

There is flexibility in this macro so that sub-runs
can be specified for fractions of a night  (not just half-nights).
Where a decimal value precedes the H, H1 or H2 values, they specify the
fraction of the night that is required for each observation. Only
one decimal place is allowed.
For example:

\begin{tabular}{ll}
& \\
0.6n=2x0.6H1     & for 2 consecutive 0.3n runs -- starting in the first half of the night, \\
1.8n=3x1.2H2    & for  3 consecutive 0.6n runs -- starting at the end of the first half of the night, \\
2.5n=2H1+2x1.6H2 & for 2 consecutive first halves of a night followed by 2 x 0.8n runs.\\
\end{tabular}

Please note that designations such as 2.0n=2x2.0H1 are invalid. 2.0H1 implies a run that is one whole night and will therefore be rejected by the proposal receiver.
In this case the user should specify 2n=2x1.
Similarly any values greater than or equal to 2.0 preceding
H, H1 or H2 will be rejected.

\bigskip

Provide the total amount of time
that is required for the observing run in the fourth pair of curly braces, with details about
possible sub-runs.  For example,

\begin{verbatim}
    \ObservingRun{A}{95}{FORS2}{8n=3x2+4H2}{}{}{}{}{}{}
\end{verbatim}
{\bf For all non-consecutive schedules, the details of the time
  intervals between the different sub-runs must be provided in Box
  12} (see Sect.~\ref{sec:schedreq}). \\

\medskip

{\bf 5. MONTH PREFERENCE.} Provide the first three letters of the
month (e.g.\ jun) that is your first preference for
scheduling (valid months are the ones included in the current period,
namely apr, may, jun, jul,
aug, sep).
%; all months are however valid for VLT-XMM proposals).  
For Service Mode runs ``any'' should be entered here with any specific 
requirements detailed in Box 12. 
For example,
\begin{verbatim}
    \ObservingRun{A}{95}{FORS2}{8n=3x2+4H2}{jun}{}{}{}{}{}
\end{verbatim}
Please note that this month preference should be consistent with the time
constraints specified in Box 12 (Sect.~\ref{sec:schedreq}).

\medskip

{\bf 6. MOON REQUIREMENT.} Provide the required phase of the moon, by
using only one of the following three characters (see the Call for
Proposals for the exact definition), namely:
\begin{itemize}
\item {\tt d} for ``dark time''
\item {\tt g} for ``grey time''
\item {\tt n} for ``no restriction''
\end{itemize}
For example,
\begin{verbatim}
    \ObservingRun{A}{95}{FORS2}{8n=3x2+4H2}{jun}{d}{}{}{}{}
\end{verbatim}

\medskip

{\bf 7. SEEING REQUIREMENT.} Provide the required maximum acceptable
seeing value (FWHM in arcseconds) at the wavelength of observation
(see the Call for Proposals for the exact definition).  Your
requirement must be one of the following values:

\smallskip

0.4, 0.6, 0.8, 1.0, 1.2, 1.4, n

\smallskip

For example,
\begin{verbatim}
    \ObservingRun{A}{95}{FORS2}{8n=3x2+4H2}{jun}{d}{0.8}{}{}{}
\end{verbatim}

\medskip

{\bf 8. TRANSPARENCY REQUIREMENT.} Provide the transparency
condition of the atmosphere required during your observations (see the
Call for Proposals for the exact definition). Your requirement must
be one of the following values:

\smallskip

\begin{tabular}{ll}
  CLR & for clear sky, although with some rare clouds \\
  PHO & for photometric, a perfect night \\
  THN & for thin cirrus, inducing absorption up to 0.2 mag \\
\end{tabular}

\smallskip

For example, 
\begin{verbatim}
    \ObservingRun{A}{95}{FORS2}{8n=3x2+4H2}{jun}{d}{0.8}{PHO}{}{}
\end{verbatim}

\medskip

{\bf 9.  OBSERVING MODE.} Provide the requested observing mode:
\verb|v| = Visitor Mode and \verb|s| = Service Mode. For example,
\begin{verbatim}
    \ObservingRun{A}{95}{FORS2}{8n=3x2+4H2}{jun}{d}{0.8}{PHO}{v}{}
\end{verbatim}

{\bf 10.  RUN TYPE.}
This should only be filled in if the programme is a TOO or GTO 
and the run is a TOO Run.  The RUN TYPE field must be left blank 
for all Normal and Calibration Programme proposals.

For programme types that allow TOO Runs (TOO, GTO and DDT), this feature
should be used only for ToO runs.
Briefly, ToO Runs are defined to be runs for which the target cannot be known more than
one week before the observation must be carried out.
If you want to specify one of your runs 
as being a ToO run, then please enter ``TOO'' in the tenth (and final) brackets.

\begin{verbatim}
    \ObservingRun{A}{95}{FORS2}{1h}{jun}{d}{0.8}{PHO}{s}{TOO}
\end{verbatim}

More details corresponding to this ToO run must be specified in the
\verb|\TOORun| macro (Sect.~\ref{sec:toopage}).
This field should be left blank for all other (non-TOO) runs.

\subsubsection*{Alternative runs}
For each requested run, you may specify one or several
``alternative runs'' for possible execution of the proposed
observations with another instrument (in general mounted on another
telescope). To this effect, add another line in Box~3, with in the
first pair of curly braces, the letter identifying your primary run,
followed by ``\verb|/alt|''. For example,
\begin{verbatim}
    \ObservingRun{A}{95}{FORS2}{20h}{jun}{d}{0.8}{CLR}{s}{}
    \ObservingRun{A/alt}{95}{VIMOS}{6n}{jun}{d}{0.8}{CLR}{v}{}
\end{verbatim}
indicates that the observations of run A could be obtained
through allocation either of 20 hours in Service Mode with FORS2
(primary choice) or of 6 nights in Visitor Mode with VIMOS (secondary
choice). You may specify several alternative runs for each primary run
(e.g., in the example above, an EFOSC2 run might be a suitable alternative).

\subsubsection*{Multiple runs}
If more than one run is needed for execution of
the programme, then fill as many lines as needed.  For example,
\begin{verbatim} 
    \ObservingRun{A}{95}{FORS2}{40h}{jun}{d}{0.8}{PHO}{s}{}
    \ObservingRun{A/alt}{95}{VIMOS}{4n=2x4H1}{jun}{d}{0.8}{PHO}{v}{}
    \ObservingRun{B}{95}{NACO}{3n=3x1}{jul}{n}{0.6}{CLR}{v}{}
    \ObservingRun{C}{95}{FORS2}{2n=2x1}{jul}{n}{0.6}{CLR}{v}{}
\end{verbatim}

APEX users should note that all observations for a given APEX
instrument must be included in a {\bf single run} for each period. The proposal
receiver 
will reject any proposal with more than one run per period per APEX instrument.

\subsubsection*{Programmes using both EFOSC-2 and SOFI at La Silla}
\label{sec:sofiefosc}

As described in the Call for Proposals,
proposals involving observations with both EFOSC-2 and SOFI,
for which the amount of time with either instrument is less than
3 nights can be carried out at La Silla.
Note that the runs should be contiguous and the requested time
summed over both the EFOSC-2 and SOFI runs is $\ge$ 3 nights.
These types of programmes should be specified
by using the pseudo-instrument SOFOSC in the Instrument field and
the total amount of time over the EFOSC-2 and SOFI runs.

For instance, if you require 3 nights on EFOSC-2 and 2 nights on SOFI, the
combined run would be specified as follows using the \verb|\ObservingRun|
 macro:
\begin{verbatim}
    \ObservingRun{A}{95}{SOFOSC}{5n}{jun}{d}{0.8}{PHO}{v}
\end{verbatim}

and in the justification for the requested amount of time
(Box 8, macro \verb|\WhyNights|), you should indicate the actual
lengths of each of the two runs. 
SOFOSC runs (i.e. combined EFOSC-2 and SOFI runs)  with a
total duration of less than 3 nights are not accepted.

Please specify the instrument configuration for the longest run
in the macro \verb|\INSconfig|. For example, if the longest run
is with the EFOSC-2 instrument:
\begin{verbatim}
   \INSconfig{A}{SOFOSC}{Spectro-long-slit}{Grism\#2:510-1100}
\end{verbatim}

\subsubsection*{Proprietary time} 
The default data proprietary time is 12 months.  Nevertheless, you can
ask to reduce it for 
your data by using the macro \verb|\ProprietaryTime{|{\it
    time\/}\verb|}|.  The {\it time\/} is expressed in months, and
only the following values can be entered:  0, 1, 2, 6, 12.  For
example, 
\begin{verbatim}
    \ProprietaryTime{6}
\end{verbatim}

Please note that this macro does not produce any printable output at
compilation, but the information that it contains will be duly stored
in ESO's database when the proposal is submitted.

\subsection{Past, Present, and Future of this Programme:  {\bf BOX 4}}

In order to allow for the evaluation of the proposal within the
broader context of the project of which it is part, taking into
account the observations already obtained in the past and the data
still to be acquired in the future, indicate in Box~4:
\begin{itemize}
\item \verb|\AwardedNights|: the amount of time (in nights or hours)
  allocated to this project in previous periods, together with the
  programme number (e.g., 093.B-1234), and the telescope
  on which this time was allocated;
\item \verb|\FutureNights|: the amount of time (in nights or hours)
  still necessary, in the future, after this proposal, to complete the
  programme, if any, and the corresponding telescope(s). 
\end{itemize}

For example,
\begin{verbatim}
    \AwardedNights{UT1}{4n in 093.B-1234}
    \FutureNights{UT3}{20h}
\end{verbatim}

\subsection{Special Remarks:  {\bf BOX 5}}

Take advantage of this box to provide any special remark (up to three
lines).  For example,
\begin{verbatim}
  \SpecialRemarks{This programme is a resubmission, in updated form, of
    proposal 093.B-1234, which had been granted 2n in VM
    with UT2+UVES and was entirely clouded out.}
\end{verbatim}

\subsection{Name and Affiliations of PI and CoI(s): {\bf BOX 6}}

The macro \verb|\PI| must be used to identify the Principal
Investigator (PI) of the proposals. Its parameters are simply the 
ESO User Portal username of the PI. Do not fill in the PI's full
name as the receiver will convert the ESO User Portal username
to the PI's full name on submission.
Usage of this macro is illustrated in the following example, please note that the username is case-sensitive:
\begin{verbatim}
    \PI{JSMITH99}
\end{verbatim}

You should use the macro \verb|\CoI| to specify also, for
all the Co-Investigators (CoIs) of this proposal, their initial(s),
last name, and institute code. Institutes and their codes are listed 
according to country at the following webpage:\\
\href{http://www.eso.org/sci/observing/phase1/countryselect.html}
{\bf\underline{http://www.eso.org/sci/observing/phase1/countryselect.html}}.\\
You should have one instance of the macro \verb|\CoI| for each CoI of the
proposal. Even if several CoIs share the same institute
please use a separate macro for each CoI. The number of CoI macros you can 
use is in principle unlimited. However, please note that 
even if all CoIs do not appear in the printed version of the form,
the entire  CoI list is stored in the ESO database, where it
can be accessed when required.

Please note: Due to the way in which the proposal receiver system parses
the CoI macro, the number of pairs of curly brackets, \verb|{}|,
in this macro must be strictly equal to 3, i.e., the
number of parameters of the macro. Accordingly, curly
brackets should not be used within the parameters (e.g.,
to protect LaTeX signs).

\noindent
For instance:
\begin{verbatim}
   \CoI{L.}{Ma\c con}{1150}
   \CoI{R.}{Men\'endez}{1098}
\end{verbatim}

\noindent
are valid, while

\begin{verbatim}
   \CoI{L.}{Ma{\c}con}{1150}
   \CoI{R.}{Men{\'}endez}{1098}
\end{verbatim}

\noindent
are not. Unfortunately the receiver does not give an
explicit error message when such invalid forms are
used in the CoI macro, but the processing of the proposal
keeps hanging indefinitely.

\vspace{0.5cm}
An example of a CoI list follows:
\begin{verbatim}
   \CoI{H.}{Cerny}{1150}
   \CoI{S.}{Bailer-Brown}{1088}
   \CoI{K.L.}{Giorgi}{1164}
   \CoI{S.}{Lichtman}{2047}
   \CoI{L.}{Men\'endez}{1150}
\end{verbatim}

When the proposal is submitted each institute code will be converted to the
corresponding country and institute by the online receiver.

\subsection{Description of the Proposed Programme: {\bf BOX 7}}

The next two pages contain the description of the proposed programme,
including all text, references and figures.
This is restricted to TWO pages *including* figures. The
text sections are composed of two required components, which are 
activated by two macros.

A -- Scientific rationale: this section should describe the scientific
background of the project, with pertinent references; any previous
work in the field plus the justification for the present proposal
should be included.  The content of this section should be placed
between the curly braces of the macro \verb|\ScientificRationale{}|.

B -- Immediate objective of the proposal: this section should state what
is actually going to be observed and what will be extracted from the
observations, so that the feasibility becomes clear.  The content of
this section should be placed between the curly braces of the
macro \verb|\ImmediateObjective{}|.
%VLT-XMM: If this is an odd period uncomment the following.
%If this is a VLT-XMM proposal, the immediate objectives to be met with the XMM observations should also be described here.

The references should preferably use the simplified abbreviations used 
in {\em Astronomy \& Astrophysics\/}.

\medskip

\noindent{\bf THE RELATIVE LENGTHS OF EACH OF THESE 
SECTIONS ARE VARIABLE. ALL TEXT AND FIGURES MUST FIT WITHIN A TOTAL
OF TWO PAGES. } Any text not fitting within the allocated page can be
added to the next page (with the figures). 
It is the responsibility of the proposers to check that their
programme description does not exceed the maximum acceptable
length. To this effect, proposers should carry out a careful visual
inspection of a print-out of their proposal prior to submitting
it. Also, when the proposal is compiled with pdf\LaTeX, the length of
the text is checked, and a warning message is issued if it is greater
than 2 pages. While this warning may easily be overlooked in the
real-time terminal window from which pdf\LaTeX\ is run because of the
continued scrolling resulting from other output, it is recorded in the
pdf\LaTeX\ logfile.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Figures: {\bf BOX 7 (cont'd)}}

Up to ONE page of figures and text can be added in addition to the 
page detailing the Scientific Justification and the Immediate Objective.
There is a size limit of 1MB for each figure to be uploaded.
This material can be included using the 
macros \verb|\MakePicture{}{}| and \verb|\MakeCaption{}|. Any
additional text from the sections in the previous page (e.g.,
Immediate Objective) can also be included using the 
\verb|\MakeCaption{}|.

{\bf NOTE THAT POSTSCRIPT PICTURES ARE NOT ACCEPTED.} Since the
proposals are compiled using the pdf\LaTeX\ package, only JPEG and PDF
file formats are accepted.  Attachments in other formats should be
converted into one of the accepted formats using appropriate tools
(such as ps2pdf, convert, or gimp).  In order to reduce the size of
the attachments, {\bf we strongly suggest the use of the PDF format for
  simple plots and graphs, and JPEG for large figures (such as
  astronomical images).}

The figure macro \verb|\MakePicture{}{}| has two arguments: the name
of the file of the picture, and a list of optional keywords specifying
formatting parameters of the image (as defined in the {\tt graphicx}
package). For example:
\begin{verbatim}
    \MakePicture{MyPic1.pdf}{width=15cm,height=8.0cm,angle=90}
    \MakePicture{MyPic2.jpg}{width=12cm}
\end{verbatim}
The filename should have a {\tt .jpg} or {\tt .jpeg} extension for
JPEG files, and a {\tt .pdf} extension for PDF files; other extensions
are not accepted. If there is no extension in the filename the receiver
will hang and submission will not be possible.

The  caption macro \verb|\MakeCaption{}| takes one single  argument,
which should contain any \LaTeX\ caption. For example:
\begin{verbatim}
    \MakeCaption{Insert caption using LaTeX.}
\end{verbatim}

These attachments will be printed on up to one page
immediately following the scientific description. You must check the
pdf output generated by pdf\LaTeX\ before submitting your proposal to
make sure that the attachments are properly included. In particular,
{\bf colour figures should still be readable if printed in  
black and white}. Also,
it is {\bf your responsibility} to check that your attachments 
{\bf fit within the allocated 1 page}. Please note that
when the proposal is compiled with pdf\LaTeX, the space required by
the attachments is checked, and a warning message is issued if it 
exceeds 1 page. While this warning may easily be overlooked in the
real-time terminal window from which \LaTeX is run because of the
continued scrolling resulting from other output, it is recorded in the
logfile generated by LaTeX. You are strongly encouraged to check
this log file.

\subsection{Justification of Requested Time: {\bf BOX 8}}

In this box, you should provide a careful justification of the
requested lunar phase and of the requested amount of time. 
To this effect, you should use the ESO Exposure
Time Calculators whenever possible; these exist for all Paranal and La Silla
instruments and are available at
\href{http://www.eso.org/observing/etc}{\bf
  \underline{http://www.eso.org/observing/etc}}.

Links to exposure time calculators for APEX instrumentation
  can be found in Section 7 of the Call for Proposals.

For each telescope and instrument to be used, please specify the
version of the ESO Exposure Time Calculator that you have
used. Do {\bf not} include any correction for unexpected
meteorological conditions. The text should be typed as arguments of
the following two macros:
\begin{verbatim}
    \WhyLunarPhase{}
    \WhyNights{}
\end{verbatim}

\subsection{Telescope Justification: {\bf BOX 8a}}
This section should provide a justification for the use of the selected telescope(s) (e.g., VLT, NTT, etc...) with respect to other available alternatives. The content of this section should be placed between the curly
braces of the macro \verb|\TelescopeJustification{}|.

\subsection{Observing Mode Justification: {\bf BOX 8b}}
This section should provide a justification for the observing mode requested (Visitor or Service). The
content of this section should be placed between the curly braces of the macro \verb|\ModeJustification{}|.

\subsection{Calibration Request: {\bf BOX 8c}}

For Service Mode runs, the calibrations foreseen in the instrument
calibration plans are 
absorbed by the Observatory; they do not need to be included in the
amount of requested time. In Visitor Mode, up to
30\,min per night is devoted to the acquisition of these calibrations
by the Observatory staff. If, in order to achieve the scientific goals
of your projects, calibrations not foreseen in the respective
calibration plan are required, you must include the additional amount
of time that is needed to obtain them in the total amount of time that
you are requesting.

The macro \verb|\Calibrations| must be used to specify the calibration
requirements of your proposal. It takes two arguments. The first one
should be set to \verb|standard| if the calibrations contemplated in
the calibration plan are sufficient. In this case, no input is
required for the second argument:
\begin{verbatim}
    \Calibrations{standard}{}
\end{verbatim} 
If, on the other hand, you need
additional calibrations, the first argument must be set to
\verb|special|, and a brief description of non-standard calibrations
that you need must be given as second argument. For example,
\begin{verbatim}
    \Calibrations{special}{Adopt a special calibration}
\end{verbatim} 
Note that non-standard
daytime calibrations must be specified here, but contrary to
additional nighttime calibrations, the corresponding time need not 
be included in the total amount of requested time.

\subsection{Last Use of ESO Facilities: {\bf BOX 9}}
The macro \verb|\LastObservationRemark| must be used to 
provide a brief report on the use of the ESO facilities during the
last 2 years. You should specify the programme identification
numbers, and describe the status of the data obtained, and the 
scientific output generated.

\subsection{ESO Archive: {\bf BOX 9a}}
You should use the \verb|\RequestedDataRemark| macro to indicate if
the data requested in the proposal are in the ESO Archive
(\href{http://archive.eso.org}{\bf
  \underline{http://archive.eso.org}}), and if so, to explain the
need for new data.

\subsection{Duplication of GTO/Public Surveys: {\bf BOX 9b}}

Specify whether there is any duplication of targets/regions covered by ongoing GTO and/or Public Survey
programmes. 
If so, please explain the need for the new data here. Details on the protected target/fields in these
ongoing programmes can be found at the following webpages for GTO programmes and Public surveys: \\
\href{http://www.eso.org/sci/observing/teles-alloc/gto.html}
{\bf  \underline{http://www.eso.org/sci/observing/teles-alloc/gto.html}}\\
\href{http://www.eso.org/sci/observing/PublicSurveys/sciencePublicSurveys.html}
{\bf \underline{http://www.eso.org/sci/observing/PublicSurveys/sciencePublicSurveys.html}}.

\subsection{Applicant's  Publications: {\bf BOX 10}}

The macro \verb|\Publications{}| should be used to 
provide a list of your publications related to the
subject of the current proposal and published during the past two
years. The A\&A simplified abbreviations for references should be
used. The individual references should be separated with a small
amount of vertical space, to be created with the standard \LaTeX\
command \verb|\smallskip\\|. For example:
\begin{verbatim}
   \Publications{
   Name1 A., Name2 B., 2001, ApJ, 518, 567: Title of article1
   \smallskip\\
   Name3 A., Name4 B., 2002, A\&A, 388, 17: Title of article2
   \smallskip\\
   Name5 A. et al., 2002, AJ, 118, 1567: Title of article3
   }
\end{verbatim}

\subsection{List of Targets: {\bf BOX 11}}

Provide the complete list of targets to be observed in this programme,
by using the macro \verb|\Target| 
with the following nine parameters: run identifier (you may use the same
target/field in more than one run), target field/name, Right Ascension
(hh mm ss.f, or hh mm.f, or hh.f) and Declination (dd mm ss, or dd
mm.f, or dd.f) for the 
J2000 equinox, requested time on target (in hours
with overheads and calibration included), magnitude, angular diameter,
additional information (see below), and reference star identifier
(see below)
for each target field.  Please use the format \{00~00~00\}
in case of unknown coordinates. There can be as many occurrences of
the macro \verb|\Target| as required to accommodate all targets of all
runs of the programme. Long lists of targets will continue
on the last page(s) of the proposal form.

{\bf Please note} that the scheduling of your programme will take
{\bf all targets} given in this list into account. {\bf Only include
 the targets requested for Period 95}.
%(except for VLT-XMM proposals).
Make sure your targets
are significantly observable during this period.  Inclusion of targets
with insufficient visibility during the current period may result in
rejection of your programme by the automatic scheduler.

The additional information field (8th argument of the \verb|\Target|
macro) may in general be used to provide any relevant piece of
information about the target that does not pertain to any other
argument of the macro (e.g. the period of a variable star). However,
for APEX targets, usage of this field is {\bf mandatory} to indicate the
requested Precipitable Water Vapour (PWV) and the acceptable range of
Local Sidereal Time (LST) for the considered observation. The format
should be similar to the one shown in the following example:
\begin{verbatim}
\Target{A}{HD 104237}{12 00 05.6}{-78 11 33}{1}{}{}{PWV<0.7mm;LST=9h00-15h00}{}
\end{verbatim}

A reference source identifier must be provided for all natural guide stars
(NGS), in the case of NGS observations with NACO, SINFONI and CRIRES,
and all tip-tilt stars (TTS), in the case of all laser guide star (LGS)
observations with NACO 
and SINFONI. For observations with the noAO modes of SINFONI and
CRIRES, you do
not need to provide this information.  {\bf The reference source
  designation has to be the exact identifier of the selected star
  either 
  from the Guide Star Catalog 2 (GSC2) or the 2MASS point source
  catalogue.} Note that GSC2 stars identifiers should NOT be preceded
by GSC2, 
but must start with either N or S. In case the reference source is not
included in either catalogue, for instance because it is a supernova or
a solar system object, ``alt'' should be entered as reference source
identifier, and additional information can be provided in the
\verb|\TargetNotes| macro.
Rules for reference star designation can be found for GSC2 at:\\
  \href{http://vizier.u-strasbg.fr/viz-bin/VizieR-n?-source=METAnot&catid=1271&notid=1&-out=text}{\bf
    \underline{http://vizier.u-strasbg.fr/viz-bin/VizieR-n?-source=METAnot\&catid=1271\&notid=1\&-out=text}}.

\smallskip

Examples of valid and invalid GSC2 identifiers are given below:

\begin{center}
\begin{tabular}{ll}
%\multicolumn{2}{c}{} \\
%\hline
    N01230121           & good\\
    S33333331           & good\\
  n01230121             & bad\\
 N012301201             & bad\\
    S01230141           & bad\\
    S3333333000001      & bad\\
    S01201201234567     & bad\\
%\hline
\end{tabular}
\end{center}

%\vspace{1cm}

For 2MASS, the rules for reference star designation are available at:\\
  \href{http://www.ipac.caltech.edu/2mass/releases/allsky/doc/sec2_2a.html}{\bf
    \underline{http://www.ipac.caltech.edu/2mass/releases/allsky/doc/sec2\_2a.html}}.

\smallskip

Here are some examples of correct and incorrect identifiers:

\begin{center}
\begin{tabular}{ll}
%\multicolumn{2}{c}{} \\
%\hline
 2MASS J01234567+7801020      & good\\
 2MASS J00000000+7801020L     & good\\
 2MASS J01234567+9000000      & good\\
 2MASS J01234567+9000000W     & good\\
 2MASX J01234567+7801020      & bad\\
 2MASS J97234567+7801020      & bad\\
%\hline
\end{tabular}
\end{center}

Thus the following examples illustrate the correct usage of the
\verb|\Target| macro when a reference star must be specified:
\begin{verbatim}
    \Target{B}{NGC 105}{22 55 00}{-47 50 30}{9.0}{}{}{}{S33333331}
    \Target{C}{NGC 106}{00 24 43}{-05 09 00}{2.0}{}{}{}{2MASS J01234567+7801020}
\end{verbatim}

The macro \verb|\TargetsNotes{}| should be used to include any
comments that apply to several or all targets (or to specify reference
stars that are not found in the GSC2 or 2MASS catalogues).
\begin{verbatim}
   \TargetsNotes{The planned grid of pointings around the targets 
                listed above will be defined during the first 
                observing night.}
\end{verbatim}

\subsection{Scheduling Requirements: {\bf BOX 12}}
\label{sec:schedreq}
If your proposal involves any of the following:
\begin{itemize}
\item observations to be executed on specific dates (e.g., for
  simultaneity with observations at other facilities);
\item observations to be executed at pre-defined time intervals (e.g.,
  at different epochs so as to achieve phase coverage of a periodically
  variable target);
\item Visitor Mode runs split into non-consecutive nights (for which
  the fourth argument specifying the number of nights in Box~3 of the
  proposal form includes a formula);
\item Visitor Mode runs mutually linked, which need to be scheduled in
  a given sequence and at specified time intervals,
\end{itemize}
you {\bf must} uncomment the macro \verb|\HasTimingConstraints|.
Please leave the brackets in \verb|\HasTimingConstraints| blank.  Details of time
 constraints can be entered in Special Remarks and using the
 other flags in Box 12.

Please note that the macro \verb|\HasTimingConstraints| should 
be {\bf commented out} in the following cases:
\begin{itemize}
\item for scheduling constraints resulting only from the genuine
  visibility window of the target sources (defined by their location
  in the sky) or from the phases of the Moon; 
\item for time series of observations acquired during a single night
  or over several consecutive nights of a contiguous Visitor Mode run; 
\item for Target of Opportunity observations.
\end{itemize}

In order to allow for the automated scheduling of all
Visitor and Service Mode observing runs, you must provide all
information related to the details of the way your programme should be
scheduled.  If you have any doubts about how to specify your particular
requirements please email OPO at \href{mailto:opo@eso.org}{\tt opo@eso.org}
well in advance of the proposal submission deadline.

\medskip

{\bf 1. RUN SPLITTING.} For Visitor Mode runs: if the fourth argument
you have provided in 
Box~3 indicates a simple number of consecutive nights or half-nights,
e.g.:
\begin{verbatim}
    \ObservingRun{A}{95}{FORS2}{4}{may}{d}{0.8}{CLR}{v}{}
\end{verbatim}
you do not have to do anything.  If the fourth argument in Box~3
indicates a more complicated requirement, with some non-consecutive
nights, e.g.: 
\begin{verbatim}
    \ObservingRun{A}{95}{FORS2}{8n=3x2+4H2}{may}{d}{0.8}{CLR}{v}{}
\end{verbatim}
you must provide some additional information.  The fourth argument above,
namely, \verb|8n=3x2+4H2|, means that you request a total of 8 nights,
made of three sub-runs of 2 consecutive nights each, followed later by
a sub-run of 4 consecutive second halves of a night.

You should use the macro \verb|\RunSplitting{}{}|, and put
the run identifier in the first argument.  The second argument should indicate
the way the run should be split into different sub-runs.  If the
interval between two sub-runs has to be exactly a given number of
days, say 20, then this is a {\bf strong} constraint and this number
of days should be followed by the letter \verb|s|.  If the interval
between two sub-runs has some tolerance on the number of days, say
20$\pm$5, then this is a {\bf weak} constraint and this number of days
should be followed by the letter \verb|w|.  Consequently, if
\verb|8n=3x2+4H2| means that you want three sub-runs of 2 consecutive
nights each, the first and the second separated by 10 nights exactly,
the second and the third separated by more or less 20 nights, followed
exactly 15 days later by one sub-run of 4 consecutive second halves of
a night, then the second argument of the macro
\verb|\RunSplitting{}{}| should contain the following expression:
\verb|2,10s,2,20w,2,15s,4H2|. Hence the following entry should appear
in your proposal file:
\begin{verbatim}
    \RunSplitting{A}{2,10s,2,20w,2,15s,4H2}
\end{verbatim}

Note that it is necessary to specify explicitly the time intervals
among all non consecutive sub-runs.

The macro \verb|\RunSplitting| is meaningless for Service Mode runs,
for which constraints of the considered type should be fully specified
at Phase~2 (but the macro \verb|\HasTimingConstraints| should be
uncommented for such runs, so that they are duly flagged). {\bf
  However please note that in Service Mode, 
  monitoring programmes are executed on a best effort 
  basis only.} In particular, a monitoring sequence can be interrupted
by unsuitable weather conditions or by runs scheduled in Visitor Mode.

\medskip

{\bf 2.  LINK FOR COORDINATED OBSERVATIONS.} If you have requested
three different runs in Box 3, e.g.:
\begin{verbatim}
    \ObservingRun{A}{95}{FORS2}{2n}{may}{d}{0.8}{CLR}{v}{}
    \ObservingRun{B}{95}{FORS2}{3n}{may}{d}{0.8}{PHO}{v}{}
    \ObservingRun{C}{95}{UVES}{20h}{may}{d}{0.8}{CLR}{v}{}
\end{verbatim}
and would like some of them to be simultaneous and some later than
others, independently of the exact period of scheduling, then use
\verb|simultaneous, after| and the macro \verb|\Link{}{}{}{}| in the
following way:
\begin{verbatim}
    \Link{B}{after}{A}{10}
    \Link{B}{after}{A}{}
    \Link{B}{simultaneous}{C}{}
\end{verbatim}

\medskip

{\bf 3. UNSUITABLE PERIOD(S) OF TIME.}  If you have requested two
nights in Box 3 and would like them to be scheduled to avoid some
unsuitable periods of time, for some reason, then use the macro
\verb|\UnsuitableTimes{}{}{}{}| in the following way:
\begin{verbatim}
    \UnsuitableTimes{A}{15-jul-15}{18-jul-15}{International Conference}
    \UnsuitableTimes{B}{15-jul-15}{18-jul-15}{International Conference}
    \UnsuitableTimes{C}{15-jul-15}{18-jul-15}{International Conference}
    \UnsuitableTimes{C}{1-jul-15}{3-jul-15}{Committee Meeting}
\end{verbatim}

Dates correspond to 12:00 noon Local Time at
the Observatory location (i.e., in Chile).
In other words the first date refers to the start of the first night
of the unsuitable period; the second date refers to the end of the last night.
As with the \verb|TimeCritical| macro, only one run can be specified in each
\verb|UnsuitableTimes macro|. 
Further explanation can also be entered in  \verb|\SpecialRemarks{}| (Box 5).

\subsection{Scheduling Requirements contd.. : {\bf BOX 12}}
\label{sec:timecrit}

{\bf 4. SPECIFIC DATE(S) FOR TIME CRITICAL  OBSERVATIONS.} If you have
requested 2 nights in Box 3, e.g.:
\begin{verbatim}
    \ObservingRun{A}{95}{FORS2}{2n}{may}{d}{0.8}{CLR}{v}{} 
\end{verbatim}
and if for some reason (e.g., specific phase of a variable object
or parallel observations with already scheduled HST observations,
etc.)  you need these two nights scheduled between some specific
dates, then use the macro \verb|\TimeCritical{}{}{}{}| in the
following way:
\begin{verbatim}
    \TimeCritical{A}{12-may-15}{14-may-15}{parallel observation with HST}
    \TimeCritical{A}{20-may-15}{25-may-15}{co-ordinated with XMM}     
    \TimeCritical{B}{20-may-15}{25-may-15}{co-ordinated with XMM}
    \TimeCritical{C}{20-may-15}{25-may-15}{co-ordinated with XMM}
\end{verbatim}

Please note the following points:
\begin{itemize}
\item As of P87, the constraints set by this macro are checked against the
time specified by the \verb|\ObservingRun| in more detail.
Some examples of correct and incorrect uses of this pair of macros are shown below.
\item The indicated dates correspond to 12:00 noon LOCAL Time at
the Observatory location (i.e., in Chile). In other words, the first
date refers to the start of the first night of the acceptable
interval, and the second to the end of the last night. Please 
convert event times from Universal Time to Local Time.
\item Only one run can be specified in each \verb|TimeCritical| field. 
However, as an observing run can have multiple Time critical dates, users
can specify multiple separate \verb|\TimeCritical| macros for a single observing run (as shown above
for Observing Run A).
\item ESO can still access the information for this macro (and all other macros in Box 12) so
do not worry if the constraints overfill the available space.
\end{itemize}

{\bf Examples of correct usages of the \verb|\TimeCritical| macro: }\\
If the user requires an observing run to be split into three observations, each to taken in the 
second half of the night over a specific 8-day period, this could be specified using the following macros:
\begin{verbatim}
    \ObservingRun{A}{87}{VIMOS}{1.5n=3H2}{jun}{n}{0.6}{CLR}{v}{}
    \RunSplitting{A}{0.5,4w,0.5,4w,0.5}
    \TimeCritical{A}{12-apr-15}{20-apr-15}{reason here}
\end{verbatim}

Conversely if the ObservingRun was specified as above but the TimeCritical macro was configured as follows
it would be rejected as the configuration 1.5n=3H2 implies there must be a minimum time separation of
three nights.
\begin{verbatim}
    \TimeCritical{A}{12-apr-15}{14-apr-15}{reason here}
\end{verbatim}

\subsection{Instrument configuration: {\bf BOX 13}}

The two template proposals ({\tt template.tex} for normal
applications, 
and {\tt templatelarge.tex} for Large Programme applications) contain
the full list of configurations for all available instruments at all
available ESO telescopes (Paranal, La Silla and Chajnantor).  In order
to provide 
general information about the setup of the ESO instrument(s) you plan
to use, please uncomment only the lines related to the instrument
modes and configurations needed for the acquisition of your desired
observations.  For some lines related, e.g., to special filters or
central wavelength, please add the required information where
appropriate (between the already existing curly
braces).

Note that you {\bf must} put the run ID within the first pair of curly
braces of the relevant lines. {\bf Do not} specify any instrument
configuration for alternative runs or visitor instruments (see Box 3). 
Note
that all parameters are {\bf mandatory} for the \verb|\INSconfig|
macro (do not use empty fields).

\subsection{Interferometry page}
\label{sec:vlti}

For VLTI runs, a separate run must be specified for each requested baseline configuration.
If your proposal includes VLTI runs, you MUST uncomment and fill in the arguments of the 
macro \verb|\VLTITarget| with run ID, target name, visual
magnitude, magnitude at wavelength of observation, wavelength of
observation (in microns), size at wavelength of observation (in mas),
baseline (see the following website for available configurations:
\href{http://www.eso.org/paranal/insnews/vlti_overview.html}{\bf
  \underline{http://www.eso.org/paranal/insnews/vlti\_overview.html}}), 
visibility for the specified
configuration (at preferred hour angle or at hour angle 0), correlated
magnitude, and time on target (ToT) in hours. For example,
\begin{verbatim}
\VLTITarget{E}{NGC 106}{-0.7}{-3.5}{10.6}{40}{UT2-UT3-47m}{0.84}{-2.5}{6}
\VLTITarget{F}{NGC 107}{-0.7}{-3.5}{2.1}{40}{UT1-UT2-UT3}{0.84/1.0/0.1}{1./0.5/2.}{6}
\VLTITarget{G}{NGC 108}{-0.7}{1.5}{2.1}{40}{E0-G0-H0-I1}{0.90/0.80/0.70}{1.6/1.7/1.9}{6}
\end{verbatim}

For UT observations, please specify one of the available
3-telescope AMBER baselines or 2-telescope MIDI baselines.

Please specify one of the 3 available AT quadruplets listed in the
template proposal files. At Phase 2, the time can then be split
among the AMBER triplets and MIDI baselines of the chosen quadruplet.

For AT observation baselines, please use one typical baseline of the
quadruplet that you have specified in order to compute
representative visibility and correlated magnitude values.

For AMBER observations using UT baselines, you should specify the three visibilities
corresponding to the various baselines as three values separated by a
slash (/); up to two of the three values can be replaced by a star
($*$).  Similarly, the correlated magnitudes for the various baselines are also
specified as three values separated by slash.

You can use the macro \verb|\VLTITargetNotes| to insert comments about
some or all of your VLTI targets. You should take advantage of this
macro to indicate suitable alternative baselines for your
observations.

\subsection{ToO page}
\label{sec:toopage}

If your programme has Target of Opportunity runs (ToO) runs, 
you should specify these in the  \verb|\ObservingRun| macro 
using ``TOO'' in the corresponding field. 
The ToO information must be filled in for the run
using the corresponding  
\verb|\TOORun| macro.  The arguments in this macro 
are, in order: the run identifier,
the nature of the observation, the number of 
targets per run, and the number of triggers per targets. There must be
one occurrence of the macro \verb|\TOORun| for each of the ToO runs specified
in Run Type field in  the \verb|\ObservingRun| macro (Box~3).
 The second argument (nature of the observation) may be one
of the following keywords:
\begin{itemize}
\item RRM, for observations to be triggered via the automated Rapid
  Response Mode system;
\item ToO-hard, for observations to be triggered manually that need to
  be carried out within 48 hours of receipt of the trigger by the
  Observatory (and in most cases, as soon as possible), or that
  involve a strict time constraint (i.e., that must be executed during
  a specific night);
\item ToO-soft, for manually triggered observations for which the
  Observatory can receive notification more than 48 hours before
  execution, and which can be scheduled for execution with a
  flexibility of at least $\pm1$\,day.
\end{itemize}
Only one keyword can be specified for each run. If observations
pertaining to different categories are needed,
several runs must be defined. The number of triggers must be indicated
for RRM, ToO-hard and ToO-soft observations only.
An occurrence of the macro \verb|\TOORun|
looks like the following example:
\begin{verbatim}
\TOORun{A}{ToO-hard}{2}{3}
\end{verbatim}

You have the opportunity to add some notes to the ToO page by using
the macro \verb|\TOONotes|.
As a rule, ToO proposals should involve at least one trigger
of one of either the RRM or ToO-hard or ToO-soft types.

\subsection{The Visitor Instrument Page}
\label{sec:visins}

The following commands are only needed for proposals involving a Visitor
Instrument, in which case they are also {\bf mandatory}.  You should
uncomment them and provide the required information between the
different pairs of curly braces. Please do not enter anything in Box 13 
(i.e. \verb|\INSconfig|) for Visitor Instrument runs.

If applying for the VLTI visitor instrument, SpecialVLTI, please remember to uncomment
and fill in the \verb|\VLTITarget| macro and this Visitor Instrument page.

\begin{verbatim}
    %\Desc{}   % Description of the instrument and its operation
    %\Comm{}   % On which telescope(s) has instrument been commissioned/used
    %\WV{}     % Total weight and value of equipment to be shipped
    %\Wfocus{} % Weight at the focus (including ancillary equipment)
    %\Interf{} % Compatibility of attachment interface with required focus
    %\Focal{}  % Back focal distance value
    %\Acqu{}   % Acquisition, focusing, and guiding procedure
    %\Softw{}  % Compatibility with ESO software standards (data handling)
    %\Suppl    % Estimate of services expected from ESO (in person days)
\end{verbatim}

\section{HOW TO FILL A LARGE PROGRAMME TEMPLATE}
\label{sec:large}

The ESOFORM package includes a specific template ({\tt
  templatelarge.tex}) that must be used to generate a Large Programme
proposal. This template is only slightly different from the
template for Normal Programmes. Hereafter is the list of the
differences between the two templates. Any feature not appearing in
this list is {\bf identical} in both templates though the box number
may differ.

\begin{table}[t]
\caption{Keywords of Available Instruments (Large Programmes)}
\label{tab:inslarge}
\medskip
\begin{center}
\begin{tabular}{@{\extracolsep{0pt}}l@{\extracolsep{40pt}}l@{\extracolsep{0pt}}}
\hline
\hline \\[-6pt]
Telescope & Instrument keywords     \\[4pt]
\hline                              \\[-6pt]
UT1       & NACO$^1$, FORS2, KMOS    \\
UT2       & FLAMES, UVES            \\
UT3       & SPHERE$^2$, VIMOS$^3$\\
UT4$^4$   & HAWK-I, SINFONI$^5$, MUSE   \\
APEX      & LABOCA, SHFI            \\
NTT       & SOFI, EFOSC2, SpecialNTT\\
3.6       & HARPS, Special3.6       \\
\hline
\end{tabular}
\end{center}
$^1$ NACO availability is subject to  successful re-commissioning in September 2014;\\
$^2$ SPHERE availability is subject to successful commissioning during Period~94;\\
$^3$ VIMOS LPs must comply with the RA restrictions outlined in the Call for Proposals;\\
$^4$ Large Programmes for UT4 instruments must be compatible with
constraints imposed by the AOF installation activities described in
Sections 1.1 and 1.3 of Call for Proposals (taking into account that the schedule
described is provisional);\\
$^5$ Large programmes using SINFONI must be compatible with the AOF installation activities described above
as well as the likely changes to SINFONI in Period~96 as described in Section 1.2 of the Call for Proposals.

\end{table}

\begin{itemize}
\item Box 2: the abstract for Large Programme proposals can be
  slightly longer (up to 13 lines) than for Normal Programme proposals
  (up to 9 lines).  
\item Box 3: proposers must specify {\bf all the runs across all the
    Periods} covered by the Large Programme.
For Paranal instruments this means a maximum of 4 periods, while for La Silla instruments
up to Period 97 is allowed (see the Call for Proposals for more information). 
  A Large Programme can have up to 26 runs. Since the space for the runs
  description in Box 3 is limited to 12 lines, a new box containing
  the observing runs beyond this limit will be created at the end of
  the proposal form if needed.
\item Box 3: the list of keywords of all instruments offered in Period
  95 for Large Programmes is given in
  Table~\ref{tab:inslarge}.
\item Box 5: while two pages (including figures) are allowed for the scientific
  description of Normal Programmes, the
  description of Large Programmes may take up to three pages followed by
  two additional pages for figures.

  The scientific description is comprised of the following four subsections:
  \verb|\ScientificRationale{}|,
  \verb|\ImmediateObjective{}|, \verb|\TelescopeJustification{}|, and\\
  \verb|\ModeJustification{}|.  
  The \verb|\TelescopeJustification{}| and \verb|\ModeJustification{}|
  macros are the same as those in Boxes 8b and 8c in the Normal Programme
  template.

  As noted above there are also there are two additional pages for
  attachments (e.g., figures, etc). 
\item Boxes 6 and 7: specific to the Large Programme proposals, these
  two boxes should contain the required information describing the
  experience of the applicants with telescopes, instrumentation, 
  data reduction  and delivery of data products to the ESO Archive 
  (\verb|\Experience{}|). The strategy for data reduction, analysis 
  and a delivery plan for data products into the ESO Archive should 
  be presented along with a brief description of the resources available
  to the observing team, such as: computing facilities and research
  assistants (\verb|\Resources{}|).
\item Box 8: Additional remarks or comments can be provided using
  the macro \verb|\SpecialRemarks{}|.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{SUBMISSION OF THE APPLICATION}
\label{sec:submission}

Proposals must be prepared as pdf\LaTeX\ source files, making use of
the {\bf latest ESOFORM} package, corresponding to the ESO Period
for which they are submitted.  Proposals received in
any other format, or with modified ESOFORM macros, will be
automatically rejected by the automated proposal handling system.

When the \LaTeX\ source file of your application is complete, {\bf
  please process it with pdf\LaTeX} so as to identify any
possible \LaTeX\ format errors. In particular, we {\bf strongly}
recommend that you
\begin{itemize} 
\item review the log file generated by \LaTeX\ so as to
  check for the presence of warning messages issued by the ESOFORM
  macros. Such messages report, among others, instances in which a text
  field is too long, so that your input is truncated in the pdf file
  that is generated, and part of the information that you
  submit will not be communicated to the OPC;
\item carefully inspect a printed copy of the output to make sure that
  all parts of the application are completed, and that their
  formatting is appropriate.
\end{itemize}

{\bf Please note} that while several checks are
performed by the ESOFORM package when running pdf\LaTeX, successful 
compilation  {\bf does not guarantee} that a proposal is
fully compliant. Indeed, many key  checks 
can only be performed by the proposal reception system after the proposal
has been uploaded within the ESO User Portal.

{\bf You should verify that your proposal complies with ESO
requirements using a ``skeleton'' version of  your proposal that
only contains the technical details of your  programme. This should be done well before the Phase 1 Deadline.}
 Please upload the ``skeleton'' proposal as if you were submitting it and follow the instructions online.
If the proposal passes all the technical checks you will get the message:
``Your proposal is verified and passed all the checks (but not submitted yet)!
''. You can then either log out of the User Portal or 
return to the submission page by selecting the option to 
``Go back to the beginning''. 
More details on the full proposal submission procedure are given below.

Proposals are submitted by logging into the ESO User Portal.
In order to do this, proposers must be registered and have activated
their User Portal accounts. To submit please go to:
\begin{center}
  \href{http://www.eso.org/UserPortal}{\bf \underline{http://www.eso.org/UserPortal}}
\end{center}
Once you have logged in, Once you have logged in, you should select the item called ``Download ESOFORM packages'' under the Phase 1 menu items.  In order to submit, select ``Submit a proposal'' from the same menu. You will then be taken to the WASP page (Web Application for Submitting Proposals). 
On choosing the relevant cycle, you will be asked to upload the \LaTeX\ file of your proposal and should follow the subsequent instructions. \\
A number
of checks are executed at the various steps of the submission process;
please follow the instructions online. If a problem is detected 
it will be clearly reported by the system: fix it in your proposal and
resubmit it.
Once the proposal passes through all the checks, you will be requested to finalise the
submission by clicking on the corresponding button. {\bf It is
  essential that you execute this final step}: your proposal will not
be submitted until this is done, even though you have uploaded all the
necessary files! \\
 Upon submission of
a correctly completed proposal, the ESO proposal validation
software will return an identifier assigned to the valid proposal. This
identifier, and the acknowledgment page in which it appears, represent
the official confirmation that the proposal successfully entered the
proposal handling system. We recommend that you take note of the
identifier; you may also want to print the acknowledgment page for
your records. In addition, an email confirmation
is sent to the submitter and to the PI of the proposal.

Some common problems are described below.
\begin{itemize}
\item BibTeX formats are not permitted within the ESOFORM package.
\item Figures without the corresponding .jpg or .pdf extensions will make the receiver hang.
\item The proposal reception system checks for {\bf the presence of text outside
 the argument fields of the ESOFORM macros} in the \LaTeX\ source of
the proposal, and rejects proposals in which such text is
found. If there is text outside the macros it appears as text or extra space above 
the ESO logo on top of the first page.  
This input {\bf must be commented out or relocated within the relevant macro} before the proposal is
submitted.

\end{itemize}

\subsubsection*{Submission Problems} 
The proposal submission
acknowledgment page normally appears within seconds of completion of a
a submission. However, during the last few hours before the Phase 1
deadline, the system may be slowed down by the high load, and the
acknowledgment process may take several minutes. Please {\bf be
  patient}: even though it may look like ``nothing is happening'', the
system most likely is actually busy processing a queue of
proposals. Please {\bf do not} abort your submission or make a new
attempt at submitting the same proposal as this would only increase the
load on the system and make it slower.

If you have not succeeded in completing your submission within 1~hour, 
please contact ESO via email  
at \href{mailto:esoform@eso.org}{\tt esoform@eso.org}.
{\bf Do not under any circumstance} attach your application (in any
format) to this email. {\bf Do not try to resubmit your application}
before receiving further instructions from ESO. {\bf Do not panic!}
Once you have uploaded the \LaTeX\ source of your proposal, your
attempt, and the time at which you initiated it, are recorded in the
ESO system, so anomalous delays due to the proposal reception
system will be duly identified. {\bf Be aware} that if you experience
difficulties due to the proposal reception system, you are most likely 
not the only user in this situation. As problem reports must be
handled sequentially it may take some time before you receive
feedback from ESO.

As mentioned above, the acknowledgment Web page providing the
identifier of your proposal is the official confirmation of its
successful submission. The subsequent email notification is only sent
to you as a secondary confirmation, and delay in its delivery should
not represent a concern. However, if you have not received
it within 24 hours of your submission, please report this anomaly to  
\href{mailto:esoform@eso.org}{\tt esoform@eso.org}.

A safe way to avoid submission problems (often related to heavy
system load  during the last few hours before the deadline) is to submit your
proposal early.  We strongly encourage you to send in your
applications and all attachments several days before the deadline.

\vspace*{1cm} {\bf All proposals and their attachments must reach the
  ESO servers via the WASP interface BEFORE 12:00 noon (Central
  European Summer Time) on the date of the deadline. Please note that  
  ESO cannot provide support beyond 11:00 on that day. 
  Responsibility for 
  verifying that ESO has correctly received, processed, and
  acknowledged your proposal before the proposal submission deadline
  rests entirely with you.  Revisions, corrections, and/or
  modifications will not be accepted after the deadline.}

\vspace*{1cm}

\begin{center}
\framebox{
\parbox[t]{135mm}{
  \noindent
  \begin{center}
    {IMPORTANT NOTICE}
  \end{center}
  Electronic proposal submission does not allow applicants to sign
  their proposals. Therefore ESO assumes that PI's take full
  responsibility for the contents of the proposal, in particular in
  regard to the names of co-investigators and the agreement to act
  according to the instructions for visiting astronomers, should
  observing time be granted.}}
\end{center}

\end{document}

### 2015-02-20 Paper0/ms.tex
initial commit

\documentclass[iop, numberedappendix]{emulateapj}
\usepackage{apjfonts}
%\documentclass[12pt,preprint]{aastex}

%\topmargin 0.6in

%\usepackage{deluxetable}
\usepackage{color}
\definecolor{citeRGB}{rgb}{0,0.1,0.7}
\usepackage[hyperfootnotes=true,naturalnames=true,letterpaper,pdfstartview=FitH,pdfpagemode=UseNone,colorlinks=true,citecolor=citeRGB]{hyperref}

%%%% The following code forces Hyperref to only link the year in a citation
%%%% Taken from http://tex.stackexchange.com/questions/23227/how-to-hyperlink-only-the-year-part-when-using-natbib-and-hyperref
% \usepackage{etoolbox}
% 
% \makeatletter
% 
% % Patch case where name and year have no delimiter
% \patchcmd{\NAT@citex}
%   {\@citea\NAT@hyper@{\NAT@nmfmt{\NAT@nm}\NAT@date}}
%   {\@citea\NAT@nmfmt{\NAT@nm}\NAT@hyper@{\NAT@date}}
%   {}% Do nothing if patch works
%   {}% Do nothing if patch fails
% 
% % Patch case where name and year have basic delimiter
% \patchcmd{\NAT@citex}
%   {\@citea\NAT@hyper@{%
%      \NAT@nmfmt{\NAT@nm}%
%      \hyper@natlinkbreak{\NAT@aysep\NAT@spacechar}{\@citeb\@extra@b@citeb}%
%      \NAT@date}}
%   {\@citea\NAT@nmfmt{\NAT@nm}%
%    \NAT@aysep\NAT@spacechar%
%    \NAT@hyper@{\NAT@date}}
%   {}% Do nothing if patch works
%   {}% Do nothing if patch fails
% 
% % Patch case where name and year are separated by a prenote
% \patchcmd{\NAT@citex}
%   {\@citea\NAT@hyper@{%
%      \NAT@nmfmt{\NAT@nm}%
%      \hyper@natlinkbreak{\NAT@spacechar\NAT@@open\if*#1*\else#1\NAT@spacechar\fi}%
%        {\@citeb\@extra@b@citeb}%
%      \NAT@date}}
%   {\@citea\NAT@nmfmt{\NAT@nm}%
%    \NAT@spacechar\NAT@@open\if*#1*\else#1\NAT@spacechar\fi%
%    \NAT@hyper@{\NAT@date}}
%   {}% Do nothing if patch works
%   {}% Do nothing if patch fails
% 
% \makeatother

%\gdef\zphot{$z_\mathrm{phot}$}
\gdef\HST{\textit{HST}}
\gdef\G141{\textit{G141}}
\gdef\F140W{\textit{F140W}}
\gdef\fluxcgs{\mathrm{erg\ s^{-1}\ cm^{-2}}}
\gdef\micront{$\mu$m}
\gdef\micronm{\mu\mathrm{m}}
\gdef\aXe{\texttt{aXe}}
\gdef\flux_radius{\textsc{flux\_radius}}
\gdef\epers{\textit{e}$^{-}$ s$^{-1}$}

\gdef\logOH{12 + \log \left(\mathrm{O/H}\right)}
\gdef\UmV{U-V}
\gdef\VmJ{V-J}
\gdef\SFRuvir{SFR_\mathrm{UV+IR}}
\gdef\peryr{\mathrm{yr}^{-1}}
\gdef\kms{km\,s$^{-1}$}
\gdef\mum{$\mu\mathrm{m}$}
\gdef\24mum{$24\,\mu\mathrm{m}$}
\gdef\arcsec{^{\prime\prime}}
\gdef\Jy{\mathrm{Jy}}
\gdef\uJy{\mu\mathrm{Jy}}
\gdef\LIR{L_\mathrm{IR}}
\gdef\LUV{L_\mathrm{UV}}
\gdef\4ang{4000\,\AA}
\gdef\msun{$M_{\odot}$}
\gdef\mathmsun{\ M_{\odot}}
\gdef\mathgyr{\ \textrm{Gyr}}
\gdef\mathmyr{\ \textrm{Myr}}
\gdef\pegase{\textsc{P\'{e}gase}}
\gdef\eazy{\texttt{Eazy}}
\gdef\galfit{\texttt{galfit}}
\gdef\emcee{\texttt{emcee}}
\gdef\interrest{\texttt{InterRest}}
\gdef\fast{\texttt{fast}}
\gdef\zCOSMOS{$z$\textsc{COSMOS}}
\gdef\zspec{z_\mathrm{spec}}
\gdef\zphot{z_\mathrm{phot}}
\gdef\zml{z_\mathrm{ML}}
\gdef\zprior{z_\mathrm{prior}}
\newcommand{\weblink}[1]{\href{#1}{#1}}
\gdef\UDFj{UDFj-39546284}
\gdef\compareID{UDF-40106456}
\gdef\Lya{\mathrm{Ly}\alpha}
\gdef\Halpha{\mathrm{H}\alpha}
\gdef\Hbeta{\mathrm{H}\beta}
\gdef\othree{[\ion{O}{3}]$\lambda$4959+5007}
\gdef\epers{electrons~s$^{-1}$}
\gdef\xxx{\textbf{xxx}}

\gdef\lineflux{$3.5\pm1.3\times10^{-18}\,\fluxcgs$}

\shortauthors{Brammer et al.}
\shorttitle{$K_s$-band imaging of the Frontier Fields}
%\slugcomment{Draft (\today)}
\slugcomment{Submitted to ApJSS}

%\citestyle{aa}

\begin{document}

\title{Ultra-deep $K_s$-band Imaging of the \textit{Hubble} Frontier Fields}

% \footnotetext[*]{Based on observations made with the NASA/ESA \textit{Hubble
% Space Telescope}, programs GO-11640, 12177 and 12328, obtained at the
% Space Telescope Science Institute, which is operated by the Association of
% Universities for Research in Astronomy, Inc., under NASA contract NAS
% 5-26555.}

\author{Gabriel B. Brammer\altaffilmark{1},
et~al.}
% Pieter G.\ van Dokkum\altaffilmark{2},
% Garth D.\ Illingworth\altaffilmark{4},
% Rychard J.\ Bouwens\altaffilmark{3},
% Ivo Labb\'e\altaffilmark{3}, 
% Marijn Franx\altaffilmark{3},
% Ivelina Momcheva\altaffilmark{2},
% Pascal A.\ Oesch\altaffilmark{4}}
% Rub\'en S\'anchez-Janssen\altaffilmark{1}, 
% Ivo Labb\'e\altaffilmark{2},
% Elisabete da Cunha\altaffilmark{3},
% Dawn K.\ Erb\altaffilmark{4},
% Marijn Franx\altaffilmark{2},
% Mattia Fumagalli\altaffilmark{2},
% Britt Lundgren\altaffilmark{5},
% Danilo Marchesini\altaffilmark{6},
% Ivelina Momcheva\altaffilmark{5},
% Erica Nelson\altaffilmark{5},
% Shannon Patel\altaffilmark{2},
% Ryan Quadri\altaffilmark{7},
% Hans-Walter Rix\altaffilmark{3},
% Rosalind E.\ Skelton\altaffilmark{5},
% Kasper B.\ Schmidt\altaffilmark{3},
% Arjen van der Wel\altaffilmark{3},
% Pieter G.\ van Dokkum\altaffilmark{5},
% David A.\ Wake\altaffilmark{5},
% Katherine E.\ Whitaker\altaffilmark{5}}

% Rachel Bezanson\altaffilmark{5},
% Xiaohui Fan\altaffilmark{6},
% Natascha F\"orster Schreiber\altaffilmark{8},
% Garth D.\ Illingworth\altaffilmark{9},
% Joel Leja\altaffilmark{5},
% Dan Magee\altaffilmark{9},
% Patrick McCarthy\altaffilmark{11},
% Adam Muzzin\altaffilmark{2},
% Ryan Quadri\altaffilmark{11},
% Charles C.\ Steidel\altaffilmark{12},
% Tomer Tal\altaffilmark{5},
% Anna Williams\altaffilmark{13}

\email{brammer@stsci.edu}

\altaffiltext{1}{ESA/AURA, Space Telescope Science Institute, 3700 San Martin Dr., Baltimore, MD 21218, USA}
% \altaffiltext{2}
% {Department of Astronomy, Yale University, New Haven, CT 06520, USA}
% \altaffiltext{3}
% {Leiden Observatory, Leiden University, NL-2300 RA Leiden, Netherlands}
% \altaffiltext{4}
% {UCO/Lick Observatory, University of California, Santa Cruz, CA 95064}
% \altaffiltext{3}
% {Max Planck Institute for Astronomy (MPIA), K\"onigstuhl 17,
% 69117, Heidelberg, Germany}
% \altaffiltext{4}
% {Department of Physics, University of Wisconsin-Milwaukee, P.O. Box 413,
% Milwaukee, WI 53201, USA}
% \altaffiltext{5}
% {Department of Astronomy, Yale University, New Haven, CT 06520, USA}
% \altaffiltext{6}
% {Physics and Astronomy Department, Tufts University, Robinson Hall,
% Room 257, Medford, MA, 02155, USA}
% \altaffiltext{7}
% {Carnegie Observatories, 813 Santa Barbara Street, Pasadena, CA 91101, USA}

\begin{abstract}

We present ultra-deep $K_s$-band imaging of the \textit{Hubble} Frontier Fields clusters Abell 2744, MACS-0416, MACS-0717 and MACS-1149 (\xxx\ include MOSFIRE???), which have recently been observed with large allocations of Directors' Discretionary Time with the \textit{HST} and \textit{Spitzer} telescopes covering $0.4 < \lambda < 4.5$\,\mum.  \textit{VLT}/HAWKI integrations of the first two fields reach 5-$\sigma$ limiting depths of $K_s=26.3$ (AB) and have seeing FWHM$\sim0\farcs4$.  Shorter MOSFIRE integrations of the latter two northern fields reach limiting depths $K_s=$\xxx\ with seeing FWHM=\xxx.  In all cases the $K_s$-band mosaics cover the primary cluster and parallel \textit{HST}/ACS+WFC3 fields, and the total area of the $K_s$-band coverage is \xxx\ arcmin$^2$.  The $K_s$-band at 2.2\,\mum\ crucially fills the gap between the reddest HST filter (1.6\,\mum\,$\sim H$) and the IRAC 3.6\,\mum\ passband.  While reaching the full depths of the space-based imaging is not currently feasible from the ground, the deep $K_s$-band images provide significant constraints on both the redshifts and the stellar-population properties of galaxies extending well below the characteristic stellar mass across most of the age of the universe, down to, and including, the redshifts of the targeted galaxy clusters ($z \lesssim 0.5$).  Reduced, aligned mosaics of the four survey fields are provided accompanying this manuscript.  Time has been awarded for similar HAWKI coverage of the final two Frontier Fields Abell 370 and Abell S1063, with observations expected by the end of 2015.

\end{abstract}

\keywords{galaxies: evolution --- galaxies: high-redshift --- surveys: xxx}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                %
%   Introduction
%                                                                %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{s:introduction}

Selecting in the reddest band possible.

K-band surveys:  ISAAC, NMBS, UltraVISTA, ...

HUGS

Frontier Fields

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                %
%   Observations
%                                                                %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Observations \& Data Reduction}
\label{s:observations}

Table~\ref{tab:summary} provides a brief summary of the survey fields and observation characteristics.  Additional details on the observing strategy and image reduction procedures are provided below.

\xxx\ currently a bit awkward since MOSFIRE reduction done quickly with Ivo's code and HAWK-I done in many additional steps (astrometry and photometry).

\subsection{Observation log}
\label{s:log}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                %
%  Summary table
%                                                                %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{deluxetable*}{lcclrrcc} 
\tabletypesize{\footnotesize}
\tablewidth{0pt} 
\tablecaption{Field summary \label{tab:summary}} 
\tablehead{ 
\colhead{Field} & \colhead{R.A.} & \colhead{Dec.} & \colhead{Instrument} & \colhead{Epoch} & \colhead{$t$} & \colhead{Depth} & \colhead{FWHM}}
\startdata 
Abell 2744 & 00:14:21.2 & -30:23:50 & \textit{VLT}/HAWKI     & 2013 Oct 24--2013 Dec 24 & 29.4 h & 25.8 & $0\farcs40$ \\ 
MACS-0416  & 04:16:08.9 & -24:04:28 & \textit{VLT}/HAWKI     & 2013 Oct 25--2014 Feb 23 & 25.5 h & 25.9 & $0\farcs38$ \\ 
MACS-0717  & 07:17:34.0 & +37:44:49 & \textit{Keck}/MOSFIRE  & 2015 Jan 26--27                 & \xxx & 25.0 & $0\farcs43$ \\ 
MACS-1149  & 11:49:36.3 & +22:23:58 & \textit{Keck}/MOSFIRE  & 2015 Feb 24                    & \xxx & \xxx & \xxx
\enddata 
%\vspace{-0.8cm} 
\tablecomments{Depth is defined as the 5$\sigma$ limiting magnitude for point sources, measured in $D=0\farcs5$ apertures (\S\ref{s:depth}).} 
\end{deluxetable*}
%\vspace{1cm}

$K_s$-band observations of the Abell 2744 and MACS-0416 fields were obtained between 2013 Oct and 2014 Feb with the High Acuity Wide-field K-band Imager \citep[HAWK-I;][]{hawki, hawki2} on the 8.2\,m UT4 telescope at the ESO \textit{Very Large Telescope} (ESO program \mbox{092.A-0472}).  HAWK-I is composed of four chips with 2048$\times$2048 $0\farcs106$ pixels, and the full $7\farcm5\times7\farcm5$ HAWKI field-of-view is perfectly suited to cover both the primary and parallel \textit{HST} ACS optical and WFC3 IR fields simultaneously in a single pointing (see Fig.\ref{fig:layout}).

The HAWK-I observations were divided into individual service mode Observation Blocks (OBs); typically one or two OBs of a given field were observed per night as conditions allowed, and occasionally both fields were observed on the same night.  The execution of each OB lasted either 60 or 90 minutes, depending on how many exposures were obtained in the block.  The exposures were roughly one minute each, with multiple coadds of shorter reads making up the exposure ($\mathrm{NDIT}\times\mathrm{DIT} = 4\times15$\,s~=~60\,s for Abell 2744 and 7$\times$8\,s~=~56\,s for MACS-0416\footnote{Longer DITs are preferable to reduce instrument overheads; the shorter DITs on the MACS-0416 field were done do accommodate \textit{VLT} service mode restrictions on bright source saturation, which have been eased since P93.}).  The telescope was offset with random dithers between each exposure to facilitate sky subtraction;  observations taken before 2013 Nov 4 were dithered within a 20$^{\prime\prime}$ box, which was subsequently increased to 40$^{\prime\prime}$ to improve sky subtraction in the cluster cores crowded with bright galaxies and intra-cluster light.  The total on-sky integration times for the Abell 2744 and MACS-0416 fields are 29.4 and 25.5 hours, respectively.

Imaging observations of the MACS-0717 and MACS-1149 fields in the $K_s$ filter were obtained on 2015 Jan 26/27 and 2015 Feb 24, respectively, with the Multi-Object Spectrometer For Infra-Red Exploration \citep[MOSFIRE;][]{mosfire} on the 10\,m \textit{Keck I} telescope.  The MOSFIRE detector consists of $2048\times2048$ $0\farcs1798$ pixels; the resulting $6\farcm1\times6\farcm1$ field-of-view is unfortunately slightly too small to cover both \textit{HST} fields simultaneously and therefore requires two pointings to cover the entire deep \textit{HST} Frontier Fields imaging area (Fig.~\ref{fig:layout}).  The 6$\times$7\,s~=~42\,s MOSFIRE exposures were obtained in a $3\times3$ ``Box9'' dither pattern spaced roughly 40$^{\prime\prime}$ between offset positions.  \xxx\ time on cluster, parallel field etc \xxx.

\subsection{Image processing}
\label{s:processing}

The HAWK-I and MOSFIRE observations were reduced with a pipeline that has been developed for previous surveys with the NEWFIRM \citep[NMBS;][]{whitaker:nmbs} and FOURSTAR \citep[zFOURGE;][, Labb\'e in prep]{spitler:12} infrared imaging instruments.  Treating each detector individually, the pipeline is easily modified for the different instrument configurations of the four HAWK-I and single MOSFIRE chips.  The primary task of the pipeline is removing the bright, time-variable sky background from the individual exposures, which is often some $10^4$ times brighter than the distant galaxies of interest in the field.  With such a bright background, we first determine an empirical ``sky flat'' that is a median of all of the \textit{science} exposures in a HAWK-I OB or MOSFIRE group, after rejecting the brightest 12 exposures at each pixel position to remove the contribution of bright objects.  We find these empirical flats to be preferable to external twilight or dome flats given the difficulty of obtaining a truly flat illumination pattern over such large detector fields-of-view.

After dividing by the flat, the background of each exposure is determined in a first pass from the simple median of the four exposures that came both immediately before and after it.  The background-subtracted exposures are combined into a mosaic from which objects are detected.  The final refined background of each exposure is determined in a second ``mask pass'' from a median again from the four exposures before and after it but now aggressively masking all pixels that contain flux from the detected objects.

\subsection{Photometric calibration}
\label{s:calib}

For the HAWK-I observations, a single 90-minute OB was obtained in each of the Abell 2744 and MACS-0416 fields requiring photometric transparency conditions, and these OBs were followed immediately by an observation of a photometric standard star at a similar airmass with single exposures on each of the four chips.  The standard star exposures were processed with empirical flats derived from the science exposures as described above, and the observed fluxes of the standard stars yield absolute photometric zeropoints for each chip.  \xxx\ We determine additional relative zeropoint offsets between the four HAWK-I chips by comparing the photometry of sources in the overlap regions between them.  \xxx\ \textit{Not sure why these chip corrections are needed beyond the standard star, but they are consistent between the two fields and are needed to improve the agreement in relative depths between the chips.} Correction factors were then computed to scale the OBs obtained at varying transparency levels to the calibrated photometric OB; the additional service mode OBs were obtained under generally satisfactory weather conditions and the scale factors typically differ from unity by only a few percent.

The 2-Micron All Sky Survey \citep[2MASS;][]{2mass} provides an additional check on the photometric calibration, though the comparison is limited due to relatively little brightness overlap between the faint end of reliable 2MASS photometry and the bright end where stars are in the linear regime of the detectors on the 8--10\,m telescopes.  For the MOSFIRE observations obtained over just one or two nights, we use the 2MASS overlap directly to determine the absolute photometric calibration.  A comparison of the observed photometry in the four survey fields to the 2MASS catalog magnitudes is shown in Fig.~\ref{fig:2mass}.  The stellar photometry on the deep $K_s$-band mosaics is measured within $1\arcsec$ apertures corrected to infinity with the curves of growth described below.  The error-bars on the points in Fig.~\ref{fig:2mass} come from the 2MASS catalog; the photometric uncertainties are negligible for such bright stars in the deep images ($\mathrm{S/N}>100$). We estimate that the systematic uncertainty of the absolute photometric calibration is $\sigma_\mathrm{sys}\lesssim 0.05$ mag.

\subsection{Astrometric alignment}
\label{s:astrometry}

We refine the relative and absolute astrometry of the $K_s$-band images using the SExtractor \citep{bertin:96} and SCAMP \citep{scamp} software tools. Reference astrometric catalogs are generated from \textit{HST} images when available (chips \#1 and \#4 of both HAWK-I cluster fields) and public \textit{Subaru} Suprime-Cam $r_c$-band images otherwise\footnote{http://archive.stsci.edu/pub/hlsp/clash/macs0416/data/subaru/}.  Next, object catalogs are extracted for each background-subtracted exposure and transformations to the reference frame are computed with SCAMP.  All of the exposures for each HAWK-I chip are processed simultaneously as a SCAMP ``instrument'', ensuring a uniform geometric distortion correction for each chip \citep[see also][for a detailed discussion of the HAWK-I geometric distortion]{libralato:14}.

For the MOSFIRE observations, we adopt a more simplified approach and refine the shifts specified by the World Coordinate System header keywords by cross-correlating each exposure in a group with an arbitrarily-defined reference exposure (\xxx\ this is what Ivo's pipeline does internally.  I found a bit better final image quality of the HAWK-I images using the SCAMP approach, and the HAWK-I distortions are non-trivial).  For the absolute astrometry of the MOSFIRE mosaic, we align the $K_s$-band image to the Suprime-Cam image\footnote{http://archive.stsci.edu/pub/hlsp/clash/macs0717/data/subaru/} allowing for a linear shift, rotation and scale using the \texttt{iraf.geomap} tool.

\xxx\ Add some sort of diagnostic figure showing the astrometric comparison between Ks and, say, ACS F814W and get a number to quote for astrometric precision

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%   Exposure Map figure
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\plotone{hawkiff_layout.pdf} % Figures/exposure_map.pdf
\caption{Layout of the Frontier Fields $K_s$-band mosaics.  The positions of the \textit{HST} cluster and parallel fields are shown in the blue (ACS optical) and red (WFC3 IR) polygons.  The area covered by the \textit{Spitzer}/IRAC 3.6 and 4.5\,\mum\ channels is shown in orange.  The bottom panels show the exposure maps, where the four HAWK-I chips and two MOSFIRE pointings can be seen.\label{fig:layout}}  
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%   Exposure Map figure
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%   2MASS comparison
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!ht]
\epsscale{1.18}
\plotone{2mass_phot.pdf} % Figures/exposure_map.pdf
\caption{Comparison of bright-star photometry to the 2MASS public catalog.  The Abell 2744 and MACS 0416 fields at high galactic latitude have few stars for cross calibration, and the photometric calibration is determined from observations of standard stars on nights with photometric transparency conditions.  At relatively low galactic latitude ($b=+21$), the MACS 0717 field has many bright stars, and the photometric calibration in that field is defined from the 2MASS comparison (hence, $\Delta_\mathrm{2mass}\equiv0$, as indicated for that field).  We estimate the overall systematic uncertainty in the photometric calibration to be $\lesssim0.05~\mathrm{mag}$.\label{fig:2mass}}  
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%   Image quality figure
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!ht]
\epsscale{1.18}
\plotone{curve_of_growth.pdf} % Figures/exposure_map.pdf
\caption{Curves of growth of stellar profiles in the Frontier Fields $K_s$-band mosaics.  The profiles of stars are well-fit by Moffat profiles with $\mathrm{FWHM}\sim0\farcs4$ and $\beta\sim2$, which have substantially more flux at radii $r>\mathrm{FWHM}$ than Gaussian profiles with the same half-widths (orange curves).  These extended profiles appear to be characteristic of deep $K_s$-band images; the stellar profiles from the HUGS GOODS-South D2 HAWK-I $K_s$ image \citep{hugs} (lower right panel), created by a different team with a different reduction pipeline, are nearly identical to those of the Frontier Fields mosaics presented here. \xxx\ I only show the parallel field region of MACS0717, which has significantly better image quality from the Jan. run.  You can see a few of the broader profiles in the figure; there are many more of these if you include the whole cluster area.  \label{fig:cog}}  
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%   Depth figure
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!ht]
\epsscale{1.18}
\plotone{aper_depth.pdf} % Figures/exposure_map.pdf
\caption{\textit{Top panel:} Noise in empty apertures as a function of aperture diameter, normalized to the smallest $D=0\farcs2$ aperture.  The dashed and dotted lines show the expectations for uncorrelated and perfectly-correlated noise within the aperture.  \textit{Bottom panel:} The product of the inverse curves-of-growth (Fig.\ref{fig:cog} and the empty aperture noise gives the point-source sensitivity as a function of aperture size.  The S/N is maximized with an aperture just larger than the FWHM.  For reference, the colored dashed lines in the bottom panel indicate the depth computed assuming uncorrelated noise within the apertures ($\sigma \propto D$) and the per-pixel variance reported in the weight maps.)
\label{fig:depth}}  
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                %
%   Mosaic properties
%                                                                %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{$K_s$-band Mosaics}
\label{s:mosaic}

We generate stacks of each OB using the SWarp tool \citep{swarp}, which takes into account the alignment to the reference catalog.  Final mosaics of each field with $0\farcs100$ pixels are generated from the OB stacks weighting the input images with weeights
\[
w^{-1} = F^2\cdot\sigma^2 \cdot \mathrm{FWHM}^\alpha, 
\]
where $F$ is the factor to scale the OB to the photometric system ($F\equiv1$ for the OB taken under photometric contidions, $F\gtrsim1$ otherwise), $\sigma^2$ is the variance map of each OB stack determined by SWarp ($\sigma^2$ accounts for both exposure time and sky background variations between OBs), FWHM is the full width at half maximum of stars identified in the field, and the parameter $\alpha$ allows for optimizing the image quality of the final mosaic.  Increasing $\alpha$ puts larger weight on the best image-quality OBs while effectively ignoring data with poorer seeing; we adopt $\alpha=2$ for a compromise between the image quality and effective exposure time of the final mosaic.  Finally, we subtract a cell-based background from the final mosaics using an algorithm based on that used by SExtractor and SWarp but that provides more aggressive masking of flux in the outer isophotes of bright galaxies. The mosaics and inverse-variance maps are shown in Fig.~\ref{fig:layout}, along with the position of the deep Frontier Fields \textit{HST} imaging fields.

\xxx Right now I just swarp the MOSFIRE groups together without any weighting

The final image and inverse-variance mosaics are provided as an ESO Phase 3 data product \xxx\ URL, as well as attached as a digital product to this manuscript (\xxx\ I think ApJSS can and wants to do this).  The images are scaled to an AB magnitude zeropoint ZP=26.0.  In the sections below we describe the characteristics of the mosaics in more detail.

\subsection{Image quality}
\label{s:fwhm}

Fig.~\ref{fig:cog} shows the curves of growth for stars identified by the tight relationship between their brightness and half-light radii \citep[see, e.g., Fig.~13 of][]{skelton:14}.  The image quality of the $K_s$-band mosaics is excellent, with stellar FWHM $0\farcs40$ and $0\farcs38$ in the Abell 2744 and MACS 0416 fields, respectively, thanks, in large part, to the execution of the HAWK-I observations in service mode ensuring optimal and uniform image quality across the many hours of integration on the survey fields.  Despite the compact cores of the stellar PSFs, the curves of growth shown in Fig.~\ref{fig:cog} deviate from Gaussian profiles with significantly higher flux in the outer wings, with a shape more consistent with a Moffat profile \citep{trujillo:01}  with $\beta\sim2$.  Also shown in Fig.~\ref{fig:cog} are stellar curves of growth measured from the HAWK-I UDS and GOODS Survey \citep[HUGS;][]{hugs}, which are nearly identical in both FWHM and profile shape.  We caution that these extended profiles will yield significantly shallower final image depths than would be predicted with Exposure Time Calculators that may assume Gaussian profiles. \xxx\ not clear what HAWK-I ETC assumes, the description just says ``FWHM of the seeing disk'' \xxx.  Nevertheless, the deep $K_s$ mosaics have excellent image quality that is much closer to the resolution of the \textit{HST} IR imaging than the redder \textit{Spitzer} IRAC bands.

\xxx\ add description of MOSFIRE clusters.  IQ not uniform across the MACS0717 field.

\subsection{Depth}
\label{s:depth}

The inverse-variance weight maps shown in Fig.~\ref{fig:layout} provide a formal measurement of the pixel-to-pixel variance.  However, for aperture photometry measurements this variance will yield an underestimate of the true photometric uncertainty due to correlated noise of neighboring pixels that results from pixel resampling during the image reduction process.  For uncorrelated noise between pixels in an aperture, the variance within the aperture scales with the number of pixels in the aperture, $N_\mathrm{pix}$.  For perfect correlation within the aperture, however, the variance scales as $N_\mathrm{pix}^2$, and the properties of the $K_s$ mosaics will lie somewhere in between these two extremes.

Following \cite{labbe:03} \citep[see also][]{quadri:musyc, whitaker:nmbs, skelton:14}, we characterize the noise properties of the $K_s$ mosaics by measuring the flux within circular apertures placed randomly throughout empty parts of the images.  The top panel of Fig.~\ref{fig:depth} shows the standard deviation of these empty-aperture fluxes, $\sigma_\mathrm{D}$, as a function of aperture diameter, which should be used in place of simple scalings of the inverse-variance maps to represent the true uncertainties within photometric apertures.

The product of the inverse curves of growth shown in Fig.~\ref{fig:cog} (i.e., aperture corrections) and the noise curves in the top panel of Fig.~\ref{fig:depth} defines an optimal photometric aperture size where this product is minimized \citep{whitaker:nmbs}.  For our $K_s$-band mosaics, this optimal aperture has $D\sim0\farcs5$, just larger than the FWHM of point-source profiles.  The $5\,\sigma$ limiting magnitudes (AB) within $0\farcs5$ diameter apertures are 25.8, 25.9, and 25.0 in the Abell 2744, MACS-0416, and MACS-0717 fields, respectively.  We note that the depths of the FF HAWK-I mosaics are nearly identical to those of the two deepest HUGS pointings in the GOODS-South Field (i.e., GOODS-South ``D1'' and ``D2'').  The image depths reported by \cite{hugs} assume ideal noise properties that can be compared to the dashed curves in Fig.~\ref{fig:depth}.  We have analyzed the HUGS GOODS-South ``D2'' image with the same empty-aperture technique as was done for the FF images, and the sensitivities are fully consistent with those reported by \cite{hugs} when accounting for the different definitions.  Along with the two HUGS deep HAWK-I fields, our images covering the Frontier Fields represent the deepest $K_s$-band images ever obtained and will provide a crucial complement to the deep \textit{HST} and \textit{Spitzer} imaging of the same fields.

\xxx\ say something about distant extended sources, perhaps scaling from Arjen's size distributions?

\subsection{Number counts}
\label{s:number}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                %
%   Constraints
%                                                                %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Photometric constraints from the $K_s$ band}
\label{s:constraints}

\subsection{Photometric redshifts}
\label{s:photoz}

\subsection{SED modeling}
\label{s:sed}

Balmer break, UV slope

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                %
%   Discussion and Summary
%                                                                %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion and Conclusions}
\label{s:discussion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%   End things
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{-0.3cm} %% emulateapj
\acknowledgements

\noindent We gratefully acknowledge funding support from the STScI Director's Discretionary Research Fund.  \xxx\ acknowledge NASA/Keck time + Mauna Kea \xxx. This research made use of Astropy, a community-developed core Python package for Astronomy \citep{astropy}, and open-source Python modules Numpy, Scipy, and Matplotlib.

% \noindent We gratefully acknowledge funding support from the Marie Curie Actions of the European Commission (FP7-COFUND) and STScI
% grant GO-12177.

{\it Facilities:} \facility{VLT:Yepun (HAWKI)}, \facility{Keck:I (MOSFIRE)} % \facility{Hubble Space Telescope (WFC3)}

\bibliographystyle{apj_url}
\bibliography{ms}

\end{document}

### 2015-02-20 Paper0/ms.tex
Add cluster redshift to table

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{deluxetable*}{lcclrrcc} 
+\begin{deluxetable*}{lrcclrrcc} 
 \tabletypesize{\footnotesize}
@@
 \tablehead{ 
-\colhead{Field} & \colhead{R.A.} & \colhead{Dec.} & \colhead{Instrument} & \colhead{Epoch} & \colhead{$t$} & \colhead{Depth} & \colhead{FWHM}}
+\colhead{Field} & \colhead{$z$} & \colhead{R.A.} & \colhead{Dec.} & \colhead{Instrument} & \colhead{Epoch} & \colhead{$t$} & \colhead{Depth} & \colhead{FWHM}}
 \startdata 
-Abell 2744 & 00:14:21.2 & -30:23:50 & \textit{VLT}/HAWKI     & 2013 Oct 24--2013 Dec 24 & 29.4 h & 25.8 & $0\farcs40$ \\ 
-MACS-0416  & 04:16:08.9 & -24:04:28 & \textit{VLT}/HAWKI     & 2013 Oct 25--2014 Feb 23 & 25.5 h & 25.9 & $0\farcs38$ \\ 
-MACS-0717  & 07:17:34.0 & +37:44:49 & \textit{Keck}/MOSFIRE  & 2015 Jan 26--27                 & \xxx & 25.0 & $0\farcs43$ \\ 
-MACS-1149  & 11:49:36.3 & +22:23:58 & \textit{Keck}/MOSFIRE  & 2015 Feb 24                    & \xxx & \xxx & \xxx
+Abell 2744 & 0.308 & 00:14:21.2 & -30:23:50 & \textit{VLT}/HAWKI     & 2013 Oct 24--2013 Dec 24 & 29.4 h & 25.8 & $0\farcs40$ \\ 
+MACS-0416  & 0.396 & 04:16:08.9 & -24:04:28 & \textit{VLT}/HAWKI     & 2013 Oct 25--2014 Feb 23 & 25.5 h & 25.9 & $0\farcs38$ \\ 
+MACS-0717  & 0.545 & 07:17:34.0 & +37:44:49 & \textit{Keck}/MOSFIRE  & 2015 Jan 26--27                 & \xxx & 25.0 & $0\farcs43$ \\ 
+MACS-1149  & 0.543 & 11:49:36.3 & +22:23:58 & \textit{Keck}/MOSFIRE  & 2015 Feb 24                    & \xxx & \xxx & \xxx
 \enddata 

### 2015-02-21 Paper0/ms.tex
add number counts

--- 2015-02-20 Paper0/ms.tex
@@
 
-K-band surveys:  ISAAC, NMBS, UltraVISTA, ...
+K-band surveys:  ISAAC \citep{labbe:03}, NMBS, UltraVISTA, ...
 
-HUGS
+Subaru Super Deep Field (AO) \cite{minowa:05}.
+
+HUGS \citep{hugs}.
 
@@
 The positions of the \textit{HST} cluster and parallel fields are shown in the blue (ACS optical) and red (WFC3 IR) polygons.
-The area covered by the \textit{Spitzer}/IRAC 3.6 and 4.5\,\mum\ channels is shown in orange.
-The bottom panels show the exposure maps, where the four HAWK-I chips and two MOSFIRE pointings can be seen.\label{fig:layout}}
+The light blue polygons in the MACS-0717 field show additional wide-field ACS imaging coverage from  programs GO-9722 and GO-10420 \citep[PI: Ebeling;][]{ma:11}.
+The area covered by deep imaging in the \textit{Spitzer}/IRAC 3.6 and 4.5\,\mum\ channels is shown in orange.
+The bottom panels show the mosaic weight maps, where the four HAWK-I chips and two MOSFIRE pointings can be seen.\label{fig:layout}}
 \end{figure*}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%   Exposure Map figure
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 We have analyzed the HUGS GOODS-South ``D2'' image with the same empty-aperture technique as was done for the FF images, and the sensitivities are fully consistent with those reported by \cite{hugs} when accounting for the different definitions.
-Along with the two HUGS deep HAWK-I fields, our images covering the Frontier Fields represent the deepest $K_s$-band images ever obtained and will provide a crucial complement to the deep \textit{HST} and \textit{Spitzer} imaging of the same fields.
+The HAWK-I images presented here are among the deepest $K_s$-band images ever obtained and will provide a crucial complement to the deep \textit{HST} and \textit{Spitzer} imaging of the Frontier Fields.
 
@@
 \label{s:number}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%  Number counts figure
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure*}
+\epsscale{1.18}
+\plotone{number_counts_all.pdf} % Figures/exposure_map.pdf
+\caption{Number counts.\label{fig:counts}}  
+\end{figure*}
 

### 2015-02-21 Paper0/ms.tex

@@
 \end{document}
+=======
+\documentclass[iop, numberedappendix]{emulateapj}
+\usepackage{apjfonts}
+%\documentclass[12pt,preprint]{aastex}
+
+%\topmargin 0.6in
+
+%\usepackage{deluxetable}
+\usepackage{color}
+\definecolor{citeRGB}{rgb}{0,0.1,0.7}
+\usepackage[hyperfootnotes=true,naturalnames=true,letterpaper,pdfstartview=FitH,pdfpagemode=UseNone,colorlinks=true,citecolor=citeRGB]{hyperref}
+
+%%%% The following code forces Hyperref to only link the year in a citation
+%%%% Taken from http://tex.stackexchange.com/questions/23227/how-to-hyperlink-only-the-year-part-when-using-natbib-and-hyperref
+% \usepackage{etoolbox}
+% 
+% \makeatletter
+% 
+% % Patch case where name and year have no delimiter
+% \patchcmd{\NAT@citex}
+%   {\@citea\NAT@hyper@{\NAT@nmfmt{\NAT@nm}\NAT@date}}
+%   {\@citea\NAT@nmfmt{\NAT@nm}\NAT@hyper@{\NAT@date}}
+%   {}% Do nothing if patch works
+%   {}% Do nothing if patch fails
+% 
+% % Patch case where name and year have basic delimiter
+% \patchcmd{\NAT@citex}
+%   {\@citea\NAT@hyper@{%
+%      \NAT@nmfmt{\NAT@nm}%
+%      \hyper@natlinkbreak{\NAT@aysep\NAT@spacechar}{\@citeb\@extra@b@citeb}%
+%      \NAT@date}}
+%   {\@citea\NAT@nmfmt{\NAT@nm}%
+%    \NAT@aysep\NAT@spacechar%
+%    \NAT@hyper@{\NAT@date}}
+%   {}% Do nothing if patch works
+%   {}% Do nothing if patch fails
+% 
+% % Patch case where name and year are separated by a prenote
+% \patchcmd{\NAT@citex}
+%   {\@citea\NAT@hyper@{%
+%      \NAT@nmfmt{\NAT@nm}%
+%      \hyper@natlinkbreak{\NAT@spacechar\NAT@@open\if*#1*\else#1\NAT@spacechar\fi}%
+%        {\@citeb\@extra@b@citeb}%
+%      \NAT@date}}
+%   {\@citea\NAT@nmfmt{\NAT@nm}%
+%    \NAT@spacechar\NAT@@open\if*#1*\else#1\NAT@spacechar\fi%
+%    \NAT@hyper@{\NAT@date}}
+%   {}% Do nothing if patch works
+%   {}% Do nothing if patch fails
+% 
+% \makeatother
+
+%\gdef\zphot{$z_\mathrm{phot}$}
+\gdef\HST{\textit{HST}}
+\gdef\G141{\textit{G141}}
+\gdef\F140W{\textit{F140W}}
+\gdef\fluxcgs{\mathrm{erg\ s^{-1}\ cm^{-2}}}
+\gdef\micront{$\mu$m}
+\gdef\micronm{\mu\mathrm{m}}
+\gdef\aXe{\texttt{aXe}}
+\gdef\flux_radius{\textsc{flux\_radius}}
+\gdef\epers{\textit{e}$^{-}$ s$^{-1}$}
+
+\gdef\logOH{12 + \log \left(\mathrm{O/H}\right)}
+\gdef\UmV{U-V}
+\gdef\VmJ{V-J}
+\gdef\SFRuvir{SFR_\mathrm{UV+IR}}
+\gdef\peryr{\mathrm{yr}^{-1}}
+\gdef\kms{km\,s$^{-1}$}
+\gdef\mum{$\mu\mathrm{m}$}
+\gdef\24mum{$24\,\mu\mathrm{m}$}
+\gdef\arcsec{^{\prime\prime}}
+\gdef\Jy{\mathrm{Jy}}
+\gdef\uJy{\mu\mathrm{Jy}}
+\gdef\LIR{L_\mathrm{IR}}
+\gdef\LUV{L_\mathrm{UV}}
+\gdef\4ang{4000\,\AA}
+\gdef\msun{$M_{\odot}$}
+\gdef\mathmsun{\ M_{\odot}}
+\gdef\mathgyr{\ \textrm{Gyr}}
+\gdef\mathmyr{\ \textrm{Myr}}
+\gdef\pegase{\textsc{P\'{e}gase}}
+\gdef\eazy{\texttt{Eazy}}
+\gdef\galfit{\texttt{galfit}}
+\gdef\emcee{\texttt{emcee}}
+\gdef\interrest{\texttt{InterRest}}
+\gdef\fast{\texttt{fast}}
+\gdef\zCOSMOS{$z$\textsc{COSMOS}}
+\gdef\zspec{z_\mathrm{spec}}
+\gdef\zphot{z_\mathrm{phot}}
+\gdef\zml{z_\mathrm{ML}}
+\gdef\zprior{z_\mathrm{prior}}
+\newcommand{\weblink}[1]{\href{#1}{#1}}
+\gdef\UDFj{UDFj-39546284}
+\gdef\compareID{UDF-40106456}
+\gdef\Lya{\mathrm{Ly}\alpha}
+\gdef\Halpha{\mathrm{H}\alpha}
+\gdef\Hbeta{\mathrm{H}\beta}
+\gdef\othree{[\ion{O}{3}]$\lambda$4959+5007}
+\gdef\epers{electrons~s$^{-1}$}
+\gdef\xxx{\textbf{xxx}}
+
+\gdef\lineflux{$3.5\pm1.3\times10^{-18}\,\fluxcgs$}
+
+\shortauthors{Brammer et al.}
+\shorttitle{$K_s$-band imaging of the Frontier Fields}
+%\slugcomment{Draft (\today)}
+\slugcomment{Submitted to ApJSS}
+
+%\citestyle{aa}
+
+\begin{document}
+
+\title{Ultra-deep $K_s$-band Imaging of the \textit{Hubble} Frontier Fields}
+
+% \footnotetext[*]{Based on observations made with the NASA/ESA \textit{Hubble
+% Space Telescope}, programs GO-11640, 12177 and 12328, obtained at the
+% Space Telescope Science Institute, which is operated by the Association of
+% Universities for Research in Astronomy, Inc., under NASA contract NAS
+% 5-26555.}
+
+\author{Gabriel B. Brammer\altaffilmark{1},
+et~al.}
+% Pieter G.\ van Dokkum\altaffilmark{2},
+% Garth D.\ Illingworth\altaffilmark{4},
+% Rychard J.\ Bouwens\altaffilmark{3},
+% Ivo Labb\'e\altaffilmark{3}, 
+% Marijn Franx\altaffilmark{3},
+% Ivelina Momcheva\altaffilmark{2},
+% Pascal A.\ Oesch\altaffilmark{4}}
+% Rub\'en S\'anchez-Janssen\altaffilmark{1}, 
+% Ivo Labb\'e\altaffilmark{2},
+% Elisabete da Cunha\altaffilmark{3},
+% Dawn K.\ Erb\altaffilmark{4},
+% Marijn Franx\altaffilmark{2},
+% Mattia Fumagalli\altaffilmark{2},
+% Britt Lundgren\altaffilmark{5},
+% Danilo Marchesini\altaffilmark{6},
+% Ivelina Momcheva\altaffilmark{5},
+% Erica Nelson\altaffilmark{5},
+% Shannon Patel\altaffilmark{2},
+% Ryan Quadri\altaffilmark{7},
+% Hans-Walter Rix\altaffilmark{3},
+% Rosalind E.\ Skelton\altaffilmark{5},
+% Kasper B.\ Schmidt\altaffilmark{3},
+% Arjen van der Wel\altaffilmark{3},
+% Pieter G.\ van Dokkum\altaffilmark{5},
+% David A.\ Wake\altaffilmark{5},
+% Katherine E.\ Whitaker\altaffilmark{5}}
+
+% Rachel Bezanson\altaffilmark{5},
+% Xiaohui Fan\altaffilmark{6},
+% Natascha F\"orster Schreiber\altaffilmark{8},
+% Garth D.\ Illingworth\altaffilmark{9},
+% Joel Leja\altaffilmark{5},
+% Dan Magee\altaffilmark{9},
+% Patrick McCarthy\altaffilmark{11},
+% Adam Muzzin\altaffilmark{2},
+% Ryan Quadri\altaffilmark{11},
+% Charles C.\ Steidel\altaffilmark{12},
+% Tomer Tal\altaffilmark{5},
+% Anna Williams\altaffilmark{13}
+
+\email{brammer@stsci.edu}
+
+\altaffiltext{1}{ESA/AURA, Space Telescope Science Institute, 3700 San Martin Dr., Baltimore, MD 21218, USA}
+% \altaffiltext{2}
+% {Department of Astronomy, Yale University, New Haven, CT 06520, USA}
+% \altaffiltext{3}
+% {Leiden Observatory, Leiden University, NL-2300 RA Leiden, Netherlands}
+% \altaffiltext{4}
+% {UCO/Lick Observatory, University of California, Santa Cruz, CA 95064}
+% \altaffiltext{3}
+% {Max Planck Institute for Astronomy (MPIA), K\"onigstuhl 17,
+% 69117, Heidelberg, Germany}
+% \altaffiltext{4}
+% {Department of Physics, University of Wisconsin-Milwaukee, P.O. Box 413,
+% Milwaukee, WI 53201, USA}
+% \altaffiltext{5}
+% {Department of Astronomy, Yale University, New Haven, CT 06520, USA}
+% \altaffiltext{6}
+% {Physics and Astronomy Department, Tufts University, Robinson Hall,
+% Room 257, Medford, MA, 02155, USA}
+% \altaffiltext{7}
+% {Carnegie Observatories, 813 Santa Barbara Street, Pasadena, CA 91101, USA}
+
+\begin{abstract}
+
+We present ultra-deep $K_s$-band imaging of the \textit{Hubble} Frontier Fields clusters Abell 2744, MACS-0416, MACS-0717 and MACS-1149 (\xxx\ include MOSFIRE???), which have recently been observed with large allocations of Directors' Discretionary Time with the \textit{HST} and \textit{Spitzer} telescopes covering $0.4 < \lambda < 4.5$\,\mum.
+\textit{VLT}/HAWKI integrations of the first two fields reach 5-$\sigma$ limiting depths of $K_s=26.3$ (AB) and have seeing FWHM$\sim0\farcs4$.
+Shorter MOSFIRE integrations of the latter two northern fields reach limiting depths $K_s=$\xxx\ with seeing FWHM=\xxx.
+In all cases the $K_s$-band mosaics cover the primary cluster and parallel \textit{HST}/ACS+WFC3 fields, and the total area of the $K_s$-band coverage is \xxx\ arcmin$^2$.
+The $K_s$-band at 2.2\,\mum\ crucially fills the gap between the reddest HST filter (1.6\,\mum\,$\sim H$) and the IRAC 3.6\,\mum\ passband.
+While reaching the full depths of the space-based imaging is not currently feasible from the ground, the deep $K_s$-band images provide significant constraints on both the redshifts and the stellar-population properties of galaxies extending well below the characteristic stellar mass across most of the age of the universe, down to, and including, the redshifts of the targeted galaxy clusters ($z \lesssim 0.5$).
+Reduced, aligned mosaics of the four survey fields are provided accompanying this manuscript.
+Time has been awarded for similar HAWKI coverage of the final two Frontier Fields Abell 370 and Abell S1063, with observations expected by the end of 2015.
+
+\end{abstract}
+
+\keywords{galaxies: evolution --- galaxies: high-redshift --- surveys: xxx}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%                                                                %
+%   Introduction
+%                                                                %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Introduction}
+\label{s:introduction}
+
+Selecting in the reddest band possible.
+
+K-band surveys:  ISAAC, NMBS, UltraVISTA, ...
+
+HUGS
+
+Frontier Fields
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%                                                                %
+%   Observations
+%                                                                %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Observations \& Data Reduction}
+\label{s:observations}
+
+Table~\ref{tab:summary} provides a brief summary of the survey fields and observation characteristics.  Additional details on the observing strategy and image reduction procedures are provided below.
+
+\xxx\ currently a bit awkward since MOSFIRE reduction done quickly with Ivo's code and HAWK-I done in many additional steps (astrometry and photometry).
+
+\subsection{Observation log}
+\label{s:log}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%                                                                %
+%  Summary table
+%                                                                %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{deluxetable*}{lrcclrrcc} 
+\tabletypesize{\footnotesize}
+\tablewidth{0pt} 
+\tablecaption{Field summary \label{tab:summary}} 
+\tablehead{ 
+\colhead{Field} & \colhead{$z$} & \colhead{R.A.} & \colhead{Dec.} & \colhead{Instrument} & \colhead{Epoch} & \colhead{$t$} & \colhead{Depth} & \colhead{FWHM}}
+\startdata 
+Abell 2744 & 0.308 & 00:14:21.2 & -30:23:50 & \textit{VLT}/HAWKI     & 2013 Oct 24--2013 Dec 24 & 29.4 h & 25.8 & $0\farcs40$ \\ 
+MACS-0416  & 0.396 & 04:16:08.9 & -24:04:28 & \textit{VLT}/HAWKI     & 2013 Oct 25--2014 Feb 23 & 25.5 h & 25.9 & $0\farcs38$ \\ 
+MACS-0717  & 0.545 & 07:17:34.0 & +37:44:49 & \textit{Keck}/MOSFIRE  & 2015 Jan 26--27                 & \xxx & 25.0 & $0\farcs43$ \\ 
+MACS-1149  & 0.543 & 11:49:36.3 & +22:23:58 & \textit{Keck}/MOSFIRE  & 2015 Feb 24                    & \xxx & \xxx & \xxx
+\enddata 
+%\vspace{-0.8cm} 
+\tablecomments{Depth is defined as the 5$\sigma$ limiting magnitude for point sources, measured in $D=0\farcs5$ apertures (\S\ref{s:depth}).} 
+\end{deluxetable*}
+%\vspace{1cm}
+
+$K_s$-band observations of the Abell 2744 and MACS-0416 fields were obtained between 2013 Oct and 2014 Feb with the High Acuity Wide-field K-band Imager \citep[HAWK-I;][]{hawki, hawki2} on the 8.2\,m UT4 telescope at the ESO \textit{Very Large Telescope} (ESO program \mbox{092.A-0472}).
+HAWK-I is composed of four chips with 2048$\times$2048 $0\farcs106$ pixels, and the full $7\farcm5\times7\farcm5$ HAWKI field-of-view is perfectly suited to cover both the primary and parallel \textit{HST} ACS optical and WFC3 IR fields simultaneously in a single pointing (see Fig.\ref{fig:layout}).
+
+The HAWK-I observations were divided into individual service mode Observation Blocks (OBs); typically one or two OBs of a given field were observed per night as conditions allowed, and occasionally both fields were observed on the same night.
+The execution of each OB lasted either 60 or 90 minutes, depending on how many exposures were obtained in the block.
+The exposures were roughly one minute each, with multiple coadds of shorter reads making up the exposure ($\mathrm{NDIT}\times\mathrm{DIT} = 4\times15$\,s~=~60\,s for Abell 2744 and 7$\times$8\,s~=~56\,s for MACS-0416\footnote{Longer DITs are preferable to reduce instrument overheads; the shorter DITs on the MACS-0416 field were done do accommodate \textit{VLT} service mode restrictions on bright source saturation, which have been eased since P93.}).
+The telescope was offset with random dithers between each exposure to facilitate sky subtraction;  observations taken before 2013 Nov 4 were dithered within a 20$^{\prime\prime}$ box, which was subsequently increased to 40$^{\prime\prime}$ to improve sky subtraction in the cluster cores crowded with bright galaxies and intra-cluster light.
+The total on-sky integration times for the Abell 2744 and MACS-0416 fields are 29.4 and 25.5 hours, respectively.
+
+Imaging observations of the MACS-0717 and MACS-1149 fields in the $K_s$ filter were obtained on 2015 Jan 26/27 and 2015 Feb 24, respectively, with the Multi-Object Spectrometer For Infra-Red Exploration \citep[MOSFIRE;][]{mosfire} on the 10\,m \textit{Keck I} telescope.
+The MOSFIRE detector consists of $2048\times2048$ $0\farcs1798$ pixels; the resulting $6\farcm1\times6\farcm1$ field-of-view is unfortunately slightly too small to cover both \textit{HST} fields simultaneously and therefore requires two pointings to cover the entire deep \textit{HST} Frontier Fields imaging area (Fig.~\ref{fig:layout}).
+The 6$\times$7\,s~=~42\,s MOSFIRE exposures were obtained in a $3\times3$ ``Box9'' dither pattern spaced roughly 40$^{\prime\prime}$ between offset positions.
+\xxx\ time on cluster, parallel field etc \xxx.
+
+\subsection{Image processing}
+\label{s:processing}
+
+The HAWK-I and MOSFIRE observations were reduced with a pipeline that has been developed for previous surveys with the NEWFIRM \citep[NMBS;][]{whitaker:nmbs} and FOURSTAR \citep[zFOURGE;][, Labb\'e in prep]{spitler:12} infrared imaging instruments.
+Treating each detector individually, the pipeline is easily modified for the different instrument configurations of the four HAWK-I and single MOSFIRE chips.
+The primary task of the pipeline is removing the bright, time-variable sky background from the individual exposures, which is often some $10^4$ times brighter than the distant galaxies of interest in the field.
+With such a bright background, we first determine an empirical ``sky flat'' that is a median of all of the \textit{science} exposures in a HAWK-I OB or MOSFIRE group, after rejecting the brightest 12 exposures at each pixel position to remove the contribution of bright objects.
+We find these empirical flats to be preferable to external twilight or dome flats given the difficulty of obtaining a truly flat illumination pattern over such large detector fields-of-view.
+
+After dividing by the flat, the background of each exposure is determined in a first pass from the simple median of the four exposures that came both immediately before and after it.
+The background-subtracted exposures are combined into a mosaic from which objects are detected.
+The final refined background of each exposure is determined in a second ``mask pass'' from a median again from the four exposures before and after it but now aggressively masking all pixels that contain flux from the detected objects.
+
+\subsection{Photometric calibration}
+\label{s:calib}
+
+For the HAWK-I observations, a single 90-minute OB was obtained in each of the Abell 2744 and MACS-0416 fields requiring photometric transparency conditions, and these OBs were followed immediately by an observation of a photometric standard star at a similar airmass with single exposures on each of the four chips.
+The standard star exposures were processed with empirical flats derived from the science exposures as described above, and the observed fluxes of the standard stars yield absolute photometric zeropoints for each chip.
+\xxx\ We determine additional relative zeropoint offsets between the four HAWK-I chips by comparing the photometry of sources in the overlap regions between them.
+\xxx\ \textit{Not sure why these chip corrections are needed beyond the standard star, but they are consistent between the two fields and are needed to improve the agreement in relative depths between the chips.} Correction factors were then computed to scale the OBs obtained at varying transparency levels to the calibrated photometric OB; the additional service mode OBs were obtained under generally satisfactory weather conditions and the scale factors typically differ from unity by only a few percent.
+
+The 2-Micron All Sky Survey \citep[2MASS;][]{2mass} provides an additional check on the photometric calibration, though the comparison is limited due to relatively little brightness overlap between the faint end of reliable 2MASS photometry and the bright end where stars are in the linear regime of the detectors on the 8--10\,m telescopes.
+For the MOSFIRE observations obtained over just one or two nights, we use the 2MASS overlap directly to determine the absolute photometric calibration.
+A comparison of the observed photometry in the four survey fields to the 2MASS catalog magnitudes is shown in Fig.~\ref{fig:2mass}.
+The stellar photometry on the deep $K_s$-band mosaics is measured within $1\arcsec$ apertures corrected to infinity with the curves of growth described below.
+The error-bars on the points in Fig.~\ref{fig:2mass} come from the 2MASS catalog; the photometric uncertainties are negligible for such bright stars in the deep images ($\mathrm{S/N}>100$).
+We estimate that the systematic uncertainty of the absolute photometric calibration is $\sigma_\mathrm{sys}\lesssim 0.05$ mag.
+
+\subsection{Astrometric alignment}
+\label{s:astrometry}
+
+We refine the relative and absolute astrometry of the $K_s$-band images using the SExtractor \citep{bertin:96} and SCAMP \citep{scamp} software tools.
+Reference astrometric catalogs are generated from \textit{HST} images when available (chips \#1 and \#4 of both HAWK-I cluster fields) and public \textit{Subaru} Suprime-Cam $r_c$-band images otherwise\footnote{http://archive.stsci.edu/pub/hlsp/clash/macs0416/data/subaru/}.
+Next, object catalogs are extracted for each background-subtracted exposure and transformations to the reference frame are computed with SCAMP.
+All of the exposures for each HAWK-I chip are processed simultaneously as a SCAMP ``instrument'', ensuring a uniform geometric distortion correction for each chip \citep[see also][for a detailed discussion of the HAWK-I geometric distortion]{libralato:14}.
+
+For the MOSFIRE observations, we adopt a more simplified approach and refine the shifts specified by the World Coordinate System header keywords by cross-correlating each exposure in a group with an arbitrarily-defined reference exposure (\xxx\ this is what Ivo's pipeline does internally.
+I found a bit better final image quality of the HAWK-I images using the SCAMP approach, and the HAWK-I distortions are non-trivial).
+For the absolute astrometry of the MOSFIRE mosaic, we align the $K_s$-band image to the Suprime-Cam image\footnote{http://archive.stsci.edu/pub/hlsp/clash/macs0717/data/subaru/} allowing for a linear shift, rotation and scale using the \texttt{iraf.geomap} tool.
+
+\xxx\ Add some sort of diagnostic figure showing the astrometric comparison between Ks and, say, ACS F814W and get a number to quote for astrometric precision
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%   Exposure Map figure
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure*}
+\plotone{hawkiff_layout.pdf} % Figures/exposure_map.pdf
+\caption{Layout of the Frontier Fields $K_s$-band mosaics.
+The positions of the \textit{HST} cluster and parallel fields are shown in the blue (ACS optical) and red (WFC3 IR) polygons.
+The area covered by the \textit{Spitzer}/IRAC 3.6 and 4.5\,\mum\ channels is shown in orange.
+The bottom panels show the exposure maps, where the four HAWK-I chips and two MOSFIRE pointings can be seen.\label{fig:layout}}
+\end{figure*}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%   Exposure Map figure
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%   2MASS comparison
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[!ht]
+\epsscale{1.18}
+\plotone{2mass_phot.pdf} % Figures/exposure_map.pdf
+\caption{Comparison of bright-star photometry to the 2MASS public catalog.
+The Abell 2744 and MACS 0416 fields at high galactic latitude have few stars for cross calibration, and the photometric calibration is determined from observations of standard stars on nights with photometric transparency conditions.
+At relatively low galactic latitude ($b=+21$), the MACS 0717 field has many bright stars, and the photometric calibration in that field is defined from the 2MASS comparison (hence, $\Delta_\mathrm{2mass}\equiv0$, as indicated for that field).
+We estimate the overall systematic uncertainty in the photometric calibration to be $\lesssim0.05~\mathrm{mag}$.\label{fig:2mass}}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%   Image quality figure
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[!ht]
+\epsscale{1.18}
+\plotone{curve_of_growth.pdf} % Figures/exposure_map.pdf
+\caption{Curves of growth of stellar profiles in the Frontier Fields $K_s$-band mosaics.
+The profiles of stars are well-fit by Moffat profiles with $\mathrm{FWHM}\sim0\farcs4$ and $\beta\sim2$, which have substantially more flux at radii $r>\mathrm{FWHM}$ than Gaussian profiles with the same half-widths (orange curves).
+These extended profiles appear to be characteristic of deep $K_s$-band images; the stellar profiles from the HUGS GOODS-South D2 HAWK-I $K_s$ image \citep{hugs} (lower right panel), created by a different team with a different reduction pipeline, are nearly identical to those of the Frontier Fields mosaics presented here.
+\xxx\ I only show the parallel field region of MACS0717, which has significantly better image quality from the Jan.
+run.
+You can see a few of the broader profiles in the figure; there are many more of these if you include the whole cluster area.
+\label{fig:cog}}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%   Depth figure
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[!ht]
+\epsscale{1.18}
+\plotone{aper_depth.pdf} % Figures/exposure_map.pdf
+\caption{\textit{Top panel:} Noise in empty apertures as a function of aperture diameter, normalized to the smallest $D=0\farcs2$ aperture.
+The dashed and dotted lines show the expectations for uncorrelated and perfectly-correlated noise within the aperture.
+\textit{Bottom panel:} The product of the inverse curves-of-growth (Fig.\ref{fig:cog} and the empty aperture noise gives the point-source sensitivity as a function of aperture size.
+The S/N is maximized with an aperture just larger than the FWHM.
+For reference, the colored dashed lines in the bottom panel indicate the depth computed assuming uncorrelated noise within the apertures ($\sigma \propto D$) and the per-pixel variance reported in the weight maps.)
+\label{fig:depth}}  
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%                                                                %
+%   Mosaic properties
+%                                                                %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{$K_s$-band Mosaics}
+\label{s:mosaic}
+
+We generate stacks of each OB using the SWarp tool \citep{swarp}, which takes into account the alignment to the reference catalog.
+Final mosaics of each field with $0\farcs100$ pixels are generated from the OB stacks weighting the input images with weeights
+\[
+w^{-1} = F^2\cdot\sigma^2 \cdot \mathrm{FWHM}^\alpha, 
+\]
+where $F$ is the factor to scale the OB to the photometric system ($F\equiv1$ for the OB taken under photometric contidions, $F\gtrsim1$ otherwise), $\sigma^2$ is the variance map of each OB stack determined by SWarp ($\sigma^2$ accounts for both exposure time and sky background variations between OBs), FWHM is the full width at half maximum of stars identified in the field, and the parameter $\alpha$ allows for optimizing the image quality of the final mosaic.
+Increasing $\alpha$ puts larger weight on the best image-quality OBs while effectively ignoring data with poorer seeing; we adopt $\alpha=2$ for a compromise between the image quality and effective exposure time of the final mosaic.
+Finally, we subtract a cell-based background from the final mosaics using an algorithm based on that used by SExtractor and SWarp but that provides more aggressive masking of flux in the outer isophotes of bright galaxies.
+The mosaics and inverse-variance maps are shown in Fig.~\ref{fig:layout}, along with the position of the deep Frontier Fields \textit{HST} imaging fields.
+
+\xxx Right now I just swarp the MOSFIRE groups together without any weighting
+
+The final image and inverse-variance mosaics are provided as an ESO Phase 3 data product \xxx\ URL, as well as attached as a digital product to this manuscript (\xxx\ I think ApJSS can and wants to do this).
+The images are scaled to an AB magnitude zeropoint ZP=26.0.
+In the sections below we describe the characteristics of the mosaics in more detail.
+
+\subsection{Image quality}
+\label{s:fwhm}
+
+Fig.~\ref{fig:cog} shows the curves of growth for stars identified by the tight relationship between their brightness and half-light radii \citep[see, e.g., Fig.~13 of][]{skelton:14}.
+The image quality of the $K_s$-band mosaics is excellent, with stellar FWHM $0\farcs40$ and $0\farcs38$ in the Abell 2744 and MACS 0416 fields, respectively, thanks, in large part, to the execution of the HAWK-I observations in service mode ensuring optimal and uniform image quality across the many hours of integration on the survey fields.
+Despite the compact cores of the stellar PSFs, the curves of growth shown in Fig.~\ref{fig:cog} deviate from Gaussian profiles with significantly higher flux in the outer wings, with a shape more consistent with a Moffat profile \citep{trujillo:01}  with $\beta\sim2$.
+Also shown in Fig.~\ref{fig:cog} are stellar curves of growth measured from the HAWK-I UDS and GOODS Survey \citep[HUGS;][]{hugs}, which are nearly identical in both FWHM and profile shape.
+We caution that these extended profiles will yield significantly shallower final image depths than would be predicted with Exposure Time Calculators that may assume Gaussian profiles.
+\xxx\ not clear what HAWK-I ETC assumes, the description just says ``FWHM of the seeing disk'' \xxx.
+Nevertheless, the deep $K_s$ mosaics have excellent image quality that is much closer to the resolution of the \textit{HST} IR imaging than the redder \textit{Spitzer} IRAC bands.
+
+\xxx\ add description of MOSFIRE clusters.  IQ not uniform across the MACS0717 field.
+
+\subsection{Depth}
+\label{s:depth}
+
+The inverse-variance weight maps shown in Fig.~\ref{fig:layout} provide a formal measurement of the pixel-to-pixel variance.
+However, for aperture photometry measurements this variance will yield an underestimate of the true photometric uncertainty due to correlated noise of neighboring pixels that results from pixel resampling during the image reduction process.
+For uncorrelated noise between pixels in an aperture, the variance within the aperture scales with the number of pixels in the aperture, $N_\mathrm{pix}$.
+For perfect correlation within the aperture, however, the variance scales as $N_\mathrm{pix}^2$, and the properties of the $K_s$ mosaics will lie somewhere in between these two extremes.
+
+Following \cite{labbe:03} \citep[see also][]{quadri:musyc, whitaker:nmbs, skelton:14}, we characterize the noise properties of the $K_s$ mosaics by measuring the flux within circular apertures placed randomly throughout empty parts of the images.
+The top panel of Fig.~\ref{fig:depth} shows the standard deviation of these empty-aperture fluxes, $\sigma_\mathrm{D}$, as a function of aperture diameter, which should be used in place of simple scalings of the inverse-variance maps to represent the true uncertainties within photometric apertures.
+
+The product of the inverse curves of growth shown in Fig.~\ref{fig:cog} (i.e., aperture corrections) and the noise curves in the top panel of Fig.~\ref{fig:depth} defines an optimal photometric aperture size where this product is minimized \citep{whitaker:nmbs}.
+For our $K_s$-band mosaics, this optimal aperture has $D\sim0\farcs5$, just larger than the FWHM of point-source profiles.
+The $5\,\sigma$ limiting magnitudes (AB) within $0\farcs5$ diameter apertures are 25.8, 25.9, and 25.0 in the Abell 2744, MACS-0416, and MACS-0717 fields, respectively.
+We note that the depths of the FF HAWK-I mosaics are nearly identical to those of the two deepest HUGS pointings in the GOODS-South Field (i.e., GOODS-South ``D1'' and ``D2'').
+The image depths reported by \cite{hugs} assume ideal noise properties that can be compared to the dashed curves in Fig.~\ref{fig:depth}.
+We have analyzed the HUGS GOODS-South ``D2'' image with the same empty-aperture technique as was done for the FF images, and the sensitivities are fully consistent with those reported by \cite{hugs} when accounting for the different definitions.
+Along with the two HUGS deep HAWK-I fields, our images covering the Frontier Fields represent the deepest $K_s$-band images ever obtained and will provide a crucial complement to the deep \textit{HST} and \textit{Spitzer} imaging of the same fields.
+
+\xxx\ say something about distant extended sources, perhaps scaling from Arjen's size distributions?
+
+\subsection{Number counts}
+\label{s:number}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%                                                                %
+%   Constraints
+%                                                                %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Photometric constraints from the $K_s$ band}
+\label{s:constraints}
+
+\subsection{Photometric redshifts}
+\label{s:photoz}
+
+\subsection{SED modeling}
+\label{s:sed}
+
+Balmer break, UV slope
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%                                                                %
+%   Discussion and Summary
+%                                                                %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Discussion and Conclusions}
+\label{s:discussion}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%   End things
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vspace{-0.3cm} %% emulateapj
+\acknowledgements
+
+\noindent We gratefully acknowledge funding support from the STScI Director's Discretionary Research Fund.
+\xxx\ acknowledge NASA/Keck time + Mauna Kea \xxx.
+This research made use of Astropy, a community-developed core Python package for Astronomy \citep{astropy}, and open-source Python modules Numpy, Scipy, and Matplotlib.
+
+% \noindent We gratefully acknowledge funding support from the Marie Curie Actions of the European Commission (FP7-COFUND) and STScI
+% grant GO-12177.
+
+{\it Facilities:} \facility{VLT:Yepun (HAWKI)}, \facility{Keck:I (MOSFIRE)} % \facility{Hubble Space Telescope (WFC3)}
+
+\bibliographystyle{apj_url}
+\bibliography{ms}
+
+\end{document}

### 2015-02-23 Paper0/ms.tex
Updates from ShareLaTeX

@@
+<<<<<<< HEAD
 \documentclass[iop, numberedappendix]{emulateapj}
@@
 \end{document}
+=======
+\documentclass[iop, numberedappendix]{emulateapj}
+\usepackage{apjfonts}
+%\documentclass[12pt,preprint]{aastex}
+
+%\topmargin 0.6in
+
+%\usepackage{deluxetable}
+\usepackage{color}
+\definecolor{citeRGB}{rgb}{0,0.1,0.7}
+\usepackage[hyperfootnotes=true,naturalnames=true,letterpaper,pdfstartview=FitH,pdfpagemode=UseNone,colorlinks=true,citecolor=citeRGB]{hyperref}
+
+%%%% The following code forces Hyperref to only link the year in a citation
+%%%% Taken from http://tex.stackexchange.com/questions/23227/how-to-hyperlink-only-the-year-part-when-using-natbib-and-hyperref
+% \usepackage{etoolbox}
+% 
+% \makeatletter
+% 
+% % Patch case where name and year have no delimiter
+% \patchcmd{\NAT@citex}
+%   {\@citea\NAT@hyper@{\NAT@nmfmt{\NAT@nm}\NAT@date}}
+%   {\@citea\NAT@nmfmt{\NAT@nm}\NAT@hyper@{\NAT@date}}
+%   {}% Do nothing if patch works
+%   {}% Do nothing if patch fails
+% 
+% % Patch case where name and year have basic delimiter
+% \patchcmd{\NAT@citex}
+%   {\@citea\NAT@hyper@{%
+%      \NAT@nmfmt{\NAT@nm}%
+%      \hyper@natlinkbreak{\NAT@aysep\NAT@spacechar}{\@citeb\@extra@b@citeb}%
+%      \NAT@date}}
+%   {\@citea\NAT@nmfmt{\NAT@nm}%
+%    \NAT@aysep\NAT@spacechar%
+%    \NAT@hyper@{\NAT@date}}
+%   {}% Do nothing if patch works
+%   {}% Do nothing if patch fails
+% 
+% % Patch case where name and year are separated by a prenote
+% \patchcmd{\NAT@citex}
+%   {\@citea\NAT@hyper@{%
+%      \NAT@nmfmt{\NAT@nm}%
+%      \hyper@natlinkbreak{\NAT@spacechar\NAT@@open\if*#1*\else#1\NAT@spacechar\fi}%
+%        {\@citeb\@extra@b@citeb}%
+%      \NAT@date}}
+%   {\@citea\NAT@nmfmt{\NAT@nm}%
+%    \NAT@spacechar\NAT@@open\if*#1*\else#1\NAT@spacechar\fi%
+%    \NAT@hyper@{\NAT@date}}
+%   {}% Do nothing if patch works
+%   {}% Do nothing if patch fails
+% 
+% \makeatother
+
+%\gdef\zphot{$z_\mathrm{phot}$}
+\gdef\HST{\textit{HST}}
+\gdef\G141{\textit{G141}}
+\gdef\F140W{\textit{F140W}}
+\gdef\fluxcgs{\mathrm{erg\ s^{-1}\ cm^{-2}}}
+\gdef\micront{$\mu$m}
+\gdef\micronm{\mu\mathrm{m}}
+\gdef\aXe{\texttt{aXe}}
+\gdef\flux_radius{\textsc{flux\_radius}}
+\gdef\epers{\textit{e}$^{-}$ s$^{-1}$}
+
+\gdef\logOH{12 + \log \left(\mathrm{O/H}\right)}
+\gdef\UmV{U-V}
+\gdef\VmJ{V-J}
+\gdef\SFRuvir{SFR_\mathrm{UV+IR}}
+\gdef\peryr{\mathrm{yr}^{-1}}
+\gdef\kms{km\,s$^{-1}$}
+\gdef\mum{$\mu\mathrm{m}$}
+\gdef\24mum{$24\,\mu\mathrm{m}$}
+\gdef\arcsec{^{\prime\prime}}
+\gdef\Jy{\mathrm{Jy}}
+\gdef\uJy{\mu\mathrm{Jy}}
+\gdef\LIR{L_\mathrm{IR}}
+\gdef\LUV{L_\mathrm{UV}}
+\gdef\4ang{4000\,\AA}
+\gdef\msun{$M_{\odot}$}
+\gdef\mathmsun{\ M_{\odot}}
+\gdef\mathgyr{\ \textrm{Gyr}}
+\gdef\mathmyr{\ \textrm{Myr}}
+\gdef\pegase{\textsc{P\'{e}gase}}
+\gdef\eazy{\texttt{Eazy}}
+\gdef\galfit{\texttt{galfit}}
+\gdef\emcee{\texttt{emcee}}
+\gdef\interrest{\texttt{InterRest}}
+\gdef\fast{\texttt{fast}}
+\gdef\zCOSMOS{$z$\textsc{COSMOS}}
+\gdef\zspec{z_\mathrm{spec}}
+\gdef\zphot{z_\mathrm{phot}}
+\gdef\zml{z_\mathrm{ML}}
+\gdef\zprior{z_\mathrm{prior}}
+\newcommand{\weblink}[1]{\href{#1}{#1}}
+\gdef\UDFj{UDFj-39546284}
+\gdef\compareID{UDF-40106456}
+\gdef\Lya{\mathrm{Ly}\alpha}
+\gdef\Halpha{\mathrm{H}\alpha}
+\gdef\Hbeta{\mathrm{H}\beta}
+\gdef\othree{[\ion{O}{3}]$\lambda$4959+5007}
+\gdef\epers{electrons~s$^{-1}$}
+\gdef\xxx{\textbf{xxx}}
+
+\gdef\lineflux{$3.5\pm1.3\times10^{-18}\,\fluxcgs$}
+
+\shortauthors{Brammer et al.}
+\shorttitle{$K_s$-band imaging of the Frontier Fields}
+%\slugcomment{Draft (\today)}
+\slugcomment{Submitted to ApJSS}
+
+%\citestyle{aa}
+
+\begin{document}
+
+\title{Ultra-deep $K_s$-band Imaging of the \textit{Hubble} Frontier Fields}
+
+% \footnotetext[*]{Based on observations made with the NASA/ESA \textit{Hubble
+% Space Telescope}, programs GO-11640, 12177 and 12328, obtained at the
+% Space Telescope Science Institute, which is operated by the Association of
+% Universities for Research in Astronomy, Inc., under NASA contract NAS
+% 5-26555.}
+
+\author{Gabriel B. Brammer\altaffilmark{1},
+et~al.}
+% Pieter G.\ van Dokkum\altaffilmark{2},
+% Garth D.\ Illingworth\altaffilmark{4},
+% Rychard J.\ Bouwens\altaffilmark{3},
+% Ivo Labb\'e\altaffilmark{3}, 
+% Marijn Franx\altaffilmark{3},
+% Ivelina Momcheva\altaffilmark{2},
+% Pascal A.\ Oesch\altaffilmark{4}}
+% Rub\'en S\'anchez-Janssen\altaffilmark{1}, 
+% Ivo Labb\'e\altaffilmark{2},
+% Elisabete da Cunha\altaffilmark{3},
+% Dawn K.\ Erb\altaffilmark{4},
+% Marijn Franx\altaffilmark{2},
+% Mattia Fumagalli\altaffilmark{2},
+% Britt Lundgren\altaffilmark{5},
+% Danilo Marchesini\altaffilmark{6},
+% Ivelina Momcheva\altaffilmark{5},
+% Erica Nelson\altaffilmark{5},
+% Shannon Patel\altaffilmark{2},
+% Ryan Quadri\altaffilmark{7},
+% Hans-Walter Rix\altaffilmark{3},
+% Rosalind E.\ Skelton\altaffilmark{5},
+% Kasper B.\ Schmidt\altaffilmark{3},
+% Arjen van der Wel\altaffilmark{3},
+% Pieter G.\ van Dokkum\altaffilmark{5},
+% David A.\ Wake\altaffilmark{5},
+% Katherine E.\ Whitaker\altaffilmark{5}}
+
+% Rachel Bezanson\altaffilmark{5},
+% Xiaohui Fan\altaffilmark{6},
+% Natascha F\"orster Schreiber\altaffilmark{8},
+% Garth D.\ Illingworth\altaffilmark{9},
+% Joel Leja\altaffilmark{5},
+% Dan Magee\altaffilmark{9},
+% Patrick McCarthy\altaffilmark{11},
+% Adam Muzzin\altaffilmark{2},
+% Ryan Quadri\altaffilmark{11},
+% Charles C.\ Steidel\altaffilmark{12},
+% Tomer Tal\altaffilmark{5},
+% Anna Williams\altaffilmark{13}
+
+\email{brammer@stsci.edu}
+
+\altaffiltext{1}{ESA/AURA, Space Telescope Science Institute, 3700 San Martin Dr., Baltimore, MD 21218, USA}
+% \altaffiltext{2}
+% {Department of Astronomy, Yale University, New Haven, CT 06520, USA}
+% \altaffiltext{3}
+% {Leiden Observatory, Leiden University, NL-2300 RA Leiden, Netherlands}
+% \altaffiltext{4}
+% {UCO/Lick Observatory, University of California, Santa Cruz, CA 95064}
+% \altaffiltext{3}
+% {Max Planck Institute for Astronomy (MPIA), K\"onigstuhl 17,
+% 69117, Heidelberg, Germany}
+% \altaffiltext{4}
+% {Department of Physics, University of Wisconsin-Milwaukee, P.O. Box 413,
+% Milwaukee, WI 53201, USA}
+% \altaffiltext{5}
+% {Department of Astronomy, Yale University, New Haven, CT 06520, USA}
+% \altaffiltext{6}
+% {Physics and Astronomy Department, Tufts University, Robinson Hall,
+% Room 257, Medford, MA, 02155, USA}
+% \altaffiltext{7}
+% {Carnegie Observatories, 813 Santa Barbara Street, Pasadena, CA 91101, USA}
+
+\begin{abstract}
+
+We present ultra-deep $K_s$-band imaging of the \textit{Hubble} Frontier Fields clusters Abell 2744, MACS-0416, MACS-0717 and MACS-1149 (\xxx\ include MOSFIRE???), which have recently been observed with large allocations of Directors' Discretionary Time with the \textit{HST} and \textit{Spitzer} telescopes covering $0.4 < \lambda < 4.5$\,\mum.
+\textit{VLT}/HAWKI integrations of the first two fields reach 5-$\sigma$ limiting depths of $K_s=26.3$ (AB) and have seeing FWHM$\sim0\farcs4$.
+Shorter MOSFIRE integrations of the latter two northern fields reach limiting depths $K_s=$\xxx\ with seeing FWHM=\xxx.
+In all cases the $K_s$-band mosaics cover the primary cluster and parallel \textit{HST}/ACS+WFC3 fields, and the total area of the $K_s$-band coverage is \xxx\ arcmin$^2$.
+The $K_s$-band at 2.2\,\mum\ crucially fills the gap between the reddest HST filter (1.6\,\mum\,$\sim H$) and the IRAC 3.6\,\mum\ passband.
+While reaching the full depths of the space-based imaging is not currently feasible from the ground, the deep $K_s$-band images provide significant constraints on both the redshifts and the stellar-population properties of galaxies extending well below the characteristic stellar mass across most of the age of the universe, down to, and including, the redshifts of the targeted galaxy clusters ($z \lesssim 0.5$).
+Reduced, aligned mosaics of the four survey fields are provided accompanying this manuscript.
+Time has been awarded for similar HAWKI coverage of the final two Frontier Fields Abell 370 and Abell S1063, with observations expected by the end of 2015.
+
+\end{abstract}
+
+\keywords{galaxies: evolution --- galaxies: high-redshift --- surveys: xxx}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%                                                                %
+%   Introduction
+%                                                                %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Introduction}
+\label{s:introduction}
+
+Selecting in the reddest band possible.
+
+K-band surveys:  ISAAC \citep{labbe:03}, NMBS, UltraVISTA, ...
+
+Subaru Super Deep Field (AO) \cite{minowa:05}.
+
+HUGS \citep{hugs}.
+
+Frontier Fields
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%                                                                %
+%   Observations
+%                                                                %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Observations \& Data Reduction}
+\label{s:observations}
+
+Table~\ref{tab:summary} provides a brief summary of the survey fields and observation characteristics.  Additional details on the observing strategy and image reduction procedures are provided below.
+
+\xxx\ currently a bit awkward since MOSFIRE reduction done quickly with Ivo's code and HAWK-I done in many additional steps (astrometry and photometry).
+
+\subsection{Observation log}
+\label{s:log}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%                                                                %
+%  Summary table
+%                                                                %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{deluxetable*}{lcclrrcc} 
+\tabletypesize{\footnotesize}
+\tablewidth{0pt} 
+\tablecaption{Field summary \label{tab:summary}} 
+\tablehead{ 
+\colhead{Field} & \colhead{R.A.} & \colhead{Dec.} & \colhead{Instrument} & \colhead{Epoch} & \colhead{$t$} & \colhead{Depth} & \colhead{FWHM}}
+\startdata 
+Abell 2744 & 00:14:21.2 & -30:23:50 & \textit{VLT}/HAWKI     & 2013 Oct 24--2013 Dec 24 & 29.4 h & 25.8 & $0\farcs40$ \\ 
+MACS-0416  & 04:16:08.9 & -24:04:28 & \textit{VLT}/HAWKI     & 2013 Oct 25--2014 Feb 23 & 25.5 h & 25.9 & $0\farcs38$ \\ 
+MACS-0717  & 07:17:34.0 & +37:44:49 & \textit{Keck}/MOSFIRE  & 2015 Jan 26--27                 & \xxx & 25.0 & $0\farcs43$ \\ 
+MACS-1149  & 11:49:36.3 & +22:23:58 & \textit{Keck}/MOSFIRE  & 2015 Feb 24                    & \xxx & \xxx & \xxx
+\enddata 
+%\vspace{-0.8cm} 
+\tablecomments{Depth is defined as the 5$\sigma$ limiting magnitude for point sources, measured in $D=0\farcs5$ apertures (\S\ref{s:depth}).} 
+\end{deluxetable*}
+%\vspace{1cm}
+
+$K_s$-band observations of the Abell 2744 and MACS-0416 fields were obtained between 2013 Oct and 2014 Feb with the High Acuity Wide-field K-band Imager \citep[HAWK-I;][]{hawki, hawki2} on the 8.2\,m UT4 telescope at the ESO \textit{Very Large Telescope} (ESO program \mbox{092.A-0472}).
+HAWK-I is composed of four chips with 2048$\times$2048 $0\farcs106$ pixels, and the full $7\farcm5\times7\farcm5$ HAWKI field-of-view is perfectly suited to cover both the primary and parallel \textit{HST} ACS optical and WFC3 IR fields simultaneously in a single pointing (see Fig.\ref{fig:layout}).
+
+The HAWK-I observations were divided into individual service mode Observation Blocks (OBs); typically one or two OBs of a given field were observed per night as conditions allowed, and occasionally both fields were observed on the same night.
+The execution of each OB lasted either 60 or 90 minutes, depending on how many exposures were obtained in the block.
+The exposures were roughly one minute each, with multiple coadds of shorter reads making up the exposure ($\mathrm{NDIT}\times\mathrm{DIT} = 4\times15$\,s~=~60\,s for Abell 2744 and 7$\times$8\,s~=~56\,s for MACS-0416\footnote{Longer DITs are preferable to reduce instrument overheads; the shorter DITs on the MACS-0416 field were done do accommodate \textit{VLT} service mode restrictions on bright source saturation, which have been eased since P93.}).
+The telescope was offset with random dithers between each exposure to facilitate sky subtraction;  observations taken before 2013 Nov 4 were dithered within a 20$^{\prime\prime}$ box, which was subsequently increased to 40$^{\prime\prime}$ to improve sky subtraction in the cluster cores crowded with bright galaxies and intra-cluster light.
+The total on-sky integration times for the Abell 2744 and MACS-0416 fields are 29.4 and 25.5 hours, respectively.
+
+Imaging observations of the MACS-0717 and MACS-1149 fields in the $K_s$ filter were obtained on 2015 Jan 26/27 and 2015 Feb 24, respectively, with the Multi-Object Spectrometer For Infra-Red Exploration \citep[MOSFIRE;][]{mosfire} on the 10\,m \textit{Keck I} telescope.
+The MOSFIRE detector consists of $2048\times2048$ $0\farcs1798$ pixels; the resulting $6\farcm1\times6\farcm1$ field-of-view is unfortunately slightly too small to cover both \textit{HST} fields simultaneously and therefore requires two pointings to cover the entire deep \textit{HST} Frontier Fields imaging area (Fig.~\ref{fig:layout}).
+The 6$\times$7\,s~=~42\,s MOSFIRE exposures were obtained in a $3\times3$ ``Box9'' dither pattern spaced roughly 40$^{\prime\prime}$ between offset positions.
+\xxx\ time on cluster, parallel field etc \xxx.
+
+\subsection{Image processing}
+\label{s:processing}
+
+The HAWK-I and MOSFIRE observations were reduced with a pipeline that has been developed for previous surveys with the NEWFIRM \citep[NMBS;][]{whitaker:nmbs} and FOURSTAR \citep[zFOURGE;][, Labb\'e in prep]{spitler:12} infrared imaging instruments.
+Treating each detector individually, the pipeline is easily modified for the different instrument configurations of the four HAWK-I and single MOSFIRE chips.
+The primary task of the pipeline is removing the bright, time-variable sky background from the individual exposures, which is often some $10^4$ times brighter than the distant galaxies of interest in the field.
+With such a bright background, we first determine an empirical ``sky flat'' that is a median of all of the \textit{science} exposures in a HAWK-I OB or MOSFIRE group, after rejecting the brightest 12 exposures at each pixel position to remove the contribution of bright objects.
+We find these empirical flats to be preferable to external twilight or dome flats given the difficulty of obtaining a truly flat illumination pattern over such large detector fields-of-view.
+
+After dividing by the flat, the background of each exposure is determined in a first pass from the simple median of the four exposures that came both immediately before and after it.
+The background-subtracted exposures are combined into a mosaic from which objects are detected.
+The final refined background of each exposure is determined in a second ``mask pass'' from a median again from the four exposures before and after it but now aggressively masking all pixels that contain flux from the detected objects.
+
+\subsection{Photometric calibration}
+\label{s:calib}
+
+For the HAWK-I observations, a single 90-minute OB was obtained in each of the Abell 2744 and MACS-0416 fields requiring photometric transparency conditions, and these OBs were followed immediately by an observation of a photometric standard star at a similar airmass with single exposures on each of the four chips.
+The standard star exposures were processed with empirical flats derived from the science exposures as described above, and the observed fluxes of the standard stars yield absolute photometric zeropoints for each chip.
+\xxx\ We determine additional relative zeropoint offsets between the four HAWK-I chips by comparing the photometry of sources in the overlap regions between them.
+\xxx\ \textit{Not sure why these chip corrections are needed beyond the standard star, but they are consistent between the two fields and are needed to improve the agreement in relative depths between the chips.} Correction factors were then computed to scale the OBs obtained at varying transparency levels to the calibrated photometric OB; the additional service mode OBs were obtained under generally satisfactory weather conditions and the scale factors typically differ from unity by only a few percent.
+
+The 2-Micron All Sky Survey \citep[2MASS;][]{2mass} provides an additional check on the photometric calibration, though the comparison is limited due to relatively little brightness overlap between the faint end of reliable 2MASS photometry and the bright end where stars are in the linear regime of the detectors on the 8--10\,m telescopes.
+For the MOSFIRE observations obtained over just one or two nights, we use the 2MASS overlap directly to determine the absolute photometric calibration.
+A comparison of the observed photometry in the four survey fields to the 2MASS catalog magnitudes is shown in Fig.~\ref{fig:2mass}.
+The stellar photometry on the deep $K_s$-band mosaics is measured within $1\arcsec$ apertures corrected to infinity with the curves of growth described below.
+The error-bars on the points in Fig.~\ref{fig:2mass} come from the 2MASS catalog; the photometric uncertainties are negligible for such bright stars in the deep images ($\mathrm{S/N}>100$).
+We estimate that the systematic uncertainty of the absolute photometric calibration is $\sigma_\mathrm{sys}\lesssim 0.05$ mag.
+
+\subsection{Astrometric alignment}
+\label{s:astrometry}
+
+We refine the relative and absolute astrometry of the $K_s$-band images using the SExtractor \citep{bertin:96} and SCAMP \citep{scamp} software tools.
+Reference astrometric catalogs are generated from \textit{HST} images when available (chips \#1 and \#4 of both HAWK-I cluster fields) and public \textit{Subaru} Suprime-Cam $r_c$-band images otherwise\footnote{http://archive.stsci.edu/pub/hlsp/clash/macs0416/data/subaru/}.
+Next, object catalogs are extracted for each background-subtracted exposure and transformations to the reference frame are computed with SCAMP.
+All of the exposures for each HAWK-I chip are processed simultaneously as a SCAMP ``instrument'', ensuring a uniform geometric distortion correction for each chip \citep[see also][for a detailed discussion of the HAWK-I geometric distortion]{libralato:14}.
+
+For the MOSFIRE observations, we adopt a more simplified approach and refine the shifts specified by the World Coordinate System header keywords by cross-correlating each exposure in a group with an arbitrarily-defined reference exposure (\xxx\ this is what Ivo's pipeline does internally.
+I found a bit better final image quality of the HAWK-I images using the SCAMP approach, and the HAWK-I distortions are non-trivial).
+For the absolute astrometry of the MOSFIRE mosaic, we align the $K_s$-band image to the Suprime-Cam image\footnote{http://archive.stsci.edu/pub/hlsp/clash/macs0717/data/subaru/} allowing for a linear shift, rotation and scale using the \texttt{iraf.geomap} tool.
+
+\xxx\ Add some sort of diagnostic figure showing the astrometric comparison between Ks and, say, ACS F814W and get a number to quote for astrometric precision
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%   Exposure Map figure
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure*}
+\plotone{hawkiff_layout.pdf} % Figures/exposure_map.pdf
+\caption{Layout of the Frontier Fields $K_s$-band mosaics.
+The positions of the \textit{HST} cluster and parallel fields are shown in the blue (ACS optical) and red (WFC3 IR) polygons.
+The light blue polygons in the MACS-0717 field show additional wide-field ACS imaging coverage from  programs GO-9722 and GO-10420 \citep[PI: Ebeling;][]{ma:11}.
+The area covered by deep imaging in the \textit{Spitzer}/IRAC 3.6 and 4.5\,\mum\ channels is shown in orange.
+The bottom panels show the mosaic weight maps, where the four HAWK-I chips and two MOSFIRE pointings can be seen.\label{fig:layout}}
+\end{figure*}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%   2MASS comparison
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[!ht]
+\epsscale{1.18}
+\plotone{2mass_phot.pdf} % Figures/exposure_map.pdf
+\caption{Comparison of bright-star photometry to the 2MASS public catalog.
+The Abell 2744 and MACS 0416 fields at high galactic latitude have few stars for cross calibration, and the photometric calibration is determined from observations of standard stars on nights with photometric transparency conditions.
+At relatively low galactic latitude ($b=+21$), the MACS 0717 field has many bright stars, and the photometric calibration in that field is defined from the 2MASS comparison (hence, $\Delta_\mathrm{2mass}\equiv0$, as indicated for that field).
+We estimate the overall systematic uncertainty in the photometric calibration to be $\lesssim0.05~\mathrm{mag}$.\label{fig:2mass}}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%   Image quality figure
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[!ht]
+\epsscale{1.18}
+\plotone{curve_of_growth.pdf} % Figures/exposure_map.pdf
+\caption{Curves of growth of stellar profiles in the Frontier Fields $K_s$-band mosaics.
+The profiles of stars are well-fit by Moffat profiles with $\mathrm{FWHM}\sim0\farcs4$ and $\beta\sim2$, which have substantially more flux at radii $r>\mathrm{FWHM}$ than Gaussian profiles with the same half-widths (orange curves).
+These extended profiles appear to be characteristic of deep $K_s$-band images; the stellar profiles from the HUGS GOODS-South D2 HAWK-I $K_s$ image \citep{hugs} (lower right panel), created by a different team with a different reduction pipeline, are nearly identical to those of the Frontier Fields mosaics presented here.
+\xxx\ I only show the parallel field region of MACS0717, which has significantly better image quality from the Jan.
+run.
+You can see a few of the broader profiles in the figure; there are many more of these if you include the whole cluster area.
+\label{fig:cog}}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%   Depth figure
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[!ht]
+\epsscale{1.18}
+\plotone{aper_depth.pdf} % Figures/exposure_map.pdf
+\caption{\textit{Top panel:} Noise in empty apertures as a function of aperture diameter, normalized to the smallest $D=0\farcs2$ aperture.
+The dashed and dotted lines show the expectations for uncorrelated and perfectly-correlated noise within the aperture.
+\textit{Bottom panel:} The product of the inverse curves-of-growth (Fig.\ref{fig:cog} and the empty aperture noise gives the point-source sensitivity as a function of aperture size.
+The S/N is maximized with an aperture just larger than the FWHM.
+For reference, the colored dashed lines in the bottom panel indicate the depth computed assuming uncorrelated noise within the apertures ($\sigma \propto D$) and the per-pixel variance reported in the weight maps.)
+\label{fig:depth}}  
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%   Depth figure
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[!ht]
+\epsscale{1.18}
+\plotone{image_cutouts.pdf} % Figures/exposure_map.pdf
+\caption{Image cutouts of the Abell 2744 field.  The \textit{RGB} and monochrome panels come from the sources as indicated; the cutouts are all $10\arcsec$ on a side.
+\label{fig:cutout}}  
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%                                                                %
+%   Mosaic properties
+%                                                                %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{$K_s$-band Mosaics}
+\label{s:mosaic}
+
+We generate stacks of each OB using the SWarp tool \citep{swarp}, which takes into account the alignment to the reference catalog.
+Final mosaics of each field with $0\farcs100$ pixels are generated from the OB stacks weighting the input images with weeights
+\[
+w^{-1} = F^2\cdot\sigma^2 \cdot \mathrm{FWHM}^\alpha, 
+\]
+where $F$ is the factor to scale the OB to the photometric system ($F\equiv1$ for the OB taken under photometric contidions, $F\gtrsim1$ otherwise), $\sigma^2$ is the variance map of each OB stack determined by SWarp ($\sigma^2$ accounts for both exposure time and sky background variations between OBs), FWHM is the full width at half maximum of stars identified in the field, and the parameter $\alpha$ allows for optimizing the image quality of the final mosaic.
+Increasing $\alpha$ puts larger weight on the best image-quality OBs while effectively ignoring data with poorer seeing; we adopt $\alpha=2$ for a compromise between the image quality and effective exposure time of the final mosaic.
+Finally, we subtract a cell-based background from the final mosaics using an algorithm based on that used by SExtractor and SWarp but that provides more aggressive masking of flux in the outer isophotes of bright galaxies.
+The mosaics and inverse-variance maps are shown in Fig.~\ref{fig:layout}, along with the position of the deep Frontier Fields \textit{HST} imaging fields.
+
+\xxx Right now I just swarp the MOSFIRE groups together without any weighting
+
+The final image and inverse-variance mosaics are provided as an ESO Phase 3 data product \xxx\ URL, as well as attached as a digital product to this manuscript (\xxx\ I think ApJSS can and wants to do this).
+The images are scaled to an AB magnitude zeropoint ZP=26.0, which gives pixel values of order unity for faint galaxies.
+In the sections below we describe the characteristics of the mosaics in more detail.
+
+\subsection{Image quality}
+\label{s:fwhm}
+
+Fig.~\ref{fig:cog} shows the curves of growth for stars identified by the tight relationship between their brightness and half-light radii \citep[see, e.g., Fig.~13 of][]{skelton:14}.
+The image quality of the $K_s$-band mosaics is excellent, with stellar FWHM $0\farcs40$ and $0\farcs38$ in the Abell 2744 and MACS 0416 fields, respectively, thanks, in large part, to the execution of the HAWK-I observations in service mode ensuring optimal and uniform image quality across the many hours of integration on the survey fields.
+Despite the compact cores of the stellar PSFs, the curves of growth shown in Fig.~\ref{fig:cog} deviate from Gaussian profiles with significantly higher flux in the outer wings, with a shape more consistent with a Moffat profile \citep{trujillo:01}  with $\beta\sim2$.
+Also shown in Fig.~\ref{fig:cog} are stellar curves of growth measured from the HAWK-I UDS and GOODS Survey \citep[HUGS;][]{hugs}, which are nearly identical in both FWHM and profile shape.
+We caution that these extended profiles will yield significantly shallower final image depths than would be predicted with Exposure Time Calculators that may assume Gaussian profiles.
+\xxx\ not clear what HAWK-I ETC assumes, the description just says ``FWHM of the seeing disk'' \xxx.
+Nevertheless, the deep $K_s$ mosaics have excellent image quality that is much closer to the resolution of the \textit{HST} IR imaging than the redder \textit{Spitzer} IRAC bands.
+
+\xxx\ add description of MOSFIRE clusters.  IQ not uniform across the MACS0717 field.
+
+\subsection{Depth}
+\label{s:depth}
+
+The inverse-variance weight maps shown in Fig.~\ref{fig:layout} provide a formal measurement of the pixel-to-pixel variance.
+However, for aperture photometry measurements this variance will yield an underestimate of the true photometric uncertainty due to correlated noise of neighboring pixels that results from pixel resampling during the image reduction process.
+For uncorrelated noise between pixels in an aperture, the variance within the aperture scales with the number of pixels in the aperture, $N_\mathrm{pix}$.
+For perfect correlation within the aperture, however, the variance scales as $N_\mathrm{pix}^2$, and the properties of the $K_s$ mosaics will lie somewhere in between these two extremes.
+
+Following \cite{labbe:03} \citep[see also][]{quadri:musyc, whitaker:nmbs, skelton:14}, we characterize the noise properties of the $K_s$ mosaics by measuring the flux within circular apertures placed randomly throughout empty parts of the images.
+The top panel of Fig.~\ref{fig:depth} shows the standard deviation of these empty-aperture fluxes, $\sigma_\mathrm{D}$, as a function of aperture diameter, which should be used in place of simple scalings of the inverse-variance maps to represent the true uncertainties within photometric apertures.
+
+The product of the inverse curves of growth shown in Fig.~\ref{fig:cog} (i.e., aperture corrections) and the noise curves in the top panel of Fig.~\ref{fig:depth} defines an optimal photometric aperture size where this product is minimized \citep{whitaker:nmbs}.
+For our $K_s$-band mosaics, this optimal aperture has $D\sim0\farcs5$, just larger than the FWHM of point-source profiles.
+The $5\,\sigma$ limiting magnitudes (AB) within $0\farcs5$ diameter apertures are 25.8, 25.9, and 25.0 in the Abell 2744, MACS-0416, and MACS-0717 fields, respectively.
+We note that the depths of the FF HAWK-I mosaics are nearly identical to those of the two deepest HUGS pointings in the GOODS-South Field (i.e., GOODS-South ``D1'' and ``D2'').
+The image depths reported by \cite{hugs} assume ideal noise properties that can be compared to the dashed curves in Fig.~\ref{fig:depth}.
+We have analyzed the HUGS GOODS-South ``D2'' image with the same empty-aperture technique as was done for the FF images, and the sensitivities are fully consistent with those reported by \cite{hugs} when accounting for the different definitions.
+The HAWK-I images presented here are among the deepest $K_s$-band images ever obtained and will provide a crucial complement to the deep \textit{HST} and \textit{Spitzer} imaging of the Frontier Fields.
+
+\xxx\ say something about distant extended sources, perhaps scaling from Arjen's size distributions?
+
+\subsection{Number counts}
+\label{s:number}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%  Number counts figure
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure*}
+\epsscale{1.18}
+\plotone{number_counts_all.pdf} % Figures/exposure_map.pdf
+\caption{Number counts.\label{fig:counts}}  
+\end{figure*}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%                                                                %
+%   Constraints
+%                                                                %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Photometric constraints from the $K_s$ band}
+\label{s:constraints}
+
+\subsection{Photometric redshifts}
+\label{s:photoz}
+
+\subsection{SED modeling}
+\label{s:sed}
+
+Balmer break, UV slope
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%                                                                %
+%   Discussion and Summary
+%                                                                %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Discussion and Conclusions}
+\label{s:discussion}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%   End things
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vspace{-0.3cm} %% emulateapj
+\acknowledgements
+
+\noindent We gratefully acknowledge funding support from the STScI Director's Discretionary Research Fund.
+\xxx\ acknowledge NASA/Keck time + Mauna Kea \xxx.
+This research made use of Astropy, a community-developed core Python package for Astronomy \citep{astropy}, and open-source Python modules Numpy, Scipy, and Matplotlib.
+
+% \noindent We gratefully acknowledge funding support from the Marie Curie Actions of the European Commission (FP7-COFUND) and STScI
+% grant GO-12177.
+
+{\it Facilities:} \facility{VLT:Yepun (HAWKI)}, \facility{Keck:I (MOSFIRE)} % \facility{Hubble Space Telescope (WFC3)}
+
+\bibliographystyle{apj_url}
+\bibliography{ms}
+
+\end{document}
+=======
+\documentclass[iop, numberedappendix]{emulateapj}
+\usepackage{apjfonts}
+%\documentclass[12pt,preprint]{aastex}
+
+%\topmargin 0.6in
+
+%\usepackage{deluxetable}
+\usepackage{color}
+\definecolor{citeRGB}{rgb}{0,0.1,0.7}
+\usepackage[hyperfootnotes=true,naturalnames=true,letterpaper,pdfstartview=FitH,pdfpagemode=UseNone,colorlinks=true,citecolor=citeRGB]{hyperref}
+
+%%%% The following code forces Hyperref to only link the year in a citation
+%%%% Taken from http://tex.stackexchange.com/questions/23227/how-to-hyperlink-only-the-year-part-when-using-natbib-and-hyperref
+% \usepackage{etoolbox}
+% 
+% \makeatletter
+% 
+% % Patch case where name and year have no delimiter
+% \patchcmd{\NAT@citex}
+%   {\@citea\NAT@hyper@{\NAT@nmfmt{\NAT@nm}\NAT@date}}
+%   {\@citea\NAT@nmfmt{\NAT@nm}\NAT@hyper@{\NAT@date}}
+%   {}% Do nothing if patch works
+%   {}% Do nothing if patch fails
+% 
+% % Patch case where name and year have basic delimiter
+% \patchcmd{\NAT@citex}
+%   {\@citea\NAT@hyper@{%
+%      \NAT@nmfmt{\NAT@nm}%
+%      \hyper@natlinkbreak{\NAT@aysep\NAT@spacechar}{\@citeb\@extra@b@citeb}%
+%      \NAT@date}}
+%   {\@citea\NAT@nmfmt{\NAT@nm}%
+%    \NAT@aysep\NAT@spacechar%
+%    \NAT@hyper@{\NAT@date}}
+%   {}% Do nothing if patch works
+%   {}% Do nothing if patch fails
+% 
+% % Patch case where name and year are separated by a prenote
+% \patchcmd{\NAT@citex}
+%   {\@citea\NAT@hyper@{%
+%      \NAT@nmfmt{\NAT@nm}%
+%      \hyper@natlinkbreak{\NAT@spacechar\NAT@@open\if*#1*\else#1\NAT@spacechar\fi}%
+%        {\@citeb\@extra@b@citeb}%
+%      \NAT@date}}
+%   {\@citea\NAT@nmfmt{\NAT@nm}%
+%    \NAT@spacechar\NAT@@open\if*#1*\else#1\NAT@spacechar\fi%
+%    \NAT@hyper@{\NAT@date}}
+%   {}% Do nothing if patch works
+%   {}% Do nothing if patch fails
+% 
+% \makeatother
+
+%\gdef\zphot{$z_\mathrm{phot}$}
+\gdef\HST{\textit{HST}}
+\gdef\G141{\textit{G141}}
+\gdef\F140W{\textit{F140W}}
+\gdef\fluxcgs{\mathrm{erg\ s^{-1}\ cm^{-2}}}
+\gdef\micront{$\mu$m}
+\gdef\micronm{\mu\mathrm{m}}
+\gdef\aXe{\texttt{aXe}}
+\gdef\flux_radius{\textsc{flux\_radius}}
+\gdef\epers{\textit{e}$^{-}$ s$^{-1}$}
+
+\gdef\logOH{12 + \log \left(\mathrm{O/H}\right)}
+\gdef\UmV{U-V}
+\gdef\VmJ{V-J}
+\gdef\SFRuvir{SFR_\mathrm{UV+IR}}
+\gdef\peryr{\mathrm{yr}^{-1}}
+\gdef\kms{km\,s$^{-1}$}
+\gdef\mum{$\mu\mathrm{m}$}
+\gdef\24mum{$24\,\mu\mathrm{m}$}
+\gdef\arcsec{^{\prime\prime}}
+\gdef\Jy{\mathrm{Jy}}
+\gdef\uJy{\mu\mathrm{Jy}}
+\gdef\LIR{L_\mathrm{IR}}
+\gdef\LUV{L_\mathrm{UV}}
+\gdef\4ang{4000\,\AA}
+\gdef\msun{$M_{\odot}$}
+\gdef\mathmsun{\ M_{\odot}}
+\gdef\mathgyr{\ \textrm{Gyr}}
+\gdef\mathmyr{\ \textrm{Myr}}
+\gdef\pegase{\textsc{P\'{e}gase}}
+\gdef\eazy{\texttt{Eazy}}
+\gdef\galfit{\texttt{galfit}}
+\gdef\emcee{\texttt{emcee}}
+\gdef\interrest{\texttt{InterRest}}
+\gdef\fast{\texttt{fast}}
+\gdef\zCOSMOS{$z$\textsc{COSMOS}}
+\gdef\zspec{z_\mathrm{spec}}
+\gdef\zphot{z_\mathrm{phot}}
+\gdef\zml{z_\mathrm{ML}}
+\gdef\zprior{z_\mathrm{prior}}
+\newcommand{\weblink}[1]{\href{#1}{#1}}
+\gdef\UDFj{UDFj-39546284}
+\gdef\compareID{UDF-40106456}
+\gdef\Lya{\mathrm{Ly}\alpha}
+\gdef\Halpha{\mathrm{H}\alpha}
+\gdef\Hbeta{\mathrm{H}\beta}
+\gdef\othree{[\ion{O}{3}]$\lambda$4959+5007}
+\gdef\epers{electrons~s$^{-1}$}
+\gdef\xxx{\textbf{xxx}}
+
+\gdef\lineflux{$3.5\pm1.3\times10^{-18}\,\fluxcgs$}
+
+\shortauthors{Brammer et al.}
+\shorttitle{$K_s$-band imaging of the Frontier Fields}
+%\slugcomment{Draft (\today)}
+\slugcomment{Submitted to ApJSS}
+
+%\citestyle{aa}
+
+\begin{document}
+
+\title{Ultra-deep $K_s$-band Imaging of the \textit{Hubble} Frontier Fields}
+
+% \footnotetext[*]{Based on observations made with the NASA/ESA \textit{Hubble
+% Space Telescope}, programs GO-11640, 12177 and 12328, obtained at the
+% Space Telescope Science Institute, which is operated by the Association of
+% Universities for Research in Astronomy, Inc., under NASA contract NAS
+% 5-26555.}
+
+\author{Gabriel B. Brammer\altaffilmark{1},
+et~al.}
+% Pieter G.\ van Dokkum\altaffilmark{2},
+% Garth D.\ Illingworth\altaffilmark{4},
+% Rychard J.\ Bouwens\altaffilmark{3},
+% Ivo Labb\'e\altaffilmark{3}, 
+% Marijn Franx\altaffilmark{3},
+% Ivelina Momcheva\altaffilmark{2},
+% Pascal A.\ Oesch\altaffilmark{4}}
+% Rub\'en S\'anchez-Janssen\altaffilmark{1}, 
+% Ivo Labb\'e\altaffilmark{2},
+% Elisabete da Cunha\altaffilmark{3},
+% Dawn K.\ Erb\altaffilmark{4},
+% Marijn Franx\altaffilmark{2},
+% Mattia Fumagalli\altaffilmark{2},
+% Britt Lundgren\altaffilmark{5},
+% Danilo Marchesini\altaffilmark{6},
+% Ivelina Momcheva\altaffilmark{5},
+% Erica Nelson\altaffilmark{5},
+% Shannon Patel\altaffilmark{2},
+% Ryan Quadri\altaffilmark{7},
+% Hans-Walter Rix\altaffilmark{3},
+% Rosalind E.\ Skelton\altaffilmark{5},
+% Kasper B.\ Schmidt\altaffilmark{3},
+% Arjen van der Wel\altaffilmark{3},
+% Pieter G.\ van Dokkum\altaffilmark{5},
+% David A.\ Wake\altaffilmark{5},
+% Katherine E.\ Whitaker\altaffilmark{5}}
+
+% Rachel Bezanson\altaffilmark{5},
+% Xiaohui Fan\altaffilmark{6},
+% Natascha F\"orster Schreiber\altaffilmark{8},
+% Garth D.\ Illingworth\altaffilmark{9},
+% Joel Leja\altaffilmark{5},
+% Dan Magee\altaffilmark{9},
+% Patrick McCarthy\altaffilmark{11},
+% Adam Muzzin\altaffilmark{2},
+% Ryan Quadri\altaffilmark{11},
+% Charles C.\ Steidel\altaffilmark{12},
+% Tomer Tal\altaffilmark{5},
+% Anna Williams\altaffilmark{13}
+
+\email{brammer@stsci.edu}
+
+\altaffiltext{1}{ESA/AURA, Space Telescope Science Institute, 3700 San Martin Dr., Baltimore, MD 21218, USA}
+% \altaffiltext{2}
+% {Department of Astronomy, Yale University, New Haven, CT 06520, USA}
+% \altaffiltext{3}
+% {Leiden Observatory, Leiden University, NL-2300 RA Leiden, Netherlands}
+% \altaffiltext{4}
+% {UCO/Lick Observatory, University of California, Santa Cruz, CA 95064}
+% \altaffiltext{3}
+% {Max Planck Institute for Astronomy (MPIA), K\"onigstuhl 17,
+% 69117, Heidelberg, Germany}
+% \altaffiltext{4}
+% {Department of Physics, University of Wisconsin-Milwaukee, P.O. Box 413,
+% Milwaukee, WI 53201, USA}
+% \altaffiltext{5}
+% {Department of Astronomy, Yale University, New Haven, CT 06520, USA}
+% \altaffiltext{6}
+% {Physics and Astronomy Department, Tufts University, Robinson Hall,
+% Room 257, Medford, MA, 02155, USA}
+% \altaffiltext{7}
+% {Carnegie Observatories, 813 Santa Barbara Street, Pasadena, CA 91101, USA}
+
+\begin{abstract}
+
+We present ultra-deep $K_s$-band imaging of the \textit{Hubble} Frontier Fields clusters Abell 2744, MACS-0416, MACS-0717 and MACS-1149 (\xxx\ include MOSFIRE???), which have recently been observed with large allocations of Directors' Discretionary Time with the \textit{HST} and \textit{Spitzer} telescopes covering $0.4 < \lambda < 4.5$\,\mum.
+\textit{VLT}/HAWKI integrations of the first two fields reach 5-$\sigma$ limiting depths of $K_s=26.3$ (AB) and have seeing FWHM$\sim0\farcs4$.
+Shorter MOSFIRE integrations of the latter two northern fields reach limiting depths $K_s=$\xxx\ with seeing FWHM=\xxx.
+In all cases the $K_s$-band mosaics cover the primary cluster and parallel \textit{HST}/ACS+WFC3 fields, and the total area of the $K_s$-band coverage is \xxx\ arcmin$^2$.
+The $K_s$-band at 2.2\,\mum\ crucially fills the gap between the reddest HST filter (1.6\,\mum\,$\sim H$) and the IRAC 3.6\,\mum\ passband.
+While reaching the full depths of the space-based imaging is not currently feasible from the ground, the deep $K_s$-band images provide significant constraints on both the redshifts and the stellar-population properties of galaxies extending well below the characteristic stellar mass across most of the age of the universe, down to, and including, the redshifts of the targeted galaxy clusters ($z \lesssim 0.5$).
+Reduced, aligned mosaics of the four survey fields are provided accompanying this manuscript.
+Time has been awarded for similar HAWKI coverage of the final two Frontier Fields Abell 370 and Abell S1063, with observations expected by the end of 2015.
+
+\end{abstract}
+
+\keywords{galaxies: evolution --- galaxies: high-redshift --- surveys: xxx}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%                                                                %
+%   Introduction
+%                                                                %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Introduction}
+\label{s:introduction}
+
+Selecting in the reddest band possible.
+
+K-band surveys:  ISAAC, NMBS, UltraVISTA, ...
+
+HUGS
+
+Frontier Fields
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%                                                                %
+%   Observations
+%                                                                %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Observations \& Data Reduction}
+\label{s:observations}
+
+Table~\ref{tab:summary} provides a brief summary of the survey fields and observation characteristics.  Additional details on the observing strategy and image reduction procedures are provided below.
+
+\xxx\ currently a bit awkward since MOSFIRE reduction done quickly with Ivo's code and HAWK-I done in many additional steps (astrometry and photometry).
+
+\subsection{Observation log}
+\label{s:log}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%                                                                %
+%  Summary table
+%                                                                %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{deluxetable*}{lrcclrrcc} 
+\tabletypesize{\footnotesize}
+\tablewidth{0pt} 
+\tablecaption{Field summary \label{tab:summary}} 
+\tablehead{ 
+\colhead{Field} & \colhead{$z$} & \colhead{R.A.} & \colhead{Dec.} & \colhead{Instrument} & \colhead{Epoch} & \colhead{$t$} & \colhead{Depth} & \colhead{FWHM}}
+\startdata 
+Abell 2744 & 0.308 & 00:14:21.2 & -30:23:50 & \textit{VLT}/HAWKI     & 2013 Oct 24--2013 Dec 24 & 29.4 h & 25.8 & $0\farcs40$ \\ 
+MACS-0416  & 0.396 & 04:16:08.9 & -24:04:28 & \textit{VLT}/HAWKI     & 2013 Oct 25--2014 Feb 23 & 25.5 h & 25.9 & $0\farcs38$ \\ 
+MACS-0717  & 0.545 & 07:17:34.0 & +37:44:49 & \textit{Keck}/MOSFIRE  & 2015 Jan 26--27                 & \xxx & 25.0 & $0\farcs43$ \\ 
+MACS-1149  & 0.543 & 11:49:36.3 & +22:23:58 & \textit{Keck}/MOSFIRE  & 2015 Feb 24                    & \xxx & \xxx & \xxx
+\enddata 
+%\vspace{-0.8cm} 
+\tablecomments{Depth is defined as the 5$\sigma$ limiting magnitude for point sources, measured in $D=0\farcs5$ apertures (\S\ref{s:depth}).} 
+\end{deluxetable*}
+%\vspace{1cm}
+
+$K_s$-band observations of the Abell 2744 and MACS-0416 fields were obtained between 2013 Oct and 2014 Feb with the High Acuity Wide-field K-band Imager \citep[HAWK-I;][]{hawki, hawki2} on the 8.2\,m UT4 telescope at the ESO \textit{Very Large Telescope} (ESO program \mbox{092.A-0472}).
+HAWK-I is composed of four chips with 2048$\times$2048 $0\farcs106$ pixels, and the full $7\farcm5\times7\farcm5$ HAWKI field-of-view is perfectly suited to cover both the primary and parallel \textit{HST} ACS optical and WFC3 IR fields simultaneously in a single pointing (see Fig.\ref{fig:layout}).
+
+The HAWK-I observations were divided into individual service mode Observation Blocks (OBs); typically one or two OBs of a given field were observed per night as conditions allowed, and occasionally both fields were observed on the same night.
+The execution of each OB lasted either 60 or 90 minutes, depending on how many exposures were obtained in the block.
+The exposures were roughly one minute each, with multiple coadds of shorter reads making up the exposure ($\mathrm{NDIT}\times\mathrm{DIT} = 4\times15$\,s~=~60\,s for Abell 2744 and 7$\times$8\,s~=~56\,s for MACS-0416\footnote{Longer DITs are preferable to reduce instrument overheads; the shorter DITs on the MACS-0416 field were done do accommodate \textit{VLT} service mode restrictions on bright source saturation, which have been eased since P93.}).
+The telescope was offset with random dithers between each exposure to facilitate sky subtraction;  observations taken before 2013 Nov 4 were dithered within a 20$^{\prime\prime}$ box, which was subsequently increased to 40$^{\prime\prime}$ to improve sky subtraction in the cluster cores crowded with bright galaxies and intra-cluster light.
+The total on-sky integration times for the Abell 2744 and MACS-0416 fields are 29.4 and 25.5 hours, respectively.
+
+Imaging observations of the MACS-0717 and MACS-1149 fields in the $K_s$ filter were obtained on 2015 Jan 26/27 and 2015 Feb 24, respectively, with the Multi-Object Spectrometer For Infra-Red Exploration \citep[MOSFIRE;][]{mosfire} on the 10\,m \textit{Keck I} telescope.
+The MOSFIRE detector consists of $2048\times2048$ $0\farcs1798$ pixels; the resulting $6\farcm1\times6\farcm1$ field-of-view is unfortunately slightly too small to cover both \textit{HST} fields simultaneously and therefore requires two pointings to cover the entire deep \textit{HST} Frontier Fields imaging area (Fig.~\ref{fig:layout}).
+The 6$\times$7\,s~=~42\,s MOSFIRE exposures were obtained in a $3\times3$ ``Box9'' dither pattern spaced roughly 40$^{\prime\prime}$ between offset positions.
+\xxx\ time on cluster, parallel field etc \xxx.
+
+\subsection{Image processing}
+\label{s:processing}
+
+The HAWK-I and MOSFIRE observations were reduced with a pipeline that has been developed for previous surveys with the NEWFIRM \citep[NMBS;][]{whitaker:nmbs} and FOURSTAR \citep[zFOURGE;][, Labb\'e in prep]{spitler:12} infrared imaging instruments.
+Treating each detector individually, the pipeline is easily modified for the different instrument configurations of the four HAWK-I and single MOSFIRE chips.
+The primary task of the pipeline is removing the bright, time-variable sky background from the individual exposures, which is often some $10^4$ times brighter than the distant galaxies of interest in the field.
+With such a bright background, we first determine an empirical ``sky flat'' that is a median of all of the \textit{science} exposures in a HAWK-I OB or MOSFIRE group, after rejecting the brightest 12 exposures at each pixel position to remove the contribution of bright objects.
+We find these empirical flats to be preferable to external twilight or dome flats given the difficulty of obtaining a truly flat illumination pattern over such large detector fields-of-view.
+
+After dividing by the flat, the background of each exposure is determined in a first pass from the simple median of the four exposures that came both immediately before and after it.
+The background-subtracted exposures are combined into a mosaic from which objects are detected.
+The final refined background of each exposure is determined in a second ``mask pass'' from a median again from the four exposures before and after it but now aggressively masking all pixels that contain flux from the detected objects.
+
+\subsection{Photometric calibration}
+\label{s:calib}
+
+For the HAWK-I observations, a single 90-minute OB was obtained in each of the Abell 2744 and MACS-0416 fields requiring photometric transparency conditions, and these OBs were followed immediately by an observation of a photometric standard star at a similar airmass with single exposures on each of the four chips.
+The standard star exposures were processed with empirical flats derived from the science exposures as described above, and the observed fluxes of the standard stars yield absolute photometric zeropoints for each chip.
+\xxx\ We determine additional relative zeropoint offsets between the four HAWK-I chips by comparing the photometry of sources in the overlap regions between them.
+\xxx\ \textit{Not sure why these chip corrections are needed beyond the standard star, but they are consistent between the two fields and are needed to improve the agreement in relative depths between the chips.} Correction factors were then computed to scale the OBs obtained at varying transparency levels to the calibrated photometric OB; the additional service mode OBs were obtained under generally satisfactory weather conditions and the scale factors typically differ from unity by only a few percent.
+
+The 2-Micron All Sky Survey \citep[2MASS;][]{2mass} provides an additional check on the photometric calibration, though the comparison is limited due to relatively little brightness overlap between the faint end of reliable 2MASS photometry and the bright end where stars are in the linear regime of the detectors on the 8--10\,m telescopes.
+For the MOSFIRE observations obtained over just one or two nights, we use the 2MASS overlap directly to determine the absolute photometric calibration.
+A comparison of the observed photometry in the four survey fields to the 2MASS catalog magnitudes is shown in Fig.~\ref{fig:2mass}.
+The stellar photometry on the deep $K_s$-band mosaics is measured within $1\arcsec$ apertures corrected to infinity with the curves of growth described below.
+The error-bars on the points in Fig.~\ref{fig:2mass} come from the 2MASS catalog; the photometric uncertainties are negligible for such bright stars in the deep images ($\mathrm{S/N}>100$).
+We estimate that the systematic uncertainty of the absolute photometric calibration is $\sigma_\mathrm{sys}\lesssim 0.05$ mag.
+
+\subsection{Astrometric alignment}
+\label{s:astrometry}
+
+We refine the relative and absolute astrometry of the $K_s$-band images using the SExtractor \citep{bertin:96} and SCAMP \citep{scamp} software tools.
+Reference astrometric catalogs are generated from \textit{HST} images when available (chips \#1 and \#4 of both HAWK-I cluster fields) and public \textit{Subaru} Suprime-Cam $r_c$-band images otherwise\footnote{http://archive.stsci.edu/pub/hlsp/clash/macs0416/data/subaru/}.
+Next, object catalogs are extracted for each background-subtracted exposure and transformations to the reference frame are computed with SCAMP.
+All of the exposures for each HAWK-I chip are processed simultaneously as a SCAMP ``instrument'', ensuring a uniform geometric distortion correction for each chip \citep[see also][for a detailed discussion of the HAWK-I geometric distortion]{libralato:14}.
+
+For the MOSFIRE observations, we adopt a more simplified approach and refine the shifts specified by the World Coordinate System header keywords by cross-correlating each exposure in a group with an arbitrarily-defined reference exposure (\xxx\ this is what Ivo's pipeline does internally.
+I found a bit better final image quality of the HAWK-I images using the SCAMP approach, and the HAWK-I distortions are non-trivial).
+For the absolute astrometry of the MOSFIRE mosaic, we align the $K_s$-band image to the Suprime-Cam image\footnote{http://archive.stsci.edu/pub/hlsp/clash/macs0717/data/subaru/} allowing for a linear shift, rotation and scale using the \texttt{iraf.geomap} tool.
+
+\xxx\ Add some sort of diagnostic figure showing the astrometric comparison between Ks and, say, ACS F814W and get a number to quote for astrometric precision
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%   Exposure Map figure
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure*}
+\plotone{hawkiff_layout.pdf} % Figures/exposure_map.pdf
+\caption{Layout of the Frontier Fields $K_s$-band mosaics.
+The positions of the \textit{HST} cluster and parallel fields are shown in the blue (ACS optical) and red (WFC3 IR) polygons.
+The area covered by the \textit{Spitzer}/IRAC 3.6 and 4.5\,\mum\ channels is shown in orange.
+The bottom panels show the exposure maps, where the four HAWK-I chips and two MOSFIRE pointings can be seen.\label{fig:layout}}
+\end{figure*}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%   Exposure Map figure
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%   2MASS comparison
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[!ht]
+\epsscale{1.18}
+\plotone{2mass_phot.pdf} % Figures/exposure_map.pdf
+\caption{Comparison of bright-star photometry to the 2MASS public catalog.
+The Abell 2744 and MACS 0416 fields at high galactic latitude have few stars for cross calibration, and the photometric calibration is determined from observations of standard stars on nights with photometric transparency conditions.
+At relatively low galactic latitude ($b=+21$), the MACS 0717 field has many bright stars, and the photometric calibration in that field is defined from the 2MASS comparison (hence, $\Delta_\mathrm{2mass}\equiv0$, as indicated for that field).
+We estimate the overall systematic uncertainty in the photometric calibration to be $\lesssim0.05~\mathrm{mag}$.\label{fig:2mass}}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%   Image quality figure
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[!ht]
+\epsscale{1.18}
+\plotone{curve_of_growth.pdf} % Figures/exposure_map.pdf
+\caption{Curves of growth of stellar profiles in the Frontier Fields $K_s$-band mosaics.
+The profiles of stars are well-fit by Moffat profiles with $\mathrm{FWHM}\sim0\farcs4$ and $\beta\sim2$, which have substantially more flux at radii $r>\mathrm{FWHM}$ than Gaussian profiles with the same half-widths (orange curves).
+These extended profiles appear to be characteristic of deep $K_s$-band images; the stellar profiles from the HUGS GOODS-South D2 HAWK-I $K_s$ image \citep{hugs} (lower right panel), created by a different team with a different reduction pipeline, are nearly identical to those of the Frontier Fields mosaics presented here.
+\xxx\ I only show the parallel field region of MACS0717, which has significantly better image quality from the Jan.
+run.
+You can see a few of the broader profiles in the figure; there are many more of these if you include the whole cluster area.
+\label{fig:cog}}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%   Depth figure
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[!ht]
+\epsscale{1.18}
+\plotone{aper_depth.pdf} % Figures/exposure_map.pdf
+\caption{\textit{Top panel:} Noise in empty apertures as a function of aperture diameter, normalized to the smallest $D=0\farcs2$ aperture.
+The dashed and dotted lines show the expectations for uncorrelated and perfectly-correlated noise within the aperture.
+\textit{Bottom panel:} The product of the inverse curves-of-growth (Fig.\ref{fig:cog} and the empty aperture noise gives the point-source sensitivity as a function of aperture size.
+The S/N is maximized with an aperture just larger than the FWHM.
+For reference, the colored dashed lines in the bottom panel indicate the depth computed assuming uncorrelated noise within the apertures ($\sigma \propto D$) and the per-pixel variance reported in the weight maps.)
+\label{fig:depth}}  
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%                                                                %
+%   Mosaic properties
+%                                                                %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{$K_s$-band Mosaics}
+\label{s:mosaic}
+
+We generate stacks of each OB using the SWarp tool \citep{swarp}, which takes into account the alignment to the reference catalog.
+Final mosaics of each field with $0\farcs100$ pixels are generated from the OB stacks weighting the input images with weeights
+\[
+w^{-1} = F^2\cdot\sigma^2 \cdot \mathrm{FWHM}^\alpha, 
+\]
+where $F$ is the factor to scale the OB to the photometric system ($F\equiv1$ for the OB taken under photometric contidions, $F\gtrsim1$ otherwise), $\sigma^2$ is the variance map of each OB stack determined by SWarp ($\sigma^2$ accounts for both exposure time and sky background variations between OBs), FWHM is the full width at half maximum of stars identified in the field, and the parameter $\alpha$ allows for optimizing the image quality of the final mosaic.
+Increasing $\alpha$ puts larger weight on the best image-quality OBs while effectively ignoring data with poorer seeing; we adopt $\alpha=2$ for a compromise between the image quality and effective exposure time of the final mosaic.
+Finally, we subtract a cell-based background from the final mosaics using an algorithm based on that used by SExtractor and SWarp but that provides more aggressive masking of flux in the outer isophotes of bright galaxies.
+The mosaics and inverse-variance maps are shown in Fig.~\ref{fig:layout}, along with the position of the deep Frontier Fields \textit{HST} imaging fields.
+
+\xxx Right now I just swarp the MOSFIRE groups together without any weighting
+
+The final image and inverse-variance mosaics are provided as an ESO Phase 3 data product \xxx\ URL, as well as attached as a digital product to this manuscript (\xxx\ I think ApJSS can and wants to do this).
+The images are scaled to an AB magnitude zeropoint ZP=26.0.
+In the sections below we describe the characteristics of the mosaics in more detail.
+
+\subsection{Image quality}
+\label{s:fwhm}
+
+Fig.~\ref{fig:cog} shows the curves of growth for stars identified by the tight relationship between their brightness and half-light radii \citep[see, e.g., Fig.~13 of][]{skelton:14}.
+The image quality of the $K_s$-band mosaics is excellent, with stellar FWHM $0\farcs40$ and $0\farcs38$ in the Abell 2744 and MACS 0416 fields, respectively, thanks, in large part, to the execution of the HAWK-I observations in service mode ensuring optimal and uniform image quality across the many hours of integration on the survey fields.
+Despite the compact cores of the stellar PSFs, the curves of growth shown in Fig.~\ref{fig:cog} deviate from Gaussian profiles with significantly higher flux in the outer wings, with a shape more consistent with a Moffat profile \citep{trujillo:01}  with $\beta\sim2$.
+Also shown in Fig.~\ref{fig:cog} are stellar curves of growth measured from the HAWK-I UDS and GOODS Survey \citep[HUGS;][]{hugs}, which are nearly identical in both FWHM and profile shape.
+We caution that these extended profiles will yield significantly shallower final image depths than would be predicted with Exposure Time Calculators that may assume Gaussian profiles.
+\xxx\ not clear what HAWK-I ETC assumes, the description just says ``FWHM of the seeing disk'' \xxx.
+Nevertheless, the deep $K_s$ mosaics have excellent image quality that is much closer to the resolution of the \textit{HST} IR imaging than the redder \textit{Spitzer} IRAC bands.
+
+\xxx\ add description of MOSFIRE clusters.  IQ not uniform across the MACS0717 field.
+
+\subsection{Depth}
+\label{s:depth}
+
+The inverse-variance weight maps shown in Fig.~\ref{fig:layout} provide a formal measurement of the pixel-to-pixel variance.
+However, for aperture photometry measurements this variance will yield an underestimate of the true photometric uncertainty due to correlated noise of neighboring pixels that results from pixel resampling during the image reduction process.
+For uncorrelated noise between pixels in an aperture, the variance within the aperture scales with the number of pixels in the aperture, $N_\mathrm{pix}$.
+For perfect correlation within the aperture, however, the variance scales as $N_\mathrm{pix}^2$, and the properties of the $K_s$ mosaics will lie somewhere in between these two extremes.
+
+Following \cite{labbe:03} \citep[see also][]{quadri:musyc, whitaker:nmbs, skelton:14}, we characterize the noise properties of the $K_s$ mosaics by measuring the flux within circular apertures placed randomly throughout empty parts of the images.
+The top panel of Fig.~\ref{fig:depth} shows the standard deviation of these empty-aperture fluxes, $\sigma_\mathrm{D}$, as a function of aperture diameter, which should be used in place of simple scalings of the inverse-variance maps to represent the true uncertainties within photometric apertures.
+
+The product of the inverse curves of growth shown in Fig.~\ref{fig:cog} (i.e., aperture corrections) and the noise curves in the top panel of Fig.~\ref{fig:depth} defines an optimal photometric aperture size where this product is minimized \citep{whitaker:nmbs}.
+For our $K_s$-band mosaics, this optimal aperture has $D\sim0\farcs5$, just larger than the FWHM of point-source profiles.
+The $5\,\sigma$ limiting magnitudes (AB) within $0\farcs5$ diameter apertures are 25.8, 25.9, and 25.0 in the Abell 2744, MACS-0416, and MACS-0717 fields, respectively.
+We note that the depths of the FF HAWK-I mosaics are nearly identical to those of the two deepest HUGS pointings in the GOODS-South Field (i.e., GOODS-South ``D1'' and ``D2'').
+The image depths reported by \cite{hugs} assume ideal noise properties that can be compared to the dashed curves in Fig.~\ref{fig:depth}.
+We have analyzed the HUGS GOODS-South ``D2'' image with the same empty-aperture technique as was done for the FF images, and the sensitivities are fully consistent with those reported by \cite{hugs} when accounting for the different definitions.
+Along with the two HUGS deep HAWK-I fields, our images covering the Frontier Fields represent the deepest $K_s$-band images ever obtained and will provide a crucial complement to the deep \textit{HST} and \textit{Spitzer} imaging of the same fields.
+
+\xxx\ say something about distant extended sources, perhaps scaling from Arjen's size distributions?
+
+\subsection{Number counts}
+\label{s:number}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%                                                                %
+%   Constraints
+%                                                                %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Photometric constraints from the $K_s$ band}
+\label{s:constraints}
+
+\subsection{Photometric redshifts}
+\label{s:photoz}
+
+\subsection{SED modeling}
+\label{s:sed}
+
+Balmer break, UV slope
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%                                                                %
+%   Discussion and Summary
+%                                                                %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Discussion and Conclusions}
+\label{s:discussion}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%   End things
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vspace{-0.3cm} %% emulateapj
+\acknowledgements
+
+\noindent We gratefully acknowledge funding support from the STScI Director's Discretionary Research Fund.
+\xxx\ acknowledge NASA/Keck time + Mauna Kea \xxx.
+This research made use of Astropy, a community-developed core Python package for Astronomy \citep{astropy}, and open-source Python modules Numpy, Scipy, and Matplotlib.
+
+% \noindent We gratefully acknowledge funding support from the Marie Curie Actions of the European Commission (FP7-COFUND) and STScI
+% grant GO-12177.
+
+{\it Facilities:} \facility{VLT:Yepun (HAWKI)}, \facility{Keck:I (MOSFIRE)} % \facility{Hubble Space Telescope (WFC3)}
+
+\bibliographystyle{apj_url}
+\bibliography{ms}
+
+\end{document}
+>>>>>>> master

### 2015-02-23 Paper0/ms.tex
merge again with new figure
Updates from ShareLaTeX
Push from sharelatex

--- 2015-02-21 Paper0/ms.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%   Depth figure
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[!ht]
+\epsscale{1.18}
+\plotone{image_cutouts.pdf} % Figures/exposure_map.pdf
+\caption{Image cutouts of the Abell 2744 field.  The \textit{RGB} and monochrome panels come from the sources as indicated; the cutouts are all $10\arcsec$ on a side.
+\label{fig:cutout}}  
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %                                                                %
@@
 The final image and inverse-variance mosaics are provided as an ESO Phase 3 data product \xxx\ URL, as well as attached as a digital product to this manuscript (\xxx\ I think ApJSS can and wants to do this).
-The images are scaled to an AB magnitude zeropoint ZP=26.0.
+The images are scaled to an AB magnitude zeropoint ZP=26.0, which gives pixel values of order unity for faint galaxies.
 In the sections below we describe the characteristics of the mosaics in more detail.

### 2015-09-25 Paper0/ms.tex
test changed text.
test change.
edits to introduction

@@
 \author{Gabriel B. Brammer\altaffilmark{1},
+Danilo Marchesini\altaffilmark{2},
 et~al.}
@@
 % Britt Lundgren\altaffilmark{5},
-% Danilo Marchesini\altaffilmark{6},
 % Ivelina Momcheva\altaffilmark{5},
@@
 \altaffiltext{1}{ESA/AURA, Space Telescope Science Institute, 3700 San Martin Dr., Baltimore, MD 21218, USA}
-% \altaffiltext{2}
+\altaffiltext{2}{Physics and Astronomy Department, Tufts University, Medford, MA, 02155, USA}
+% \altaffiltext{6}
 % {Department of Astronomy, Yale University, New Haven, CT 06520, USA}
@@
 % {Department of Astronomy, Yale University, New Haven, CT 06520, USA}
-% \altaffiltext{6}
-% {Physics and Astronomy Department, Tufts University, Robinson Hall,
-% Room 257, Medford, MA, 02155, USA}
 % \altaffiltext{7}
@@
 
-Selecting in the reddest band possible.
+During the last two decades, near-infrared (NIR) imaging has gained a dominant role in studies of galaxy formation and evolution.
 
-K-band surveys:  ISAAC \citep{labbe:03}, NMBS, UltraVISTA, ...
+Detection of galaxies in the $K$-band ($\lambda_{\rm eff} \sim 2.2$~$\mu$m) provided the first opportunity to construct a comprehensive picture of the population of galaxies in the early universe, as it enabled the discovery of galaxies at $z > 2$ with suppressed emission in the optical (i.e., rest-frame UV) due to evolved stellar populations and/or significant amount of dust extinction (e.g., \textcolor{red}{refs}).
+In fact, these galaxies, which dominate the high-mass end of the high-$z$ galaxy population, were previously missed by rest-frame UV selection techniques (e.g., $U$-dropout galaxies; \textcolor{red}{refs}).
+Imaging in the $K$-band allows for the direct probing out to $z\approx5$ of rest-frame wavelengths longer than the Balmer break.
+Sampling the rest-frame optical wavelength regime is crucial for high-$z$ studies, as significantly less affected by dust obscuration and a better probe of the galaxy stellar mass compared to the rest-frame UV (more sensitive to unobscured star formation).
+While imaging at even longer wavelengths ($\lambda>3$~$\mu$m) are needed to probe the rest-frame optical of galaxies in the first billion years of cosmic history, deep sub-arsec resolution imaging at these wavelengths will not be available until the advent of the {\it James Webb Space Telescope} ({\it JWST}).
+Deep $K$-band data with good enough (i.e., FWHM$<1^{\prime \prime}$) image quality have been fundamental to fully exploit {\it Spitzer}-IRAC imaging data characterized by much poorer spatial resolution, allowing for the correction of the resulting blending of sources in IRAC images.
+Finally, for space-based studies, $K$ band crucially fills the gap between the reddest {\it Hubble Space Telescope} ({\it HST}) filter (i.e., F160W, with $\lambda \sim 1.6$~$\mu$m) and the IRAC 3.6~$\mu$m pass-band, greatly improving the constraints on both the redshifts and the stellar-population properties of galaxies.
 
-Subaru Super Deep Field (AO) \cite{minowa:05}.
+It is not surprising that, given the aforementioned reasons, all successful extragalactic surveys performed in the last fifteen years to investigate galaxies populations in the early universe invested significant resources to obtain deep NIR imaging data with good image quality, especially in the $K$ band, usually adopted as the detection band.
+Indeed, our understanding of galaxy formation and evolution in the last couple of decades has impressively proceeded forward mostly driven by NIR progressively deeper, wider, and with better image quality, surveys, such as ...
+FIRES, FIREWORKS, NMBS, UltraVISTA, etc...
+Subaru Super Deep Field (AO) \cite{minowa:05}...
+HUGS \citep{hugs}.\textcolor{red}{-- we should provide a complete list of NIR surveys here -- }
 
-HUGS \citep{hugs}.
-
-Frontier Fields
+The latest effort to further our knowledge of galaxy formation and evolution is represented by the HST Frontier Fields program.
+\textcolor{red}{-- brief description of HFF, followed by the need for mega-deep K-band data --}
 
@@
 \plotone{number_counts_all.pdf} % Figures/exposure_map.pdf
-\caption{Number counts.\label{fig:counts}}  
+\caption{$K_s$-band number counts in the Frontier Fields. \label{fig:counts}}  
 \end{figure*}

### 2016-02-20 Paper0/ms.tex
Updated with latest drizzled mosaics in all fields

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{deluxetable*}{lcclrrcc} 
+\begin{deluxetable*}{ccccrccc} 
 \tabletypesize{\footnotesize}
@@
 \tablehead{ 
-\colhead{Field} & \colhead{R.A.} & \colhead{Dec.} & \colhead{Instrument} & \colhead{Epoch} & \colhead{$t$} & \colhead{Depth} & \colhead{FWHM}}
+\colhead{Field} & \colhead{R.A.} & \colhead{Dec.} & \colhead{Instrument} & \colhead{Epoch} & \colhead{$t$} & \colhead{Depth$^\mathrm{a}$} & \colhead{FWHM}}
 \startdata 
-Abell 2744 & 00:14:21.2 & -30:23:50 & \textit{VLT}/HAWKI     & 2013 Oct 24--2013 Dec 24 & 29.4 h & 25.8 & $0\farcs40$ \\ 
-MACS-0416  & 04:16:08.9 & -24:04:28 & \textit{VLT}/HAWKI     & 2013 Oct 25--2014 Feb 23 & 25.5 h & 25.9 & $0\farcs38$ \\ 
-MACS-0717  & 07:17:34.0 & +37:44:49 & \textit{Keck}/MOSFIRE  & 2015 Jan 26--27                 & \xxx & 25.0 & $0\farcs43$ \\ 
-MACS-1149  & 11:49:36.3 & +22:23:58 & \textit{Keck}/MOSFIRE  & 2015 Feb 24                    & \xxx & \xxx & \xxx
+Abell 2744   & $00:14:21.2$ & $-30:23:50$ & \textit{VLT}/HAWKI     & 2013 Oct 24--2013 Dec 24 & 29.4 h & 25.8 & $0\farcs40$ \\ 
+MACS-0416    & $04:16:08.9$ & $-24:04:28$ & \textit{VLT}/HAWKI     & 2013 Oct 25--2014 Feb 23 & 25.5 h & 25.9 & $0\farcs38$ \\ 
+Abell S1063  & $22:49:01.1$ & $-44:32:13$ & \textit{VLT}/HAWKI     & \xxx & 25.5 h & 25.9 & $0\farcs38$ \\ 
+Abell 370    & $02:40:03.3$ & $-01:36:23$ & \textit{VLT}/HAWKI     & \xxx & 25.5 h & 25.9 & $0\farcs38$ \\ 
+MACS-0717    & $07:17:34.0$ & $+37:44:49$ & \textit{Keck}/MOSFIRE  & 2015 Jan 26--27  & 0/0 h\footnotemark[2] & 25.0/25.0 & $0\farcs43$/$0\farcs00$ \\ 
+MACS-1149    & $11:49:36.3$ & $+22:23:58$ & \textit{Keck}/MOSFIRE  & 2015 Feb 24      & \xxx\footnotemark[2] & \xxx & \xxx \\
+''           & ''           & ''          & \textit{VLT}/HAWKI\footnotemark[3]     & 2013--2014       & \xxx & \xxx & \xxx
 \enddata 
 %\vspace{-0.8cm} 
-\tablecomments{Depth is defined as the 5$\sigma$ limiting magnitude for point sources, measured in $D=0\farcs5$ apertures (\S\ref{s:depth}).} 
+\footnotetext[1]{Depth is defined as the 5$\sigma$ limiting magnitude for point sources, measured in $D=0\farcs6$ apertures (\S\ref{s:depth}).} 
+\footnotetext[2]{The MOSFIRE observations of MACS-0717 and MACS-1149 require two separate pointings of the instrument to cover the cluster/parallel \textit{HST} fields, which have slightly different characteristics as indicated.}
+\footnotetext[3]{Archival observations from ESO program \href{http://archive.eso.org/wdb/wdb/eso/eso_archive_main/query?prog_id=090.A-0458(A)}{090.A-0458(A)}.}
 \end{deluxetable*}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%   Drizzle figure
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[!t]
+\epsscale{1.18}
+\plotone{drizzle_cutouts.pdf} % Figures/exposure_map.pdf
+\caption{\xxx
+\label{fig:drizzle_cutout}}  
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%   correlated noise figure
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[!t]
+\epsscale{1.18}
+\plotone{correlated_noise_figure.pdf} % Figures/exposure_map.pdf
+\caption{\xxx
+\label{fig:corr_noise}}  
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %   Exposure Map figure
@@
 The area covered by deep imaging in the \textit{Spitzer}/IRAC 3.6 and 4.5\,\mum\ channels is shown in orange.
-The bottom panels show the mosaic weight maps, where the four HAWK-I chips and two MOSFIRE pointings can be seen.\label{fig:layout}}
+The HAWK-I field of view is perfectly suited for simultaneous imaging of the cluster+parallel field pairs, which require two separate pointings with MOSFIRE.\label{fig:layout}}
 \end{figure*}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[!ht]
+\begin{figure*}[!ht]
 \epsscale{1.18}
@@
 \label{fig:cog}}
-\end{figure}
+\end{figure*}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%   Depth figure
+%   Cutout figure
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2016-02-21 Paper0/ms.tex
more consistent with drizzled

@@
 
-Table~\ref{tab:summary} provides a brief summary of the survey fields and observation characteristics.  Additional details on the observing strategy and image reduction procedures are provided below.
+Table~\ref{tab:summary} provides a brief summary of the survey fields and observation characteristics.
+Additional details on the observing strategy and image reduction procedures are provided in the subsections below.
 
@@
 MACS-0416    & $04:16:08.9$ & $-24:04:28$ & \textit{VLT}/HAWKI     & 2013 Oct 25--2014 Feb 23 & 25.5 h & 25.9 & $0\farcs38$ \\ 
-Abell S1063  & $22:49:01.1$ & $-44:32:13$ & \textit{VLT}/HAWKI     & \xxx & 25.5 h & 25.9 & $0\farcs38$ \\ 
-Abell 370    & $02:40:03.3$ & $-01:36:23$ & \textit{VLT}/HAWKI     & \xxx & 25.5 h & 25.9 & $0\farcs38$ \\ 
-MACS-0717    & $07:17:34.0$ & $+37:44:49$ & \textit{Keck}/MOSFIRE  & 2015 Jan 26--27  & 0/0 h\footnotemark[2] & 25.0/25.0 & $0\farcs43$/$0\farcs00$ \\ 
-MACS-1149    & $11:49:36.3$ & $+22:23:58$ & \textit{Keck}/MOSFIRE  & 2015 Feb 24      & \xxx\footnotemark[2] & \xxx & \xxx \\
-''           & ''           & ''          & \textit{VLT}/HAWKI\footnotemark[3]     & 2013--2014       & \xxx & \xxx & \xxx
+Abell S1063  & $22:49:01.1$ & $-44:32:13$ & \textit{VLT}/HAWKI     & 2015 Jul 8--Sep 22 & 25.5 h & 25.9 & $0\farcs38$ \\ 
+Abell 370    & $02:40:03.3$ & $-01:36:23$ & \textit{VLT}/HAWKI     & 2015 Jul 26--2016 Jan 28 & 25.5 h & 25.9 & $0\farcs38$ \\ 
+MACS-0717    & $07:17:34.0$ & $+37:44:49$ & \textit{Keck}/MOSFIRE  & 2015 Jan 26--27, 2016 Jan 21--22  & 0/0 h\footnotemark[2] & 25.0/25.0 & $0\farcs43$/$0\farcs00$ \\ 
+MACS-1149    & $11:49:36.3$ & $+22:23:58$ & \textit{Keck}/MOSFIRE  & 2015 Feb 24, 2016 Jan 21--22      & \xxx\footnotemark[2] & \xxx & \xxx \\
+''           & ''           & ''          & \textit{VLT}/HAWKI\footnotemark[3]     & 2013 Mar 21--2014 Jun 9      & \xxx & \xxx & \xxx
 \enddata 
@@
 \footnotetext[1]{Depth is defined as the 5$\sigma$ limiting magnitude for point sources, measured in $D=0\farcs6$ apertures (\S\ref{s:depth}).} 
-\footnotetext[2]{The MOSFIRE observations of MACS-0717 and MACS-1149 require two separate pointings of the instrument to cover the cluster/parallel \textit{HST} fields, which have slightly different characteristics as indicated.}
-\footnotetext[3]{Archival observations from ESO program \href{http://archive.eso.org/wdb/wdb/eso/eso_archive_main/query?prog_id=090.A-0458(A)}{090.A-0458(A)}.}
+\footnotetext[2]{The MOSFIRE observations of MACS-0717 and MACS-1149 require two separate pointings of the instrument to cover the cluster/parallel \textit{HST} fields, which have the different characteristics as indicated, respectively.}
+\footnotetext[3]{Archival observations from ESO program \href{http://archive.eso.org/wdb/wdb/eso/eso_archive_main/query?prog_id=090.A-0458(A)}{090.A-0458} (PI: Infante).}
 \end{deluxetable*}
@@
 $K_s$-band observations of the Abell 2744 and MACS-0416 fields were obtained between 2013 Oct and 2014 Feb with the High Acuity Wide-field K-band Imager \citep[HAWK-I;][]{hawki, hawki2} on the 8.2\,m UT4 telescope at the ESO \textit{Very Large Telescope} (ESO program \mbox{092.A-0472}).
+HAWK-I $K_s$-badn bservations of the Abell S1063 and Abell 370 fields were obtained between 2015 Jul and 2016 Jan from a subsequent program (095.A-0533).
 HAWK-I is composed of four chips with 2048$\times$2048 $0\farcs106$ pixels, and the full $7\farcm5\times7\farcm5$ HAWKI field-of-view is perfectly suited to cover both the primary and parallel \textit{HST} ACS optical and WFC3 IR fields simultaneously in a single pointing (see Fig.\ref{fig:layout}).
 
-The HAWK-I observations were divided into individual service mode Observation Blocks (OBs); typically one or two OBs of a given field were observed per night as conditions allowed, and occasionally both fields were observed on the same night.
+The HAWK-I observations were divided into individual service mode Observation Blocks (OBs); typically one or two OBs of a given field were observed per night as conditions allowed, and occasionally multiple fields were observed on the same night.
 The execution of each OB lasted either 60 or 90 minutes, depending on how many exposures were obtained in the block.
-The exposures were roughly one minute each, with multiple coadds of shorter reads making up the exposure ($\mathrm{NDIT}\times\mathrm{DIT} = 4\times15$\,s~=~60\,s for Abell 2744 and 7$\times$8\,s~=~56\,s for MACS-0416\footnote{Longer DITs are preferable to reduce instrument overheads; the shorter DITs on the MACS-0416 field were done do accommodate \textit{VLT} service mode restrictions on bright source saturation, which have been eased since P93.}).
+The exposures were roughly one minute each, with multiple coadds of shorter reads making up the exposure ($\mathrm{NDIT}\times\mathrm{DIT} = 4\times15$\,s~=~60\,s for Abell 2744, 7$\times$8\,s~=~56\,s for MACS-0416, 7$\times$8\,s~=~56\,s for Abell S1063 and 7$\times$8\,s~=~56\,s for Abell 370\footnote{Longer DITs are preferable to reduce instrument overheads; the shorter DITs on the MACS-0416 field were done do accommodate \textit{VLT} service mode restrictions on bright source saturation, which have been eased since ESO Period 93.}).
 The telescope was offset with random dithers between each exposure to facilitate sky subtraction;  observations taken before 2013 Nov 4 were dithered within a 20$^{\prime\prime}$ box, which was subsequently increased to 40$^{\prime\prime}$ to improve sky subtraction in the cluster cores crowded with bright galaxies and intra-cluster light.
-The total on-sky integration times for the Abell 2744 and MACS-0416 fields are 29.4 and 25.5 hours, respectively.
+The total on-sky integration times for the Abell 2744, MACS-0416, Abell S1063 and Abell 370 fields are 29.4, 25.5, \xxx, and \xxx\ hours, respectively.
 
-Imaging observations of the MACS-0717 and MACS-1149 fields in the $K_s$ filter were obtained on 2015 Jan 26/27 and 2015 Feb 24, respectively, with the Multi-Object Spectrometer For Infra-Red Exploration \citep[MOSFIRE;][]{mosfire} on the 10\,m \textit{Keck I} telescope.
+Imaging observations of the MACS-0717 and MACS-1149 fields in the $K_s$ filter were obtained on 2015 Jan 26/27 and 2015 Feb 24 (program N097M), respectively, and both fields again on 2016 Jan 21/22 (program N135M), with the Multi-Object Spectrometer For Infra-Red Exploration \citep[MOSFIRE;][]{mosfire} on the 10\,m \mbox{\textit{Keck I}} telescope.
 The MOSFIRE detector consists of $2048\times2048$ $0\farcs1798$ pixels; the resulting $6\farcm1\times6\farcm1$ field-of-view is unfortunately slightly too small to cover both \textit{HST} fields simultaneously and therefore requires two pointings to cover the entire deep \textit{HST} Frontier Fields imaging area (Fig.~\ref{fig:layout}).
 The 6$\times$7\,s~=~42\,s MOSFIRE exposures were obtained in a $3\times3$ ``Box9'' dither pattern spaced roughly 40$^{\prime\prime}$ between offset positions.
-\xxx\ time on cluster, parallel field etc \xxx.
+The exposure times on the MOSFIRE survey fields are summarized in Table~\ref{tab:summary}.
+
+Additional archival \mbox{HAWK-I} imaging of the MACS-1149 fields was obtained from the program 090.A-0458.
+These data were obtained with one \mbox{HAWK-I} chip centered on the $z=9.7$ candidate from \cite{zheng:12}; they largely cover the HFF cluster field but the pointing was not optimized to also include the parallel field whose location was defined later (see Fig.~\ref{fig:layout}).
+The MACS-1149 HAWK-I observations were obtained in Service Mode at the ESO/\textit{VLT} in \xxx\ separate OBs between 2013 Mar 21 and 2014 Jun 9.
+The 7$\times$8\,s~=~56\,s \xxx\ exposures were taken at nine dithered positions with regular offsets separated by \xxx\ arcsec; the on-source exposure time of the final mosaic is \xxx\ hours.
 
@@
 
-\subsection{Photometric calibration}
-\label{s:calib}
-
-For the HAWK-I observations, a single 90-minute OB was obtained in each of the Abell 2744 and MACS-0416 fields requiring photometric transparency conditions, and these OBs were followed immediately by an observation of a photometric standard star at a similar airmass with single exposures on each of the four chips.
-The standard star exposures were processed with empirical flats derived from the science exposures as described above, and the observed fluxes of the standard stars yield absolute photometric zeropoints for each chip.
-\xxx\ We determine additional relative zeropoint offsets between the four HAWK-I chips by comparing the photometry of sources in the overlap regions between them.
-\xxx\ \textit{Not sure why these chip corrections are needed beyond the standard star, but they are consistent between the two fields and are needed to improve the agreement in relative depths between the chips.} Correction factors were then computed to scale the OBs obtained at varying transparency levels to the calibrated photometric OB; the additional service mode OBs were obtained under generally satisfactory weather conditions and the scale factors typically differ from unity by only a few percent.
-
-The 2-Micron All Sky Survey \citep[2MASS;][]{2mass} provides an additional check on the photometric calibration, though the comparison is limited due to relatively little brightness overlap between the faint end of reliable 2MASS photometry and the bright end where stars are in the linear regime of the detectors on the 8--10\,m telescopes.
-For the MOSFIRE observations obtained over just one or two nights, we use the 2MASS overlap directly to determine the absolute photometric calibration.
-A comparison of the observed photometry in the four survey fields to the 2MASS catalog magnitudes is shown in Fig.~\ref{fig:2mass}.
-The stellar photometry on the deep $K_s$-band mosaics is measured within $1\arcsec$ apertures corrected to infinity with the curves of growth described below.
-The error-bars on the points in Fig.~\ref{fig:2mass} come from the 2MASS catalog; the photometric uncertainties are negligible for such bright stars in the deep images ($\mathrm{S/N}>100$).
-We estimate that the systematic uncertainty of the absolute photometric calibration is $\sigma_\mathrm{sys}\lesssim 0.05$ mag.
-
-\subsection{Astrometric alignment}
-\label{s:astrometry}
-
-We refine the relative and absolute astrometry of the $K_s$-band images using the SExtractor \citep{bertin:96} and SCAMP \citep{scamp} software tools.
-Reference astrometric catalogs are generated from \textit{HST} images when available (chips \#1 and \#4 of both HAWK-I cluster fields) and public \textit{Subaru} Suprime-Cam $r_c$-band images otherwise\footnote{http://archive.stsci.edu/pub/hlsp/clash/macs0416/data/subaru/}.
-Next, object catalogs are extracted for each background-subtracted exposure and transformations to the reference frame are computed with SCAMP.
-All of the exposures for each HAWK-I chip are processed simultaneously as a SCAMP ``instrument'', ensuring a uniform geometric distortion correction for each chip \citep[see also][for a detailed discussion of the HAWK-I geometric distortion]{libralato:14}.
-
-For the MOSFIRE observations, we adopt a more simplified approach and refine the shifts specified by the World Coordinate System header keywords by cross-correlating each exposure in a group with an arbitrarily-defined reference exposure (\xxx\ this is what Ivo's pipeline does internally.
-I found a bit better final image quality of the HAWK-I images using the SCAMP approach, and the HAWK-I distortions are non-trivial).
-For the absolute astrometry of the MOSFIRE mosaic, we align the $K_s$-band image to the Suprime-Cam image\footnote{http://archive.stsci.edu/pub/hlsp/clash/macs0717/data/subaru/} allowing for a linear shift, rotation and scale using the \texttt{iraf.geomap} tool.
-
-\xxx\ Add some sort of diagnostic figure showing the astrometric comparison between Ks and, say, ACS F814W and get a number to quote for astrometric precision
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%   Drizzle figure
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[!t]
-\epsscale{1.18}
-\plotone{drizzle_cutouts.pdf} % Figures/exposure_map.pdf
-\caption{\xxx
-\label{fig:drizzle_cutout}}  
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%   correlated noise figure
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[!t]
-\epsscale{1.18}
-\plotone{correlated_noise_figure.pdf} % Figures/exposure_map.pdf
-\caption{\xxx
-\label{fig:corr_noise}}  
-\end{figure}
+As the archival HAWK-I images of the MACS-1149 field were obtained with longer individual exposures and with fewer exposures per sequence, the mask-pass technique described above did not produce satisfactory results.
+In this case, for each raw exposure we divide by the sky flat and then subtract a third-order polynomial fit to the background.
 
@@
 The area covered by deep imaging in the \textit{Spitzer}/IRAC 3.6 and 4.5\,\mum\ channels is shown in orange.
-The HAWK-I field of view is perfectly suited for simultaneous imaging of the cluster+parallel field pairs, which require two separate pointings with MOSFIRE.\label{fig:layout}}
+The HAWK-I field of view is perfectly suited (Abell 2744, MACS-0416, Abell S1063 and Abell 370) for simultaneous imaging of the cluster+parallel field pairs, which require two separate pointings with MOSFIRE (MACS-0717 and MACS-1149).
+Additional archival HAWK-I coverage of the MACS-1149 field is indicated with the gray square in the lower right panel.\label{fig:layout}}
 \end{figure*}
+
+\subsection{Photometric calibration}
+\label{s:calib}
+
+For the HAWK-I observations, a single 90-minute Service Mode OB was obtained in each of the fields requiring photometric transparency conditions, and these OBs were followed immediately by an observation of a photometric standard star at a similar airmass with single exposures on each of the four chips.
+The standard star exposures were processed with empirical sky flats derived from the science exposures as described above, and the observed fluxes of the standard stars yield absolute photometric zeropoints for each chip.
+Correction factors were then computed to scale the OBs obtained at varying transparency levels to the calibrated photometric OB; the additional service mode OBs were obtained under generally satisfactory (i.e., ``clear'') weather conditions and the scale factors typically differ from unity by only a few percent.
+The photometric calibration of the MOSFIRE observations was determined from standard-star observations taken the same night as the science exposures, again with the standard exposures reduced in the same way as the science exposures.
+
+The 2-Micron All Sky Survey \citep[2MASS;][]{2mass} provides an additional check on the photometric calibration, though the comparison is limited due to relatively little brightness overlap between the faint end of reliable 2MASS photometry and the bright end where stars are in the linear regime of the detectors on the 8--10\,m telescopes.
+A comparison of the observed photometry in the four survey fields to the 2MASS catalog magnitudes is shown in Fig.~\ref{fig:2mass}.
+The stellar photometry on the deep $K_s$-band mosaics is measured within $1\arcsec$ apertures corrected to infinity with the curves of growth described below.
+The error-bars on the points in Fig.~\ref{fig:2mass} come from the 2MASS catalog; the photometric uncertainties are negligible for such bright stars in the deep images ($\mathrm{S/N}>100$).
+The 2MASS comparison suggests a systematic offset of $-0.09$~mag for the Abell S1063 field.
+However, the exposure time for that field was similar to those of the additional deep HAWK-I fields and the derived depths of \textit{all} deep HAWK-I fields are nearly identical (Table~\ref{tab:summary}).
+Given the overall agreement with the 2MASS photometry, we estimate that the systematic uncertainty of the absolute photometric calibration is $\sigma_\mathrm{sys}\leq0.05$~mag.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[!ht]
+\begin{figure}[!t]
 \epsscale{1.18}
@@
 \end{figure}
+
+\subsection{Astrometric alignment}
+\label{s:astrometry}
+
+In order to simplify measurements from the $K_s$-band images in consort with space-based \textit{HST} and \textit{Spitzer} mosaics, the astrometry of the ground-based mosaics must be refined.
+Reference absolute astrometric catalogs were generated from \textit{HST} images when available (i.e., the chips overlapping the cluster and parallel \textit{HST} fields) and public \textit{Subaru} Suprime-Cam $r_c$-band images otherwise\footnote{http://archive.stsci.edu/pub/hlsp/clash/}.
+Next, object catalogs are generated with the SEXtractor software \citep{bertin:96} for each individual background-subtracted exposure and transformations to the reference frame (shift, rotation and scale) are computed with with the \texttt{stsci.stimage.xyxymatch} and \texttt{scikit-image.transform}\footnote{\url{http://scikit-image.org}} Python software tools.
+For the HAWK-I exposures, we fit a third-order polynomial to the geometric distortion model determined by \cite{libralato:14} and specify the polynomial terms as ``Simple Imaging Polynomial'' (SIP) coefficients \citep{fits:sip} in the individual exposure FITS files.
+The SIP FITS header generally applicable for HAWK-I is provided in Appendix \xxx.
+We note that the HAWK-I distortion is generally small, reaching $\sim$2 pixels ($0\farcs2$) at the image corners, but that the distortion corrections are necessary in particular given the excellent overall image quality of the observations.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%   Depth figure
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[!ht]
-\epsscale{1.18}
-\plotone{aper_depth.pdf} % Figures/exposure_map.pdf
-\caption{\textit{Top panel:} Noise in empty apertures as a function of aperture diameter, normalized to the smallest $D=0\farcs2$ aperture.
-The dashed and dotted lines show the expectations for uncorrelated and perfectly-correlated noise within the aperture.
-\textit{Bottom panel:} The product of the inverse curves-of-growth (Fig.\ref{fig:cog} and the empty aperture noise gives the point-source sensitivity as a function of aperture size.
-The S/N is maximized with an aperture just larger than the FWHM.
-For reference, the colored dashed lines in the bottom panel indicate the depth computed assuming uncorrelated noise within the apertures ($\sigma \propto D$) and the per-pixel variance reported in the weight maps.)
-\label{fig:depth}}  
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%   Cutout figure
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[!ht]
-\epsscale{1.18}
-\plotone{image_cutouts.pdf} % Figures/exposure_map.pdf
-\caption{Image cutouts of the Abell 2744 field.  The \textit{RGB} and monochrome panels come from the sources as indicated; the cutouts are all $10\arcsec$ on a side.
-\label{fig:cutout}}  
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %                                                                %
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{$K_s$-band Mosaics}
+\section{Drizzled $K_s$-band Mosaics}
 \label{s:mosaic}
 
-We generate stacks of each OB using the SWarp tool \citep{swarp}, which takes into account the alignment to the reference catalog.
-Final mosaics of each field with $0\farcs100$ pixels are generated from the OB stacks weighting the input images with weeights
+A crucial innovation of the image processing of the present analysis compared to previous works is that we combine all of the individual background-subtracted raw images (e.g., \xxx\ 4000 files for 1000 exposures times 4 detectors for the \mbox{Abell 2744} field) into the final output mosaic with the ``drizzle'' algorithm (\citealp{fruchter:02}, as implemented in the DrizzlePac/AstroDrizzle package; \citealp{drizzlepac}).
+The benefits of the drizzle algorithm  preserving robust noise properties of the output mosaics will be discussed in detail below in \S\xxx.
+Here we simply indicate that the drizzle implementation provides additional benefits in 1) trivial application of relative weights of the input images in creating the final mosaics, 2) easily applying the astrometric alignment and geometric distortion stored in the FITS headers, and 3) allowing the definition of an arbitrary output pixel grid without any additional image resampling.
+
+Final mosaics of each field with $0\farcs1$ pixels are drizzled from the input exposures, $i$, with weights
 \[
-w^{-1} = F^2\cdot\sigma^2 \cdot \mathrm{FWHM}^\alpha, 
+w^{-1} = F^2\cdot B \cdot \mathrm{FWHM}^\alpha / \sqrt{1-e^2}, 
 \]
-where $F$ is the factor to scale the OB to the photometric system ($F\equiv1$ for the OB taken under photometric contidions, $F\gtrsim1$ otherwise), $\sigma^2$ is the variance map of each OB stack determined by SWarp ($\sigma^2$ accounts for both exposure time and sky background variations between OBs), FWHM is the full width at half maximum of stars identified in the field, and the parameter $\alpha$ allows for optimizing the image quality of the final mosaic.
-Increasing $\alpha$ puts larger weight on the best image-quality OBs while effectively ignoring data with poorer seeing; we adopt $\alpha=2$ for a compromise between the image quality and effective exposure time of the final mosaic.
+where $F$ is the factor to scale the exposure to the photometric system ($F\equiv1$ for the OB taken under photometric conditions, $F\gtrsim1$ otherwise), $B$ is the measured background, FWHM is the full width at half maximum of stars identified in the field, and $e$ is the ellipticity of the stellar point spread function (PSF).
+The parameter $\alpha$ allows for optimizing the image quality of the final mosaic.
+Increasing $\alpha$ puts larger weight on the exposures with the best image-quality while effectively ignoring data with poorer seeing; we adopt $\alpha=2$ for a compromise between the image quality and effective exposure time of the final mosaic.
+
 Finally, we subtract a cell-based background from the final mosaics using an algorithm based on that used by SExtractor and SWarp but that provides more aggressive masking of flux in the outer isophotes of bright galaxies.
-The mosaics and inverse-variance maps are shown in Fig.~\ref{fig:layout}, along with the position of the deep Frontier Fields \textit{HST} imaging fields.
-
-\xxx Right now I just swarp the MOSFIRE groups together without any weighting
-
+The mosaics are shown in Fig.~\ref{fig:layout}, along with the position of the deep Frontier Fields \textit{HST} imaging fields.
+Inverse variance maps are shown in Appendix \xxx\.
 The final image and inverse-variance mosaics are provided as an ESO Phase 3 data product \xxx\ URL, as well as attached as a digital product to this manuscript (\xxx\ I think ApJSS can and wants to do this).
-The images are scaled to an AB magnitude zeropoint ZP=26.0, which gives pixel values of order unity for faint galaxies.
+The released images are all scaled to a common zeropoint of 26.0 (AB mag), which gives pixel values of order unity for faint galaxies.
 In the sections below we describe the characteristics of the mosaics in more detail.
@@
 
-\subsection{Depth}
+\subsection{Noise properties \& Depth}
 \label{s:depth}
 
-The inverse-variance weight maps shown in Fig.~\ref{fig:layout} provide a formal measurement of the pixel-to-pixel variance.
-However, for aperture photometry measurements this variance will yield an underestimate of the true photometric uncertainty due to correlated noise of neighboring pixels that results from pixel resampling during the image reduction process.
-For uncorrelated noise between pixels in an aperture, the variance within the aperture scales with the number of pixels in the aperture, $N_\mathrm{pix}$.
-For perfect correlation within the aperture, however, the variance scales as $N_\mathrm{pix}^2$, and the properties of the $K_s$ mosaics will lie somewhere in between these two extremes.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%   Drizzle figure
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[!t]
+\epsscale{1.18}
+\plotone{drizzle_cutouts.pdf} % Figures/exposure_map.pdf
+\caption{Comparison of small regions of the final MACS 0416 mosaics with two output pixel sizes for images combined with the drizzle algorithm (top panels) and with the SWarp software (bottom panels).
+With drizzle we can decrease the size of the input pixels before dropping them into the output grid, which significantly reduces the correlations between neighboring pixels.
+Combining images with simple shift-and-add resampling such as with SWarp results in an effective smoothing of the pixel-to-pixel noise in the final combined image.
+\label{fig:drizzle_cutout}}  
+\end{figure}
 
-Following \cite{labbe:03} \citep[see also][]{quadri:musyc, whitaker:nmbs, skelton:14}, we characterize the noise properties of the $K_s$ mosaics by measuring the flux within circular apertures placed randomly throughout empty parts of the images.
-The top panel of Fig.~\ref{fig:depth} shows the standard deviation of these empty-aperture fluxes, $\sigma_\mathrm{D}$, as a function of aperture diameter, which should be used in place of simple scalings of the inverse-variance maps to represent the true uncertainties within photometric apertures.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%   correlated noise figure
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[!t]
+\epsscale{1.18}
+\plotone{correlated_noise_figure.pdf} % Figures/exposure_map.pdf
+\caption{The standard deviation of the pairwise pixel differences, sorted by pixel separation, provides a quantitative measure of the correlations between pixels.
+The black curves determined from the drizzled images show very little variation with pixel separation, suggesting minimal correlations between adjacent pixels, as expected.
+In contrast, the statistics of the SWarped images show a depression at small separations indicative of correlated pixels.
+The magnitude and shape of the depression is reproduced exactly by the orange curves, which show the pairwise differences for the drizzled images convolved with small Gaussian kernels.
+The effect is larger in the right panel, where the ratio of the input and output pixel sizes is larger by almost a factor of two.
+\label{fig:corr_noise}}  
+\end{figure}
 
-The product of the inverse curves of growth shown in Fig.~\ref{fig:cog} (i.e., aperture corrections) and the noise curves in the top panel of Fig.~\ref{fig:depth} defines an optimal photometric aperture size where this product is minimized \citep{whitaker:nmbs}.
-For our $K_s$-band mosaics, this optimal aperture has $D\sim0\farcs5$, just larger than the FWHM of point-source profiles.
-The $5\,\sigma$ limiting magnitudes (AB) within $0\farcs5$ diameter apertures are 25.8, 25.9, and 25.0 in the Abell 2744, MACS-0416, and MACS-0717 fields, respectively.
-We note that the depths of the FF HAWK-I mosaics are nearly identical to those of the two deepest HUGS pointings in the GOODS-South Field (i.e., GOODS-South ``D1'' and ``D2'').
-The image depths reported by \cite{hugs} assume ideal noise properties that can be compared to the dashed curves in Fig.~\ref{fig:depth}.
-We have analyzed the HUGS GOODS-South ``D2'' image with the same empty-aperture technique as was done for the FF images, and the sensitivities are fully consistent with those reported by \cite{hugs} when accounting for the different definitions.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%   Depth figure
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[!ht]
+\epsscale{1.18}
+\plotone{aper_depth.pdf} % Figures/exposure_map.pdf
+\caption{\textit{Top panel:} \xxx Noise in empty apertures as a function of aperture diameter, normalized to the smallest $D=0\farcs2$ aperture.
+The dashed and dotted lines show the expectations for uncorrelated and perfectly-correlated noise within the aperture.
+\textit{Bottom panel:} The product of the inverse curves-of-growth (Fig.\ref{fig:cog} and the empty aperture noise gives the point-source sensitivity as a function of aperture size.
+The S/N is maximized with an aperture just larger than the FWHM.
+For reference, the colored dashed lines in the bottom panel indicate the depth computed assuming uncorrelated noise within the apertures ($\sigma \propto D$) and the per-pixel variance reported in the weight maps.)
+\label{fig:depth}}  
+\end{figure}
+%%%%%%
+
+Along with practical benefits mentioned above, the most dramatic benefit of using the drizzle algorithm comes from the fact that the raw detector pixels are only resampled once in the process of generating the output mosaic.
+Furthermore, with a large number of dithered raw images we can use drizzle to shrink the raw input pixels by a factor of ten before dropping them into the output grid,\footnote{The pixels are shrunk by adopting pixfrac=0.1 as defined by \cite{fruchter:02}.} and the result is dramatically reduced correlations between neighboring pixels, particularly for output pixel sizes that are somewhat smaller than the native HAWK-I or MOSFIRE detector pixels.
+
+A comparison of final combined images created with output $0\farcs1$ and $0\farcs06$ pixel grids using the drizzle implementation to versions created with the SWarp software \citep{swarp} is shown in Fig.~\ref{fig:drizzle_cutout}.
+In the later case, the raw pixels are resampled twice before making the final mosaic, once combining the raw exposures in each OB and a second time combining the separate OBs into the final stack with an arbitrary pixel grid.
+One could avoid the second resampling and combine all of the raw exposures at once, though SWarp is still limited to dropping the original pixels into the output mosaic preserving the input native pixel grid and therefore a single input pixel maps onto multiple (perhaps many) output pixels.
+
+The fact that the drizzle-combined images in Fig.~\ref{fig:drizzle_cutout} (panels \textit{a} and \textit{b}) appear to the eye to be noisier than their SWarped counterparts (panels \textit{c} and \textit{d}) is simply the result of a smoothing of the pixel noise in the latter case rather than reflecting true underlying differences in their pixel-to-pixel variance.
+\cite{casertano:00} discuss how the noise properties of correlated output pixels are affected by the relative sizes of the input and output pixel grids and the adopted drizzle parameters.
+With the drizzle algorithm and a small pixfrac, a given input pixel will map to only a single output pixel and therefore correlations between the output pixels should be minimal\footnote{Inter-pixel capacitance and other forms of crosstalk intrinsic to the IR detectors effectively place a lower limit on the correlations between adjacent pixels \citep{hilbert:ipc}.}.
+
+We explore these correlations between the output pixels in Fig.~\ref{fig:corr_noise}.
+For the black and red curves in each panel, we measure value differences between random pairs of pixels sorted by the separation between the pixels for the drizzled and SWarped images, respectively.
+For perfectly uncorrelated pixels, these curves will be flat and will reflect the intrinsic noise of the image pixels.
+In the presence of correlations between adjacent pixels, however, the curves will show a depression at small separations.
+As expected, the SWarp-combined images show just such a depression for pairs of pixels 1--2 pixels apart.
+The yellow curves in Fig.~\ref{fig:corr_noise} show the same pairwise differences for the drizzled images now convolved by small Gaussian kernels as indicated, which agree very well with the curves for the images with correlated pixels.
+
+In terms of understanding the noise properties of the images, a key point here is not only that depression at small separations indicates the presence of correlations between adjacent pixels but also the fact that correlations cause the apparent overall r.m.s.
+to be decreased at all separations.
+That the pixel variances can't be trusted in the presence of such correlations is the primary reason why significant effort has been devoted to placing random ``empty apertures'' across images in order to characterize their noise properties \citep[e.g., ]{quadri:musyc, whitaker:nmbs, skelton:14}.
+Now by eliminating the pixel correlations with the drizzle combination (with a small pixfrac) we have shown that the pixel statistics are robust and it is now trivial to compute the expected variance within an arbitrary photometric aperture, for example:
+\begin{equation}
+\sigma_D^2 = \sigma_\mathrm{pix}^2\cdot \pi/4 \cdot D^2,
+\label{eq:aper_var}
+\end{equation}
+for circular apertures with diameter, $D$, in pixels, and where $\sigma_\mathrm{pix}^2$ is the per-pixel variance determined from an analysis such as that in Fig.~\ref{fig:corr_noise}.
+
+The product of Eq.~\ref{eq:aper_var} and the inverse of the curves of growth shown in Fig.~\ref{fig:cog} (i.e., aperture corrections) defines an optimal photometric aperture size where this product is minimized \citep{whitaker:nmbs}.
+For our $K_s$-band mosaics, this optimal aperture has $D\sim0\farcs6$, just larger than the FWHM of point-source profiles.
+The $5\,\sigma$ limiting magnitudes (AB) within $0\farcs6$ diameter apertures are \xxx\ 25.8, 25.9, and 25.0 in the Abell 2744, MACS-0416, and MACS-0717 fields, respectively.
 The HAWK-I images presented here are among the deepest $K_s$-band images ever obtained and will provide a crucial complement to the deep \textit{HST} and \textit{Spitzer} imaging of the Frontier Fields.
+
+We conclude here with a brief comment on the ``HUGS'' ultra-deep HAWK-I images of the CANDELS survey fields.
+\cite{hugs} report a final depth of 26.5 AB for point sources within $D=0\farcs4$ apertures for their deepest field, ``GOODS-D1'', which has an exposure time of 31.5 hours similar to the deep HAWK-I fields summarized in Table~\ref{tab:summary}.
+\cite{hugs} assume ideal noise properties that likely suffer some residual correlations between adjacent pixels such as that indicated by the red curves in Fig.~\ref{fig:corr_noise}, where we found that the drizzled rms was some 40\% higher than that measured in the presence of the pixel correlations for $0\farcs1$ pixel mosaics.
+This effect ($\sim$0.4 mag) along with the relative exposure times ($\sim$0.1 mag) can account for much of the difference between the reported depths of the deepest HUGS pointings and the \textit{Hubble} Frontier Fields HAWK-I images described here.
+The comparison must be made because any additional half magnitude increase depth is exponentially more expensive to obtain!
 
@@
 \label{s:number}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%   Cutout figure
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[!ht]
+\epsscale{1.18}
+\plotone{image_cutouts.pdf} % Figures/exposure_map.pdf
+\caption{Image cutouts of the MACS 0416 parallel field.  The \textit{RGB} and monochrome panels come from the sources as indicated; the cutouts are all $10\arcsec$ \xxx\ [check] on a side.
+\label{fig:cutout}}  
+\end{figure}
 

### 2016-02-22 Paper0/ms.tex
further fix text for latest analysis

@@
 
-We present ultra-deep $K_s$-band imaging of the \textit{Hubble} Frontier Fields clusters Abell 2744, MACS-0416, MACS-0717 and MACS-1149 (\xxx\ include MOSFIRE???), which have recently been observed with large allocations of Directors' Discretionary Time with the \textit{HST} and \textit{Spitzer} telescopes covering $0.4 < \lambda < 4.5$\,\mum.
-\textit{VLT}/HAWKI integrations of the first two fields reach 5-$\sigma$ limiting depths of $K_s=26.3$ (AB) and have seeing FWHM$\sim0\farcs4$.
-Shorter MOSFIRE integrations of the latter two northern fields reach limiting depths $K_s=$\xxx\ with seeing FWHM=\xxx.
+We present ultra-deep $K_s$-band imaging of all six of the \textit{Hubble} Frontier Fields clusters Abell 2744, MACS-0416, Abell S1063, Abell 370, MACS-0717 and MACS-1149, which have recently been observed with large allocations of Directors' Discretionary Time with the \textit{HST} and \textit{Spitzer} telescopes covering $0.4 < \lambda < 1.6$\,\mum and 3.6--4.5\,\mum, respectively.
+\textit{VLT}/HAWKI integrations of the first four fields reach 5-$\sigma$ limiting depths of $K_s\sim26.0$ (AB, point sources) and have seeing FWHM$\sim0\farcs4$.
+Shorter MOSFIRE integrations of the MACS-0717 (MACS-1149) field better observable in the north reach limiting depths $K_s$=25.5 (25.1) with seeing FWHM$\sim$$0\farcs4$ ($0\farcs5$).
 In all cases the $K_s$-band mosaics cover the primary cluster and parallel \textit{HST}/ACS+WFC3 fields, and the total area of the $K_s$-band coverage is \xxx\ arcmin$^2$.
 The $K_s$-band at 2.2\,\mum\ crucially fills the gap between the reddest HST filter (1.6\,\mum\,$\sim H$) and the IRAC 3.6\,\mum\ passband.
-While reaching the full depths of the space-based imaging is not currently feasible from the ground, the deep $K_s$-band images provide significant constraints on both the redshifts and the stellar-population properties of galaxies extending well below the characteristic stellar mass across most of the age of the universe, down to, and including, the redshifts of the targeted galaxy clusters ($z \lesssim 0.5$).
-Reduced, aligned mosaics of the four survey fields are provided accompanying this manuscript.
-Time has been awarded for similar HAWKI coverage of the final two Frontier Fields Abell 370 and Abell S1063, with observations expected by the end of 2015.
+While reaching the full depths of the space-based imaging is not currently feasible from the ground, the deep $K_s$-band images provide significant constraints on both the redshifts and the stellar population properties of galaxies extending well below the characteristic stellar mass across most of the age of the universe, down to, and including, the redshifts of the targeted galaxy clusters ($z \lesssim 0.5$).
+Reduced, aligned mosaics of all six survey fields are provided accompanying this manuscript.
 
@@
 MACS-0416    & $04:16:08.9$ & $-24:04:28$ & \textit{VLT}/HAWKI     & 2013 Oct 25--2014 Feb 23 & 25.5 h & 25.9 & $0\farcs38$ \\ 
-Abell S1063  & $22:49:01.1$ & $-44:32:13$ & \textit{VLT}/HAWKI     & 2015 Jul 8--Sep 22 & 25.5 h & 25.9 & $0\farcs38$ \\ 
-Abell 370    & $02:40:03.3$ & $-01:36:23$ & \textit{VLT}/HAWKI     & 2015 Jul 26--2016 Jan 28 & 25.5 h & 25.9 & $0\farcs38$ \\ 
+Abell S1063  & $22:49:01.1$ & $-44:32:13$ & \textit{VLT}/HAWKI     & 2015 Jul 8--Sep 22       & \xxx & \xxx & \xxx \\ 
+Abell 370    & $02:40:03.3$ & $-01:36:23$ & \textit{VLT}/HAWKI     & 2015 Jul 26--2016 Jan 28 & \xxx & \xxx & \xxx \\ 
 MACS-0717    & $07:17:34.0$ & $+37:44:49$ & \textit{Keck}/MOSFIRE  & 2015 Jan 26--27, 2016 Jan 21--22  & 0/0 h\footnotemark[2] & 25.0/25.0 & $0\farcs43$/$0\farcs00$ \\ 
@@
 \caption{Comparison of bright-star photometry to the 2MASS public catalog.
-The Abell 2744 and MACS 0416 fields at high galactic latitude have few stars for cross calibration, and the photometric calibration is determined from observations of standard stars on nights with photometric transparency conditions.
-At relatively low galactic latitude ($b=+21$), the MACS 0717 field has many bright stars, and the photometric calibration in that field is defined from the 2MASS comparison (hence, $\Delta_\mathrm{2mass}\equiv0$, as indicated for that field).
+The photometric calibration of the $K_s$-band images were determined from standard star observations taken commensurate with the science exposures.
+The uncertainties shown are taken from the 2MASS catalog; these stars are measured at very high S/N in the deep $K_s$-band mosaics.
 We estimate the overall systematic uncertainty in the photometric calibration to be $\lesssim0.05~\mathrm{mag}$.\label{fig:2mass}}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%   Image quality figure
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}[!ht]
-\epsscale{1.18}
-\plotone{curve_of_growth.pdf} % Figures/exposure_map.pdf
-\caption{Curves of growth of stellar profiles in the Frontier Fields $K_s$-band mosaics.
-The profiles of stars are well-fit by Moffat profiles with $\mathrm{FWHM}\sim0\farcs4$ and $\beta\sim2$, which have substantially more flux at radii $r>\mathrm{FWHM}$ than Gaussian profiles with the same half-widths (orange curves).
-These extended profiles appear to be characteristic of deep $K_s$-band images; the stellar profiles from the HUGS GOODS-South D2 HAWK-I $K_s$ image \citep{hugs} (lower right panel), created by a different team with a different reduction pipeline, are nearly identical to those of the Frontier Fields mosaics presented here.
-\xxx\ I only show the parallel field region of MACS0717, which has significantly better image quality from the Jan.
-run.
-You can see a few of the broader profiles in the figure; there are many more of these if you include the whole cluster area.
-\label{fig:cog}}
-\end{figure*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %                                                                %
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%   Drizzle figure
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[!t]
+\epsscale{1.18}
+\plotone{ivar_mosaics.pdf} % Figures/exposure_map.pdf
+\caption{Inverse variance maps of the $K_s$-band mosaics.
+The orientation of the fields is the same as that shown in Fig.~\ref{fig:layout}.
+The four HAWK-I chips and two MOSFIRE pointings are clearly visible for the first four and last two fields, respectively.
+\label{fig:ivar_mosaics}}  
+\end{figure}
+
 A crucial innovation of the image processing of the present analysis compared to previous works is that we combine all of the individual background-subtracted raw images (e.g., \xxx\ 4000 files for 1000 exposures times 4 detectors for the \mbox{Abell 2744} field) into the final output mosaic with the ``drizzle'' algorithm (\citealp{fruchter:02}, as implemented in the DrizzlePac/AstroDrizzle package; \citealp{drizzlepac}).
-The benefits of the drizzle algorithm  preserving robust noise properties of the output mosaics will be discussed in detail below in \S\xxx.
-Here we simply indicate that the drizzle implementation provides additional benefits in 1) trivial application of relative weights of the input images in creating the final mosaics, 2) easily applying the astrometric alignment and geometric distortion stored in the FITS headers, and 3) allowing the definition of an arbitrary output pixel grid without any additional image resampling.
+The benefits of the drizzle algorithm  preserving robust noise properties of the output mosaics will be discussed in detail below in \S\ref{s:depth}.
+Here we simply indicate that the drizzle implementation provides additional benefits in 1) trivial application of relative weights of the input images in creating the final mosaics, 2) providing infrastructure to apply the astrometric alignment and geometric distortion stored in the FITS headers, and 3) allowing the definition of an arbitrary output pixel grid without any additional image resampling.
 
-Final mosaics of each field with $0\farcs1$ pixels are drizzled from the input exposures, $i$, with weights
-\[
+Final mosaics of each field with $0\farcs1$ pixels are drizzled from the input exposures with weights
+\begin{equation}
 w^{-1} = F^2\cdot B \cdot \mathrm{FWHM}^\alpha / \sqrt{1-e^2}, 
-\]
-where $F$ is the factor to scale the exposure to the photometric system ($F\equiv1$ for the OB taken under photometric conditions, $F\gtrsim1$ otherwise), $B$ is the measured background, FWHM is the full width at half maximum of stars identified in the field, and $e$ is the ellipticity of the stellar point spread function (PSF).
+\label{eq:weight}
+\end{equation}
+where $F$ is the factor to scale the exposure to the photometric system ($F\equiv1$ for the average of exposures in the OB taken under photometric conditions, $F\gtrsim1$ otherwise), $B$ is the measured background, FWHM is the full width at half maximum of stars identified in the field, and $e$ is the ellipticity of the stellar point spread function (PSF).
 The parameter $\alpha$ allows for optimizing the image quality of the final mosaic.
@@
 The mosaics are shown in Fig.~\ref{fig:layout}, along with the position of the deep Frontier Fields \textit{HST} imaging fields.
-Inverse variance maps are shown in Appendix \xxx\.
-The final image and inverse-variance mosaics are provided as an ESO Phase 3 data product \xxx\ URL, as well as attached as a digital product to this manuscript (\xxx\ I think ApJSS can and wants to do this).
+We also compute the robust NMAD scatter \citep{brammer:08} of empty sky pixels within the same cells to empirically calibrate the inverse variance maps that are shown in Fig.~\ref{fig:ivar_mosaics} (see also \S\ref{s:depth}).
+The final image and inverse variance mosaics are provided as an ESO Phase 3 data product \xxx\ URL, as well as attached as a digital product to this manuscript (\xxx\ I think ApJSS can and wants to do this).
 The released images are all scaled to a common zeropoint of 26.0 (AB mag), which gives pixel values of order unity for faint galaxies.
 In the sections below we describe the characteristics of the mosaics in more detail.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%   Image quality figure
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure*}[!ht]
+\epsscale{1.18}
+\plotone{curve_of_growth.pdf} % Figures/exposure_map.pdf
+\caption{Curves of growth of stellar profiles in the Frontier Fields $K_s$-band mosaics.
+The inverse aperture corrections for point sources in $D=0\farcs6$ apertures are indicated in each panel.
+The profiles of stars are well-fit by Moffat profiles with $\mathrm{FWHM}\sim0\farcs4$ in most cases and $\beta\sim2.1$.
+These extended Moffat profiles appear to be characteristic of deep $K_s$-band images and they have substantially more flux at large radii ($r>\mathrm{FWHM}/2$) than Gaussian profiles with the same FWHM (orange curves in the left panels).
+\label{fig:cog}}
+\end{figure*}
 
@@
 Fig.~\ref{fig:cog} shows the curves of growth for stars identified by the tight relationship between their brightness and half-light radii \citep[see, e.g., Fig.~13 of][]{skelton:14}.
-The image quality of the $K_s$-band mosaics is excellent, with stellar FWHM $0\farcs40$ and $0\farcs38$ in the Abell 2744 and MACS 0416 fields, respectively, thanks, in large part, to the execution of the HAWK-I observations in service mode ensuring optimal and uniform image quality across the many hours of integration on the survey fields.
+The image quality of the $K_s$-band mosaics is excellent with stellar FWHM$\lesssim0\farcs4$ for the deep HAWK-I fields, thanks in large part to the execution of the HAWK-I observations in service mode ensuring optimal and uniform image quality across the many hours of integration on the survey fields.
+The classical scheduling of the MOSFIRE observations does not allow such control over the seeing conditions.
+The result is that the image quality of the MACS-0717 and MACS-1149 fields is somewhat degraded in comparison, at $0\farcs42$--$0\farcs54$, and the image quality differs measurably between the two MOSFIRE pointings in the MACS-0717 field.
 Despite the compact cores of the stellar PSFs, the curves of growth shown in Fig.~\ref{fig:cog} deviate from Gaussian profiles with significantly higher flux in the outer wings, with a shape more consistent with a Moffat profile \citep{trujillo:01}  with $\beta\sim2$.
-Also shown in Fig.~\ref{fig:cog} are stellar curves of growth measured from the HAWK-I UDS and GOODS Survey \citep[HUGS;][]{hugs}, which are nearly identical in both FWHM and profile shape.
 We caution that these extended profiles will yield significantly shallower final image depths than would be predicted with Exposure Time Calculators that may assume Gaussian profiles.
-\xxx\ not clear what HAWK-I ETC assumes, the description just says ``FWHM of the seeing disk'' \xxx.
-Nevertheless, the deep $K_s$ mosaics have excellent image quality that is much closer to the resolution of the \textit{HST} IR imaging than the redder \textit{Spitzer} IRAC bands.
-
-\xxx\ add description of MOSFIRE clusters.  IQ not uniform across the MACS0717 field.
+Nevertheless, the deep $K_s$ mosaics have excellent image quality that is much closer to the resolution of the \textit{HST} IR imaging and the typical apparent size of distant galaxies \citep[median $r_e\sim0\farcs1$--$0\farcs2$ at $z>4$;][]{shibuya:15} than the redder \textit{Spitzer} IRAC bands (cf.
+FWHM$\sim$$1\farcs7$--$2^{\prime\prime}$).
 
@@
 \plotone{aper_depth.pdf} % Figures/exposure_map.pdf
-\caption{\textit{Top panel:} \xxx Noise in empty apertures as a function of aperture diameter, normalized to the smallest $D=0\farcs2$ aperture.
-The dashed and dotted lines show the expectations for uncorrelated and perfectly-correlated noise within the aperture.
-\textit{Bottom panel:} The product of the inverse curves-of-growth (Fig.\ref{fig:cog} and the empty aperture noise gives the point-source sensitivity as a function of aperture size.
-The S/N is maximized with an aperture just larger than the FWHM.
-For reference, the colored dashed lines in the bottom panel indicate the depth computed assuming uncorrelated noise within the apertures ($\sigma \propto D$) and the per-pixel variance reported in the weight maps.)
+\caption{\textit{Top panel:} The product of the inverse curves-of-growth (Fig.~\ref{fig:cog}) and the predicted variance within a photometric aperture (Eq.\ref{eq:aper_var}) gives the point-source sensitivity as a function of aperture size.
+The S/N is maximized with an aperture somewhat larger than the FWHM; depths evaluated at $D=0\farcs6$ near the maxima are indicated.)
 \label{fig:depth}}  
@@
 \cite{casertano:00} discuss how the noise properties of correlated output pixels are affected by the relative sizes of the input and output pixel grids and the adopted drizzle parameters.
-With the drizzle algorithm and a small pixfrac, a given input pixel will map to only a single output pixel and therefore correlations between the output pixels should be minimal\footnote{Inter-pixel capacitance and other forms of crosstalk intrinsic to the IR detectors effectively place a lower limit on the correlations between adjacent pixels \citep{hilbert:ipc}.}.
+With the drizzle algorithm and a small pixfrac, a given input pixel will map to only a single output pixel and therefore correlations between the output pixels should be minimal\footnote{Inter-pixel capacitance and other forms of crosstalk intrinsic to the IR detectors place a lower limit on the correlations between adjacent pixels \citep{finger:05, hilbert:ipc}.}.
 
@@
 to be decreased at all separations.
-That the pixel variances can't be trusted in the presence of such correlations is the primary reason why significant effort has been devoted to placing random ``empty apertures'' across images in order to characterize their noise properties \citep[e.g., ]{quadri:musyc, whitaker:nmbs, skelton:14}.
-Now by eliminating the pixel correlations with the drizzle combination (with a small pixfrac) we have shown that the pixel statistics are robust and it is now trivial to compute the expected variance within an arbitrary photometric aperture, for example:
+That the pixel variances can't be used directly in the presence of such correlations \citep{casertano:00} is the primary reason why significant effort has been devoted to placing random ``empty apertures'' across images in order to characterize their noise properties \citep[e.g.,][]{quadri:musyc, whitaker:nmbs, skelton:14}.
+Now by eliminating the pixel correlations with the drizzle image combination technique (with a small pixfrac) we have shown that the pixel statistics are robust and it is now trivial to compute the expected variance within an arbitrary photometric aperture, for example:
 \begin{equation}
-\sigma_D^2 = \sigma_\mathrm{pix}^2\cdot \pi/4 \cdot D^2,
+\sigma_{\mathrm{aper}, D}^2 = \sigma_\mathrm{pix}^2\cdot \pi/4 \cdot D^2,
 \label{eq:aper_var}
@@
 For our $K_s$-band mosaics, this optimal aperture has $D\sim0\farcs6$, just larger than the FWHM of point-source profiles.
-The $5\,\sigma$ limiting magnitudes (AB) within $0\farcs6$ diameter apertures are \xxx\ 25.8, 25.9, and 25.0 in the Abell 2744, MACS-0416, and MACS-0717 fields, respectively.
-The HAWK-I images presented here are among the deepest $K_s$-band images ever obtained and will provide a crucial complement to the deep \textit{HST} and \textit{Spitzer} imaging of the Frontier Fields.
+The $5\,\sigma$ limiting magnitudes (AB) within $0\farcs6$ diameter apertures are indicated in Fig.~\ref{fig:depth} and listed in Table~\ref{tab:summary}.
+Reaching 26th magnitude (AB), the HAWK-I images presented here are among the deepest $K_s$-band images ever obtained and will provide a crucial complement to the deep \textit{HST} and \textit{Spitzer} imaging of the Frontier Fields.
 
-We conclude here with a brief comment on the ``HUGS'' ultra-deep HAWK-I images of the CANDELS survey fields.
+We conclude here with a brief comment on the ``HUGS'' ultra-deep HAWK-I images of the CANDELS GOODS-S and UDS survey fields.
 \cite{hugs} report a final depth of 26.5 AB for point sources within $D=0\farcs4$ apertures for their deepest field, ``GOODS-D1'', which has an exposure time of 31.5 hours similar to the deep HAWK-I fields summarized in Table~\ref{tab:summary}.
-\cite{hugs} assume ideal noise properties that likely suffer some residual correlations between adjacent pixels such as that indicated by the red curves in Fig.~\ref{fig:corr_noise}, where we found that the drizzled rms was some 40\% higher than that measured in the presence of the pixel correlations for $0\farcs1$ pixel mosaics.
-This effect ($\sim$0.4 mag) along with the relative exposure times ($\sim$0.1 mag) can account for much of the difference between the reported depths of the deepest HUGS pointings and the \textit{Hubble} Frontier Fields HAWK-I images described here.
+\cite{hugs} assume ideal noise properties that likely suffer some residual correlation between adjacent pixels such as that indicated by the red curves in Fig.~\ref{fig:corr_noise}, where we found that the drizzled rms was some 40\% higher than that measured in the presence of the pixel correlations for $0\farcs1$ pixel mosaics.
+This effect ($\sim$0.4 mag) along with the relative exposure time difference ($\sim$0.1 mag) can account for much of the difference between the reported depths of the deepest HUGS pointings and the \textit{Hubble} Frontier Fields HAWK-I images described here.
 The comparison must be made because any additional half magnitude increase depth is exponentially more expensive to obtain!
@@
 \plotone{image_cutouts.pdf} % Figures/exposure_map.pdf
-\caption{Image cutouts of the MACS 0416 parallel field.  The \textit{RGB} and monochrome panels come from the sources as indicated; the cutouts are all $10\arcsec$ \xxx\ [check] on a side.
+\caption{Image cutouts of the MACS 0416 parallel field.
+The \textit{RGB} and monochrome panels come from the sources as indicated; the cutouts are all $10\arcsec$ \xxx\ \textcolor{red}{check, different now} on a side.
 \label{fig:cutout}}  

### 2016-02-23 Paper0/ms.tex
minor updates
Updated table

@@
 %\topmargin 0.6in
+
+\usepackage{multirow}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{deluxetable*}{ccccrccc} 
+\begin{deluxetable*}{lcccrccccc} 
 \tabletypesize{\footnotesize}
@@
 \tablehead{ 
-\colhead{Field} & \colhead{R.A.} & \colhead{Dec.} & \colhead{Instrument} & \colhead{Epoch} & \colhead{$t$} & \colhead{Depth$^\mathrm{a}$} & \colhead{FWHM}}
+\colhead{Field} & \colhead{R.A.} & \colhead{Dec.} & \colhead{Instrument} & \colhead{Epoch} & Field\footnotemark[1] & \colhead{$t$, hours} & \colhead{$t_{0.4}$\footnotemark[2]} & \colhead{Depth\footnotemark[3]} & \colhead{FWHM}}
 \startdata 
-Abell 2744   & $00:14:21.2$ & $-30:23:50$ & \textit{VLT}/HAWKI     & 2013 Oct 24--2013 Dec 24 & 29.4 h & 25.8 & $0\farcs40$ \\ 
-MACS-0416    & $04:16:08.9$ & $-24:04:28$ & \textit{VLT}/HAWKI     & 2013 Oct 25--2014 Feb 23 & 25.5 h & 25.9 & $0\farcs38$ \\ 
-Abell S1063  & $22:49:01.1$ & $-44:32:13$ & \textit{VLT}/HAWKI     & 2015 Jul 8--Sep 22       & \xxx & \xxx & \xxx \\ 
-Abell 370    & $02:40:03.3$ & $-01:36:23$ & \textit{VLT}/HAWKI     & 2015 Jul 26--2016 Jan 28 & \xxx & \xxx & \xxx \\ 
-MACS-0717    & $07:17:34.0$ & $+37:44:49$ & \textit{Keck}/MOSFIRE  & 2015 Jan 26--27, 2016 Jan 21--22  & 0/0 h\footnotemark[2] & 25.0/25.0 & $0\farcs43$/$0\farcs00$ \\ 
-MACS-1149    & $11:49:36.3$ & $+22:23:58$ & \textit{Keck}/MOSFIRE  & 2015 Feb 24, 2016 Jan 21--22      & \xxx\footnotemark[2] & \xxx & \xxx \\
-''           & ''           & ''          & \textit{VLT}/HAWKI\footnotemark[3]     & 2013 Mar 21--2014 Jun 9      & \xxx & \xxx & \xxx
+Abell 2744   & $00:14:21.2$ & $-30:23:50$ & \textit{VLT}/HAWKI     & 2013 Oct 24--2013 Dec 24 & C+P & 29.3 & 20.5 & 26.0 & $0\farcs39$ \\ 
+MACS-0416    & $04:16:08.9$ & $-24:04:28$ & \textit{VLT}/HAWKI     & 2013 Oct 25--2014 Feb 23 & C+P & 25.8 & 21.7 & 26.0 & $0\farcs36$ \\ 
+Abell S1063  & $22:49:01.1$ & $-44:32:13$ & \textit{VLT}/HAWKI     & 2015 Jul 8--Sep 22       & C+P & 27.9 & 21.0 & 26.0 & $0\farcs39$ \\ 
+Abell 370    & $02:40:03.3$ & $-01:36:23$ & \textit{VLT}/HAWKI     & 2015 Jul 26--2016 Jan 28 & C+P & 29.1 & \xxx & 25.9 & $0\farcs36$ \\ 
+\vspace {-0.2cm}\\ 
+\multirow{2}{*}{MACS-0717}    & \multirow{2}{*}{$07:17:34.0$} & \multirow{2}{*}{$+37:44:49$} &  \multirow{2}{*}{\textit{Keck}/MOSFIRE}  &  \multirow{2}{*}{2015 Jan 26--27, 2016 Jan 21--22}  & C & 4.3 & 1.9 & 25.3 & $0\farcs42$ \\
+                              &                             &                             &                                          &      & P & 3.8 & 2.2 & 25.5 & $0\farcs49$ \\ 
+\vspace {-0.2cm}\\ 
+\multirow{4}{*}{MACS-1149}    & \multirow{4}{*}{$11:49:36.3$} & \multirow{4}{*}{$+22:23:58$} & \multirow{2}{*}{\textit{Keck}/MOSFIRE}  & \multirow{2}{*}{2015 Feb 24, 2016 Jan 21--22}      & C & 5.5 & 1.9 & 25.2 & $0\farcs53$ \\
+&  &         &      &   & P & 4.8 & 2.2 & 25.1 & $0\farcs54$ \\
+\vspace {-0.3cm}\\ 
+&  &         & \textit{VLT}/HAWKI\footnotemark[4]     & 2013 Mar 21--2014 Jun 9      & C & 5.3 & 3.1 & 25.0 & $0\farcs41$
 \enddata 
 %\vspace{-0.8cm} 
-\footnotetext[1]{Depth is defined as the 5$\sigma$ limiting magnitude for point sources, measured in $D=0\farcs6$ apertures (\S\ref{s:depth}).} 
-\footnotetext[2]{The MOSFIRE observations of MACS-0717 and MACS-1149 require two separate pointings of the instrument to cover the cluster/parallel \textit{HST} fields, which have the different characteristics as indicated, respectively.}
-\footnotetext[3]{Archival observations from ESO program \href{http://archive.eso.org/wdb/wdb/eso/eso_archive_main/query?prog_id=090.A-0458(A)}{090.A-0458} (PI: Infante).}
+\footnotetext[1]{Coverage of \textit{Hubble} survey fields: C=Cluster, P=Parallel.
+Most HAWKI pointings cover the two \textit{HST} fields simultaneously (C+P; Fig.~\ref{fig:layout}).
+The MOSFIRE observations of MACS-0717 and MACS-1149 require two separate pointings of the instrument to cover the cluster and parallel \textit{HST} fields, which have the different characteristics as indicated.}
+\footnotetext[2]{Exposure time equivalent at $0\farcs4$ seeing:  $t_{0.4}=\sum t \cdot (0.4/\mathrm{FWHM})^2$.}
+\footnotetext[3]{Depth is defined as the 5$\sigma$ limiting magnitude for point sources, measured in $D=0\farcs6$ apertures (\S\ref{s:depth}).} 
+\footnotetext[4]{Archival observations from ESO program \href{http://archive.eso.org/wdb/wdb/eso/eso_archive_main/query?prog_id=090.A-0458(A)}{090.A-0458} (PI: Infante).}
 \end{deluxetable*}
@@
 \caption{Comparison of bright-star photometry to the 2MASS public catalog.
-The photometric calibration of the $K_s$-band images were determined from standard star observations taken commensurate with the science exposures.
+The photometric calibration of the $K_s$-band images were determined from standard star observations taken concurrently with the science exposures.
 The uncertainties shown are taken from the 2MASS catalog; these stars are measured at very high S/N in the deep $K_s$-band mosaics.

### 2016-02-24 Paper0/ms.tex
updated A370 mosaic will all data

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{deluxetable*}{lcccrccccc} 
+\begin{deluxetable*}{ccccrccccc} 
 \tabletypesize{\footnotesize}
@@
 \tablehead{ 
-\colhead{Field} & \colhead{R.A.} & \colhead{Dec.} & \colhead{Instrument} & \colhead{Epoch} & Field\footnotemark[1] & \colhead{$t$, hours} & \colhead{$t_{0.4}$\footnotemark[2]} & \colhead{Depth\footnotemark[3]} & \colhead{FWHM}}
+\colhead{Field} & \colhead{R.A.} & \colhead{Dec.} & \colhead{Instrument} & \colhead{Epoch} & \textit{HST}\footnotemark[1] & \colhead{$t$, hours} & \colhead{Depth\footnotemark[2]} & \colhead{FWHM}}
 \startdata 
-Abell 2744   & $00:14:21.2$ & $-30:23:50$ & \textit{VLT}/HAWKI     & 2013 Oct 24--2013 Dec 24 & C+P & 29.3 & 20.5 & 26.0 & $0\farcs39$ \\ 
-MACS-0416    & $04:16:08.9$ & $-24:04:28$ & \textit{VLT}/HAWKI     & 2013 Oct 25--2014 Feb 23 & C+P & 25.8 & 21.7 & 26.0 & $0\farcs36$ \\ 
-Abell S1063  & $22:49:01.1$ & $-44:32:13$ & \textit{VLT}/HAWKI     & 2015 Jul 8--Sep 22       & C+P & 27.9 & 21.0 & 26.0 & $0\farcs39$ \\ 
-Abell 370    & $02:40:03.3$ & $-01:36:23$ & \textit{VLT}/HAWKI     & 2015 Jul 26--2016 Jan 28 & C+P & 29.1 & \xxx & 25.9 & $0\farcs36$ \\ 
+Abell 2744   & $00:14:21.2$ & $-30:23:50$ & \textit{VLT}/HAWKI     & 2013 Oct 24--2013 Dec 24 & C+P & 29.3 &  26.0 & $0\farcs39$ \\ 
+MACS-0416    & $04:16:08.9$ & $-24:04:28$ & \textit{VLT}/HAWKI     & 2013 Oct 25--2014 Feb 23 & C+P & 25.8 &  26.0 & $0\farcs36$ \\ 
+Abell S1063  & $22:49:01.1$ & $-44:32:13$ & \textit{VLT}/HAWKI     & 2015 Jul 8--Sep 22       & C+P & 27.9 &  26.0 & $0\farcs39$ \\ 
+Abell 370    & $02:40:03.3$ & $-01:36:23$ & \textit{VLT}/HAWKI     & 2015 Jul 26--2016 Jan 28 & C+P & 28.3 &  26.0 & $0\farcs35$ \\ 
 \vspace {-0.2cm}\\ 
-\multirow{2}{*}{MACS-0717}    & \multirow{2}{*}{$07:17:34.0$} & \multirow{2}{*}{$+37:44:49$} &  \multirow{2}{*}{\textit{Keck}/MOSFIRE}  &  \multirow{2}{*}{2015 Jan 26--27, 2016 Jan 21--22}  & C & 4.3 & 1.9 & 25.3 & $0\farcs42$ \\
-                              &                             &                             &                                          &      & P & 3.8 & 2.2 & 25.5 & $0\farcs49$ \\ 
+\multirow{2}{*}{MACS-0717}    & \multirow{2}{*}{$07:17:34.0$} & \multirow{2}{*}{$+37:44:49$} &  \multirow{2}{*}{\textit{Keck}/MOSFIRE}  &  \multirow{2}{*}{2015 Jan 26--27, 2016 Jan 21--22}  & C & 4.3 & 25.3 & $0\farcs42$ \\
+                              &                             &                             &                                          &      & P & 3.8 & 25.5 & $0\farcs49$ \\ 
 \vspace {-0.2cm}\\ 
-\multirow{4}{*}{MACS-1149}    & \multirow{4}{*}{$11:49:36.3$} & \multirow{4}{*}{$+22:23:58$} & \multirow{2}{*}{\textit{Keck}/MOSFIRE}  & \multirow{2}{*}{2015 Feb 24, 2016 Jan 21--22}      & C & 5.5 & 1.9 & 25.2 & $0\farcs53$ \\
-&  &         &      &   & P & 4.8 & 2.2 & 25.1 & $0\farcs54$ \\
+\multirow{4}{*}{MACS-1149}    & \multirow{4}{*}{$11:49:36.3$} & \multirow{4}{*}{$+22:23:58$} & \multirow{2}{*}{\textit{Keck}/MOSFIRE}  & \multirow{2}{*}{2015 Feb 24, 2016 Jan 21--22}      & C & 5.5 & 25.2 & $0\farcs53$ \\
+&  &         &      &   & P & 4.8 & 25.1 & $0\farcs54$ \\
 \vspace {-0.3cm}\\ 
-&  &         & \textit{VLT}/HAWKI\footnotemark[4]     & 2013 Mar 21--2014 Jun 9      & C & 5.3 & 3.1 & 25.0 & $0\farcs41$
+&  &         & \textit{VLT}/HAWKI\footnotemark[3]     & 2013 Mar 21--2014 Jun 9      & C & 5.3 & 25.0 & $0\farcs41$
 \enddata 
@@
 The MOSFIRE observations of MACS-0717 and MACS-1149 require two separate pointings of the instrument to cover the cluster and parallel \textit{HST} fields, which have the different characteristics as indicated.}
-\footnotetext[2]{Exposure time equivalent at $0\farcs4$ seeing:  $t_{0.4}=\sum t \cdot (0.4/\mathrm{FWHM})^2$.}
-\footnotetext[3]{Depth is defined as the 5$\sigma$ limiting magnitude for point sources, measured in $D=0\farcs6$ apertures (\S\ref{s:depth}).} 
-\footnotetext[4]{Archival observations from ESO program \href{http://archive.eso.org/wdb/wdb/eso/eso_archive_main/query?prog_id=090.A-0458(A)}{090.A-0458} (PI: Infante).}
+%\footnotetext[2]{Approximate exposure time equivalent at $0\farcs4$ seeing:  $t_{0.4}=\sum_i t_i \cdot (0.4/\mathrm{FWHM}_i)^2$.}
+\footnotetext[2]{Depth is defined as the 5$\sigma$ limiting magnitude for point sources, measured in $D=0\farcs6$ apertures (\S\ref{s:depth}).} 
+\footnotetext[3]{Archival observations from ESO program \href{http://archive.eso.org/wdb/wdb/eso/eso_archive_main/query?prog_id=090.A-0458(A)}{090.A-0458} (PI: Infante).}
 \end{deluxetable*}
@@
 \begin{figure*}
+\epsscale{1.18}
 \plotone{hawkiff_layout.pdf} % Figures/exposure_map.pdf
@@
 The HAWK-I field of view is perfectly suited (Abell 2744, MACS-0416, Abell S1063 and Abell 370) for simultaneous imaging of the cluster+parallel field pairs, which require two separate pointings with MOSFIRE (MACS-0717 and MACS-1149).
-Additional archival HAWK-I coverage of the MACS-1149 field is indicated with the gray square in the lower right panel.\label{fig:layout}}
+The footprint of additional archival HAWK-I coverage of the MACS-1149 field is indicated by the gray square in the lower right panel.\label{fig:layout}}
 \end{figure*}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%   Image quality figure
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure*}[!ht]
+\epsscale{1.18}
+\plotone{curve_of_growth.pdf} % Figures/exposure_map.pdf
+\caption{Curves of growth of stellar profiles in the Frontier Fields $K_s$-band mosaics.
+The inverse aperture corrections for point sources in $D=0\farcs6$ apertures are indicated in each panel.
+The profiles of stars are well-fit by Moffat profiles with $\mathrm{FWHM}\sim0\farcs4$ in most cases and $\beta\sim2.1$.
+These extended Moffat profiles appear to be characteristic of deep $K_s$-band images and they have substantially more flux at large radii ($r>\mathrm{FWHM}/2$) than Gaussian profiles with the same FWHM (orange curves in the left panels).
+\label{fig:cog}}
+\end{figure*}
+
 A crucial innovation of the image processing of the present analysis compared to previous works is that we combine all of the individual background-subtracted raw images (e.g., \xxx\ 4000 files for 1000 exposures times 4 detectors for the \mbox{Abell 2744} field) into the final output mosaic with the ``drizzle'' algorithm (\citealp{fruchter:02}, as implemented in the DrizzlePac/AstroDrizzle package; \citealp{drizzlepac}).
@@
 In the sections below we describe the characteristics of the mosaics in more detail.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%   Image quality figure
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}[!ht]
-\epsscale{1.18}
-\plotone{curve_of_growth.pdf} % Figures/exposure_map.pdf
-\caption{Curves of growth of stellar profiles in the Frontier Fields $K_s$-band mosaics.
-The inverse aperture corrections for point sources in $D=0\farcs6$ apertures are indicated in each panel.
-The profiles of stars are well-fit by Moffat profiles with $\mathrm{FWHM}\sim0\farcs4$ in most cases and $\beta\sim2.1$.
-These extended Moffat profiles appear to be characteristic of deep $K_s$-band images and they have substantially more flux at large radii ($r>\mathrm{FWHM}/2$) than Gaussian profiles with the same FWHM (orange curves in the left panels).
-\label{fig:cog}}
-\end{figure*}
 

### 2016-02-25 Paper0/ms.tex
minor change
remove old comment

@@
 
-\xxx\ currently a bit awkward since MOSFIRE reduction done quickly with Ivo's code and HAWK-I done in many additional steps (astrometry and photometry).
-
 \subsection{Observation log}
@@
 The light blue polygons in the MACS-0717 field show additional wide-field ACS imaging coverage from  programs GO-9722 and GO-10420 \citep[PI: Ebeling;][]{ma:11}.
-The area covered by deep imaging in the \textit{Spitzer}/IRAC 3.6 and 4.5\,\mum\ channels is shown in orange.
+The area covered by deep imaging in the \textit{Spitzer}/IRAC 3.6 and 4.5\,\mum\ channels is shown in orange, with the cluster fields indicated by the dashed lines.
 The HAWK-I field of view is perfectly suited (Abell 2744, MACS-0416, Abell S1063 and Abell 370) for simultaneous imaging of the cluster+parallel field pairs, which require two separate pointings with MOSFIRE (MACS-0717 and MACS-1149).

### 2016-02-25 Paper0/ms.tex
change DITs

@@
 The execution of each OB lasted either 60 or 90 minutes, depending on how many exposures were obtained in the block.
-The exposures were roughly one minute each, with multiple coadds of shorter reads making up the exposure ($\mathrm{NDIT}\times\mathrm{DIT} = 4\times15$\,s~=~60\,s for Abell 2744, 7$\times$8\,s~=~56\,s for MACS-0416, 7$\times$8\,s~=~56\,s for Abell S1063 and 7$\times$8\,s~=~56\,s for Abell 370\footnote{Longer DITs are preferable to reduce instrument overheads; the shorter DITs on the MACS-0416 field were done do accommodate \textit{VLT} service mode restrictions on bright source saturation, which have been eased since ESO Period 93.}).
+The exposures were roughly one minute each, with multiple coadds of shorter reads making up the exposure ($\mathrm{NDIT}\times\mathrm{DIT} = 4\times15$\,s~=~60\,s for Abell 2744, 7$\times$8\,s~=~56\,s for MACS-0416, 8$\times$12\,s~=~96\,s for Abell S1063 and 8$\times$12\,s~=~96\,s for Abell 370\footnote{Longer DITs are preferable to reduce instrument overheads; the shorter DITs on the MACS-0416 field were done do accommodate \textit{VLT} service mode restrictions on bright source saturation, which have been eased since ESO Period 93.}).
 The telescope was offset with random dithers between each exposure to facilitate sky subtraction;  observations taken before 2013 Nov 4 were dithered within a 20$^{\prime\prime}$ box, which was subsequently increased to 40$^{\prime\prime}$ to improve sky subtraction in the cluster cores crowded with bright galaxies and intra-cluster light.
@@
 The MOSFIRE detector consists of $2048\times2048$ $0\farcs1798$ pixels; the resulting $6\farcm1\times6\farcm1$ field-of-view is unfortunately slightly too small to cover both \textit{HST} fields simultaneously and therefore requires two pointings to cover the entire deep \textit{HST} Frontier Fields imaging area (Fig.~\ref{fig:layout}).
-The 6$\times$7\,s~=~42\,s MOSFIRE exposures were obtained in a $3\times3$ ``Box9'' dither pattern spaced roughly 40$^{\prime\prime}$ between offset positions.
-The exposure times on the MOSFIRE survey fields are summarized in Table~\ref{tab:summary}.
+MOSFIRE exposures were obtained in a $3\times3$ ``Box9'' dither pattern spaced roughly 40$^{\prime\prime}$ between offset positions.
+The total integration times on the MOSFIRE survey fields are summarized in Table~\ref{tab:summary}.
+The individual detector integration times (DIT) were adjusted on the fly during the observing run to keep the total counts within the linear regime of the detector, while maintaining $\mathrm{NDIT}\times\mathrm{DIT} \sim 40$\,s.
 
@@
 The MACS-1149 HAWK-I observations were obtained in Service Mode at the ESO/\textit{VLT} in \xxx\ separate OBs between 2013 Mar 21 and 2014 Jun 9.
-The 7$\times$8\,s~=~56\,s \xxx\ exposures were taken at nine dithered positions with regular offsets separated by \xxx\ arcsec; the on-source exposure time of the final mosaic is \xxx\ hours.
+The 20$\times$15\,s~=~300\,s exposures were taken at nine dithered positions with regular offsets separated by \xxx\ arcsec; the on-source exposure time of the final mosaic is \xxx\ hours.
 

### 2016-04-12 Paper0/ms.tex

@@
 %\usepackage{deluxetable}
-\usepackage{color}
+%\usepackage{color}
+\usepackage[usenames, dvipsnames]{color}
 \definecolor{citeRGB}{rgb}{0,0.1,0.7}
@@
 \gdef\epers{electrons~s$^{-1}$}
-\gdef\xxx{\textbf{xxx}}
+%\gdef\xxx{\textbf{xxx}}
+\newcommand\xxx{{\textcolor{red}{\bf xxx}}}
+
+\gdef\HAWKI{\mbox{HAWK-I}}
 
@@
 
-We present ultra-deep $K_s$-band imaging of all six of the \textit{Hubble} Frontier Fields clusters Abell 2744, MACS-0416, Abell S1063, Abell 370, MACS-0717 and MACS-1149, which have recently been observed with large allocations of Directors' Discretionary Time with the \textit{HST} and \textit{Spitzer} telescopes covering $0.4 < \lambda < 1.6$\,\mum and 3.6--4.5\,\mum, respectively.
-\textit{VLT}/HAWKI integrations of the first four fields reach 5-$\sigma$ limiting depths of $K_s\sim26.0$ (AB, point sources) and have seeing FWHM$\sim0\farcs4$.
+We present ultra-deep $K_s$-band imaging of all six of the \textit{Hubble} Frontier Fields clusters Abell 2744, MACS-0416, Abell S1063, Abell 370, MACS-0717 and MACS-1149, which have recently been observed with large allocations of Directors' Discretionary Time with the \textit{HST} and \textit{Spitzer} telescopes covering $0.4 < \lambda < 1.6$\,\mum\ and 3.6--4.5\,\mum, respectively.
+\textit{VLT}/\HAWKI\ integrations of the first four fields reach 5$\sigma$ limiting depths of $K_s\sim26.0$ (AB, point sources) and have excellent seeing FWHM$\sim0\farcs4$.
 Shorter MOSFIRE integrations of the MACS-0717 (MACS-1149) field better observable in the north reach limiting depths $K_s$=25.5 (25.1) with seeing FWHM$\sim$$0\farcs4$ ($0\farcs5$).
@@
 \startdata 
-Abell 2744   & $00:14:21.2$ & $-30:23:50$ & \textit{VLT}/HAWKI     & 2013 Oct 24--2013 Dec 24 & C+P & 29.3 &  26.0 & $0\farcs39$ \\ 
-MACS-0416    & $04:16:08.9$ & $-24:04:28$ & \textit{VLT}/HAWKI     & 2013 Oct 25--2014 Feb 23 & C+P & 25.8 &  26.0 & $0\farcs36$ \\ 
-Abell S1063  & $22:49:01.1$ & $-44:32:13$ & \textit{VLT}/HAWKI     & 2015 Jul 8--Sep 22       & C+P & 27.9 &  26.0 & $0\farcs39$ \\ 
-Abell 370    & $02:40:03.3$ & $-01:36:23$ & \textit{VLT}/HAWKI     & 2015 Jul 26--2016 Jan 28 & C+P & 28.3 &  26.0 & $0\farcs35$ \\ 
+Abell 2744   & $00:14:21.2$ & $-30:23:50$ & \textit{VLT}/\HAWKI     & 2013 Oct 24--2013 Dec 24 & C+P & 29.3 &  26.0 & $0\farcs39$ \\ 
+MACS-0416    & $04:16:08.9$ & $-24:04:28$ & \textit{VLT}/\HAWKI     & 2013 Oct 25--2014 Feb 23 & C+P & 25.8 &  26.0 & $0\farcs36$ \\ 
+Abell S1063  & $22:49:01.1$ & $-44:32:13$ & \textit{VLT}/\HAWKI     & 2015 Jul 8--Sep 22       & C+P & 27.9 &  26.0 & $0\farcs39$ \\ 
+Abell 370    & $02:40:03.3$ & $-01:36:23$ & \textit{VLT}/\HAWKI     & 2015 Jul 26--2016 Jan 28 & C+P & 28.3 &  26.0 & $0\farcs35$ \\ 
 \vspace {-0.2cm}\\ 
@@
 \vspace {-0.3cm}\\ 
-&  &         & \textit{VLT}/HAWKI\footnotemark[3]     & 2013 Mar 21--2014 Jun 9      & C & 5.3 & 25.0 & $0\farcs41$
+&  &         & \textit{VLT}/\HAWKI\footnotemark[3]     & 2013 Mar 21--2014 Jun 9      & C & 5.3 & 25.0 & $0\farcs41$
 \enddata 
@@
 \footnotetext[1]{Coverage of \textit{Hubble} survey fields: C=Cluster, P=Parallel.
-Most HAWKI pointings cover the two \textit{HST} fields simultaneously (C+P; Fig.~\ref{fig:layout}).
+Most \HAWKI\ pointings cover the two \textit{HST} fields simultaneously (C+P; Fig.~\ref{fig:layout}).
 The MOSFIRE observations of MACS-0717 and MACS-1149 require two separate pointings of the instrument to cover the cluster and parallel \textit{HST} fields, which have the different characteristics as indicated.}
@@
 
-$K_s$-band observations of the Abell 2744 and MACS-0416 fields were obtained between 2013 Oct and 2014 Feb with the High Acuity Wide-field K-band Imager \citep[HAWK-I;][]{hawki, hawki2} on the 8.2\,m UT4 telescope at the ESO \textit{Very Large Telescope} (ESO program \mbox{092.A-0472}).
-HAWK-I $K_s$-badn bservations of the Abell S1063 and Abell 370 fields were obtained between 2015 Jul and 2016 Jan from a subsequent program (095.A-0533).
-HAWK-I is composed of four chips with 2048$\times$2048 $0\farcs106$ pixels, and the full $7\farcm5\times7\farcm5$ HAWKI field-of-view is perfectly suited to cover both the primary and parallel \textit{HST} ACS optical and WFC3 IR fields simultaneously in a single pointing (see Fig.\ref{fig:layout}).
+$K_s$-band observations of the Abell 2744 and MACS-0416 fields were obtained between 2013 Oct and 2014 Feb with the High Acuity Wide-field K-band Imager \citep[\HAWKI;][]{hawki, hawki2} on the 8.2\,m UT4 telescope at the ESO \textit{Very Large Telescope} (ESO program \mbox{092.A-0472}).
+\HAWKI\ $K_s$-badn bservations of the Abell S1063 and Abell 370 fields were obtained between 2015 Jul and 2016 Jan from a subsequent program (095.A-0533).
+\HAWKI\ is composed of four chips with 2048$\times$2048 $0\farcs106$ pixels, and the full $7\farcm5\times7\farcm5$ \HAWKI\ field-of-view is perfectly suited to cover both the primary and parallel \textit{HST} ACS optical and WFC3 IR fields simultaneously in a single pointing (see Fig.\ref{fig:layout}).
 
-The HAWK-I observations were divided into individual service mode Observation Blocks (OBs); typically one or two OBs of a given field were observed per night as conditions allowed, and occasionally multiple fields were observed on the same night.
+The \HAWKI\ observations were divided into individual service mode Observation Blocks (OBs); typically one or two OBs of a given field were observed per night as conditions allowed, and occasionally multiple fields were observed on the same night.
 The execution of each OB lasted either 60 or 90 minutes, depending on how many exposures were obtained in the block.
@@
 
-Additional archival \mbox{HAWK-I} imaging of the MACS-1149 fields was obtained from the program 090.A-0458.
-These data were obtained with one \mbox{HAWK-I} chip centered on the $z=9.7$ candidate from \cite{zheng:12}; they largely cover the HFF cluster field but the pointing was not optimized to also include the parallel field whose location was defined later (see Fig.~\ref{fig:layout}).
-The MACS-1149 HAWK-I observations were obtained in Service Mode at the ESO/\textit{VLT} in \xxx\ separate OBs between 2013 Mar 21 and 2014 Jun 9.
-The 20$\times$15\,s~=~300\,s exposures were taken at nine dithered positions with regular offsets separated by \xxx\ arcsec; the on-source exposure time of the final mosaic is \xxx\ hours.
+Additional archival \HAWKI\ imaging of the MACS-1149 fields was obtained from the program 090.A-0458.
+These data were obtained with one \HAWKI\ chip centered on the $z=9.7$ candidate from \cite{zheng:12}; they largely cover the HFF cluster field but the pointing was not optimized to also include the parallel field whose location was defined later (see Fig.~\ref{fig:layout}).
+The MACS-1149 \HAWKI\ observations were obtained in Service Mode at the ESO/\textit{VLT} in \xxx\ separate OBs between 2013 Mar 21 and 2014 Jun 9.
+The 20$\times$15\,s~=~300\,s exposures were taken at nine dithered positions with regular offsets separated by \xxx\ arcsec; the on-source exposure time of the final MACS-1149 \HAWKI\ mosaic is \xxx\ hours.
 
@@
 
-The HAWK-I and MOSFIRE observations were reduced with a pipeline that has been developed for previous surveys with the NEWFIRM \citep[NMBS;][]{whitaker:nmbs} and FOURSTAR \citep[zFOURGE;][, Labb\'e in prep]{spitler:12} infrared imaging instruments.
-Treating each detector individually, the pipeline is easily modified for the different instrument configurations of the four HAWK-I and single MOSFIRE chips.
+The \HAWKI\ and MOSFIRE observations were reduced with a pipeline that has been developed for previous surveys with the NEWFIRM \citep[NMBS;][]{whitaker:nmbs} and FOURSTAR \citep[zFOURGE;][, Labb\'e in prep]{spitler:12} infrared imaging instruments.
+Treating each detector individually, the pipeline is easily modified for the different instrument configurations of the four \HAWKI\ and single MOSFIRE chips.
 The primary task of the pipeline is removing the bright, time-variable sky background from the individual exposures, which is often some $10^4$ times brighter than the distant galaxies of interest in the field.
-With such a bright background, we first determine an empirical ``sky flat'' that is a median of all of the \textit{science} exposures in a HAWK-I OB or MOSFIRE group, after rejecting the brightest 12 exposures at each pixel position to remove the contribution of bright objects.
+With such a bright background, we first determine an empirical ``sky flat'' that is a median of all of the \textit{science} exposures in a \HAWKI\ OB or MOSFIRE group, after rejecting the brightest 12 exposures at each pixel position to remove the contribution of bright objects.
 We find these empirical flats to be preferable to external twilight or dome flats given the difficulty of obtaining a truly flat illumination pattern over such large detector fields-of-view.
@@
 
-As the archival HAWK-I images of the MACS-1149 field were obtained with longer individual exposures and with fewer exposures per sequence, the mask-pass technique described above did not produce satisfactory results.
+As the archival \HAWKI\ images of the MACS-1149 field were obtained with longer individual exposures and with fewer exposures per sequence, the mask-pass technique described above did not produce satisfactory results.
 In this case, for each raw exposure we divide by the sky flat and then subtract a third-order polynomial fit to the background.
@@
 The area covered by deep imaging in the \textit{Spitzer}/IRAC 3.6 and 4.5\,\mum\ channels is shown in orange, with the cluster fields indicated by the dashed lines.
-The HAWK-I field of view is perfectly suited (Abell 2744, MACS-0416, Abell S1063 and Abell 370) for simultaneous imaging of the cluster+parallel field pairs, which require two separate pointings with MOSFIRE (MACS-0717 and MACS-1149).
-The footprint of additional archival HAWK-I coverage of the MACS-1149 field is indicated by the gray square in the lower right panel.\label{fig:layout}}
+The \HAWKI\ field of view is perfectly suited (Abell 2744, MACS-0416, Abell S1063 and Abell 370) for simultaneous imaging of the cluster+parallel field pairs, which require two separate pointings with MOSFIRE (MACS-0717 and MACS-1149).
+The footprint of additional archival \HAWKI\ coverage of the MACS-1149 field is indicated by the gray square in the lower right panel.\label{fig:layout}}
 \end{figure*}
@@
 
-For the HAWK-I observations, a single 90-minute Service Mode OB was obtained in each of the fields requiring photometric transparency conditions, and these OBs were followed immediately by an observation of a photometric standard star at a similar airmass with single exposures on each of the four chips.
+For the \HAWKI\ observations, a single 90-minute Service Mode OB was obtained in each of the fields requiring photometric transparency conditions, and these OBs were followed immediately by an observation of a photometric standard star at a similar airmass with single exposures on each of the four chips.
 The standard star exposures were processed with empirical sky flats derived from the science exposures as described above, and the observed fluxes of the standard stars yield absolute photometric zeropoints for each chip.
@@
 The 2MASS comparison suggests a systematic offset of $-0.09$~mag for the Abell S1063 field.
-However, the exposure time for that field was similar to those of the additional deep HAWK-I fields and the derived depths of \textit{all} deep HAWK-I fields are nearly identical (Table~\ref{tab:summary}).
+However, the exposure time for that field was similar to those of the additional deep \HAWKI\ fields and the derived depths of \textit{all} deep \HAWKI\ fields are nearly identical (Table~\ref{tab:summary}).
 Given the overall agreement with the 2MASS photometry, we estimate that the systematic uncertainty of the absolute photometric calibration is $\sigma_\mathrm{sys}\leq0.05$~mag.
@@
 Next, object catalogs are generated with the SEXtractor software \citep{bertin:96} for each individual background-subtracted exposure and transformations to the reference frame (shift, rotation and scale) are computed with with the \texttt{stsci.stimage.xyxymatch} and \texttt{scikit-image.transform}\footnote{\url{http://scikit-image.org}} Python software tools.
-For the HAWK-I exposures, we fit a third-order polynomial to the geometric distortion model determined by \cite{libralato:14} and specify the polynomial terms as ``Simple Imaging Polynomial'' (SIP) coefficients \citep{fits:sip} in the individual exposure FITS files.
-The SIP FITS header generally applicable for HAWK-I is provided in Appendix \xxx.
-We note that the HAWK-I distortion is generally small, reaching $\sim$2 pixels ($0\farcs2$) at the image corners, but that the distortion corrections are necessary in particular given the excellent overall image quality of the observations.
+For the \HAWKI\ exposures, we fit a third-order polynomial to the geometric distortion model determined by \cite{libralato:14} and specify the polynomial terms as ``Simple Imaging Polynomial'' (SIP) coefficients \citep{fits:sip} in the individual exposure FITS files.
+The SIP FITS header generally applicable for \HAWKI\ is provided in Appendix \xxx.
+We note that the \HAWKI\ distortion is generally small, reaching $\sim$2 pixels ($0\farcs2$) at the image corners, but that the distortion corrections are necessary, in particular, given the excellent overall image quality of the observations.
 
@@
 The orientation of the fields is the same as that shown in Fig.~\ref{fig:layout}.
-The four HAWK-I chips and two MOSFIRE pointings are clearly visible for the first four and last two fields, respectively.
+The four \HAWKI\ chips and two MOSFIRE pointings are clearly visible for the first four and last two fields, respectively.
 \label{fig:ivar_mosaics}}  
@@
 
-A crucial innovation of the image processing of the present analysis compared to previous works is that we combine all of the individual background-subtracted raw images (e.g., \xxx\ 4000 files for 1000 exposures times 4 detectors for the \mbox{Abell 2744} field) into the final output mosaic with the ``drizzle'' algorithm (\citealp{fruchter:02}, as implemented in the DrizzlePac/AstroDrizzle package; \citealp{drizzlepac}).
+A crucial innovation of the image processing of the present analysis compared to previous works is that we combine all of the individual background-subtracted raw images (e.g., \xxx\ 4000 files for 1000 exposures $\times$ 4 detectors for the \mbox{Abell 2744} field) into the final output mosaic with the ``drizzle'' algorithm (\citealp{fruchter:02}, as implemented in the DrizzlePac/AstroDrizzle package; \citealp{drizzlepac}).
 The benefits of the drizzle algorithm  preserving robust noise properties of the output mosaics will be discussed in detail below in \S\ref{s:depth}.
@@
 The final image and inverse variance mosaics are provided as an ESO Phase 3 data product \xxx\ URL, as well as attached as a digital product to this manuscript (\xxx\ I think ApJSS can and wants to do this).
-The released images are all scaled to a common zeropoint of 26.0 (AB mag), which gives pixel values of order unity for faint galaxies.
+The released images are all scaled to a common zeropoint of 26.0 (AB mag), which gives pixel values of order unity faint galaxies near the detection threshold.
 In the sections below we describe the characteristics of the mosaics in more detail.
@@
 Fig.~\ref{fig:cog} shows the curves of growth for stars identified by the tight relationship between their brightness and half-light radii \citep[see, e.g., Fig.~13 of][]{skelton:14}.
-The image quality of the $K_s$-band mosaics is excellent with stellar FWHM$\lesssim0\farcs4$ for the deep HAWK-I fields, thanks in large part to the execution of the HAWK-I observations in service mode ensuring optimal and uniform image quality across the many hours of integration on the survey fields.
+The image quality of the $K_s$-band mosaics is excellent with stellar FWHM$\lesssim0\farcs4$ for the deep \HAWKI\ fields, thanks in large part to the execution of the \HAWKI\ observations in service mode ensuring optimal and uniform image quality across the many hours of integration on the survey fields.
 The classical scheduling of the MOSFIRE observations does not allow such control over the seeing conditions.
@@
 Along with practical benefits mentioned above, the most dramatic benefit of using the drizzle algorithm comes from the fact that the raw detector pixels are only resampled once in the process of generating the output mosaic.
-Furthermore, with a large number of dithered raw images we can use drizzle to shrink the raw input pixels by a factor of ten before dropping them into the output grid,\footnote{The pixels are shrunk by adopting pixfrac=0.1 as defined by \cite{fruchter:02}.} and the result is dramatically reduced correlations between neighboring pixels, particularly for output pixel sizes that are somewhat smaller than the native HAWK-I or MOSFIRE detector pixels.
+Furthermore, with a large number of dithered raw images we can use drizzle to shrink the raw input pixels by a factor of ten before dropping them into the output grid,\footnote{The pixels are shrunk by adopting pixfrac=0.1 as defined by \cite{fruchter:02}.} and the result is dramatically reduced correlations between neighboring pixels, particularly for output pixel sizes that are somewhat smaller than the native \HAWKI\ or MOSFIRE detector pixels.
 
@@
 The $5\,\sigma$ limiting magnitudes (AB) within $0\farcs6$ diameter apertures are indicated in Fig.~\ref{fig:depth} and listed in Table~\ref{tab:summary}.
-Reaching 26th magnitude (AB), the HAWK-I images presented here are among the deepest $K_s$-band images ever obtained and will provide a crucial complement to the deep \textit{HST} and \textit{Spitzer} imaging of the Frontier Fields.
+Reaching 26th magnitude (AB), the \HAWKI\ images presented here are among the deepest $K_s$-band images ever obtained and will provide a crucial complement to the deep \textit{HST} and \textit{Spitzer} imaging of the Frontier Fields.
 
-We conclude here with a brief comment on the ``HUGS'' ultra-deep HAWK-I images of the CANDELS GOODS-S and UDS survey fields.
-\cite{hugs} report a final depth of 26.5 AB for point sources within $D=0\farcs4$ apertures for their deepest field, ``GOODS-D1'', which has an exposure time of 31.5 hours similar to the deep HAWK-I fields summarized in Table~\ref{tab:summary}.
+We conclude here with a brief comment on the ``HUGS'' ultra-deep \HAWKI\ images of the CANDELS GOODS-S and UDS survey fields.
+\cite{hugs} report a final depth of 26.5 AB for point sources within $D=0\farcs4$ apertures for their deepest field, ``GOODS-D1'', which has an exposure time of 31.5 hours similar to the deep \HAWKI\ fields summarized in Table~\ref{tab:summary}.
 \cite{hugs} assume ideal noise properties that likely suffer some residual correlation between adjacent pixels such as that indicated by the red curves in Fig.~\ref{fig:corr_noise}, where we found that the drizzled rms was some 40\% higher than that measured in the presence of the pixel correlations for $0\farcs1$ pixel mosaics.
-This effect ($\sim$0.4 mag) along with the relative exposure time difference ($\sim$0.1 mag) can account for much of the difference between the reported depths of the deepest HUGS pointings and the \textit{Hubble} Frontier Fields HAWK-I images described here.
+This effect ($\sim$0.4 mag) along with the relative exposure time difference ($\sim$0.1 mag) can account for much of the difference between the reported depths of the deepest HUGS pointings and the \textit{Hubble} Frontier Fields \HAWKI\ images described here.
 The comparison must be made because any additional half magnitude increase depth is exponentially more expensive to obtain!
 
-\xxx\ say something about distant extended sources, perhaps scaling from Arjen's size distributions?
+\subsection{Source detection \& completeness}
+\label{s:completeness}
+
+We complete source detection completeness curves as a function of source brightness following the techniques described by \cite{whitaker:nmbs} and \cite{muzzin:13}.
+Briefly, we insert artificial sources in blank regions of the portions of the images covering the \textit{HST} ``parallel'' fields and compute the fraction of sources recovered with SExtractor as a function of the source magnitude.
+The resulting completeness curves are shown in Fig.~\ref{fig:completeness}.
+As in the sensitivity analysis above, we find that the detection completeness curves are nearly identical for the deep \HAWKI\ fields, with 50\% (90\%) source completeness at $K_s\sim25.9$ (25.7) AB.
+The completeness thresholds for the shallower northern MACS-1149 and MACS-0717 fields are $\sim$0.75 mag brighter.
+
+We perform this completeness analysis to provide a general characterization of the $K_s$-band image mosaics and to provide a point of comparison for earlier surveys, such as UltraVISTA \citep{muzzin:13}.
+However, in the case of the Frontier Fields cluster and parallel fields with much deeper photometry available from \textit{HST} (c.f.
+28.7 AB in $H_{160}$), the $K_s$ band images can be more effectively exploited for most applications by detecting objects in the deeper \textit{HST} images and performing forced $K_s$-band photometry at the positions of the \textit{HST} sources.
+The generation of these catalogs is beyond the scope of the present work and will be by Shipley et al.
+(in prep).
+
+\begin{figure}[!ht]
+\epsscale{1.18}
+\plotone{completeness.pdf} % Figures/exposure_map.pdf
+\caption{$K_s$-band detection completeness in the six Frontier Fields survey fields.
+The four deep \HAWKI\ fields are remarkably uniform, with 50\% detection completeness at $K_s=25.9$--26.0 AB.
+The shallower northern MACS-1149 and MACS-0717 fields are 50\% complete at $K_s=24.9$ and 25.3, respectively.
+Only MACS-1149 differs appreciably in the area around the \textit{HST} cluster area, with the curve as shown;  the completeness curves in the cluster areas of the other fields differ from the curves shown by $\lesssim0.1$ mag and are omitted for clarity.
+\xxx{Danilo please check}.
+\label{fig:completeness}}  
+\end{figure}
 

### 2016-04-12 Paper0/ms.tex
new figures

@@
 \epsscale{1.18}
+\plotone{delta_UV.pdf} % Figures/exposure_map.pdf
+\caption{\textit{Left:} Difference in the derived rest-frame $U-V$ colors as a function of photometric redshift for galaxies in the MACS-0416 parallel field, before and after including $K_s$ photometry along with the \textit{HST} and \textit{Spitzer} measurements.
+The scatter in the derived $U-V$ colors at $z>2.5$ is large ($\sim$0.2 mag) where the Balmer/4000~\AA\ break is in the gap between \textit{HST} $H_{160}$ and the \textit{Spitzer} IRAC 3.6~\mum\ bands, which is well then constrained by including $K_s$.
+This large scatter translates directly into large scatter on other quantities of interest, such as the stellar mass to light ratio and the age of the stellar populations.
+The center and right panels show a dramatic example of a galaxy with a strong break between the \textit{HST} and IRAC bands.
+The addition of the $K_s$ band pinpoints the location of the break and shrinks the photometric redshift uncertainties by a factor of two and suggesting $z_\mathrm{phot} > 3$ (right panel).
+\label{fig:delta_UV}}
+\end{figure*}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%  Number counts figure
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure*}
+\epsscale{1.18}
 \plotone{number_counts_all.pdf} % Figures/exposure_map.pdf

### 2016-04-15 Paper0/ms.tex
aastex6 formatting

@@
-\documentclass[iop, numberedappendix]{emulateapj}
+%\documentclass[iop, numberedappendix]{emulateapj}
+\documentclass[preprint2]{aastex6}
 \usepackage{apjfonts}
@@
 %\usepackage{color}
-\usepackage[usenames, dvipsnames]{color}
-\definecolor{citeRGB}{rgb}{0,0.1,0.7}
-\usepackage[hyperfootnotes=true,naturalnames=true,letterpaper,pdfstartview=FitH,pdfpagemode=UseNone,colorlinks=true,citecolor=citeRGB]{hyperref}
+%\usepackage[usenames, dvipsnames]{color}
+%\definecolor{citeRGB}{rgb}{0,0.1,0.7}
+%\usepackage[hyperfootnotes=true,naturalnames=true,letterpaper,pdfstartview=FitH,pdfpagemode=UseNone,colorlinks=true,citecolor=citeRGB]{hyperref}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%                                                                %
+%  Summary table
+%                                                                %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{deluxetable*}{ccccrccccc}
+\tabletypesize{\footnotesize}
+\tablewidth{0pt} 
+\tablecaption{Field summary \label{tab:summary}} 
+\tablehead{ 
+\colhead{Field} & \colhead{R.A.} & \colhead{Dec.} & \colhead{Instrument} & \colhead{Epoch} & \textit{HST}\footnotemark[1] & \colhead{$t$, hours} & {Depth\footnotemark[2]} & \colhead{FWHM}}
+\startdata 
+Abell 2744   & $00:14:21.2$ & $-30:23:50$ & \textit{VLT}/\HAWKI     & 2013 Oct 24--2013 Dec 24 & C+P & 29.3 &  26.0 & $0\farcs39$ \\ 
+MACS-0416    & $04:16:08.9$ & $-24:04:28$ & \textit{VLT}/\HAWKI     & 2013 Oct 25--2014 Feb 23 & C+P & 25.8 &  26.0 & $0\farcs36$ \\ 
+Abell S1063  & $22:49:01.1$ & $-44:32:13$ & \textit{VLT}/\HAWKI     & 2015 Jul 8--Sep 22       & C+P & 27.9 &  26.0 & $0\farcs39$ \\ 
+Abell 370    & $02:40:03.3$ & $-01:36:23$ & \textit{VLT}/\HAWKI     & 2015 Jul 26--2016 Jan 28 & C+P & 28.3 &  26.0 & $0\farcs35$ \\ 
+\vspace {-0.2cm}\\ 
+\multirow{2}{*}{MACS-0717}    & \multirow{2}{*}{$07:17:34.0$} & \multirow{2}{*}{$+37:44:49$} &  \multirow{2}{*}{\textit{Keck}/MOSFIRE}  &  \multirow{2}{*}{2015 Jan 26--27, 2016 Jan 21--22}  & C & 4.3 & 25.3 & $0\farcs42$ \\
+                              &                             &                             &                                          &      & P & 3.8 & 25.5 & $0\farcs49$ \\ 
+\vspace {-0.2cm}\\ 
+\multirow{4}{*}{MACS-1149}    & \multirow{4}{*}{$11:49:36.3$} & \multirow{4}{*}{$+22:23:58$} & \multirow{2}{*}{\textit{Keck}/MOSFIRE}  & \multirow{2}{*}{2015 Feb 24, 2016 Jan 21--22}      & C & 5.5 & 25.2 & $0\farcs53$ \\
+&  &         &      &   & P & 4.8 & 25.1 & $0\farcs54$ \\
+\vspace {-0.4cm}\\ 
+&  &         & \textit{VLT}/\HAWKI\footnotemark[3]     & 2013 Mar 21--2014 Jun 9      & C & 5.3 & 25.0 & $0\farcs41$
+\enddata 
+%\vspace{-0.8cm} 
+\footnotetext[1]{Coverage of \textit{Hubble} survey fields: C=Cluster, P=Parallel.
+Most \HAWKI\ pointings cover the two \textit{HST} fields simultaneously (C+P; Fig.~\ref{fig:layout}).
+The MOSFIRE observations of MACS-0717 and MACS-1149 require two separate pointings of the instrument to cover the cluster and parallel \textit{HST} fields, which have the different characteristics as indicated.}
+%\footnotetext[2]{Approximate exposure time equivalent at $0\farcs4$ seeing:  $t_{0.4}=\sum_i t_i \cdot (0.4/\mathrm{FWHM}_i)^2$.}
+\footnotetext[2]{Depth is defined as the 5$\sigma$ limiting magnitude for point sources, measured in $D=0\farcs6$ apertures (\S\ref{s:depth}).} 
+\footnotetext[3]{Archival observations from ESO program \href{http://archive.eso.org/wdb/wdb/eso/eso_archive_main/query?prog_id=090.A-0458(A)}{090.A-0458} (PI: Infante).}
+\end{deluxetable*}
+%\vspace{1cm}
+
 The latest effort to further our knowledge of galaxy formation and evolution is represented by the HST Frontier Fields program.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%                                                                %
-%  Summary table
-%                                                                %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{deluxetable*}{ccccrccccc} 
-\tabletypesize{\footnotesize}
-\tablewidth{0pt} 
-\tablecaption{Field summary \label{tab:summary}} 
-\tablehead{ 
-\colhead{Field} & \colhead{R.A.} & \colhead{Dec.} & \colhead{Instrument} & \colhead{Epoch} & \textit{HST}\footnotemark[1] & \colhead{$t$, hours} & \colhead{Depth\footnotemark[2]} & \colhead{FWHM}}
-\startdata 
-Abell 2744   & $00:14:21.2$ & $-30:23:50$ & \textit{VLT}/\HAWKI     & 2013 Oct 24--2013 Dec 24 & C+P & 29.3 &  26.0 & $0\farcs39$ \\ 
-MACS-0416    & $04:16:08.9$ & $-24:04:28$ & \textit{VLT}/\HAWKI     & 2013 Oct 25--2014 Feb 23 & C+P & 25.8 &  26.0 & $0\farcs36$ \\ 
-Abell S1063  & $22:49:01.1$ & $-44:32:13$ & \textit{VLT}/\HAWKI     & 2015 Jul 8--Sep 22       & C+P & 27.9 &  26.0 & $0\farcs39$ \\ 
-Abell 370    & $02:40:03.3$ & $-01:36:23$ & \textit{VLT}/\HAWKI     & 2015 Jul 26--2016 Jan 28 & C+P & 28.3 &  26.0 & $0\farcs35$ \\ 
-\vspace {-0.2cm}\\ 
-\multirow{2}{*}{MACS-0717}    & \multirow{2}{*}{$07:17:34.0$} & \multirow{2}{*}{$+37:44:49$} &  \multirow{2}{*}{\textit{Keck}/MOSFIRE}  &  \multirow{2}{*}{2015 Jan 26--27, 2016 Jan 21--22}  & C & 4.3 & 25.3 & $0\farcs42$ \\
-                              &                             &                             &                                          &      & P & 3.8 & 25.5 & $0\farcs49$ \\ 
-\vspace {-0.2cm}\\ 
-\multirow{4}{*}{MACS-1149}    & \multirow{4}{*}{$11:49:36.3$} & \multirow{4}{*}{$+22:23:58$} & \multirow{2}{*}{\textit{Keck}/MOSFIRE}  & \multirow{2}{*}{2015 Feb 24, 2016 Jan 21--22}      & C & 5.5 & 25.2 & $0\farcs53$ \\
-&  &         &      &   & P & 4.8 & 25.1 & $0\farcs54$ \\
-\vspace {-0.3cm}\\ 
-&  &         & \textit{VLT}/\HAWKI\footnotemark[3]     & 2013 Mar 21--2014 Jun 9      & C & 5.3 & 25.0 & $0\farcs41$
-\enddata 
-%\vspace{-0.8cm} 
-\footnotetext[1]{Coverage of \textit{Hubble} survey fields: C=Cluster, P=Parallel.
-Most \HAWKI\ pointings cover the two \textit{HST} fields simultaneously (C+P; Fig.~\ref{fig:layout}).
-The MOSFIRE observations of MACS-0717 and MACS-1149 require two separate pointings of the instrument to cover the cluster and parallel \textit{HST} fields, which have the different characteristics as indicated.}
-%\footnotetext[2]{Approximate exposure time equivalent at $0\farcs4$ seeing:  $t_{0.4}=\sum_i t_i \cdot (0.4/\mathrm{FWHM}_i)^2$.}
-\footnotetext[2]{Depth is defined as the 5$\sigma$ limiting magnitude for point sources, measured in $D=0\farcs6$ apertures (\S\ref{s:depth}).} 
-\footnotetext[3]{Archival observations from ESO program \href{http://archive.eso.org/wdb/wdb/eso/eso_archive_main/query?prog_id=090.A-0458(A)}{090.A-0458} (PI: Infante).}
-\end{deluxetable*}
-%\vspace{1cm}
-
 $K_s$-band observations of the Abell 2744 and MACS-0416 fields were obtained between 2013 Oct and 2014 Feb with the High Acuity Wide-field K-band Imager \citep[\HAWKI;][]{hawki, hawki2} on the 8.2\,m UT4 telescope at the ESO \textit{Very Large Telescope} (ESO program \mbox{092.A-0472}).
@@
 The individual detector integration times (DIT) were adjusted on the fly during the observing run to keep the total counts within the linear regime of the detector, while maintaining $\mathrm{NDIT}\times\mathrm{DIT} \sim 40$\,s.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%   Exposure Map figure
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure*}
+\epsscale{1.18}
+\plotone{hawkiff_layout.pdf} % Figures/exposure_map.pdf
+\caption{Layout of the Frontier Fields $K_s$-band mosaics.
+The positions of the \textit{HST} cluster and parallel fields are shown in the blue (ACS optical) and red (WFC3 IR) polygons.
+The light blue polygons in the MACS-0717 field show additional wide-field ACS imaging coverage from  programs GO-9722 and GO-10420 \citep[PI: Ebeling;][]{ma:11}.
+The area covered by deep imaging in the \textit{Spitzer}/IRAC 3.6 and 4.5\,\mum\ channels is shown in orange, with the cluster fields indicated by the dashed lines.
+The \HAWKI\ field of view is perfectly suited (Abell 2744, MACS-0416, Abell S1063 and Abell 370) for simultaneous imaging of the cluster+parallel field pairs, which require two separate pointings with MOSFIRE (MACS-0717 and MACS-1149).
+The footprint of additional archival \HAWKI\ coverage of the MACS-1149 field is indicated by the gray square in the lower right panel.\label{fig:layout}}
+\end{figure*}
+%%%%%%%%%%%%%%%%
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%   Exposure Map figure
+%   2MASS comparison
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}
+\begin{figure}[!t]
 \epsscale{1.18}
-\plotone{hawkiff_layout.pdf} % Figures/exposure_map.pdf
-\caption{Layout of the Frontier Fields $K_s$-band mosaics.
-The positions of the \textit{HST} cluster and parallel fields are shown in the blue (ACS optical) and red (WFC3 IR) polygons.
-The light blue polygons in the MACS-0717 field show additional wide-field ACS imaging coverage from  programs GO-9722 and GO-10420 \citep[PI: Ebeling;][]{ma:11}.
-The area covered by deep imaging in the \textit{Spitzer}/IRAC 3.6 and 4.5\,\mum\ channels is shown in orange, with the cluster fields indicated by the dashed lines.
-The \HAWKI\ field of view is perfectly suited (Abell 2744, MACS-0416, Abell S1063 and Abell 370) for simultaneous imaging of the cluster+parallel field pairs, which require two separate pointings with MOSFIRE (MACS-0717 and MACS-1149).
-The footprint of additional archival \HAWKI\ coverage of the MACS-1149 field is indicated by the gray square in the lower right panel.\label{fig:layout}}
-\end{figure*}
+\plotone{2mass_phot.pdf} % Figures/exposure_map.pdf
+\caption{Comparison of bright-star photometry to the 2MASS public catalog.
+The photometric calibration of the $K_s$-band images were determined from standard star observations taken concurrently with the science exposures.
+The uncertainties shown are taken from the 2MASS catalog; these stars are measured at very high S/N in the deep $K_s$-band mosaics.
+We estimate the overall systematic uncertainty in the photometric calibration to be $\lesssim0.05~\mathrm{mag}$.\label{fig:2mass}}
+\end{figure}
+%%%%%%%%%%%%%
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%   2MASS comparison
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[!t]
-\epsscale{1.18}
-\plotone{2mass_phot.pdf} % Figures/exposure_map.pdf
-\caption{Comparison of bright-star photometry to the 2MASS public catalog.
-The photometric calibration of the $K_s$-band images were determined from standard star observations taken concurrently with the science exposures.
-The uncertainties shown are taken from the 2MASS catalog; these stars are measured at very high S/N in the deep $K_s$-band mosaics.
-We estimate the overall systematic uncertainty in the photometric calibration to be $\lesssim0.05~\mathrm{mag}$.\label{fig:2mass}}
-\end{figure}
-
 \subsection{Astrometric alignment}
@@
 In order to simplify measurements from the $K_s$-band images in consort with space-based \textit{HST} and \textit{Spitzer} mosaics, the astrometry of the ground-based mosaics must be refined.
-Reference absolute astrometric catalogs were generated from \textit{HST} images when available (i.e., the chips overlapping the cluster and parallel \textit{HST} fields) and public \textit{Subaru} Suprime-Cam $r_c$-band images otherwise\footnote{http://archive.stsci.edu/pub/hlsp/clash/}.
+Reference absolute astrometric catalogs were generated from \textit{HST} images when available (i.e., the chips overlapping the cluster and parallel \textit{HST} fields) and public \textit{Subaru} Suprime-Cam $r_c$-band images otherwise\footnote{\url{http://archive.stsci.edu/pub/hlsp/clash/}}.
 Next, object catalogs are generated with the SEXtractor software \citep{bertin:96} for each individual background-subtracted exposure and transformations to the reference frame (shift, rotation and scale) are computed with with the \texttt{stsci.stimage.xyxymatch} and \texttt{scikit-image.transform}\footnote{\url{http://scikit-image.org}} Python software tools.
@@
 We note that the \HAWKI\ distortion is generally small, reaching $\sim$2 pixels ($0\farcs2$) at the image corners, but that the distortion corrections are necessary, in particular, given the excellent overall image quality of the observations.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%   Drizzle figure
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[!ht]
+\epsscale{1.18}
+\plotone{ivar_mosaics.pdf} % Figures/exposure_map.pdf
+\caption{Inverse variance maps of the $K_s$-band mosaics.
+The orientation of the fields is the same as that shown in Fig.~\ref{fig:layout}.
+The four \HAWKI\ chips and two MOSFIRE pointings are clearly visible for the first four and last two fields, respectively.
+\label{fig:ivar_mosaics}}  
+\end{figure}
+%%%%%%%%%%%%%%%%%%%
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%   Drizzle figure
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[!t]
-\epsscale{1.18}
-\plotone{ivar_mosaics.pdf} % Figures/exposure_map.pdf
-\caption{Inverse variance maps of the $K_s$-band mosaics.
-The orientation of the fields is the same as that shown in Fig.~\ref{fig:layout}.
-The four \HAWKI\ chips and two MOSFIRE pointings are clearly visible for the first four and last two fields, respectively.
-\label{fig:ivar_mosaics}}  
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %   Image quality figure
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}[!ht]
+\begin{figure*}[!t]
 \epsscale{1.18}
@@
 \end{figure*}
+%%%%%%%%%%%%%%
 
@@
 In the sections below we describe the characteristics of the mosaics in more detail.
-
-\subsection{Image quality}
-\label{s:fwhm}
-
-Fig.~\ref{fig:cog} shows the curves of growth for stars identified by the tight relationship between their brightness and half-light radii \citep[see, e.g., Fig.~13 of][]{skelton:14}.
-The image quality of the $K_s$-band mosaics is excellent with stellar FWHM$\lesssim0\farcs4$ for the deep \HAWKI\ fields, thanks in large part to the execution of the \HAWKI\ observations in service mode ensuring optimal and uniform image quality across the many hours of integration on the survey fields.
-The classical scheduling of the MOSFIRE observations does not allow such control over the seeing conditions.
-The result is that the image quality of the MACS-0717 and MACS-1149 fields is somewhat degraded in comparison, at $0\farcs42$--$0\farcs54$, and the image quality differs measurably between the two MOSFIRE pointings in the MACS-0717 field.
-Despite the compact cores of the stellar PSFs, the curves of growth shown in Fig.~\ref{fig:cog} deviate from Gaussian profiles with significantly higher flux in the outer wings, with a shape more consistent with a Moffat profile \citep{trujillo:01}  with $\beta\sim2$.
-We caution that these extended profiles will yield significantly shallower final image depths than would be predicted with Exposure Time Calculators that may assume Gaussian profiles.
-Nevertheless, the deep $K_s$ mosaics have excellent image quality that is much closer to the resolution of the \textit{HST} IR imaging and the typical apparent size of distant galaxies \citep[median $r_e\sim0\farcs1$--$0\farcs2$ at $z>4$;][]{shibuya:15} than the redder \textit{Spitzer} IRAC bands (cf.
-FWHM$\sim$$1\farcs7$--$2^{\prime\prime}$).
-
-\subsection{Noise properties \& Depth}
-\label{s:depth}
 
@@
 
+\subsection{Image quality}
+\label{s:fwhm}
+
+Fig.~\ref{fig:cog} shows the curves of growth for stars identified by the tight relationship between their brightness and half-light radii \citep[see, e.g., Fig.~13 of][]{skelton:14}.
+The image quality of the $K_s$-band mosaics is excellent with stellar FWHM$\lesssim0\farcs4$ for the deep \HAWKI\ fields, thanks in large part to the execution of the \HAWKI\ observations in service mode ensuring optimal and uniform image quality across the many hours of integration on the survey fields.
+The classical scheduling of the MOSFIRE observations does not allow such control over the seeing conditions.
+The result is that the image quality of the MACS-0717 and MACS-1149 fields is somewhat degraded in comparison, at $0\farcs42$--$0\farcs54$, and the image quality differs measurably between the two MOSFIRE pointings in the MACS-0717 field.
+Despite the compact cores of the stellar PSFs, the curves of growth shown in Fig.~\ref{fig:cog} deviate from Gaussian profiles with significantly higher flux in the outer wings, with a shape more consistent with a Moffat profile \citep{trujillo:01}  with $\beta\sim2$.
+We caution that these extended profiles will yield significantly shallower final image depths than would be predicted with Exposure Time Calculators that may assume Gaussian profiles.
+Nevertheless, the deep $K_s$ mosaics have excellent image quality that is much closer to the resolution of the \textit{HST} IR imaging and the typical apparent size of distant galaxies \citep[median $r_e\sim0\farcs1$--$0\farcs2$ at $z>4$;][]{shibuya:15} than the redder \textit{Spitzer} IRAC bands (cf.
+FWHM$\sim$$1\farcs7$--$2^{\prime\prime}$).
+
+\subsection{Noise properties \& Depth}
+\label{s:depth}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%   Completeness figure
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[!t]
+\epsscale{1.18}
+\plotone{completeness.pdf} % Figures/exposure_map.pdf
+\caption{$K_s$-band detection completeness in the six Frontier Fields survey fields.
+The four deep \HAWKI\ fields are remarkably uniform, with 50\% detection completeness at $K_s=25.9$--26.0 AB.
+The shallower northern MACS-1149 and MACS-0717 fields are 50\% complete at $K_s=24.9$ and 25.3, respectively.
+Only MACS-1149 differs appreciably in the area around the \textit{HST} cluster area, with the curve as shown;  the completeness curves in the cluster areas of the other fields differ from the curves shown by $\lesssim0.1$ mag and are omitted for clarity.
+\xxx{Danilo please check}.
+\label{fig:completeness}}  
+\end{figure}
+
 \subsection{Source detection \& completeness}
@@
 (in prep).
-
-\begin{figure}[!ht]
-\epsscale{1.18}
-\plotone{completeness.pdf} % Figures/exposure_map.pdf
-\caption{$K_s$-band detection completeness in the six Frontier Fields survey fields.
-The four deep \HAWKI\ fields are remarkably uniform, with 50\% detection completeness at $K_s=25.9$--26.0 AB.
-The shallower northern MACS-1149 and MACS-0717 fields are 50\% complete at $K_s=24.9$ and 25.3, respectively.
-Only MACS-1149 differs appreciably in the area around the \textit{HST} cluster area, with the curve as shown;  the completeness curves in the cluster areas of the other fields differ from the curves shown by $\lesssim0.1$ mag and are omitted for clarity.
-\xxx{Danilo please check}.
-\label{fig:completeness}}  
-\end{figure}
 
@@
 \bibliographystyle{apj_url}
+%\bibliographystyle{aasjournal}
 \bibliography{ms}

### 2016-04-15 Paper0/ms.tex
fill out discussion

@@
 \newcommand\xxx{{\textcolor{red}{\bf xxx}}}
+\newcommand\XXX[1]{{\textcolor{red}{xxx #1}}}
 
@@
 28.7 AB in $H_{160}$), the $K_s$ band images can be more effectively exploited for most applications by detecting objects in the deeper \textit{HST} images and performing forced $K_s$-band photometry at the positions of the \textit{HST} sources.
-The generation of these catalogs is beyond the scope of the present work and will be by Shipley et al.
+The generation of such catalogs is beyond the scope of the present work and will be by Shipley et al.
 (in prep).
-
-\subsection{Number counts}
-\label{s:number}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%  Number counts figure
+%  U-V
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%  Number counts figure
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}
-\epsscale{1.18}
-\plotone{number_counts_all.pdf} % Figures/exposure_map.pdf
-\caption{$K_s$-band number counts in the Frontier Fields. \label{fig:counts}}  
-\end{figure*}
+% \subsection{Number counts}
+% \label{s:number}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Photometric constraints from the $K_s$ band}
-\label{s:constraints}
-
-\subsection{Photometric redshifts}
-\label{s:photoz}
-
-\subsection{SED modeling}
-\label{s:sed}
-
-Balmer break, UV slope
+% \section{Photometric constraints from the $K_s$ band}
+% \label{s:constraints}
 
@@
 \label{s:discussion}
+
+We have presented the observations and reduction for a collection of extremely deep $K_s$-band images covering the lensing clusters observed as part of the \textit{Hubble} and \textit{Spitzer} Frontier Fields program (Lotz et al., in prep; Capak et al., in prep).
+Service mode scheduling of $>$25~hour \HAWKI\ integrations on the southern hemisphere clusters visible from the ESO/\textit{VLT} result in remarkably uniform, high-quality mosaics across the separate survey fields, with superb image quality ($0\farcs4$~FWHM) and photometric depths reaching AB=26.0 mag (5$\sigma$).
+The $K_s$ band mosaics fill the gap in the infrared wavelength coverage between the WFC3/IR and IRAC instruments at depths commensurate with the deep space-based imaging and with image quality that suffers significantly less from source crowding and blending than the redder IRAC bands.
+
+A comparison of the $K_s$-band and space-based Frontier Fields imaging is shown in Fig.~\ref{fig:cutout}, showing just a small cutout of the MACS-0416 ``parallel'' field.
+The deep optical and near-infrared \textit{Hubble} imaging provides spectacular multiwavelength spatially-resolved information on physical scales of just $\sim$1~kpc (Fig.~\ref{fig:cutout}a).
+However, there are many red galaxies in the indicated areas, predominantly at redshifts $z\gtrsim2$.
+At $z>3$ even the reddest WFC3/IR filter, $H_{160}$ (Fig.~\ref{fig:cutout}b), only probes rest-frame ultraviolet wavelengths $<4000$~\AA\ and is therefore most sensitive to young, UV-bright star-forming galaxies.
+Sampling of more evolved stellar populations at $z>3$ with redder colors requires deep imaging at longer wavelengths.
+This can be obtained with imaging in the \textit{Spitzer} IRAC bands \citep[e.g.,][]{marchesini:10, stefanon:13}, though at the cost of low spatial resolution ($\sim1\farcs6$, Fig.~\ref{fig:cutout}d).
+This resolution is insufficient for spatially resolving the distant galaxies and is prone to blending of faint sources, particularly in the crowded Frontier Fields clusters where the surface density of both foreground cluster galaxies and faint background galaxies is high at the depths of interest.
+The deep \HAWKI\ imaging described here (Fig.~\ref{fig:cutout}c) is able to bridge this gap, clearly detecting and resolving all but the faintest blue galaxies seen in the deep \textit{Hubble} images.
+
+Here we explore the quantitative constraints provided by the deep $K_s$ band imaging in terms of the photometric redshifts and derived intrinsic properties of galaxies in the survey field.
+We construct preliminary PSF-matched photometric catalogs of the \textit{HST} imaging and deblended photometry from the IRAC bands following the methodology described by \XXX{ref, Labb\'e}.
+We combine then combine the space-based catalogs with aperture photometry matched directly from $K_s$-detected catalogs derived from the completeness simulations described above\footnote{Full robust \textit{HST}+$K$+IRAC catalogs will be presented by Shipley et al., in prep.}.
+We then compute photometric redshifts with EAZY \citep{brammer:08} for the catalogs with and without the $K_s$-band photometry included.
+The left panel of Fig.~\ref{fig:delta_UV} shows the difference in the rest-frame $U-V$ color derived from the photometric redshift fit, which probes the strength of the Balmer/4000~\AA\ break and is a proxy for the age and mass-to-light ratio of the underlying stellar population.
+
+The scatter in the $U-V$ colors with and without including the $K_s$-band information is low at $z<2$ where fit is constrained predominantly by the deep \textit{HST} photometry.
+At $z>2$, however, as the rest-frame $V$ band is redshifted beyond the red $H_{160}$ filter, the scatter increases dramatically, reaching $\sigma>0.1$~mag at $z\sim3$.
+This is much larger than the photometric uncertainties in the adjacent space-based photometric bands would suggest, as all of these galaxies at $H\lesssim26$ are detected in the deep WFC3 and IRAC images at $\gg10$$\sigma$.
+Therefore, the large scatter is likely dominated by systematic differences in the photometric redshifts and therefore also in the derived intrinsic properties of galaxies in the survey.
+
+The right two panels of Fig.~\ref{fig:delta_UV} show a single galaxy that illustrates how these systematic effects are not trivial and will likely result in biases in the interpretation of the galaxy population properties derived from the \textit{HST} and IRAC observations alone.
+The spectral energy distribution (SED) shown is steadily rising through the reddest WFC3/IR bands, and then shows a sharp break with bright detections in the IRAC bands.
+The $K_s$-band measurement at 2.1~$\mu$m reduces the range of allowed photometric redshifts by a factor of two by pinpointing a strong Balmer break at $z\sim3.4$.
+Even though the \textit{measured} $H_{160}-K_s$ color is redder than that inferred from the \textit{HST} + IRAC photometry alone, the final rest-frame $U-V$ color is actually \textit{bluer} as a result of the higher preferred redshift.
+Evolved galaxies at $z>3$ such as the one shown are an intriguing population deserving of detailed study in their own right, and the combined Frontier Fields \textit{Hubble}+$K_s$+IRAC dataset is ideally suited for this purpose.
+
+\XXX{blah blah, few more sentences wrapping up.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%  Number counts figure
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \begin{figure*}
+% \epsscale{1.18}
+% \plotone{number_counts_all.pdf} % Figures/exposure_map.pdf
+% \caption{$K_s$-band number counts in the Frontier Fields. \label{fig:counts}}  
+% \end{figure*}
+
+% \subsection{Photometric redshifts}
+% \label{s:photoz}
+% 
+% \subsection{SED modeling}
+% \label{s:sed}
+
+%Balmer break, UV slope
 
@@
 
+\appendix
+
+\section{SIP keywords for \HAWKI\ geometric distortion}
+
+\section{Description of provided data products}
+
 \end{document}

### 2016-04-16 Paper0/ms.tex
add appendix table for SIP coefficients

@@
 \begin{deluxetable*}{ccccrccccc}
-\tabletypesize{\footnotesize}
+%\tabletypesize{\footnotesize}
 \tablewidth{0pt} 
@@
 \tablehead{ 
-\colhead{Field} & \colhead{R.A.} & \colhead{Dec.} & \colhead{Instrument} & \colhead{Epoch} & \textit{HST}\footnotemark[1] & \colhead{$t$, hours} & {Depth\footnotemark[2]} & \colhead{FWHM}}
+\colhead{Field} & \colhead{R.A.} & \colhead{Dec.} & \colhead{Instrument} & \colhead{Epoch} & \textit{HST}\tablenotemark{a} & \colhead{$t$, hours} & {Depth\tablenotemark{b}} & \colhead{FWHM}}
 \startdata 
@@
 \vspace {-0.4cm}\\ 
-&  &         & \textit{VLT}/\HAWKI\footnotemark[3]     & 2013 Mar 21--2014 Jun 9      & C & 5.3 & 25.0 & $0\farcs41$
+&  &         & \textit{VLT}/\HAWKI\tablenotemark{c}     & 2013 Mar 21--2014 Jun 9      & C & 5.3 & 25.0 & $0\farcs41$
 \enddata 
 %\vspace{-0.8cm} 
-\footnotetext[1]{Coverage of \textit{Hubble} survey fields: C=Cluster, P=Parallel.
+\tablenotetext{a}{Coverage of \textit{Hubble} survey fields: C=Cluster, P=Parallel.
 Most \HAWKI\ pointings cover the two \textit{HST} fields simultaneously (C+P; Fig.~\ref{fig:layout}).
@@
 %\footnotetext[2]{Approximate exposure time equivalent at $0\farcs4$ seeing:  $t_{0.4}=\sum_i t_i \cdot (0.4/\mathrm{FWHM}_i)^2$.}
-\footnotetext[2]{Depth is defined as the 5$\sigma$ limiting magnitude for point sources, measured in $D=0\farcs6$ apertures (\S\ref{s:depth}).} 
-\footnotetext[3]{Archival observations from ESO program \href{http://archive.eso.org/wdb/wdb/eso/eso_archive_main/query?prog_id=090.A-0458(A)}{090.A-0458} (PI: Infante).}
+\tablenotetext{b}{Depth is defined as the 5$\sigma$ limiting magnitude for point sources, measured in $D=0\farcs6$ apertures (\S\ref{s:depth}).} 
+\tablenotetext{c}{Archival observations from ESO program \href{http://archive.eso.org/wdb/wdb/eso/eso_archive_main/query?prog_id=090.A-0458(A)}{090.A-0458} (PI: Infante).}
 \end{deluxetable*}
@@
 
+The \HAWKI\ instrument has small but non-negligible geometric distortion, resulting in astrometric shifts of order 2 pixels at the corners of the detector relative to the center.
+\cite{libralato:14} provide a precise determination of this distortion based on observations of star clusters.
+They present a full pixel map for the distortions in $x$ and $y$ across each of the detectors.
+To use this distortion model with the \textsc{drizzle} software, we fit the pixel offsets from \cite{libralato:14} with a third-order two-dimensional polynomial.
+The polynomial coefficients and additional header keywords needed to create a SIP FITS distortion model \citep{fits:sip} for the four \HAWKI\ detectors are provided in Table~\ref{tab:sip}.
+We adopt the \cite{libralato:14} definition of the \HAWKI\ chips, with the chip numbers set by the order they are found in the multi-extension FITS files.
+That is, ``Chip \#1'' is the first image extension of the raw FITS files provided by the ESO archive, with \textsc{extname}$=$CHIP1.INT1.
+The \textsc{extname} values for chips 2,3,4 are CHIP2.INT1, CHIP4.INT1, CHIP3.INT1, note the switched names of the last two chips.
+
+\begin{deluxetable}{lrrrr}
+\tabletypesize{\footnotesize}
+\tablewidth{0pt} 
+\tablecaption{\HAWKI\ SIP Header Keywords \label{tab:sip}} 
+\tablehead{ \colhead{Keyword} & \colhead{Chip \#1\tablenotemark{a}} & \colhead{Chip \#2} & \colhead{Chip \#3} & \colhead{Chip \#4} }
+\startdata 
+% EXTNAME & CHIP1.INT1 & CHIP2.INT1 & CHIP4.INT1 & CHIP3.INT1 \\
+% DET CHIP ID & ESO-Hawaii2RG-chip66 & ESO-Hawaii2RG-chip78 & ESO-Hawaii2RG-chip88 & ESO-Hawaii2RG-chip79 \\
+% \hline
+A\_ORDER & \multicolumn{4}{c}{3} \\
+B\_ORDER & \multicolumn{4}{c}{3} \\
+CTYPE1 & \multicolumn{4}{c}{RA{-}{-}{-}TAN-SIP} \\
+CTYPE2 & \multicolumn{4}{c}{DEC{-}{-}TAN-SIP} \\
+CRPIX1 & \multicolumn{4}{c}{1024\tablenotemark{b}} \\
+CRPIX2 & \multicolumn{4}{c}{1024\tablenotemark{b}} \\
+A\_0\_0 &   $ 1.751\mathrm{e}-02 $ &  $  8.098\mathrm{e}-03$ &  $  9.104\mathrm{e}-04 $ &  $  9.096\mathrm{e}-03 $ \\
+B\_0\_0 &   $-2.649\mathrm{e}-02 $ &  $ -9.338\mathrm{e}-03$ &  $  4.709\mathrm{e}-03 $ &  $  1.043\mathrm{e}-02 $ \\
+A\_1\_0 &   $-3.924\mathrm{e}-04 $ &  $ -4.964\mathrm{e}-05$ &  $ -1.035\mathrm{e}-04 $ &  $ -6.559\mathrm{e}-05 $ \\
+B\_1\_0 &   $-9.546\mathrm{e}-04 $ &  $  9.804\mathrm{e}-04$ &  $  1.228\mathrm{e}-03 $ &  $ -7.145\mathrm{e}-04 $ \\
+A\_2\_0 &   $ 8.676\mathrm{e}-07 $ &  $ -7.943\mathrm{e}-07$ &  $  8.461\mathrm{e}-07 $ &  $ -8.579\mathrm{e}-07 $ \\
+B\_2\_0 &   $ 2.026\mathrm{e}-07 $ &  $  2.347\mathrm{e}-07$ &  $ -1.896\mathrm{e}-07 $ &  $ -2.628\mathrm{e}-07 $ \\
+A\_3\_0 &   $-2.059\mathrm{e}-10 $ &  $ -3.724\mathrm{e}-10$ &  $ -2.350\mathrm{e}-10 $ &  $ -3.941\mathrm{e}-10 $ \\
+B\_3\_0 &   $-6.040\mathrm{e}-11 $ &  $  4.753\mathrm{e}-11$ &  $  1.045\mathrm{e}-12 $ &  $ -1.550\mathrm{e}-11 $ \\
+A\_0\_1 &   $-4.413\mathrm{e}-05 $ &  $  6.173\mathrm{e}-05$ &  $  1.025\mathrm{e}-04 $ &  $ -5.038\mathrm{e}-05 $ \\
+B\_0\_1 &   $-4.422\mathrm{e}-04 $ &  $ -5.841\mathrm{e}-04$ &  $ -1.974\mathrm{e}-04 $ &  $ -4.552\mathrm{e}-04 $ \\
+A\_0\_2 &   $ 2.978\mathrm{e}-07 $ &  $ -3.050\mathrm{e}-07$ &  $  3.071\mathrm{e}-07 $ &  $ -2.771\mathrm{e}-07 $ \\
+B\_0\_2 &   $ 7.265\mathrm{e}-07 $ &  $  7.453\mathrm{e}-07$ &  $ -6.841\mathrm{e}-07 $ &  $ -7.129\mathrm{e}-07 $ \\
+A\_0\_3 &   $-1.277\mathrm{e}-11 $ &  $  3.299\mathrm{e}-11$ &  $ -4.466\mathrm{e}-13 $ &  $ -1.834\mathrm{e}-11 $ \\
+B\_0\_3 &   $-2.177\mathrm{e}-10 $ &  $ -2.529\mathrm{e}-10$ &  $ -2.448\mathrm{e}-10 $ &  $ -2.646\mathrm{e}-10 $ \\
+A\_1\_1 &   $ 6.053\mathrm{e}-07 $ &  $  6.554\mathrm{e}-07$ &  $ -3.966\mathrm{e}-07 $ &  $ -6.455\mathrm{e}-07 $ \\
+B\_1\_1 &   $ 5.875\mathrm{e}-07 $ &  $ -4.337\mathrm{e}-07$ &  $  5.017\mathrm{e}-07 $ &  $ -4.399\mathrm{e}-07 $ \\
+A\_1\_2 &   $-2.199\mathrm{e}-10 $ &  $ -2.919\mathrm{e}-10$ &  $ -1.881\mathrm{e}-10 $ &  $ -3.229\mathrm{e}-10 $ \\
+B\_1\_2 &   $ 9.213\mathrm{e}-12 $ &  $  7.527\mathrm{e}-11$ &  $  3.588\mathrm{e}-11 $ &  $ -4.918\mathrm{e}-11 $ \\
+A\_2\_1 &   $-4.114\mathrm{e}-11 $ &  $  7.218\mathrm{e}-11$ &  $ -2.636\mathrm{e}-11 $ &  $ -6.355\mathrm{e}-11 $ \\
+B\_2\_1 &   $-2.118\mathrm{e}-10 $ &  $ -2.506\mathrm{e}-10$ &  $ -1.433\mathrm{e}-10 $ &  $ -2.753\mathrm{e}-10 $
+\enddata
+\tablenotetext{a}{We adopt the \cite{libralato:14} definition of the \HAWKI\ chips, see text.}
+\tablenotetext{b}{The SIP distortion polynomial is defined relative to the reference pixel \textsc{crpix}, which we set to be the center of each detector.
+This is different from the default reference pixel in the raw images so the \textsc{crval} values also have to be shifted accordingly.}
+%\footnotetext[2]{Approximate exposure time equivalent at $0\farcs4$ seeing:  $t_{0.4}=\sum_i t_i \cdot (0.4/\mathrm{FWHM}_i)^2$.}
+%\footnotetext[2]{Depth is defined as the 5$\sigma$ limiting magnitude for point sources, measured in $D=0\farcs6$ apertures (\S\ref{s:depth}).} 
+%\footnotetext[3]{Archival observations from ESO program \href{http://archive.eso.org/wdb/wdb/eso/eso_archive_main/query?prog_id=090.A-0458(A)}{090.A-0458} (PI: Infante).}
+\end{deluxetable}
+
 \section{Description of provided data products}

### 2016-04-17 Paper0/ms.tex
small edit
Danilo's edits

@@
 
-During the last two decades, near-infrared (NIR) imaging has gained a dominant role in studies of galaxy formation and evolution.
+During the last two decades, near-infrared (NIR) imaging has gained a dominant role in studies of galaxy formation and evolution, enabling transformational advances in our understanding of galaxy populations at early cosmic times.
 
@@
 While imaging at even longer wavelengths ($\lambda>3$~$\mu$m) are needed to probe the rest-frame optical of galaxies in the first billion years of cosmic history, deep sub-arsec resolution imaging at these wavelengths will not be available until the advent of the {\it James Webb Space Telescope} ({\it JWST}).
-Deep $K$-band data with good enough (i.e., FWHM$<1^{\prime \prime}$) image quality have been fundamental to fully exploit {\it Spitzer}-IRAC imaging data characterized by much poorer spatial resolution, allowing for the correction of the resulting blending of sources in IRAC images.
+Deep $K$-band data with good enough (i.e., FWHM$<1^{\prime \prime}$) image quality have been fundamental to fully exploit {\it Spitzer}-IRAC imaging data characterized by much poorer spatial resolution, allowing for the mitigation of the resulting blending of sources in IRAC images.
 Finally, for space-based studies, $K$ band crucially fills the gap between the reddest {\it Hubble Space Telescope} ({\it HST}) filter (i.e., F160W, with $\lambda \sim 1.6$~$\mu$m) and the IRAC 3.6~$\mu$m pass-band, greatly improving the constraints on both the redshifts and the stellar-population properties of galaxies.
@@
 It is not surprising that, given the aforementioned reasons, all successful extragalactic surveys performed in the last fifteen years to investigate galaxies populations in the early universe invested significant resources to obtain deep NIR imaging data with good image quality, especially in the $K$ band, usually adopted as the detection band.
-Indeed, our understanding of galaxy formation and evolution in the last couple of decades has impressively proceeded forward mostly driven by NIR progressively deeper, wider, and with better image quality, surveys, such as ...
+Indeed, our understanding of galaxy formation and evolution in the last couple of decades has impressively proceeded forward mostly driven by progressively deeper, wider, and with better image quality, NIR surveys, such as ...
 FIRES, FIREWORKS, NMBS, UltraVISTA, etc...
 Subaru Super Deep Field (AO) \cite{minowa:05}...
-HUGS \citep{hugs}.\textcolor{red}{-- we should provide a complete list of NIR surveys here -- }
+HUGS \citep{hugs}.\textcolor{red}{xxx --  we should provide a complete list of NIR surveys here -- }
 
@@
 Most \HAWKI\ pointings cover the two \textit{HST} fields simultaneously (C+P; Fig.~\ref{fig:layout}).
-The MOSFIRE observations of MACS-0717 and MACS-1149 require two separate pointings of the instrument to cover the cluster and parallel \textit{HST} fields, which have the different characteristics as indicated.}
+\\The MOSFIRE observations of MACS-0717 and MACS-1149 require two separate pointings of the instrument to cover the cluster and pa-\\rallel \textit{HST} fields, which have the different characteristics as indicated.}
 %\footnotetext[2]{Approximate exposure time equivalent at $0\farcs4$ seeing:  $t_{0.4}=\sum_i t_i \cdot (0.4/\mathrm{FWHM}_i)^2$.}
@@
 
-The latest effort to further our knowledge of galaxy formation and evolution is represented by the HST Frontier Fields program.
-\textcolor{red}{-- brief description of HFF, followed by the need for mega-deep K-band data --}
+The latest effort to further our knowledge of galaxy formation and evolution is represented by the {\it HST} Frontier Fields (HFF) program.
+The HFF program is a multi-cycle program consisting of $\sim$560 orbits of Director's Discretionary (DD) time that is imaging six deep fields centered on strong lensing galaxy clusters in parallel with six deep blank fields.
+The primary science goals of the twelve HFF fields are to 1) reveal the population of galaxies at $z=5-10$ that are 10-50 times fainter intrinsically than any presently known, 2) solidify our understanding of the stellar masses and star formation histories of faint galaxies, 3) provide the first statistically meaningful morphological characterization of star-forming galaxies at $z>5$, and 4) to find $z>8$ galaxies magnified by the cluster lensing, with some bright enough to make them accessible to spectroscopic follow-up.
+Along with {\it HST}, the {\it Spitzer Space Telescope} has devoted 1000 hours of DD time to image the HFF fields at 3.6$\mu$m and 4.5$\mu$m with IRAC.
+
+Whereas the main goal of the HFF is to explore the galaxy population in the first billion year of cosmic history, this dataset is also unique for its combination of surveyed area and depth for studies of galaxy evolution across most of the age of the universe, down to, and including, the redshifts of the targeted galaxy clusters ($z\approx0.3-0.5$).
+
+The HFF data alone is however not sufficient to robustly characterize red galaxies at $z>3$ because the H$_{\rm 160}$ band lies on the UV side of the rest-frame optical Balmer/4000\AA~breaks at these redshifts, resulting in sub-optimal accuracies in the photometric redshifts and stellar population properties (e.g., stellar mass and rest-frame optical color) of galaxies at $z>2$ and even at $z>4$.
+Very deep K-band imaging is required to significantly improve the precision of both photometric redshifts and derived stellar population properties (see \S\ref{s:discussion}).
+Moreover, at $z>8-9$, the K-band data helps constraining the Lyman-break redshifts (e.g., \citealt{bouwens:13}) and increases the wavelength lever arm for measuring the redshift evolution of the rest-frame UV slopes (i.e., dust content and/or metallicity) of the first galaxies (\citealt{bouwens:12a}; \citealt{bouwens:13}).
+
+To resolve this issue, i.e., the lack of K-band data over the HFF, we executed a program (dubbed {\it HFF-K$_s$} \textcolor{red}{\bf xxx how about finding a name for this program?}) to image all twelve HFF pointings in the $K_s$ band down to a comparable depth of the {\it HST} data using the instruments HAWK-I (mounted on the {\it VLT}) and MOSFIRE (mounted on the {\it Keck} telescope).
+This paper describes the {\it HFF-K$_s$} program and accompanies the data release.
+This paper is structured as follows.
+The observations and data reduction are presented in \S\ref{s:observations}, while \S\ref{s:mosaic} presents the drizzled K$_s$ mosaics, along with the quantification of the image quality, the noise properties and depth, and the completeness.
+Finally, \S\ref{s:discussion} concludes the paper showing a few crucial improvements introduced by the K$_s$ data.
 
@@
 $K_s$-band observations of the Abell 2744 and MACS-0416 fields were obtained between 2013 Oct and 2014 Feb with the High Acuity Wide-field K-band Imager \citep[\HAWKI;][]{hawki, hawki2} on the 8.2\,m UT4 telescope at the ESO \textit{Very Large Telescope} (ESO program \mbox{092.A-0472}).
-\HAWKI\ $K_s$-badn bservations of the Abell S1063 and Abell 370 fields were obtained between 2015 Jul and 2016 Jan from a subsequent program (095.A-0533).
+\HAWKI\ $K_s$-band observations of the Abell S1063 and Abell 370 fields were obtained between 2015 Jul and 2016 Jan from a subsequent program (095.A-0533).
 \HAWKI\ is composed of four chips with 2048$\times$2048 $0\farcs106$ pixels, and the full $7\farcm5\times7\farcm5$ \HAWKI\ field-of-view is perfectly suited to cover both the primary and parallel \textit{HST} ACS optical and WFC3 IR fields simultaneously in a single pointing (see Fig.\ref{fig:layout}).
@@
 \begin{figure*}
-\epsscale{1.18}
+\epsscale{1.17}
 \plotone{hawkiff_layout.pdf} % Figures/exposure_map.pdf
@@
 
-The \HAWKI\ and MOSFIRE observations were reduced with a pipeline that has been developed for previous surveys with the NEWFIRM \citep[NMBS;][]{whitaker:nmbs} and FOURSTAR \citep[zFOURGE;][, Labb\'e in prep]{spitler:12} infrared imaging instruments.
+The \HAWKI\ and MOSFIRE observations were reduced with a pipeline that has been developed for previous surveys with the NEWFIRM \citep[NMBS;][]{whitaker:nmbs} and FOURSTAR (zFOURGE; \citealt{spitler:12}; Labb\'e et al., in prep.) infrared imaging instruments.
 Treating each detector individually, the pipeline is easily modified for the different instrument configurations of the four \HAWKI\ and single MOSFIRE chips.
@@
 \begin{figure}[!t]
-\epsscale{1.18}
+\epsscale{1.17}
 \plotone{2mass_phot.pdf} % Figures/exposure_map.pdf
@@
 The 2-Micron All Sky Survey \citep[2MASS;][]{2mass} provides an additional check on the photometric calibration, though the comparison is limited due to relatively little brightness overlap between the faint end of reliable 2MASS photometry and the bright end where stars are in the linear regime of the detectors on the 8--10\,m telescopes.
-A comparison of the observed photometry in the four survey fields to the 2MASS catalog magnitudes is shown in Fig.~\ref{fig:2mass}.
+A comparison of the observed photometry in the four survey fields to the 2MASS catalog magnitudes is shown in Figure~\ref{fig:2mass}.
 The stellar photometry on the deep $K_s$-band mosaics is measured within $1\arcsec$ apertures corrected to infinity with the curves of growth described below.
-The error-bars on the points in Fig.~\ref{fig:2mass} come from the 2MASS catalog; the photometric uncertainties are negligible for such bright stars in the deep images ($\mathrm{S/N}>100$).
+The error-bars on the points in Figure~\ref{fig:2mass} come from the 2MASS catalog; the photometric uncertainties are negligible for such bright stars in the deep images ($\mathrm{S/N}>100$).
 The 2MASS comparison suggests a systematic offset of $-0.09$~mag for the Abell S1063 field.
-However, the exposure time for that field was similar to those of the additional deep \HAWKI\ fields and the derived depths of \textit{all} deep \HAWKI\ fields are nearly identical (Table~\ref{tab:summary}).
+However, the exposure time for that field was similar to those of the additional deep \HAWKI\ fields and the derived depths of \textit{all} deep \HAWKI\ fields are nearly identical (Tab.~\ref{tab:summary}).
 Given the overall agreement with the 2MASS photometry, we estimate that the systematic uncertainty of the absolute photometric calibration is $\sigma_\mathrm{sys}\leq0.05$~mag.
+\textcolor{red}{\bf xxx It is not clear here if the offset of -0.09 mag found for Abell S1063 was actually applied to the K-band mosaic...
+Given your argument, I am guessing it was not, approach that I would agree with, but it is not clear...
+probably worth being explicitly...
+}
 
@@
 Reference absolute astrometric catalogs were generated from \textit{HST} images when available (i.e., the chips overlapping the cluster and parallel \textit{HST} fields) and public \textit{Subaru} Suprime-Cam $r_c$-band images otherwise\footnote{\url{http://archive.stsci.edu/pub/hlsp/clash/}}.
-Next, object catalogs are generated with the SEXtractor software \citep{bertin:96} for each individual background-subtracted exposure and transformations to the reference frame (shift, rotation and scale) are computed with with the \texttt{stsci.stimage.xyxymatch} and \texttt{scikit-image.transform}\footnote{\url{http://scikit-image.org}} Python software tools.
+Next, object catalogs are generated with the SEXtractor software \citep{bertin:96} for each individual background-subtracted exposure and transformations to the reference frame (shift, rotation and scale) are computed with the \texttt{stsci.stimage.xyxymatch} and \texttt{scikit-image.transform}\footnote{\url{http://scikit-image.org}} Python software tools.
 For the \HAWKI\ exposures, we fit a third-order polynomial to the geometric distortion model determined by \cite{libralato:14} and specify the polynomial terms as ``Simple Imaging Polynomial'' (SIP) coefficients \citep{fits:sip} in the individual exposure FITS files.
@@
 \begin{figure}[!ht]
-\epsscale{1.18}
+\epsscale{1.17}
 \plotone{ivar_mosaics.pdf} % Figures/exposure_map.pdf
@@
 \begin{figure*}[!t]
-\epsscale{1.18}
+\epsscale{1.17}
 \plotone{curve_of_growth.pdf} % Figures/exposure_map.pdf
@@
 These extended Moffat profiles appear to be characteristic of deep $K_s$-band images and they have substantially more flux at large radii ($r>\mathrm{FWHM}/2$) than Gaussian profiles with the same FWHM (orange curves in the left panels).
-\label{fig:cog}}
+\textcolor{red}{\bf xxx In the bottom right panel, the inverse aperture correction of the parallel field is impossible to read, behind the other number...you may want to shift it upward a but}  \label{fig:cog}}
 \end{figure*}
@@
 Finally, we subtract a cell-based background from the final mosaics using an algorithm based on that used by SExtractor and SWarp but that provides more aggressive masking of flux in the outer isophotes of bright galaxies.
-The mosaics are shown in Fig.~\ref{fig:layout}, along with the position of the deep Frontier Fields \textit{HST} imaging fields.
-We also compute the robust NMAD scatter \citep{brammer:08} of empty sky pixels within the same cells to empirically calibrate the inverse variance maps that are shown in Fig.~\ref{fig:ivar_mosaics} (see also \S\ref{s:depth}).
+The mosaics are shown in Figure~\ref{fig:layout}, along with the position of the deep Frontier Fields \textit{HST} imaging fields.
+We also compute the robust NMAD scatter \citep{brammer:08} of empty sky pixels within the same cells to empirically calibrate the inverse variance maps that are shown in Figure~\ref{fig:ivar_mosaics} (see also \S\ref{s:depth}).
 The final image and inverse variance mosaics are provided as an ESO Phase 3 data product \xxx\ URL, as well as attached as a digital product to this manuscript (\xxx\ I think ApJSS can and wants to do this).
@@
 \begin{figure}[!t]
-\epsscale{1.18}
+\epsscale{1.17}
 \plotone{drizzle_cutouts.pdf} % Figures/exposure_map.pdf
@@
 \begin{figure}[!t]
-\epsscale{1.18}
+\epsscale{1.17}
 \plotone{correlated_noise_figure.pdf} % Figures/exposure_map.pdf
@@
 
-Fig.~\ref{fig:cog} shows the curves of growth for stars identified by the tight relationship between their brightness and half-light radii \citep[see, e.g., Fig.~13 of][]{skelton:14}.
+Figure~\ref{fig:cog} shows the curves of growth for stars identified by the tight relationship between their brightness and half-light radii \citep[see, e.g., Fig.~13 of][]{skelton:14}.
 The image quality of the $K_s$-band mosaics is excellent with stellar FWHM$\lesssim0\farcs4$ for the deep \HAWKI\ fields, thanks in large part to the execution of the \HAWKI\ observations in service mode ensuring optimal and uniform image quality across the many hours of integration on the survey fields.
@@
 The result is that the image quality of the MACS-0717 and MACS-1149 fields is somewhat degraded in comparison, at $0\farcs42$--$0\farcs54$, and the image quality differs measurably between the two MOSFIRE pointings in the MACS-0717 field.
-Despite the compact cores of the stellar PSFs, the curves of growth shown in Fig.~\ref{fig:cog} deviate from Gaussian profiles with significantly higher flux in the outer wings, with a shape more consistent with a Moffat profile \citep{trujillo:01}  with $\beta\sim2$.
+Despite the compact cores of the stellar PSFs, the curves of growth shown in Figure~\ref{fig:cog} deviate from Gaussian profiles with significantly higher flux in the outer wings, with a shape more consistent with a Moffat profile \citep{trujillo:01}  with $\beta\sim2$.
 We caution that these extended profiles will yield significantly shallower final image depths than would be predicted with Exposure Time Calculators that may assume Gaussian profiles.
@@
 \begin{figure}[!t]
-\epsscale{1.18}
+\epsscale{1.17}
 \plotone{aper_depth.pdf} % Figures/exposure_map.pdf
-\caption{\textit{Top panel:} The product of the inverse curves-of-growth (Fig.~\ref{fig:cog}) and the predicted variance within a photometric aperture (Eq.\ref{eq:aper_var}) gives the point-source sensitivity as a function of aperture size.
+\caption{\textit{Top panel:} The product of the inverse curves-of-growth (Fig.~\ref{fig:cog}) and the predicted variance within a photometric aperture (Eq.~\ref{eq:aper_var}) gives the point-source sensitivity as a function of aperture size.
 The S/N is maximized with an aperture somewhat larger than the FWHM; depths evaluated at $D=0\farcs6$ near the maxima are indicated.)
@@
 
-A comparison of final combined images created with output $0\farcs1$ and $0\farcs06$ pixel grids using the drizzle implementation to versions created with the SWarp software \citep{swarp} is shown in Fig.~\ref{fig:drizzle_cutout}.
+A comparison of final combined images created with output $0\farcs1$ and $0\farcs06$ pixel grids using the drizzle implementation to versions created with the SWarp software \citep{swarp} is shown in Figure~\ref{fig:drizzle_cutout}.
 In the later case, the raw pixels are resampled twice before making the final mosaic, once combining the raw exposures in each OB and a second time combining the separate OBs into the final stack with an arbitrary pixel grid.
@@
 
-The fact that the drizzle-combined images in Fig.~\ref{fig:drizzle_cutout} (panels \textit{a} and \textit{b}) appear to the eye to be noisier than their SWarped counterparts (panels \textit{c} and \textit{d}) is simply the result of a smoothing of the pixel noise in the latter case rather than reflecting true underlying differences in their pixel-to-pixel variance.
+The fact that the drizzle-combined images in Figure~\ref{fig:drizzle_cutout} (panels \textit{a} and \textit{b}) appear to the eye to be noisier than their SWarped counterparts (panels \textit{c} and \textit{d}) is simply the result of a smoothing of the pixel noise in the latter case rather than reflecting true underlying differences in their pixel-to-pixel variance.
 \cite{casertano:00} discuss how the noise properties of correlated output pixels are affected by the relative sizes of the input and output pixel grids and the adopted drizzle parameters.
@@
 
-We explore these correlations between the output pixels in Fig.~\ref{fig:corr_noise}.
+We explore these correlations between the output pixels in Figure~\ref{fig:corr_noise}.
 For the black and red curves in each panel, we measure value differences between random pairs of pixels sorted by the separation between the pixels for the drizzled and SWarped images, respectively.
@@
 As expected, the SWarp-combined images show just such a depression for pairs of pixels 1--2 pixels apart.
-The yellow curves in Fig.~\ref{fig:corr_noise} show the same pairwise differences for the drizzled images now convolved by small Gaussian kernels as indicated, which agree very well with the curves for the images with correlated pixels.
+The yellow curves in Figure~\ref{fig:corr_noise} show the same pairwise differences for the drizzled images now convolved by small Gaussian kernels as indicated, which agree very well with the curves for the images with correlated pixels.
 
@@
 to be decreased at all separations.
-That the pixel variances can't be used directly in the presence of such correlations \citep{casertano:00} is the primary reason why significant effort has been devoted to placing random ``empty apertures'' across images in order to characterize their noise properties \citep[e.g.,][]{quadri:musyc, whitaker:nmbs, skelton:14}.
+That the pixel variances cannot be used directly in the presence of such correlations \citep{casertano:00} is the primary reason why significant effort has been devoted to placing random ``empty apertures'' across images in order to characterize their noise properties \citep[e.g.,][]{labbe:03,forster:06,quadri:musyc, whitaker:nmbs, skelton:14}.
 Now by eliminating the pixel correlations with the drizzle image combination technique (with a small pixfrac) we have shown that the pixel statistics are robust and it is now trivial to compute the expected variance within an arbitrary photometric aperture, for example:
@@
 \end{equation}
-for circular apertures with diameter, $D$, in pixels, and where $\sigma_\mathrm{pix}^2$ is the per-pixel variance determined from an analysis such as that in Fig.~\ref{fig:corr_noise}.
+for circular apertures with diameter, $D$, in pixels, and where $\sigma_\mathrm{pix}^2$ is the per-pixel variance determined from an analysis such as that in Figure~\ref{fig:corr_noise}.
 
-The product of Eq.~\ref{eq:aper_var} and the inverse of the curves of growth shown in Fig.~\ref{fig:cog} (i.e., aperture corrections) defines an optimal photometric aperture size where this product is minimized \citep{whitaker:nmbs}.
+The product of Equation~\ref{eq:aper_var} and the inverse of the curves of growth shown in Figure~\ref{fig:cog} (i.e., aperture corrections) defines an optimal photometric aperture size where this product is minimized \citep{whitaker:nmbs}.
 For our $K_s$-band mosaics, this optimal aperture has $D\sim0\farcs6$, just larger than the FWHM of point-source profiles.
-The $5\,\sigma$ limiting magnitudes (AB) within $0\farcs6$ diameter apertures are indicated in Fig.~\ref{fig:depth} and listed in Table~\ref{tab:summary}.
+The $5\,\sigma$ limiting magnitudes (AB) within $0\farcs6$ diameter apertures are indicated in Figure~\ref{fig:depth} and listed in Table~\ref{tab:summary}.
 Reaching 26th magnitude (AB), the \HAWKI\ images presented here are among the deepest $K_s$-band images ever obtained and will provide a crucial complement to the deep \textit{HST} and \textit{Spitzer} imaging of the Frontier Fields.
@@
 \cite{hugs} report a final depth of 26.5 AB for point sources within $D=0\farcs4$ apertures for their deepest field, ``GOODS-D1'', which has an exposure time of 31.5 hours similar to the deep \HAWKI\ fields summarized in Table~\ref{tab:summary}.
-\cite{hugs} assume ideal noise properties that likely suffer some residual correlation between adjacent pixels such as that indicated by the red curves in Fig.~\ref{fig:corr_noise}, where we found that the drizzled rms was some 40\% higher than that measured in the presence of the pixel correlations for $0\farcs1$ pixel mosaics.
+\cite{hugs} assume ideal noise properties that likely suffer some residual correlation between adjacent pixels such as that indicated by the red curves in Figure~\ref{fig:corr_noise}, where we found that the drizzled rms was some 40\% higher than that measured in the presence of the pixel correlations for $0\farcs1$ pixel mosaics.
 This effect ($\sim$0.4 mag) along with the relative exposure time difference ($\sim$0.1 mag) can account for much of the difference between the reported depths of the deepest HUGS pointings and the \textit{Hubble} Frontier Fields \HAWKI\ images described here.
@@
 \begin{figure}[!t]
-\epsscale{1.18}
+\epsscale{1.17}
 \plotone{completeness.pdf} % Figures/exposure_map.pdf
@@
 Only MACS-1149 differs appreciably in the area around the \textit{HST} cluster area, with the curve as shown;  the completeness curves in the cluster areas of the other fields differ from the curves shown by $\lesssim0.1$ mag and are omitted for clarity.
-\xxx{Danilo please check}.
+\textcolor{red}{\bf xxx I would probably plot also the macs0717 cluster completeness curve, given it is actually different data...
+}
 \label{fig:completeness}}  
@@
 
-We complete source detection completeness curves as a function of source brightness following the techniques described by \cite{whitaker:nmbs} and \cite{muzzin:13}.
+We compute source detection completeness curves as a function of source brightness following the techniques described by \cite{whitaker:nmbs} and \cite{muzzin:13}.
 Briefly, we insert artificial sources in blank regions of the portions of the images covering the \textit{HST} ``parallel'' fields and compute the fraction of sources recovered with SExtractor as a function of the source magnitude.
-The resulting completeness curves are shown in Fig.~\ref{fig:completeness}.
+The resulting completeness curves are shown in Figure~\ref{fig:completeness}.
 As in the sensitivity analysis above, we find that the detection completeness curves are nearly identical for the deep \HAWKI\ fields, with 50\% (90\%) source completeness at $K_s\sim25.9$ (25.7) AB.
@@
 28.7 AB in $H_{160}$), the $K_s$ band images can be more effectively exploited for most applications by detecting objects in the deeper \textit{HST} images and performing forced $K_s$-band photometry at the positions of the \textit{HST} sources.
-The generation of such catalogs is beyond the scope of the present work and will be by Shipley et al.
-(in prep).
+The generation of such catalogs is beyond the scope of the present work and will be presented in Shipley et al.
+(2016; in preparation).
 
@@
 \begin{figure}[!ht]
-\epsscale{1.18}
+\epsscale{1.17}
 \plotone{image_cutouts.pdf} % Figures/exposure_map.pdf
@@
 \begin{figure*}
-\epsscale{1.18}
+\epsscale{1.17}
 \plotone{delta_UV.pdf} % Figures/exposure_map.pdf
 \caption{\textit{Left:} Difference in the derived rest-frame $U-V$ colors as a function of photometric redshift for galaxies in the MACS-0416 parallel field, before and after including $K_s$ photometry along with the \textit{HST} and \textit{Spitzer} measurements.
-The scatter in the derived $U-V$ colors at $z>2.5$ is large ($\sim$0.2 mag) where the Balmer/4000~\AA\ break is in the gap between \textit{HST} $H_{160}$ and the \textit{Spitzer} IRAC 3.6~\mum\ bands, which is well then constrained by including $K_s$.
+The scatter in the derived $U-V$ colors at $z>2.5$ is large ($\sim$0.2 mag) where the Balmer/4000~\AA\ break is in the gap between \textit{HST} $H_{160}$ and the \textit{Spitzer} IRAC 3.6~\mum\ bands, which is then well constrained by including $K_s$.
 This large scatter translates directly into large scatter on other quantities of interest, such as the stellar mass to light ratio and the age of the stellar populations.
@@
 
-We have presented the observations and reduction for a collection of extremely deep $K_s$-band images covering the lensing clusters observed as part of the \textit{Hubble} and \textit{Spitzer} Frontier Fields program (Lotz et al., in prep; Capak et al., in prep).
+We have presented the observations and reduction for a collection of extremely deep $K_s$-band images covering the lensing clusters observed as part of the \textit{Hubble} and \textit{Spitzer} Frontier Fields program (Lotz et al., in prep.; Capak et al., in prep.).
 Service mode scheduling of $>$25~hour \HAWKI\ integrations on the southern hemisphere clusters visible from the ESO/\textit{VLT} result in remarkably uniform, high-quality mosaics across the separate survey fields, with superb image quality ($0\farcs4$~FWHM) and photometric depths reaching AB=26.0 mag (5$\sigma$).
@@
 
-A comparison of the $K_s$-band and space-based Frontier Fields imaging is shown in Fig.~\ref{fig:cutout}, showing just a small cutout of the MACS-0416 ``parallel'' field.
+A comparison of the $K_s$-band and space-based Frontier Fields imaging is shown in Figure~\ref{fig:cutout}, showing just a small cutout of the MACS-0416 ``parallel'' field.
 The deep optical and near-infrared \textit{Hubble} imaging provides spectacular multiwavelength spatially-resolved information on physical scales of just $\sim$1~kpc (Fig.~\ref{fig:cutout}a).
@@
 Sampling of more evolved stellar populations at $z>3$ with redder colors requires deep imaging at longer wavelengths.
-This can be obtained with imaging in the \textit{Spitzer} IRAC bands \citep[e.g.,][]{marchesini:10, stefanon:13}, though at the cost of low spatial resolution ($\sim1\farcs6$, Fig.~\ref{fig:cutout}d).
+This can be obtained with imaging in the \textit{Spitzer} IRAC bands \citep[e.g.,][]{marchesini:10, stefanon:13, stefanon:15}, though at the cost of low spatial resolution ($\sim1\farcs6$, Fig.~\ref{fig:cutout}d).
 This resolution is insufficient for spatially resolving the distant galaxies and is prone to blending of faint sources, particularly in the crowded Frontier Fields clusters where the surface density of both foreground cluster galaxies and faint background galaxies is high at the depths of interest.
@@
 Here we explore the quantitative constraints provided by the deep $K_s$ band imaging in terms of the photometric redshifts and derived intrinsic properties of galaxies in the survey field.
-We construct preliminary PSF-matched photometric catalogs of the \textit{HST} imaging and deblended photometry from the IRAC bands following the methodology described by \XXX{ref, Labb\'e}.
-We combine then combine the space-based catalogs with aperture photometry matched directly from $K_s$-detected catalogs derived from the completeness simulations described above\footnote{Full robust \textit{HST}+$K$+IRAC catalogs will be presented by Shipley et al., in prep.}.
+We construct preliminary PSF-matched photometric catalogs of the \textit{HST} imaging and deblended photometry from the IRAC bands following the methodology described by \cite{skelton:14}.
+We then combine the space-based catalogs with aperture photometry matched directly from $K_s$-detected catalogs derived from the completeness simulations described above\footnote{Full robust \textit{HST}+$K$+IRAC catalogs will be presented by Shipley et al.
+(2016), in preparation.}.
 We then compute photometric redshifts with EAZY \citep{brammer:08} for the catalogs with and without the $K_s$-band photometry included.
-The left panel of Fig.~\ref{fig:delta_UV} shows the difference in the rest-frame $U-V$ color derived from the photometric redshift fit, which probes the strength of the Balmer/4000~\AA\ break and is a proxy for the age and mass-to-light ratio of the underlying stellar population.
+The left panel of Figure~\ref{fig:delta_UV} shows the difference in the rest-frame $U-V$ color derived from the photometric redshift fit, which probes the strength of the Balmer/4000~\AA\ break and is a proxy for the age and mass-to-light ratio of the underlying stellar population.
 
@@
 
-The right two panels of Fig.~\ref{fig:delta_UV} show a single galaxy that illustrates how these systematic effects are not trivial and will likely result in biases in the interpretation of the galaxy population properties derived from the \textit{HST} and IRAC observations alone.
+The right two panels of Figure~\ref{fig:delta_UV} show a single galaxy that illustrates how these systematic effects are not trivial and will likely result in biases in the interpretation of the galaxy population properties derived from the \textit{HST} and IRAC observations alone.
 The spectral energy distribution (SED) shown is steadily rising through the reddest WFC3/IR bands, and then shows a sharp break with bright detections in the IRAC bands.
@@
 
-\XXX{blah blah, few more sentences wrapping up.}
+Our ultra-deep K$_{s}$-band mosaics will complement the {\it HST} and {\it Spitzer} data of the HFF, ensuring estimation of the most accurate photometric redshifts, rest-frame colors and luminosities, and stellar population properties (e.g., stellar mass and dust extinction) for galaxies at $z>2$, enhancing enormously the scientific impact of the Hubble Frontier Fields program.
+For example, they will allow for the investigation of the spectral energy distribution of $z\approx4$ galaxies or the evolution of the stellar mass function of galaxies at high redshift.
+Finally, we stress that, our mosaics more than double the total area of the sky imaged in the K$_{s}$ band down to these extreme depths.
 
@@
 \noindent We gratefully acknowledge funding support from the STScI Director's Discretionary Research Fund.
-\xxx\ acknowledge NASA/Keck time + Mauna Kea \xxx.
+D.M.
+acknowledges the support of the Research Corporation for Science Advancement's Cottrell Scholarship, and of the National Science Foundation under Grant No.
+1513473.
+D.M.
+and D.L.-V.
+acknowledge support from program HST-AR-14302, provided by NASA through a grant from the Space Telescope Science Institute, which is operated by the Association of Universities for Research in Astronomy, Incorporated, under NASA contract NAS5-26555.
+This work was supported by NASA Keck PI Data Awards, administered by the NASA Exoplanet Science Institute.
+Some of the data presented in this study were obtained at the W.
+M.
+Keck Observatory from telescope time allocated to the National Aeronautics and Space Administration through the agency's scientific partnership with the California Institute of Technology and the University of California.
+The Observatory was made possible by the generous financial support of the W.
+M.
+Keck Foundation.
+The authors wish to recognize and acknowledge the very significant cultural role and reverence that the summit of Mauna Kea has always had within the indigenous Hawaiian community.
+We are most fortunate to have the opportunity to conduct observations from this mountain.
 This research made use of Astropy, a community-developed core Python package for Astronomy \citep{astropy}, and open-source Python modules Numpy, Scipy, and Matplotlib.

### 2016-04-17 Paper0/ms.tex
update cog figure, nearly complete draft

@@
 Danilo Marchesini\altaffilmark{2},
-et~al.}
+Ivo Labb\'e\altaffilmark{3}, 
+Lee Spitler\altaffilmark{4}, 
+Daniel Lange-Vagle\altaffilmark{2},
+Elizbeth Barker\altaffilmark{1},
+Adriano Fontana\altaffilmark{5},
+Audrey Galametz\altaffilmark{6},
+Anna Ferr\'e-Mateu\altaffilmark{7},
+Tadayuki Kodama\altaffilmark{8},
+Britt Lundgren\altaffilmark{9},
+Nicholas Martis\altaffilmark{2},
+Adam Muzzin\altaffilmark{10},
+Mauro Stefanon\altaffilmark{3},
+Masayuki Tanaka\altaffilmark{11},
+Sune Toft\altaffilmark{12},
+Arjen van der Wel\altaffilmark{13},
+Benedetta Vulcani\altaffilmark{14}, 
+Katherine E. Whitaker\altaffilmark{15}
+}
 % Pieter G.\ van Dokkum\altaffilmark{2},
@@
 \altaffiltext{2}{Physics and Astronomy Department, Tufts University, Medford, MA, 02155, USA}
+\altaffiltext{3}{Leiden Observatory, Leiden University, NL-2300 RA Leiden, Netherlands}
+\altaffiltext{4}{\xxx}
+\altaffiltext{5}{\xxx}
+\altaffiltext{6}{\xxx}
+\altaffiltext{7}{\xxx}
+\altaffiltext{8}{\xxx}
+\altaffiltext{9}{\xxx}
+\altaffiltext{10}{\xxx}
+\altaffiltext{11}{\xxx}
+\altaffiltext{12}{\xxx}
+\altaffiltext{13}{Max Planck Institute for Astronomy (MPIA), K\"onigstuhl 17, 69117, Heidelberg, Germany}
+\altaffiltext{14}{\xxx}
+\altaffiltext{15}{\xxx}
 % \altaffiltext{6}
 % {Department of Astronomy, Yale University, New Haven, CT 06520, USA}
-% \altaffiltext{3}
-% {Leiden Observatory, Leiden University, NL-2300 RA Leiden, Netherlands}
 % \altaffiltext{4}
 % {UCO/Lick Observatory, University of California, Santa Cruz, CA 95064}
-% \altaffiltext{3}
-% {Max Planck Institute for Astronomy (MPIA), K\"onigstuhl 17,
-% 69117, Heidelberg, Germany}
 % \altaffiltext{4}
@@
 
-Detection of galaxies in the $K$-band ($\lambda_{\rm eff} \sim 2.2$~$\mu$m) provided the first opportunity to construct a comprehensive picture of the population of galaxies in the early universe, as it enabled the discovery of galaxies at $z > 2$ with suppressed emission in the optical (i.e., rest-frame UV) due to evolved stellar populations and/or significant amount of dust extinction (e.g., \textcolor{red}{refs}).
-In fact, these galaxies, which dominate the high-mass end of the high-$z$ galaxy population, were previously missed by rest-frame UV selection techniques (e.g., $U$-dropout galaxies; \textcolor{red}{refs}).
-Imaging in the $K$-band allows for the direct probing out to $z\approx5$ of rest-frame wavelengths longer than the Balmer break.
-Sampling the rest-frame optical wavelength regime is crucial for high-$z$ studies, as significantly less affected by dust obscuration and a better probe of the galaxy stellar mass compared to the rest-frame UV (more sensitive to unobscured star formation).
-While imaging at even longer wavelengths ($\lambda>3$~$\mu$m) are needed to probe the rest-frame optical of galaxies in the first billion years of cosmic history, deep sub-arsec resolution imaging at these wavelengths will not be available until the advent of the {\it James Webb Space Telescope} ({\it JWST}).
-Deep $K$-band data with good enough (i.e., FWHM$<1^{\prime \prime}$) image quality have been fundamental to fully exploit {\it Spitzer}-IRAC imaging data characterized by much poorer spatial resolution, allowing for the mitigation of the resulting blending of sources in IRAC images.
-Finally, for space-based studies, $K$ band crucially fills the gap between the reddest {\it Hubble Space Telescope} ({\it HST}) filter (i.e., F160W, with $\lambda \sim 1.6$~$\mu$m) and the IRAC 3.6~$\mu$m pass-band, greatly improving the constraints on both the redshifts and the stellar-population properties of galaxies.
+Detection of galaxies in the $K$-band ($\lambda_{\rm eff} \sim 2.2$~$\mu$m) provided the first opportunity to construct a comprehensive picture of the population of galaxies in the early universe, as it enabled the discovery of galaxies at $z > 2$ that are faint at observed optical (rest-frame ultraviolet) wavelengths due to evolved stellar populations and/or significant amount of dust extinction \citep[e.g.,][]{franx:03, labbe:03, forster:04, minowa:05, kajisawa:06, brammer:07, taylor:09a}.
+In fact, these galaxies, which dominate the high-mass end of the high-$z$ galaxy population, were previously missed by rest-frame UV selection techniques (e.g., $U$-dropout galaxies) \citep[e.g.,][]{vandokkum:06}.
+Imaging in the $K$-band allows for the direct sampling of rest-frame wavelengths longer than the Balmer break out to $z\approx5$.
+Sampling the rest-frame optical wavelength regime is crucial for high-$z$ studies, as it is significantly less affected by dust obscuration and a better probe of the galaxy stellar mass compared to the rest-frame UV, which is more sensitive to unobscured star formation \citep[e.g., ][]{fontana:06, marchesini:09}.
+
+While imaging at even longer wavelengths ($\lambda>3$~$\mu$m) is needed to probe the rest-frame optical of galaxies in the first billion years of cosmic history, deep sub-arsec resolution imaging at these wavelengths will not be available until the advent of the {\it James Webb Space Telescope} ({\it JWST}).
+Deep $K$-band data with good  (i.e., FWHM~$<1\arcsec$) image quality have been fundamental to fully exploit {\it Spitzer}-IRAC imaging data characterized by much poorer spatial resolution, allowing for the mitigation of the resulting blending of sources in IRAC images \citep[e.g.,][]{labbe:05}.
+Finally, for space-based studies, the $K$ band crucially fills the gap between the reddest {\it Hubble Space Telescope} ({\it HST}) filter (i.e., F160W, with $\lambda \sim 1.6$~$\mu$m) and the IRAC 3.6~$\mu$m passband, greatly improving the constraints on both the photometric redshifts and the inferred stellar-population properties of galaxies in the survey areas.
 
 It is not surprising that, given the aforementioned reasons, all successful extragalactic surveys performed in the last fifteen years to investigate galaxies populations in the early universe invested significant resources to obtain deep NIR imaging data with good image quality, especially in the $K$ band, usually adopted as the detection band.
-Indeed, our understanding of galaxy formation and evolution in the last couple of decades has impressively proceeded forward mostly driven by progressively deeper, wider, and with better image quality, NIR surveys, such as ...
-FIRES, FIREWORKS, NMBS, UltraVISTA, etc...
-Subaru Super Deep Field (AO) \cite{minowa:05}...
-HUGS \citep{hugs}.\textcolor{red}{xxx --  we should provide a complete list of NIR surveys here -- }
+Indeed, our understanding of galaxy formation and evolution in the last couple of decades has impressively proceeded forward mostly driven by progressively deeper, wider, and with better image quality, NIR surveys, such as FIRES \citep{labbe:03}, Subaru Super Deep Field (AO) \citep{minowa:05}, MUSYC \citep{quadri:07, taylor:09b}, FIREWORKS \citep{wuyts:fireworks}, MODS \citep{kajisawa:11}, NMBS \citep{whitaker:nmbs}, UltraVISTA \citep{mccracken:ultravista}, TENIS \citep{hsieh:tenis}, WIRDS \citep{bielby:12}, ZFOURGE \citep{spitler:12}, and HUGS \citep{hugs}.
+
+The latest effort to further our knowledge of galaxy formation and evolution is represented by the {\it HST} Frontier Fields (HFF) program (Lotz et al., in prep).
+The HFF program is a multi-cycle program consisting of $\sim$560 orbits of Director's Discretionary (DD) time that is imaging six deep fields centered on strong lensing galaxy clusters in parallel with six deep blank fields.
+The primary science goals of the twelve HFF fields are to 1) reveal the population of galaxies at $z=5-10$ that are 10--50 times fainter intrinsically than any presently known, 2) solidify our understanding of the stellar masses and star formation histories of faint galaxies, 3) provide the first statistically meaningful morphological characterization of star-forming galaxies at $z>5$, and 4) to find $z>8$ galaxies magnified by the cluster lensing, with some bright enough to make them accessible to spectroscopic follow-up.
+Along with {\it HST}, the {\it Spitzer Space Telescope} has devoted 1000 hours of DD time to image the HFF fields at 3.6$\mu$m and 4.5$\mu$m with IRAC (Capak et al., in prep).
+
+Whereas the main goal of the HFF is to explore the galaxy population in the first billion year of cosmic history, this dataset is also unique for its combination of surveyed area and depth for studies of galaxy evolution across most of the age of the universe, down to, and including, the redshifts of the targeted galaxy clusters ($z\approx$~0.3--0.5).
+
+The space-based HFF data alone, however, are not sufficient to robustly characterize red galaxies at $z>3$ because the WFC3/IR $H_{160}$ band lies on the UV side of the rest-frame optical Balmer/4000\AA~break at these redshifts, resulting in sub-optimal accuracies in the photometric redshifts and stellar population properties (e.g., stellar mass and rest-frame optical color) of galaxies at $z>2$ and even at $z>4$.
+Very deep $K$-band imaging is required to significantly improve the precision of both photometric redshifts and derived stellar population properties (see \S\ref{s:discussion}).
+Moreover, at $z>8-9$, the K-band data helps constraining the Lyman-break redshifts (e.g., \citealt{bouwens:13}) and increases the wavelength lever arm for measuring the redshift evolution of the rest-frame UV slopes (i.e., dust content and/or metallicity) of the first galaxies (\citealt{bouwens:12a}; \citealt{bouwens:13}).
+
+To resolve this issue, i.e., the lack of $K$-band data over the HFF, we executed a program---``KIFF'': ``$K$-band Imaging of the Frontier Fields''---to image all twelve HFF pointings in the $K_s$ band down to a comparable depth of the {\it HST} data using the instruments HAWK-I (mounted on the {\it VLT}) and MOSFIRE (mounted on the {\it Keck I} telescope).
+This paper describes the observations and analysis of $K_s$ band data obtained by the KIFF program and is accompanied by the first full public release of the reduced full-depth $K_s$ band imaging mosaics.
+This paper is structured as follows.
+The observations and data reduction are presented in \S\ref{s:observations}, while \S\ref{s:mosaic} presents the drizzled $K_s$ mosaics, along with the quantification of the image quality, the noise properties and depth, and the completeness.
+Finally, \S\ref{s:discussion} concludes showing a few crucial improvements enabled by the deep $K_s$ data in consort with the space-based \textit{HST} and \textit{Spitzer} Frontier Fields imaging.
+AB magnitudes are used throughout, with $m_{K, \mathrm{AB}} - m_{K, \mathrm{Vega}} = 1.826$ and 1.821 and pivot wavelengths \citep{tokunaga:05} of 2.152~\mum\ and 2.147~\mum\ computed for the \HAWKI\ and MOSFIRE $K_s$ filter bandpasses, respectively.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%                                                                %
+%   Observations
+%                                                                %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Observations \& Data Reduction}
+\label{s:observations}
+
+Table~\ref{tab:summary} provides a summary of the survey fields and the characteristics of the $K_s$-band observational program.
+Additional details on the observing strategy and image reduction procedures are provided in the subsections below.
 
@@
 %\vspace{1cm}
-
-The latest effort to further our knowledge of galaxy formation and evolution is represented by the {\it HST} Frontier Fields (HFF) program.
-The HFF program is a multi-cycle program consisting of $\sim$560 orbits of Director's Discretionary (DD) time that is imaging six deep fields centered on strong lensing galaxy clusters in parallel with six deep blank fields.
-The primary science goals of the twelve HFF fields are to 1) reveal the population of galaxies at $z=5-10$ that are 10-50 times fainter intrinsically than any presently known, 2) solidify our understanding of the stellar masses and star formation histories of faint galaxies, 3) provide the first statistically meaningful morphological characterization of star-forming galaxies at $z>5$, and 4) to find $z>8$ galaxies magnified by the cluster lensing, with some bright enough to make them accessible to spectroscopic follow-up.
-Along with {\it HST}, the {\it Spitzer Space Telescope} has devoted 1000 hours of DD time to image the HFF fields at 3.6$\mu$m and 4.5$\mu$m with IRAC.
-
-Whereas the main goal of the HFF is to explore the galaxy population in the first billion year of cosmic history, this dataset is also unique for its combination of surveyed area and depth for studies of galaxy evolution across most of the age of the universe, down to, and including, the redshifts of the targeted galaxy clusters ($z\approx0.3-0.5$).
-
-The HFF data alone is however not sufficient to robustly characterize red galaxies at $z>3$ because the H$_{\rm 160}$ band lies on the UV side of the rest-frame optical Balmer/4000\AA~breaks at these redshifts, resulting in sub-optimal accuracies in the photometric redshifts and stellar population properties (e.g., stellar mass and rest-frame optical color) of galaxies at $z>2$ and even at $z>4$.
-Very deep K-band imaging is required to significantly improve the precision of both photometric redshifts and derived stellar population properties (see \S\ref{s:discussion}).
-Moreover, at $z>8-9$, the K-band data helps constraining the Lyman-break redshifts (e.g., \citealt{bouwens:13}) and increases the wavelength lever arm for measuring the redshift evolution of the rest-frame UV slopes (i.e., dust content and/or metallicity) of the first galaxies (\citealt{bouwens:12a}; \citealt{bouwens:13}).
-
-To resolve this issue, i.e., the lack of K-band data over the HFF, we executed a program (dubbed {\it HFF-K$_s$} \textcolor{red}{\bf xxx how about finding a name for this program?}) to image all twelve HFF pointings in the $K_s$ band down to a comparable depth of the {\it HST} data using the instruments HAWK-I (mounted on the {\it VLT}) and MOSFIRE (mounted on the {\it Keck} telescope).
-This paper describes the {\it HFF-K$_s$} program and accompanies the data release.
-This paper is structured as follows.
-The observations and data reduction are presented in \S\ref{s:observations}, while \S\ref{s:mosaic} presents the drizzled K$_s$ mosaics, along with the quantification of the image quality, the noise properties and depth, and the completeness.
-Finally, \S\ref{s:discussion} concludes the paper showing a few crucial improvements introduced by the K$_s$ data.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%                                                                %
-%   Observations
-%                                                                %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Observations \& Data Reduction}
-\label{s:observations}
-
-Table~\ref{tab:summary} provides a brief summary of the survey fields and observation characteristics.
-Additional details on the observing strategy and image reduction procedures are provided in the subsections below.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-$K_s$-band observations of the Abell 2744 and MACS-0416 fields were obtained between 2013 Oct and 2014 Feb with the High Acuity Wide-field K-band Imager \citep[\HAWKI;][]{hawki, hawki2} on the 8.2\,m UT4 telescope at the ESO \textit{Very Large Telescope} (ESO program \mbox{092.A-0472}).
-\HAWKI\ $K_s$-band observations of the Abell S1063 and Abell 370 fields were obtained between 2015 Jul and 2016 Jan from a subsequent program (095.A-0533).
-\HAWKI\ is composed of four chips with 2048$\times$2048 $0\farcs106$ pixels, and the full $7\farcm5\times7\farcm5$ \HAWKI\ field-of-view is perfectly suited to cover both the primary and parallel \textit{HST} ACS optical and WFC3 IR fields simultaneously in a single pointing (see Fig.\ref{fig:layout}).
-
-The \HAWKI\ observations were divided into individual service mode Observation Blocks (OBs); typically one or two OBs of a given field were observed per night as conditions allowed, and occasionally multiple fields were observed on the same night.
-The execution of each OB lasted either 60 or 90 minutes, depending on how many exposures were obtained in the block.
-The exposures were roughly one minute each, with multiple coadds of shorter reads making up the exposure ($\mathrm{NDIT}\times\mathrm{DIT} = 4\times15$\,s~=~60\,s for Abell 2744, 7$\times$8\,s~=~56\,s for MACS-0416, 8$\times$12\,s~=~96\,s for Abell S1063 and 8$\times$12\,s~=~96\,s for Abell 370\footnote{Longer DITs are preferable to reduce instrument overheads; the shorter DITs on the MACS-0416 field were done do accommodate \textit{VLT} service mode restrictions on bright source saturation, which have been eased since ESO Period 93.}).
-The telescope was offset with random dithers between each exposure to facilitate sky subtraction;  observations taken before 2013 Nov 4 were dithered within a 20$^{\prime\prime}$ box, which was subsequently increased to 40$^{\prime\prime}$ to improve sky subtraction in the cluster cores crowded with bright galaxies and intra-cluster light.
-The total on-sky integration times for the Abell 2744, MACS-0416, Abell S1063 and Abell 370 fields are 29.4, 25.5, \xxx, and \xxx\ hours, respectively.
-
-Imaging observations of the MACS-0717 and MACS-1149 fields in the $K_s$ filter were obtained on 2015 Jan 26/27 and 2015 Feb 24 (program N097M), respectively, and both fields again on 2016 Jan 21/22 (program N135M), with the Multi-Object Spectrometer For Infra-Red Exploration \citep[MOSFIRE;][]{mosfire} on the 10\,m \mbox{\textit{Keck I}} telescope.
-The MOSFIRE detector consists of $2048\times2048$ $0\farcs1798$ pixels; the resulting $6\farcm1\times6\farcm1$ field-of-view is unfortunately slightly too small to cover both \textit{HST} fields simultaneously and therefore requires two pointings to cover the entire deep \textit{HST} Frontier Fields imaging area (Fig.~\ref{fig:layout}).
-MOSFIRE exposures were obtained in a $3\times3$ ``Box9'' dither pattern spaced roughly 40$^{\prime\prime}$ between offset positions.
-The total integration times on the MOSFIRE survey fields are summarized in Table~\ref{tab:summary}.
-The individual detector integration times (DIT) were adjusted on the fly during the observing run to keep the total counts within the linear regime of the detector, while maintaining $\mathrm{NDIT}\times\mathrm{DIT} \sim 40$\,s.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}
+\begin{figure*}[!t]
 \epsscale{1.17}
@@
 %%%%%%%%%%%%%%%%
+
+$K_s$-band observations of the Abell 2744 and MACS-0416 fields were obtained between 2013 Oct and 2014 Feb with the High Acuity Wide-field K-band Imager \citep[\HAWKI;][]{hawki, hawki2} on the 8.2\,m UT4 telescope at the ESO \textit{Very Large Telescope} (ESO program \mbox{092.A-0472}).
+\HAWKI\ $K_s$-band observations of the Abell S1063 and Abell 370 fields were obtained between 2015 Jul and 2016 Jan from a subsequent program (095.A-0533).
+\HAWKI\ is composed of four chips with 2048$\times$2048 $0\farcs106$ pixels, and the full $7\farcm5\times7\farcm5$ \HAWKI\ field-of-view is perfectly suited to cover both the primary and parallel \textit{HST} ACS optical and WFC3 IR fields simultaneously in a single pointing (see Fig.\ref{fig:layout}).
+
+The \HAWKI\ observations were divided into individual service mode Observation Blocks (OBs); typically one or two OBs of a given field were observed per night as conditions allowed, and occasionally multiple fields were observed on the same night.
+The execution of each OB lasted either 60 or 90 minutes, depending on how many exposures were obtained in the block.
+The exposures were roughly one minute each, with multiple coadds of shorter reads making up the exposure ($\mathrm{NDIT}\times\mathrm{DIT} = 4\times15$\,s~=~60\,s for Abell 2744, 7$\times$8\,s~=~56\,s for MACS-0416, 8$\times$12\,s~=~96\,s for Abell S1063 and 8$\times$12\,s~=~96\,s for Abell 370\footnote{Longer DITs are preferable to reduce instrument overheads; the shorter DITs on the MACS-0416 field were done do accommodate \textit{VLT} service mode restrictions on bright source saturation, which have been eased since ESO Period 93.}).
+The telescope was offset with random dithers between each exposure to facilitate sky subtraction;  observations taken before 2013 Nov 4 were dithered within a 20$^{\prime\prime}$ box, which was subsequently increased to 40$^{\prime\prime}$ to improve sky subtraction in the cluster cores crowded with bright galaxies and intra-cluster light.
+The total on-sky integration times for the Abell 2744, MACS-0416, Abell S1063 and Abell 370 fields are 29.3, 25.8, 27.9, and 28.3 hours, respectively.
+This includes some time taken with unfavorable image quality or transparency conditions in service mode;  all on-sky exposures are included in the final mosaics with relative weighting designed to favor the optimal observing conditions (see \S\ref{s:mosaic} and Eq.~\ref{eq:weight}).
+
+Imaging observations of the MACS-0717 and MACS-1149 fields in the $K_s$ filter were obtained on 2015 Jan 26/27 and 2015 Feb 24 (program N097M), respectively, and both fields again on 2016 Jan 21/22 (program N135M), with the Multi-Object Spectrometer For Infra-Red Exploration \citep[MOSFIRE;][]{mosfire} on the 10\,m \mbox{\textit{Keck I}} telescope.
+The MOSFIRE detector consists of $2048\times2048$ $0\farcs1798$ pixels; the resulting $6\farcm1\times6\farcm1$ field-of-view is unfortunately slightly too small to cover both \textit{HST} fields simultaneously and therefore requires two pointings to cover the entire deep \textit{HST} Frontier Fields imaging area (Fig.~\ref{fig:layout}).
+MOSFIRE exposures were obtained in a $3\times3$ ``Box9'' dither pattern spaced roughly 40$^{\prime\prime}$ between offset positions.
+The total integration times on the MOSFIRE survey fields are summarized in Table~\ref{tab:summary}.
+The individual detector integration times (DIT) were adjusted on the fly during the observing run to keep the total counts within the linear regime of the detector, while maintaining $\mathrm{NDIT}\times\mathrm{DIT} \sim 40$\,s.
 
@@
 The uncertainties shown are taken from the 2MASS catalog; these stars are measured at very high S/N in the deep $K_s$-band mosaics.
-We estimate the overall systematic uncertainty in the photometric calibration to be $\lesssim0.05~\mathrm{mag}$.\label{fig:2mass}}
+\label{fig:2mass}}
 \end{figure}
@@
 The error-bars on the points in Figure~\ref{fig:2mass} come from the 2MASS catalog; the photometric uncertainties are negligible for such bright stars in the deep images ($\mathrm{S/N}>100$).
-The 2MASS comparison suggests a systematic offset of $-0.09$~mag for the Abell S1063 field.
+The 2MASS comparison suggests a systematic zeropoint offset of $-0.09$~mag for the Abell S1063 field.
 However, the exposure time for that field was similar to those of the additional deep \HAWKI\ fields and the derived depths of \textit{all} deep \HAWKI\ fields are nearly identical (Tab.~\ref{tab:summary}).
-Given the overall agreement with the 2MASS photometry, we estimate that the systematic uncertainty of the absolute photometric calibration is $\sigma_\mathrm{sys}\leq0.05$~mag.
-\textcolor{red}{\bf xxx It is not clear here if the offset of -0.09 mag found for Abell S1063 was actually applied to the K-band mosaic...
-Given your argument, I am guessing it was not, approach that I would agree with, but it is not clear...
-probably worth being explicitly...
-}
+Therefore, we do not apply this offset to the S1063 zeropoint.
+Given the overall good agreement with the 2MASS photometry, we estimate that the systematic uncertainty of the absolute photometric calibration is $\sigma_\mathrm{sys}\leq0.05$~mag.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%   Drizzle figure
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[!ht]
-\epsscale{1.17}
-\plotone{ivar_mosaics.pdf} % Figures/exposure_map.pdf
-\caption{Inverse variance maps of the $K_s$-band mosaics.
-The orientation of the fields is the same as that shown in Fig.~\ref{fig:layout}.
-The four \HAWKI\ chips and two MOSFIRE pointings are clearly visible for the first four and last two fields, respectively.
-\label{fig:ivar_mosaics}}  
-\end{figure}
-%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %                                                                %
@@
 \label{s:mosaic}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%   Image quality figure
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}[!t]
-\epsscale{1.17}
-\plotone{curve_of_growth.pdf} % Figures/exposure_map.pdf
-\caption{Curves of growth of stellar profiles in the Frontier Fields $K_s$-band mosaics.
-The inverse aperture corrections for point sources in $D=0\farcs6$ apertures are indicated in each panel.
-The profiles of stars are well-fit by Moffat profiles with $\mathrm{FWHM}\sim0\farcs4$ in most cases and $\beta\sim2.1$.
-These extended Moffat profiles appear to be characteristic of deep $K_s$-band images and they have substantially more flux at large radii ($r>\mathrm{FWHM}/2$) than Gaussian profiles with the same FWHM (orange curves in the left panels).
-\textcolor{red}{\bf xxx In the bottom right panel, the inverse aperture correction of the parallel field is impossible to read, behind the other number...you may want to shift it upward a but}  \label{fig:cog}}
-\end{figure*}
-%%%%%%%%%%%%%%
 
@@
 We also compute the robust NMAD scatter \citep{brammer:08} of empty sky pixels within the same cells to empirically calibrate the inverse variance maps that are shown in Figure~\ref{fig:ivar_mosaics} (see also \S\ref{s:depth}).
-The final image and inverse variance mosaics are provided as an ESO Phase 3 data product \xxx\ URL, as well as attached as a digital product to this manuscript (\xxx\ I think ApJSS can and wants to do this).
+The final image and inverse variance mosaics are provided as an ESO Phase 3 data product \XXX{URL}, as well as attached as a digital product to this manuscript \XXX{(I think ApJSS can and wants to do this)}.
 The released images are all scaled to a common zeropoint of 26.0 (AB mag), which gives pixel values of order unity faint galaxies near the detection threshold.
 In the sections below we describe the characteristics of the mosaics in more detail.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%   Drizzle figure
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[!t]
+\epsscale{1.17}
+\plotone{ivar_mosaics.pdf} % Figures/exposure_map.pdf
+\caption{Inverse variance maps of the $K_s$-band mosaics.
+The orientation of the fields is the same as that shown in Fig.~\ref{fig:layout}.
+The four \HAWKI\ chips and two MOSFIRE pointings are clearly visible for the first four and last two fields, respectively.
+\label{fig:ivar_mosaics}}  
+\end{figure}
+%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%   Image quality figure
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure*}[!t]
+\epsscale{1.17}
+\plotone{curve_of_growth.pdf} % Figures/exposure_map.pdf
+\caption{Curves of growth of stellar profiles in the Frontier Fields $K_s$-band mosaics.
+The inverse aperture corrections for point sources in $D=0\farcs6$ apertures are indicated in each panel.
+The profiles of stars are well-fit by Moffat profiles with $\mathrm{FWHM}\sim0\farcs4$ in most cases and $\beta\sim2.1$.
+These extended Moffat profiles appear to be characteristic of deep $K_s$-band images and they have substantially more flux at large radii ($r>\mathrm{FWHM}/2$) than Gaussian profiles with the same FWHM (orange curves in the left panels).
+\label{fig:cog}}
+\end{figure*}
+%%%%%%%%%%%%%%
+
+\subsection{Image quality}
+\label{s:fwhm}
+
+Figure~\ref{fig:cog} shows the curves of growth for stars identified by the tight relationship between their brightness and half-light radii \citep[see, e.g., Fig.~13 of][]{skelton:14}.
+The image quality of the $K_s$-band mosaics is excellent with stellar FWHM$\lesssim0\farcs4$ for the deep \HAWKI\ fields, thanks in large part to the execution of the \HAWKI\ observations in service mode ensuring optimal and uniform image quality across the many hours of integration on the survey fields.
+The classical scheduling of the MOSFIRE observations does not allow such control over the seeing conditions.
+The result is that the image quality of the MACS-0717 and MACS-1149 fields is somewhat degraded in comparison, at $0\farcs42$--$0\farcs54$, and the image quality differs measurably between the two MOSFIRE pointings in the MACS-0717 field.
+Despite the compact cores of the stellar PSFs, the curves of growth shown in Figure~\ref{fig:cog} deviate from Gaussian profiles with significantly higher flux in the outer wings, with a shape more consistent with a Moffat profile \citep{trujillo:01}  with $\beta\sim2$.
+We caution that these extended profiles will yield significantly shallower final image depths than would be predicted with Exposure Time Calculators that may assume Gaussian profiles.
+Nevertheless, the deep $K_s$ mosaics have excellent image quality that is much closer to the resolution of the \textit{HST} IR imaging and the typical apparent size of distant galaxies \citep[median $r_e\sim0\farcs1$--$0\farcs2$ at $z>4$;][]{shibuya:15} than the redder \textit{Spitzer} IRAC bands (cf.
+FWHM$\sim$$1\farcs7$--$2^{\prime\prime}$).
+
+\subsection{Noise properties \& Depth}
+\label{s:depth}
+
+Along with practical benefits mentioned above, the most dramatic benefit of using the drizzle algorithm comes from the fact that the raw detector pixels are only resampled once in the process of generating the output mosaic.
+Furthermore, with a large number of dithered raw images we can use drizzle to shrink the raw input pixels by a factor of ten before dropping them into the output grid,\footnote{The pixels are shrunk by adopting pixfrac=0.1 as defined by \cite{fruchter:02}.} and the result is dramatically reduced correlations between neighboring pixels, particularly for output pixel sizes that are somewhat smaller than the native \HAWKI\ or MOSFIRE detector pixels.
+
+A comparison of final combined images created with output $0\farcs1$ and $0\farcs06$ pixel grids using the drizzle implementation to versions created with the SWarp software \citep{swarp} is shown in Figure~\ref{fig:drizzle_cutout}.
+In the later case, the raw pixels are resampled twice before making the final mosaic, once combining the raw exposures in each OB and a second time combining the separate OBs into the final stack with an arbitrary pixel grid.
+One could avoid the second resampling and combine all of the raw exposures at once, though SWarp is still limited to dropping the original pixels into the output mosaic preserving the input native pixel grid and therefore a single input pixel maps onto multiple (perhaps many) output pixels.
+
+The fact that the drizzle-combined images in Figure~\ref{fig:drizzle_cutout} (panels \textit{a} and \textit{b}) appear to the eye to be noisier than their SWarped counterparts (panels \textit{c} and \textit{d}) is simply the result of a smoothing of the pixel noise in the latter case rather than reflecting true underlying differences in their pixel-to-pixel variance.
+\cite{casertano:00} discuss how the noise properties of correlated output pixels are affected by the relative sizes of the input and output pixel grids and the adopted drizzle parameters.
+With the drizzle algorithm and a small pixfrac, a given input pixel will map to only a single output pixel and therefore correlations between the output pixels should be minimal\footnote{Inter-pixel capacitance and other forms of crosstalk intrinsic to the IR detectors place a lower limit on the correlations between adjacent pixels \citep{finger:05, hilbert:ipc}.}.
+
+We explore these correlations between the output pixels in Figure~\ref{fig:corr_noise}.
+For the black and red curves in each panel, we measure value differences between random pairs of pixels sorted by the separation between the pixels for the drizzled and SWarped images, respectively.
+For perfectly uncorrelated pixels, these curves will be flat and will reflect the intrinsic noise of the image pixels.
+In the presence of correlations between adjacent pixels, however, the curves will show a depression at small separations.
+As expected, the SWarp-combined images show just such a depression for pairs of pixels 1--2 pixels apart.
+The yellow curves in Figure~\ref{fig:corr_noise} show the same pairwise differences for the drizzled images now convolved by small Gaussian kernels as indicated, which agree very well with the curves for the images with correlated pixels.
 
@@
 
-\subsection{Image quality}
-\label{s:fwhm}
-
-Figure~\ref{fig:cog} shows the curves of growth for stars identified by the tight relationship between their brightness and half-light radii \citep[see, e.g., Fig.~13 of][]{skelton:14}.
-The image quality of the $K_s$-band mosaics is excellent with stellar FWHM$\lesssim0\farcs4$ for the deep \HAWKI\ fields, thanks in large part to the execution of the \HAWKI\ observations in service mode ensuring optimal and uniform image quality across the many hours of integration on the survey fields.
-The classical scheduling of the MOSFIRE observations does not allow such control over the seeing conditions.
-The result is that the image quality of the MACS-0717 and MACS-1149 fields is somewhat degraded in comparison, at $0\farcs42$--$0\farcs54$, and the image quality differs measurably between the two MOSFIRE pointings in the MACS-0717 field.
-Despite the compact cores of the stellar PSFs, the curves of growth shown in Figure~\ref{fig:cog} deviate from Gaussian profiles with significantly higher flux in the outer wings, with a shape more consistent with a Moffat profile \citep{trujillo:01}  with $\beta\sim2$.
-We caution that these extended profiles will yield significantly shallower final image depths than would be predicted with Exposure Time Calculators that may assume Gaussian profiles.
-Nevertheless, the deep $K_s$ mosaics have excellent image quality that is much closer to the resolution of the \textit{HST} IR imaging and the typical apparent size of distant galaxies \citep[median $r_e\sim0\farcs1$--$0\farcs2$ at $z>4$;][]{shibuya:15} than the redder \textit{Spitzer} IRAC bands (cf.
-FWHM$\sim$$1\farcs7$--$2^{\prime\prime}$).
-
-\subsection{Noise properties \& Depth}
-\label{s:depth}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%   Depth figure
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[!t]
-\epsscale{1.17}
-\plotone{aper_depth.pdf} % Figures/exposure_map.pdf
-\caption{\textit{Top panel:} The product of the inverse curves-of-growth (Fig.~\ref{fig:cog}) and the predicted variance within a photometric aperture (Eq.~\ref{eq:aper_var}) gives the point-source sensitivity as a function of aperture size.
-The S/N is maximized with an aperture somewhat larger than the FWHM; depths evaluated at $D=0\farcs6$ near the maxima are indicated.)
-\label{fig:depth}}  
-\end{figure}
-%%%%%%
-
-Along with practical benefits mentioned above, the most dramatic benefit of using the drizzle algorithm comes from the fact that the raw detector pixels are only resampled once in the process of generating the output mosaic.
-Furthermore, with a large number of dithered raw images we can use drizzle to shrink the raw input pixels by a factor of ten before dropping them into the output grid,\footnote{The pixels are shrunk by adopting pixfrac=0.1 as defined by \cite{fruchter:02}.} and the result is dramatically reduced correlations between neighboring pixels, particularly for output pixel sizes that are somewhat smaller than the native \HAWKI\ or MOSFIRE detector pixels.
-
-A comparison of final combined images created with output $0\farcs1$ and $0\farcs06$ pixel grids using the drizzle implementation to versions created with the SWarp software \citep{swarp} is shown in Figure~\ref{fig:drizzle_cutout}.
-In the later case, the raw pixels are resampled twice before making the final mosaic, once combining the raw exposures in each OB and a second time combining the separate OBs into the final stack with an arbitrary pixel grid.
-One could avoid the second resampling and combine all of the raw exposures at once, though SWarp is still limited to dropping the original pixels into the output mosaic preserving the input native pixel grid and therefore a single input pixel maps onto multiple (perhaps many) output pixels.
-
-The fact that the drizzle-combined images in Figure~\ref{fig:drizzle_cutout} (panels \textit{a} and \textit{b}) appear to the eye to be noisier than their SWarped counterparts (panels \textit{c} and \textit{d}) is simply the result of a smoothing of the pixel noise in the latter case rather than reflecting true underlying differences in their pixel-to-pixel variance.
-\cite{casertano:00} discuss how the noise properties of correlated output pixels are affected by the relative sizes of the input and output pixel grids and the adopted drizzle parameters.
-With the drizzle algorithm and a small pixfrac, a given input pixel will map to only a single output pixel and therefore correlations between the output pixels should be minimal\footnote{Inter-pixel capacitance and other forms of crosstalk intrinsic to the IR detectors place a lower limit on the correlations between adjacent pixels \citep{finger:05, hilbert:ipc}.}.
-
-We explore these correlations between the output pixels in Figure~\ref{fig:corr_noise}.
-For the black and red curves in each panel, we measure value differences between random pairs of pixels sorted by the separation between the pixels for the drizzled and SWarped images, respectively.
-For perfectly uncorrelated pixels, these curves will be flat and will reflect the intrinsic noise of the image pixels.
-In the presence of correlations between adjacent pixels, however, the curves will show a depression at small separations.
-As expected, the SWarp-combined images show just such a depression for pairs of pixels 1--2 pixels apart.
-The yellow curves in Figure~\ref{fig:corr_noise} show the same pairwise differences for the drizzled images now convolved by small Gaussian kernels as indicated, which agree very well with the curves for the images with correlated pixels.
-
 In terms of understanding the noise properties of the images, a key point here is not only that depression at small separations indicates the presence of correlations between adjacent pixels but also the fact that correlations cause the apparent overall r.m.s.
@@
 The comparison must be made because any additional half magnitude increase depth is exponentially more expensive to obtain!
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%   Depth figure
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[!t]
+\epsscale{1.17}
+\plotone{aper_depth.pdf} % Figures/exposure_map.pdf
+\caption{\textit{Top panel:} The product of the inverse curves-of-growth (Fig.~\ref{fig:cog}) and the predicted variance within a photometric aperture (Eq.~\ref{eq:aper_var}) gives the point-source sensitivity as a function of aperture size.
+The S/N is maximized with an aperture somewhat larger than the FWHM; depths evaluated at $D=0\farcs6$ near the maxima are indicated.)
+\label{fig:depth}}  
+\end{figure}
+%%%%%%
 
@@
 \plotone{image_cutouts.pdf} % Figures/exposure_map.pdf
-\caption{Image cutouts of the MACS 0416 parallel field.
-The \textit{RGB} and monochrome panels come from the sources as indicated; the cutouts are all $10\arcsec$ \xxx\ \textcolor{red}{check, different now} on a side.
+\caption{Image cutouts of the MACS 0416 parallel field.  The \textit{RGB} and monochrome panels come from the sources as indicated; the cutouts are all $20\arcsec$ on a side.
 \label{fig:cutout}}  
@@
 That is, ``Chip \#1'' is the first image extension of the raw FITS files provided by the ESO archive, with \textsc{extname}$=$CHIP1.INT1.
-The \textsc{extname} values for chips 2,3,4 are CHIP2.INT1, CHIP4.INT1, CHIP3.INT1, note the switched names of the last two chips.
+The \textsc{extname} values for chips 2, 3, and 4 are CHIP2.INT1, CHIP4.INT1, and CHIP3.INT1, respectively; note the switched names of the last two chips.
+
+\section{Description of provided data products}
+
+\clearpage
 
@@
 
-\section{Description of provided data products}
-
 \end{document}

### 2016-04-17 Paper0/ms.tex
modified completeness, cleanup

@@
 The shallower northern MACS-1149 and MACS-0717 fields are 50\% complete at $K_s=24.9$ and 25.3, respectively.
-Only MACS-1149 differs appreciably in the area around the \textit{HST} cluster area, with the curve as shown;  the completeness curves in the cluster areas of the other fields differ from the curves shown by $\lesssim0.1$ mag and are omitted for clarity.
-\textcolor{red}{\bf xxx I would probably plot also the macs0717 cluster completeness curve, given it is actually different data...
-}
+Separate curves are shown for the cluster and parallel areas in the MACS-0717 and MACS-1149 fields, which required two separate pointings with the MOSFIRE instrument;  the completeness curves in the cluster areas of the four \HAWKI\ fields differ from the curves shown by $\lesssim0.1$ mag and are omitted for clarity.
 \label{fig:completeness}}  
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\vspace{-0.3cm} %% emulateapj
+%\vspace{-0.3cm} %% emulateapj
+
 \acknowledgements
@@
 We are most fortunate to have the opportunity to conduct observations from this mountain.
-This research made use of Astropy, a community-developed core Python package for Astronomy \citep{astropy}, and open-source Python modules Numpy, Scipy, and Matplotlib.
+This research made use of Astropy, a community-developed core Python package for Astronomy \citep{astropy}, open-source Python modules Numpy, Scipy, Matplotlib, and scikit-image, and NASA's Astrophysics Data System (ADS).
+\\
 
@@
 
-{\it Facilities:} \facility{VLT:Yepun (HAWKI)}, \facility{Keck:I (MOSFIRE)} % \facility{Hubble Space Telescope (WFC3)}
+\noindent {\normalsize \it Facilities:}
+\facility{VLT:Yepun (HAWKI)}
+\facility{Keck:I (MOSFIRE)} % \facility{Hubble Space Telescope (WFC3)}
 

### 2016-04-21 Paper0/ms.tex
v1 draft
Add a few affiliations

@@
 
-%\gdef\zphot{$z_\mathrm{phot}$}
 \gdef\HST{\textit{HST}}
-\gdef\G141{\textit{G141}}
-\gdef\F140W{\textit{F140W}}
-\gdef\fluxcgs{\mathrm{erg\ s^{-1}\ cm^{-2}}}
+\gdef\fluxcgs{\mathrm{erg~s^{-1}~cm^{-2}}}
 \gdef\micront{$\mu$m}
@@
 \gdef\logOH{12 + \log \left(\mathrm{O/H}\right)}
-\gdef\UmV{U-V}
-\gdef\VmJ{V-J}
 \gdef\SFRuvir{SFR_\mathrm{UV+IR}}
@@
 \gdef\arcsec{^{\prime\prime}}
-\gdef\Jy{\mathrm{Jy}}
-\gdef\uJy{\mu\mathrm{Jy}}
-\gdef\LIR{L_\mathrm{IR}}
-\gdef\LUV{L_\mathrm{UV}}
-\gdef\4ang{4000\,\AA}
-\gdef\msun{$M_{\odot}$}
-\gdef\mathmsun{\ M_{\odot}}
-\gdef\mathgyr{\ \textrm{Gyr}}
-\gdef\mathmyr{\ \textrm{Myr}}
-\gdef\pegase{\textsc{P\'{e}gase}}
-\gdef\eazy{\texttt{Eazy}}
-\gdef\galfit{\texttt{galfit}}
-\gdef\emcee{\texttt{emcee}}
-\gdef\interrest{\texttt{InterRest}}
-\gdef\fast{\texttt{fast}}
-\gdef\zCOSMOS{$z$\textsc{COSMOS}}
-\gdef\zspec{z_\mathrm{spec}}
-\gdef\zphot{z_\mathrm{phot}}
-\gdef\zml{z_\mathrm{ML}}
-\gdef\zprior{z_\mathrm{prior}}
-\newcommand{\weblink}[1]{\href{#1}{#1}}
 \gdef\UDFj{UDFj-39546284}
@@
 \gdef\Hbeta{\mathrm{H}\beta}
-\gdef\othree{[\ion{O}{3}]$\lambda$4959+5007}
-\gdef\epers{electrons~s$^{-1}$}
-%\gdef\xxx{\textbf{xxx}}
+
 \newcommand\xxx{{\textcolor{red}{\bf xxx}}}
@@
 
-\gdef\lineflux{$3.5\pm1.3\times10^{-18}\,\fluxcgs$}
-
 \shortauthors{Brammer et al.}
 \shorttitle{$K_s$-band imaging of the Frontier Fields}
-%\slugcomment{Draft (\today)}
-\slugcomment{Submitted to ApJSS}
+\slugcomment{Draft (\today)}
+%\slugcomment{Submitted to ApJSS}
+
+% \watermark{Draft v1.0 (\today)}
+% \setwatermarkfontsize{50pt}
 
@@
 Lee Spitler\altaffilmark{4}, 
-Daniel Lange-Vagle\altaffilmark{2},
-Elizbeth Barker\altaffilmark{1},
+Daniel Lange-Vagle\altaffilmark{1,2},
+Elizbeth A. Barker\altaffilmark{1},
 Adriano Fontana\altaffilmark{5},
@@
 }
-% Pieter G.\ van Dokkum\altaffilmark{2},
-% Garth D.\ Illingworth\altaffilmark{4},
-% Rychard J.\ Bouwens\altaffilmark{3},
-% Ivo Labb\'e\altaffilmark{3}, 
-% Marijn Franx\altaffilmark{3},
-% Ivelina Momcheva\altaffilmark{2},
-% Pascal A.\ Oesch\altaffilmark{4}}
-% Rub\'en S\'anchez-Janssen\altaffilmark{1}, 
-% Ivo Labb\'e\altaffilmark{2},
-% Elisabete da Cunha\altaffilmark{3},
-% Dawn K.\ Erb\altaffilmark{4},
-% Marijn Franx\altaffilmark{2},
-% Mattia Fumagalli\altaffilmark{2},
-% Britt Lundgren\altaffilmark{5},
-% Ivelina Momcheva\altaffilmark{5},
-% Erica Nelson\altaffilmark{5},
-% Shannon Patel\altaffilmark{2},
-% Ryan Quadri\altaffilmark{7},
-% Hans-Walter Rix\altaffilmark{3},
-% Rosalind E.\ Skelton\altaffilmark{5},
-% Kasper B.\ Schmidt\altaffilmark{3},
-% Arjen van der Wel\altaffilmark{3},
-% Pieter G.\ van Dokkum\altaffilmark{5},
-% David A.\ Wake\altaffilmark{5},
-% Katherine E.\ Whitaker\altaffilmark{5}}
-
-% Rachel Bezanson\altaffilmark{5},
-% Xiaohui Fan\altaffilmark{6},
-% Natascha F\"orster Schreiber\altaffilmark{8},
-% Garth D.\ Illingworth\altaffilmark{9},
-% Joel Leja\altaffilmark{5},
-% Dan Magee\altaffilmark{9},
-% Patrick McCarthy\altaffilmark{11},
-% Adam Muzzin\altaffilmark{2},
-% Ryan Quadri\altaffilmark{11},
-% Charles C.\ Steidel\altaffilmark{12},
-% Tomer Tal\altaffilmark{5},
-% Anna Williams\altaffilmark{13}
 
@@
 
-\altaffiltext{1}{ESA/AURA, Space Telescope Science Institute, 3700 San Martin Dr., Baltimore, MD 21218, USA}
+\altaffiltext{1}{Space Telescope Science Institute, 3700 San Martin Dr., Baltimore, MD 21218, USA}
 \altaffiltext{2}{Physics and Astronomy Department, Tufts University, Medford, MA, 02155, USA}
@@
 \altaffiltext{6}{\xxx}
-\altaffiltext{7}{\xxx}
-\altaffiltext{8}{\xxx}
+\altaffiltext{7}{Subaru Telescope, 650 N. Aohoku Pl, 96720 Hilo, HI, USA}
+\altaffiltext{8}{National Astronomical Observatory of Japan
+2-21-1 Osawa, Mitaka, Tokyo 181-8588, Japan}
 \altaffiltext{9}{\xxx}
@@
 % \altaffiltext{4}
-% {UCO/Lick Observatory, University of California, Santa Cruz, CA 95064}
-% \altaffiltext{4}
 % {Department of Physics, University of Wisconsin-Milwaukee, P.O. Box 413,
 % Milwaukee, WI 53201, USA}
-% \altaffiltext{5}
-% {Department of Astronomy, Yale University, New Haven, CT 06520, USA}
-% \altaffiltext{7}
-% {Carnegie Observatories, 813 Santa Barbara Street, Pasadena, CA 91101, USA}
 

### 2016-04-21 Paper0/ms.tex
Incorporate B.V.'s comments.

@@
 \altaffiltext{13}{Max Planck Institute for Astronomy (MPIA), K\"onigstuhl 17, 69117, Heidelberg, Germany}
-\altaffiltext{14}{\xxx}
+\altaffiltext{14}{School of Physics, Tin Alley, University of Melbourne VIC 3010, Australia}
 \altaffiltext{15}{\xxx}
@@
 The HFF program is a multi-cycle program consisting of $\sim$560 orbits of Director's Discretionary (DD) time that is imaging six deep fields centered on strong lensing galaxy clusters in parallel with six deep blank fields.
-The primary science goals of the twelve HFF fields are to 1) reveal the population of galaxies at $z=5-10$ that are 10--50 times fainter intrinsically than any presently known, 2) solidify our understanding of the stellar masses and star formation histories of faint galaxies, 3) provide the first statistically meaningful morphological characterization of star-forming galaxies at $z>5$, and 4) to find $z>8$ galaxies magnified by the cluster lensing, with some bright enough to make them accessible to spectroscopic follow-up.
+The primary science goals of the twelve HFF fields are to 1) reveal the population of galaxies at $z=5-10$ that are 10--50 times fainter intrinsically than any presently known, 2) solidify our understanding of the stellar masses and star formation histories of faint galaxies, 3) provide the first statistically meaningful morphological characterization of star-forming galaxies at $z>5$, and 4) find $z>8$ galaxies magnified by the cluster lensing, with some bright enough to make them accessible to spectroscopic follow-up.
 Along with {\it HST}, the {\it Spitzer Space Telescope} has devoted 1000 hours of DD time to image the HFF fields at 3.6$\mu$m and 4.5$\mu$m with IRAC (Capak et al., in prep).
 
-Whereas the main goal of the HFF is to explore the galaxy population in the first billion year of cosmic history, this dataset is also unique for its combination of surveyed area and depth for studies of galaxy evolution across most of the age of the universe, down to, and including, the redshifts of the targeted galaxy clusters ($z\approx$~0.3--0.5).
+The Frontier Fields initiative has inspired a number of separate supporting general observer programs, such as deep \textit{HST} ultraviolet imaging (Siana et al., in prep) and grism spectroscopy \citep[GLASS, ][]{treu:glass}, and deep far-infrared imaging with \textit{Herschel} \citep{rawle:16}.
+Whereas the main goal of the HFF is to explore the galaxy population in the first billion year of cosmic history, this dataset is also unique for its combination of surveyed area, multiwavelength coverage and depth for studies of galaxy evolution across most of the age of the universe, down to, and including, the redshifts of the targeted galaxy clusters ($z\approx$~0.3--0.5).
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Observations \& Data Reduction}
-\label{s:observations}
-
-Table~\ref{tab:summary} provides a summary of the survey fields and the characteristics of the $K_s$-band observational program.
-Additional details on the observing strategy and image reduction procedures are provided in the subsections below.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{deluxetable*}{ccccrccccc}
-%\tabletypesize{\footnotesize}
+\begin{deluxetable*}{clrrclcrrr}
+\tabletypesize{\small}
 \tablewidth{0pt} 
@@
 \tablehead{ 
-\colhead{Field} & \colhead{R.A.} & \colhead{Dec.} & \colhead{Instrument} & \colhead{Epoch} & \textit{HST}\tablenotemark{a} & \colhead{$t$, hours} & {Depth\tablenotemark{b}} & \colhead{FWHM}}
+\colhead{Field} & \colhead{$z$\tablenotemark{a}} & \colhead{R.A.} & \colhead{Dec.} & \colhead{Instrument} & \colhead{Epoch} & \textit{HST}\tablenotemark{b} & \colhead{$t$, hours} & {Depth\tablenotemark{c}} & \colhead{FWHM}}
 \startdata 
-Abell 2744   & $00:14:21.2$ & $-30:23:50$ & \textit{VLT}/\HAWKI     & 2013 Oct 24--2013 Dec 24 & C+P & 29.3 &  26.0 & $0\farcs39$ \\ 
-MACS-0416    & $04:16:08.9$ & $-24:04:28$ & \textit{VLT}/\HAWKI     & 2013 Oct 25--2014 Feb 23 & C+P & 25.8 &  26.0 & $0\farcs36$ \\ 
-Abell S1063  & $22:49:01.1$ & $-44:32:13$ & \textit{VLT}/\HAWKI     & 2015 Jul 8--Sep 22       & C+P & 27.9 &  26.0 & $0\farcs39$ \\ 
-Abell 370    & $02:40:03.3$ & $-01:36:23$ & \textit{VLT}/\HAWKI     & 2015 Jul 26--2016 Jan 28 & C+P & 28.3 &  26.0 & $0\farcs35$ \\ 
+Abell 2744   & 0.31 & 00:14:21.2 & $-$30:23:50 & \textit{VLT}/\HAWKI     & 2013 Oct 24--2013 Dec 24 &  C+P & 29.3 &  26.0 & $0\farcs39$ \\ 
+MACS-0416    & 0.40 & 04:16:08.9 & $-$24:04:28 & \textit{VLT}/\HAWKI     & 2013 Oct 25--2014 Feb 23 &  C+P & 25.8 &  26.0 & $0\farcs36$ \\ 
+Abell S1063  & 0.35 & 22:49:01.1 & $-$44:32:13 & \textit{VLT}/\HAWKI     & 2015 Jul 8--Sep 22  &  C+P & 27.9 &  26.0 & $0\farcs39$ \\ 
+Abell 370    & 0.38 & 02:40:03.3 & $-$01:36:23 & \textit{VLT}/\HAWKI     & 2015 Jul 26--2016 Jan 28 & C+P & 28.3 &  26.0 & $0\farcs35$ \\ 
 \vspace {-0.2cm}\\ 
-\multirow{2}{*}{MACS-0717}    & \multirow{2}{*}{$07:17:34.0$} & \multirow{2}{*}{$+37:44:49$} &  \multirow{2}{*}{\textit{Keck}/MOSFIRE}  &  \multirow{2}{*}{2015 Jan 26--27, 2016 Jan 21--22}  & C & 4.3 & 25.3 & $0\farcs42$ \\
-                              &                             &                             &                                          &      & P & 3.8 & 25.5 & $0\farcs49$ \\ 
+\multirow{2}{*}{MACS-0717}    & \multirow{2}{*}{0.55} & \multirow{2}{*}{07:17:34.0} & \multirow{2}{*}{$+$37:44:49} &  \multirow{2}{*}{\textit{Keck}/MOSFIRE}  &  \multirow{2}{*}{2015 Jan 26/27, 2016 Jan 21/22}  & C & 4.3 & 25.3 & $0\farcs42$ \\
+        &                      &                             &                             &                                          &      & P & 3.8 & 25.5 & $0\farcs49$ \\ 
 \vspace {-0.2cm}\\ 
-\multirow{4}{*}{MACS-1149}    & \multirow{4}{*}{$11:49:36.3$} & \multirow{4}{*}{$+22:23:58$} & \multirow{2}{*}{\textit{Keck}/MOSFIRE}  & \multirow{2}{*}{2015 Feb 24, 2016 Jan 21--22}      & C & 5.5 & 25.2 & $0\farcs53$ \\
-&  &         &      &   & P & 4.8 & 25.1 & $0\farcs54$ \\
+\multirow{4}{*}{MACS-1149}    & \multirow{4}{*}{0.54} & \multirow{4}{*}{11:49:36.3} & \multirow{4}{*}{$+$22:23:58} & \multirow{2}{*}{\textit{Keck}/MOSFIRE}  & \multirow{2}{*}{2015 Feb 24, 2016 Jan 21/22}      & C & 5.5 & 25.2 & $0\farcs53$ \\
+&  & &         &      &   & P & 4.8 & 25.1 & $0\farcs54$ \\
 \vspace {-0.4cm}\\ 
-&  &         & \textit{VLT}/\HAWKI\tablenotemark{c}     & 2013 Mar 21--2014 Jun 9      & C & 5.3 & 25.0 & $0\farcs41$
+&  & &         & \textit{VLT}/\HAWKI\tablenotemark{d}     & 2013 Mar 21--2014 Jun 9      & C & 5.3 & 25.0 & $0\farcs41$ \\
 \enddata 
 %\vspace{-0.8cm} 
-\tablenotetext{a}{Coverage of \textit{Hubble} survey fields: C=Cluster, P=Parallel.
+\tablenotetext{a}{Cluster redshift.}
+\tablenotetext{b}{Coverage of \textit{Hubble} survey fields: C=Cluster, P=Parallel.
 Most \HAWKI\ pointings cover the two \textit{HST} fields simultaneously (C+P; Fig.~\ref{fig:layout}).
@@
 %\footnotetext[2]{Approximate exposure time equivalent at $0\farcs4$ seeing:  $t_{0.4}=\sum_i t_i \cdot (0.4/\mathrm{FWHM}_i)^2$.}
-\tablenotetext{b}{Depth is defined as the 5$\sigma$ limiting magnitude for point sources, measured in $D=0\farcs6$ apertures (\S\ref{s:depth}).} 
-\tablenotetext{c}{Archival observations from ESO program \href{http://archive.eso.org/wdb/wdb/eso/eso_archive_main/query?prog_id=090.A-0458(A)}{090.A-0458} (PI: Infante).}
+\tablenotetext{c}{Depth is defined as the 5$\sigma$ limiting magnitude for point sources, measured in $D=0\farcs6$ apertures (\S\ref{s:depth}).} 
+\tablenotetext{d}{Archival observations from ESO program \href{http://archive.eso.org/wdb/wdb/eso/eso_archive_main/query?prog_id=090.A-0458(A)}{090.A-0458} (PI: Infante).}
 \end{deluxetable*}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Observations \& Data Reduction}
+\label{s:observations}
+
+Table~\ref{tab:summary} provides a summary of the survey fields and the characteristics of the $K_s$-band observational program.
+Additional details on the observing strategy and image reduction procedures are provided in the subsections below.
 
@@
 \caption{Curves of growth of stellar profiles in the Frontier Fields $K_s$-band mosaics.
-The inverse aperture corrections for point sources in $D=0\farcs6$ apertures are indicated in each panel.
+The \HAWKI\ fields include point sources across the full mosaics that cover both \textit{HST} cluster and parallel fields, while separate curves are given for the cluster and parallel areas of the northern MOSFIRE fields, which required two independent pointings of that instrument (Fig.~\ref{fig:layout}).
+The inverse aperture corrections for point sources in $D=0\farcs6$ apertures are indicated in each panel with the small labels.
 The profiles of stars are well-fit by Moffat profiles with $\mathrm{FWHM}\sim0\farcs4$ in most cases and $\beta\sim2.1$.
-These extended Moffat profiles appear to be characteristic of deep $K_s$-band images and they have substantially more flux at large radii ($r>\mathrm{FWHM}/2$) than Gaussian profiles with the same FWHM (orange curves in the left panels).
+These extended Moffat profiles appear to be characteristic of deep $K_s$-band images and they have substantially more flux at large radii ($r>\mathrm{FWHM}/2$) than Gaussian profiles with the same FWHM, as shown in by the orange curve in the upper-left panel).
 \label{fig:cog}}
@@
 28.7 AB in $H_{160}$), the $K_s$ band images can be more effectively exploited for most applications by detecting objects in the deeper \textit{HST} images and performing forced $K_s$-band photometry at the positions of the \textit{HST} sources.
-The generation of such catalogs is beyond the scope of the present work and will be presented in Shipley et al.
+\cite{merlin:16} present multiwavelength catalogs of the Abell~2744 and MACS-0416 fields constructed in this way, including incorporating an early version of the \HAWKI\ $K_s$-band mosaics.
+The generation of these catalogs is beyond the scope of the present work and will be presented in further detail by Shipley et al.
 (2016; in preparation).
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[!ht]
+\begin{figure}[!t]
 \epsscale{1.17}
@@
 acknowledge support from program HST-AR-14302, provided by NASA through a grant from the Space Telescope Science Institute, which is operated by the Association of Universities for Research in Astronomy, Incorporated, under NASA contract NAS5-26555.
+B.V.
+acknowledges the support from an Australian Re- search Council Discovery Early Career Researcher Award (PD0028506).
 This work was supported by NASA Keck PI Data Awards, administered by the NASA Exoplanet Science Institute.

### 2016-04-29 Paper0/ms.tex
submit

@@
 Ivo Labb\'e\altaffilmark{3}, 
-Lee Spitler\altaffilmark{4}, 
+Lee Spitler\altaffilmark{4,5}, 
 Daniel Lange-Vagle\altaffilmark{1,2},
 Elizbeth A. Barker\altaffilmark{1},
-Adriano Fontana\altaffilmark{5},
-Audrey Galametz\altaffilmark{6},
-Anna Ferr\'e-Mateu\altaffilmark{7},
-Tadayuki Kodama\altaffilmark{8},
-Britt Lundgren\altaffilmark{9},
+Adriano Fontana\altaffilmark{6},
+Audrey Galametz\altaffilmark{7},
+Anna Ferr\'e-Mateu\altaffilmark{8},
+Tadayuki Kodama\altaffilmark{9},
+Britt Lundgren\altaffilmark{10},
 Nicholas Martis\altaffilmark{2},
-Adam Muzzin\altaffilmark{10},
+Adam Muzzin\altaffilmark{11},
 Mauro Stefanon\altaffilmark{3},
-Masayuki Tanaka\altaffilmark{11},
+Masayuki Tanaka\altaffilmark{8},
 Sune Toft\altaffilmark{12},
@@
 Benedetta Vulcani\altaffilmark{14}, 
-Katherine E. Whitaker\altaffilmark{15}
+Katherine E. Whitaker\altaffilmark{15,16}
 }
@@
 \altaffiltext{1}{Space Telescope Science Institute, 3700 San Martin Dr., Baltimore, MD 21218, USA}
-\altaffiltext{2}{Physics and Astronomy Department, Tufts University, Medford, MA, 02155, USA}
+\altaffiltext{2}{Physics and Astronomy Department, Tufts University, Medford, MA 02155, USA}
 \altaffiltext{3}{Leiden Observatory, Leiden University, NL-2300 RA Leiden, Netherlands}
-\altaffiltext{4}{\xxx}
-\altaffiltext{5}{\xxx}
-\altaffiltext{6}{\xxx}
-\altaffiltext{7}{Subaru Telescope, 650 N. Aohoku Pl, 96720 Hilo, HI, USA}
-\altaffiltext{8}{National Astronomical Observatory of Japan
+\altaffiltext{4}{Department of Physics \& Astronomy, Macquarie University, Sydney, NSW 2109, Australia}
+\altaffiltext{5}{Australian Astronomical Observatories, PO Box 915 North Ryde NSW 1670, Australia}
+\altaffiltext{6}{INAF, Osservatorio Astronomico di Roma, Via Frascati 33, I-00040, Monteporzio, Italy}
+\altaffiltext{7}{Max-Planck-Institut fur extraterrestrische Physik (MPE), Giessenbachstr. 1, D-85748 Garching, Germany}
+\altaffiltext{8}{Subaru Telescope, 650 N. Aohoku Pl, 96720 Hilo, HI, USA}
+\altaffiltext{9}{National Astronomical Observatory of Japan
 2-21-1 Osawa, Mitaka, Tokyo 181-8588, Japan}
-\altaffiltext{9}{\xxx}
-\altaffiltext{10}{\xxx}
-\altaffiltext{11}{\xxx}
-\altaffiltext{12}{\xxx}
+\altaffiltext{10}{Department of Astronomy, University of Wisconsin, Madison, WI 53706, USA}
+\altaffiltext{11}{Kavli Institute for Cosmology, Cambridge University, Madingley Rd, Cambridge, UK, CB3 0HA}
+\altaffiltext{12}{Dark Cosmology Centre, Niels Bohr Institute, University of Copenhagen, Juliane Maries Vej 30, DK-2100 Copenhagen, Denmark}
 \altaffiltext{13}{Max Planck Institute for Astronomy (MPIA), K\"onigstuhl 17, 69117, Heidelberg, Germany}
 \altaffiltext{14}{School of Physics, Tin Alley, University of Melbourne VIC 3010, Australia}
-\altaffiltext{15}{\xxx}
+\altaffiltext{15}{Department of Astronomy, University of Massachusetts Amherst, Amherst, MA 01003, USA}
+\altaffiltext{16}{Hubble Fellow}
+
 % \altaffiltext{6}
@@
 
-We present ultra-deep $K_s$-band imaging of all six of the \textit{Hubble} Frontier Fields clusters Abell 2744, MACS-0416, Abell S1063, Abell 370, MACS-0717 and MACS-1149, which have recently been observed with large allocations of Directors' Discretionary Time with the \textit{HST} and \textit{Spitzer} telescopes covering $0.4 < \lambda < 1.6$\,\mum\ and 3.6--4.5\,\mum, respectively.
-\textit{VLT}/\HAWKI\ integrations of the first four fields reach 5$\sigma$ limiting depths of $K_s\sim26.0$ (AB, point sources) and have excellent seeing FWHM$\sim0\farcs4$.
-Shorter MOSFIRE integrations of the MACS-0717 (MACS-1149) field better observable in the north reach limiting depths $K_s$=25.5 (25.1) with seeing FWHM$\sim$$0\farcs4$ ($0\farcs5$).
-In all cases the $K_s$-band mosaics cover the primary cluster and parallel \textit{HST}/ACS+WFC3 fields, and the total area of the $K_s$-band coverage is \xxx\ arcmin$^2$.
-The $K_s$-band at 2.2\,\mum\ crucially fills the gap between the reddest HST filter (1.6\,\mum\,$\sim H$) and the IRAC 3.6\,\mum\ passband.
-While reaching the full depths of the space-based imaging is not currently feasible from the ground, the deep $K_s$-band images provide significant constraints on both the redshifts and the stellar population properties of galaxies extending well below the characteristic stellar mass across most of the age of the universe, down to, and including, the redshifts of the targeted galaxy clusters ($z \lesssim 0.5$).
+We present an overview of the ``KIFF'' project, which provides ultra-deep $K_s$-band imaging of all six of the \textit{Hubble} Frontier Fields clusters Abell 2744, MACS-0416, Abell S1063, Abell 370, MACS-0717 and MACS-1149.
+All of these fields have recently been observed with large allocations of Directors' Discretionary Time with the \textit{HST} and \textit{Spitzer} telescopes covering $0.4 < \lambda < 1.6$\,\mum\ and 3.6--4.5\,\mum, respectively.
+\textit{VLT}/\HAWKI\ integrations of the first four fields reach 5$\sigma$ limiting depths of $K_s\sim26.0$ (AB, point sources) and have excellent image quality (FWHM$\sim0\farcs4$).
+Shorter \textit{Keck}/MOSFIRE integrations of the MACS-0717 (MACS-1149) field better observable in the north reach limiting depths $K_s$=25.5 (25.1) with seeing FWHM$\sim$$0\farcs4$ ($0\farcs5$).
+In all cases the $K_s$-band mosaics cover the primary cluster and parallel \textit{HST}/ACS+WFC3 fields.
+The total area of the $K_s$-band coverage is 490~arcmin$^2$.
+The $K_s$-band at 2.2\,\mum\ crucially fills the gap between the reddest HST filter (1.6\,\mum\,$\sim H$~band) and the IRAC 3.6\,\mum\ passband.
+While reaching the full depths of the space-based imaging is not currently feasible from the ground, the deep $K_s$-band images provide important constraints on both the redshifts and the stellar population properties of galaxies extending well below the characteristic stellar mass across most of the age of the universe, down to, and including, the redshifts of the targeted galaxy clusters ($z \lesssim 0.5$).
 Reduced, aligned mosaics of all six survey fields are provided accompanying this manuscript.
@@
 
-\keywords{galaxies: evolution --- galaxies: high-redshift --- surveys: xxx}
+\keywords{galaxies: evolution --- galaxies: high-redshift --- surveys}
 
@@
 Imaging in the $K$-band allows for the direct sampling of rest-frame wavelengths longer than the Balmer break out to $z\approx5$.
-Sampling the rest-frame optical wavelength regime is crucial for high-$z$ studies, as it is significantly less affected by dust obscuration and a better probe of the galaxy stellar mass compared to the rest-frame UV, which is more sensitive to unobscured star formation \citep[e.g., ][]{fontana:06, marchesini:09}.
+Sampling the rest-frame optical wavelength regime is critical for high-$z$ studies, as it is significantly less affected by dust obscuration and a better probe of the galaxy stellar mass compared to the rest-frame UV, which is more sensitive to unobscured star formation \citep[e.g., ][]{fontana:06, marchesini:09}.
 
-While imaging at even longer wavelengths ($\lambda>3$~$\mu$m) is needed to probe the rest-frame optical of galaxies in the first billion years of cosmic history, deep sub-arsec resolution imaging at these wavelengths will not be available until the advent of the {\it James Webb Space Telescope} ({\it JWST}).
+While imaging at even longer wavelengths ($\lambda>3$~$\mu$m) is needed to probe the rest-frame optical emission of galaxies in the first billion years of cosmic history, deep sub-arsec resolution imaging at these wavelengths will not be available until the advent of the {\it James Webb Space Telescope} ({\it JWST}).
 Deep $K$-band data with good  (i.e., FWHM~$<1\arcsec$) image quality have been fundamental to fully exploit {\it Spitzer}-IRAC imaging data characterized by much poorer spatial resolution, allowing for the mitigation of the resulting blending of sources in IRAC images \citep[e.g.,][]{labbe:05}.
-Finally, for space-based studies, the $K$ band crucially fills the gap between the reddest {\it Hubble Space Telescope} ({\it HST}) filter (i.e., F160W, with $\lambda \sim 1.6$~$\mu$m) and the IRAC 3.6~$\mu$m passband, greatly improving the constraints on both the photometric redshifts and the inferred stellar-population properties of galaxies in the survey areas.
+Finally, for space-based studies, the $K$ band fills the gap between the reddest {\it Hubble Space Telescope} ({\it HST}) filter (i.e., F160W, with $\lambda \sim 1.6$~$\mu$m) and the IRAC 3.6~$\mu$m passband, greatly improving the constraints on both the photometric redshifts and the inferred stellar-population properties of galaxies in the survey areas.
 
-It is not surprising that, given the aforementioned reasons, all successful extragalactic surveys performed in the last fifteen years to investigate galaxies populations in the early universe invested significant resources to obtain deep NIR imaging data with good image quality, especially in the $K$ band, usually adopted as the detection band.
+It is not surprising that, given the aforementioned reasons, all successful extragalactic surveys performed in the last fifteen years to investigate galaxies populations in the early universe invested significant resources to obtain deep NIR imaging data with good image quality, especially in the $K$ band.
 Indeed, our understanding of galaxy formation and evolution in the last couple of decades has impressively proceeded forward mostly driven by progressively deeper, wider, and with better image quality, NIR surveys, such as FIRES \citep{labbe:03}, Subaru Super Deep Field (AO) \citep{minowa:05}, MUSYC \citep{quadri:07, taylor:09b}, FIREWORKS \citep{wuyts:fireworks}, MODS \citep{kajisawa:11}, NMBS \citep{whitaker:nmbs}, UltraVISTA \citep{mccracken:ultravista}, TENIS \citep{hsieh:tenis}, WIRDS \citep{bielby:12}, ZFOURGE \citep{spitler:12}, and HUGS \citep{hugs}.
@@
 The Frontier Fields initiative has inspired a number of separate supporting general observer programs, such as deep \textit{HST} ultraviolet imaging (Siana et al., in prep) and grism spectroscopy \citep[GLASS, ][]{treu:glass}, and deep far-infrared imaging with \textit{Herschel} \citep{rawle:16}.
-Whereas the main goal of the HFF is to explore the galaxy population in the first billion year of cosmic history, this dataset is also unique for its combination of surveyed area, multiwavelength coverage and depth for studies of galaxy evolution across most of the age of the universe, down to, and including, the redshifts of the targeted galaxy clusters ($z\approx$~0.3--0.5).
+Whereas the main goal of the HFF is to explore the galaxy population in the first billion years of cosmic history, this dataset is also unique for its combination of surveyed area, multiwavelength coverage and depth for studies of galaxy evolution across most of the age of the universe, down to, and including, the redshifts of the targeted galaxy clusters ($z\approx$~0.3--0.5).
 
-The space-based HFF data alone, however, are not sufficient to robustly characterize red galaxies at $z>3$ because the WFC3/IR $H_{160}$ band lies on the UV side of the rest-frame optical Balmer/4000\AA~break at these redshifts, resulting in sub-optimal accuracies in the photometric redshifts and stellar population properties (e.g., stellar mass and rest-frame optical color) of galaxies at $z>2$ and even at $z>4$.
+The space-based HFF data alone, however, are not sufficient to robustly characterize red galaxies at $z\gtrsim3$ because the WFC3/IR $H_{160}$ band lies on the UV side of the rest-frame optical Balmer/4000\AA~break at these redshifts, resulting in sub-optimal accuracies in the photometric redshifts and stellar population properties \citep[e.g., stellar mass and rest-frame optical color;][]{muzzin:09}.
 Very deep $K$-band imaging is required to significantly improve the precision of both photometric redshifts and derived stellar population properties (see \S\ref{s:discussion}).
@@
 Finally, \S\ref{s:discussion} concludes showing a few crucial improvements enabled by the deep $K_s$ data in consort with the space-based \textit{HST} and \textit{Spitzer} Frontier Fields imaging.
-AB magnitudes are used throughout, with $m_{K, \mathrm{AB}} - m_{K, \mathrm{Vega}} = 1.826$ and 1.821 and pivot wavelengths \citep{tokunaga:05} of 2.152~\mum\ and 2.147~\mum\ computed for the \HAWKI\ and MOSFIRE $K_s$ filter bandpasses, respectively.
+AB magnitudes are used throughout, with $m_{K, \mathrm{AB}} - m_{K, \mathrm{Vega}} = 1.826$ and 1.821 and pivot wavelengths \citep{tokunaga:05} of 2.152~\mum\ and 2.147~\mum\ for the \HAWKI\ and MOSFIRE $K_s$ filter bandpasses, respectively.
 
@@
 These data were obtained with one \HAWKI\ chip centered on the $z=9.7$ candidate from \cite{zheng:12}; they largely cover the HFF cluster field but the pointing was not optimized to also include the parallel field whose location was defined later (see Fig.~\ref{fig:layout}).
-The MACS-1149 \HAWKI\ observations were obtained in Service Mode at the ESO/\textit{VLT} in \xxx\ separate OBs between 2013 Mar 21 and 2014 Jun 9.
-The 20$\times$15\,s~=~300\,s exposures were taken at nine dithered positions with regular offsets separated by \xxx\ arcsec; the on-source exposure time of the final MACS-1149 \HAWKI\ mosaic is \xxx\ hours.
+The MACS-1149 \HAWKI\ observations were obtained in Service Mode at the ESO/\textit{VLT} in between 2013 Mar 21 and 2014 Jun 9.
+The 20$\times$15\,s~=~300\,s exposures were taken at nine dithered positions offset within a $30\arcsec$ box; the on-source exposure time of the final MACS-1149 \HAWKI\ mosaic is 5.3 hours.
 
@@
 
-The \HAWKI\ and MOSFIRE observations were reduced with a pipeline that has been developed for previous surveys with the NEWFIRM \citep[NMBS;][]{whitaker:nmbs} and FOURSTAR (zFOURGE; \citealt{spitler:12}; Labb\'e et al., in prep.) infrared imaging instruments.
+The \HAWKI\ and MOSFIRE observations were reduced with a pipeline that has been developed for previous surveys with the NEWFIRM \citep[NMBS;][]{whitaker:nmbs} and FOURSTAR (ZFOURGE; \citealt{spitler:12}; Straatman et~al, submitted) infrared imaging instruments.
 Treating each detector individually, the pipeline is easily modified for the different instrument configurations of the four \HAWKI\ and single MOSFIRE chips.
@@
 As the archival \HAWKI\ images of the MACS-1149 field were obtained with longer individual exposures and with fewer exposures per sequence, the mask-pass technique described above did not produce satisfactory results.
-In this case, for each raw exposure we divide by the sky flat and then subtract a third-order polynomial fit to the background.
+In this case, for each raw exposure we divide by the empirical sky flat and then subtract a third-order polynomial fit to the background.
 
@@
 The 2-Micron All Sky Survey \citep[2MASS;][]{2mass} provides an additional check on the photometric calibration, though the comparison is limited due to relatively little brightness overlap between the faint end of reliable 2MASS photometry and the bright end where stars are in the linear regime of the detectors on the 8--10\,m telescopes.
-A comparison of the observed photometry in the four survey fields to the 2MASS catalog magnitudes is shown in Figure~\ref{fig:2mass}.
+A comparison of the observed photometry in the four survey fields to the 2MASS catalog magnitudes is shown in Fig.~\ref{fig:2mass}.
 The stellar photometry on the deep $K_s$-band mosaics is measured within $1\arcsec$ apertures corrected to infinity with the curves of growth described below.
-The error-bars on the points in Figure~\ref{fig:2mass} come from the 2MASS catalog; the photometric uncertainties are negligible for such bright stars in the deep images ($\mathrm{S/N}>100$).
+The error-bars on the points in Fig.~\ref{fig:2mass} come from the 2MASS catalog; the photometric uncertainties are negligible for such bright stars in the deep images ($\mathrm{S/N}>100$).
 The 2MASS comparison suggests a systematic zeropoint offset of $-0.09$~mag for the Abell S1063 field.
 However, the exposure time for that field was similar to those of the additional deep \HAWKI\ fields and the derived depths of \textit{all} deep \HAWKI\ fields are nearly identical (Tab.~\ref{tab:summary}).
-Therefore, we do not apply this offset to the S1063 zeropoint.
+Therefore, we do not apply this offset to the S1063 zeropoint since if real it should be reflected in the derived depth of that field as compared to the others.
 Given the overall good agreement with the 2MASS photometry, we estimate that the systematic uncertainty of the absolute photometric calibration is $\sigma_\mathrm{sys}\leq0.05$~mag.
+
+Similar to \cite{hugs}, we do not apply any correction for non-linearity effects of the \HAWKI\ or MOSFIRE detectors.
+For the case of \HAWKI\, the user manual suggests that the detector is linear at the 1\% level up to 30~000 detector counts (ADUs).
+For a 2MASS star in the Abell~2744 field with $K=17.3$, the brightest pixel reaches $\sim$31~000~ADUs including the bright sky background.
+Therefore, there could be a linearity correction of up to a few percent between the bright 2MASS stars used for the photometric comparison and galaxies in the field some 8--9 magnitudes fainter, though the stars in Fig.~\ref{fig:2mass} do not show any significant slope in the 2MASS magnitude residuals across $\sim$1.5~mag of dynamic range.
+For the fainter galaxies, empirical ``zeropoint corrections'' are often computed as part the photometric redshift analysis \citep[e.g., ][]{skelton:14}, and these would compensate for relatively small linearity effects.
 
@@
 For the \HAWKI\ exposures, we fit a third-order polynomial to the geometric distortion model determined by \cite{libralato:14} and specify the polynomial terms as ``Simple Imaging Polynomial'' (SIP) coefficients \citep{fits:sip} in the individual exposure FITS files.
-The SIP FITS header generally applicable for \HAWKI\ is provided in Appendix \xxx.
+The SIP FITS header generally applicable for \HAWKI\ is provided in Appendix~\ref{appendix:sip}.
 We note that the \HAWKI\ distortion is generally small, reaching $\sim$2 pixels ($0\farcs2$) at the image corners, but that the distortion corrections are necessary, in particular, given the excellent overall image quality of the observations.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%                                                                %
-%   Mosaic properties
-%                                                                %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Drizzled $K_s$-band Mosaics}
-\label{s:mosaic}
-
-A crucial innovation of the image processing of the present analysis compared to previous works is that we combine all of the individual background-subtracted raw images (e.g., \xxx\ 4000 files for 1000 exposures $\times$ 4 detectors for the \mbox{Abell 2744} field) into the final output mosaic with the ``drizzle'' algorithm (\citealp{fruchter:02}, as implemented in the DrizzlePac/AstroDrizzle package; \citealp{drizzlepac}).
-The benefits of the drizzle algorithm  preserving robust noise properties of the output mosaics will be discussed in detail below in \S\ref{s:depth}.
-Here we simply indicate that the drizzle implementation provides additional benefits in 1) trivial application of relative weights of the input images in creating the final mosaics, 2) providing infrastructure to apply the astrometric alignment and geometric distortion stored in the FITS headers, and 3) allowing the definition of an arbitrary output pixel grid without any additional image resampling.
-
-Final mosaics of each field with $0\farcs1$ pixels are drizzled from the input exposures with weights
-\begin{equation}
-w^{-1} = F^2\cdot B \cdot \mathrm{FWHM}^\alpha / \sqrt{1-e^2}, 
-\label{eq:weight}
-\end{equation}
-where $F$ is the factor to scale the exposure to the photometric system ($F\equiv1$ for the average of exposures in the OB taken under photometric conditions, $F\gtrsim1$ otherwise), $B$ is the measured background, FWHM is the full width at half maximum of stars identified in the field, and $e$ is the ellipticity of the stellar point spread function (PSF).
-The parameter $\alpha$ allows for optimizing the image quality of the final mosaic.
-Increasing $\alpha$ puts larger weight on the exposures with the best image-quality while effectively ignoring data with poorer seeing; we adopt $\alpha=2$ for a compromise between the image quality and effective exposure time of the final mosaic.
-
-Finally, we subtract a cell-based background from the final mosaics using an algorithm based on that used by SExtractor and SWarp but that provides more aggressive masking of flux in the outer isophotes of bright galaxies.
-The mosaics are shown in Figure~\ref{fig:layout}, along with the position of the deep Frontier Fields \textit{HST} imaging fields.
-We also compute the robust NMAD scatter \citep{brammer:08} of empty sky pixels within the same cells to empirically calibrate the inverse variance maps that are shown in Figure~\ref{fig:ivar_mosaics} (see also \S\ref{s:depth}).
-The final image and inverse variance mosaics are provided as an ESO Phase 3 data product \XXX{URL}, as well as attached as a digital product to this manuscript \XXX{(I think ApJSS can and wants to do this)}.
-The released images are all scaled to a common zeropoint of 26.0 (AB mag), which gives pixel values of order unity faint galaxies near the detection threshold.
-In the sections below we describe the characteristics of the mosaics in more detail.
 
@@
 %%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%                                                                %
+%   Mosaic properties
+%                                                                %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Drizzled $K_s$-band Mosaics}
+\label{s:mosaic}
 
@@
 
+A crucial innovation of the image processing of the present analysis compared to previous works is that we combine all of the individual background-subtracted raw images (e.g., 7040 files for 1760 exposures $\times$ 4 detectors for the \mbox{Abell 2744} field) into the final output mosaic with the ``drizzle'' algorithm (\citealp{fruchter:02}, as implemented in the DrizzlePac/AstroDrizzle package; \citealp{drizzlepac}).
+The benefits of the drizzle algorithm  preserving robust noise properties of the output mosaics will be discussed in detail below in \S\ref{s:depth}.
+Here we simply indicate that the drizzle implementation provides additional benefits in 1) trivial application of relative weights of the input images in creating the final mosaics, 2) providing infrastructure to apply the astrometric alignment and geometric distortion stored in the FITS headers, and 3) allowing the definition of an arbitrary output pixel grid without any additional image resampling.
+
+Final mosaics of each field with $0\farcs1$ pixels are drizzled from the input exposures with weights
+\begin{equation}
+w^{-1} = F^2\cdot B \cdot \mathrm{FWHM}^\alpha / \sqrt{1-e^2}, 
+\label{eq:weight}
+\end{equation}
+where $F$ is the factor to scale the exposure to the photometric system ($F\equiv1$ for the average of exposures in the OB taken under photometric conditions, $F\gtrsim1$ otherwise), $B$ is the measured background, FWHM is the full width at half maximum of stars identified in the field, and $e$ is the ellipticity of the stellar point spread function (PSF).
+The parameter $\alpha$ allows for optimizing the image quality of the final mosaic.
+Increasing $\alpha$ puts larger weight on the exposures with the best image-quality while effectively ignoring data with poorer seeing; we adopt $\alpha=2$ for a compromise between the image quality and effective exposure time of the final mosaic.
+
+Finally, we subtract a cell-based background from the final mosaics using an algorithm based on that used by SExtractor and SWarp but that provides more aggressive masking of flux in the outer isophotes of bright galaxies.
+The mosaics are shown in Fig.~\ref{fig:layout}, along with the position of the deep Frontier Fields \textit{HST} imaging fields.
+We also compute the robust NMAD scatter \citep{brammer:08} of empty sky pixels within the same cells to empirically calibrate the inverse variance maps that are shown in Fig.~\ref{fig:ivar_mosaics} (see also \S\ref{s:depth}).
+The final image and inverse variance mosaics are provided as an ESO Phase 3 data product\footnote{\url{http://www.eso.org/sci/observing/phase3.html}} and are also currently made available at \url{http://gbrammer.github.io/HAWKI-FF/}\XXX{Note to editor/referee, the final products aren't posted yet but the images are available now and the web releases will be complete by the time of publication.}.
+The released images are all scaled to a common zeropoint of 26.0 (AB mag), which gives pixel values of order unity faint galaxies near the detection threshold.
+In the sections below we describe the characteristics of the mosaics in more detail.
+
 \subsection{Image quality}
@@
 
-Figure~\ref{fig:cog} shows the curves of growth for stars identified by the tight relationship between their brightness and half-light radii \citep[see, e.g., Fig.~13 of][]{skelton:14}.
+Fig.~\ref{fig:cog} shows the curves of growth for stars identified by the tight relationship between their brightness and half-light radii \citep[see, e.g., Fig.~13 of][]{skelton:14}.
 The image quality of the $K_s$-band mosaics is excellent with stellar FWHM$\lesssim0\farcs4$ for the deep \HAWKI\ fields, thanks in large part to the execution of the \HAWKI\ observations in service mode ensuring optimal and uniform image quality across the many hours of integration on the survey fields.
@@
 The result is that the image quality of the MACS-0717 and MACS-1149 fields is somewhat degraded in comparison, at $0\farcs42$--$0\farcs54$, and the image quality differs measurably between the two MOSFIRE pointings in the MACS-0717 field.
-Despite the compact cores of the stellar PSFs, the curves of growth shown in Figure~\ref{fig:cog} deviate from Gaussian profiles with significantly higher flux in the outer wings, with a shape more consistent with a Moffat profile \citep{trujillo:01}  with $\beta\sim2$.
+Despite the compact cores of the stellar PSFs, the curves of growth shown in Fig.~\ref{fig:cog} deviate from Gaussian profiles with significantly higher flux in the outer wings, with a shape more consistent with a Moffat profile \citep{trujillo:01}  with $\beta\sim2$.
 We caution that these extended profiles will yield significantly shallower final image depths than would be predicted with Exposure Time Calculators that may assume Gaussian profiles.
@@
 
-A comparison of final combined images created with output $0\farcs1$ and $0\farcs06$ pixel grids using the drizzle implementation to versions created with the SWarp software \citep{swarp} is shown in Figure~\ref{fig:drizzle_cutout}.
+A comparison of final combined images created with output $0\farcs1$ and $0\farcs06$ pixel grids using the drizzle implementation to versions created with the SWarp software \citep{swarp} is shown in Fig.~\ref{fig:drizzle_cutout}.
 In the later case, the raw pixels are resampled twice before making the final mosaic, once combining the raw exposures in each OB and a second time combining the separate OBs into the final stack with an arbitrary pixel grid.
@@
 
-The fact that the drizzle-combined images in Figure~\ref{fig:drizzle_cutout} (panels \textit{a} and \textit{b}) appear to the eye to be noisier than their SWarped counterparts (panels \textit{c} and \textit{d}) is simply the result of a smoothing of the pixel noise in the latter case rather than reflecting true underlying differences in their pixel-to-pixel variance.
+The fact that the drizzle-combined images in Fig.~\ref{fig:drizzle_cutout} (panels \textit{a} and \textit{b}) appear to the eye to be noisier than their SWarped counterparts (panels \textit{c} and \textit{d}) is simply the result of a smoothing of the pixel noise in the latter case rather than reflecting true underlying differences in their pixel-to-pixel variance.
 \cite{casertano:00} discuss how the noise properties of correlated output pixels are affected by the relative sizes of the input and output pixel grids and the adopted drizzle parameters.
 With the drizzle algorithm and a small pixfrac, a given input pixel will map to only a single output pixel and therefore correlations between the output pixels should be minimal\footnote{Inter-pixel capacitance and other forms of crosstalk intrinsic to the IR detectors place a lower limit on the correlations between adjacent pixels \citep{finger:05, hilbert:ipc}.}.
-
-We explore these correlations between the output pixels in Figure~\ref{fig:corr_noise}.
-For the black and red curves in each panel, we measure value differences between random pairs of pixels sorted by the separation between the pixels for the drizzled and SWarped images, respectively.
-For perfectly uncorrelated pixels, these curves will be flat and will reflect the intrinsic noise of the image pixels.
-In the presence of correlations between adjacent pixels, however, the curves will show a depression at small separations.
-As expected, the SWarp-combined images show just such a depression for pairs of pixels 1--2 pixels apart.
-The yellow curves in Figure~\ref{fig:corr_noise} show the same pairwise differences for the drizzled images now convolved by small Gaussian kernels as indicated, which agree very well with the curves for the images with correlated pixels.
 
@@
 
+We explore these correlations between the output pixels in Fig.~\ref{fig:corr_noise}.
+For the black and red curves in each panel, we measure value differences between random pairs of pixels sorted by the separation between the pixels for the drizzled and SWarped images, respectively.
+For perfectly uncorrelated pixels, these curves will be flat and will reflect the intrinsic noise of the image pixels.
+In the presence of correlations between adjacent pixels, however, the curves will show a depression at small separations.
+As expected, the SWarp-combined images show exactly such a depression for pairs of pixels 1--2 pixels apart.
+The yellow curves in Fig.~\ref{fig:corr_noise} show the same pairwise differences for the drizzled images now convolved by small Gaussian kernels as indicated, which agree very well with the curves for the images with correlated pixels.
+
 In terms of understanding the noise properties of the images, a key point here is not only that depression at small separations indicates the presence of correlations between adjacent pixels but also the fact that correlations cause the apparent overall r.m.s.
@@
 \end{equation}
-for circular apertures with diameter, $D$, in pixels, and where $\sigma_\mathrm{pix}^2$ is the per-pixel variance determined from an analysis such as that in Figure~\ref{fig:corr_noise}.
-
-The product of Equation~\ref{eq:aper_var} and the inverse of the curves of growth shown in Figure~\ref{fig:cog} (i.e., aperture corrections) defines an optimal photometric aperture size where this product is minimized \citep{whitaker:nmbs}.
-For our $K_s$-band mosaics, this optimal aperture has $D\sim0\farcs6$, just larger than the FWHM of point-source profiles.
-The $5\,\sigma$ limiting magnitudes (AB) within $0\farcs6$ diameter apertures are indicated in Figure~\ref{fig:depth} and listed in Table~\ref{tab:summary}.
-Reaching 26th magnitude (AB), the \HAWKI\ images presented here are among the deepest $K_s$-band images ever obtained and will provide a crucial complement to the deep \textit{HST} and \textit{Spitzer} imaging of the Frontier Fields.
-
-We conclude here with a brief comment on the ``HUGS'' ultra-deep \HAWKI\ images of the CANDELS GOODS-S and UDS survey fields.
-\cite{hugs} report a final depth of 26.5 AB for point sources within $D=0\farcs4$ apertures for their deepest field, ``GOODS-D1'', which has an exposure time of 31.5 hours similar to the deep \HAWKI\ fields summarized in Table~\ref{tab:summary}.
-\cite{hugs} assume ideal noise properties that likely suffer some residual correlation between adjacent pixels such as that indicated by the red curves in Figure~\ref{fig:corr_noise}, where we found that the drizzled rms was some 40\% higher than that measured in the presence of the pixel correlations for $0\farcs1$ pixel mosaics.
-This effect ($\sim$0.4 mag) along with the relative exposure time difference ($\sim$0.1 mag) can account for much of the difference between the reported depths of the deepest HUGS pointings and the \textit{Hubble} Frontier Fields \HAWKI\ images described here.
-The comparison must be made because any additional half magnitude increase depth is exponentially more expensive to obtain!
+for circular apertures with diameter, $D$, in pixels, and where $\sigma_\mathrm{pix}^2$ is the per-pixel variance determined from an analysis such as that in Fig.~\ref{fig:corr_noise}.
 
@@
 
+The product of Equation~\ref{eq:aper_var} and the inverse of the curves of growth shown in Fig.~\ref{fig:cog} (i.e., aperture corrections) defines an optimal photometric aperture size where this product is minimized \citep{whitaker:nmbs}.
+For our $K_s$-band mosaics, this optimal aperture has $D\sim0\farcs6$, just larger than the FWHM of point-source profiles.
+The $5\,\sigma$ limiting magnitudes (AB) within $0\farcs6$ diameter apertures are indicated in Fig.~\ref{fig:depth} and listed in Table~\ref{tab:summary}.
+Reaching 26th magnitude (AB), the \HAWKI\ images presented here are among the deepest $K_s$-band images ever obtained and will provide an important complement to the deep \textit{HST} and \textit{Spitzer} imaging of the Frontier Fields.
+
+We conclude here with a brief comment on the ``HUGS'' ultra-deep \HAWKI\ images of the CANDELS GOODS-S and UDS survey fields.
+\cite{hugs} report a final depth of 26.5 AB for point sources within $D=0\farcs4$ apertures for their deepest field, ``GOODS-D1''.
+This field has an exposure time of 31.5 hours, similar to the deep \HAWKI\ fields summarized in Table~\ref{tab:summary}.
+\cite{hugs} assume ideal noise properties that likely suffer some residual correlation between adjacent pixels such as that indicated by the red curves in Fig.~\ref{fig:corr_noise}, where we found that the drizzled rms was some 40\% higher than that measured in the presence of the pixel correlations for $0\farcs1$ pixel mosaics.
+This effect ($\sim$0.4 mag) along with the relative exposure time difference ($\sim$0.1 mag) can account for much of the difference between the reported depths of the deepest HUGS pointings and the \textit{Hubble} Frontier Fields \HAWKI\ images described here.
+The comparison must be made because any additional half magnitude increase depth is exponentially more expensive to obtain!
+
 \subsection{Source detection \& completeness}
@@
 Briefly, we insert artificial sources in blank regions of the portions of the images covering the \textit{HST} ``parallel'' fields and compute the fraction of sources recovered with SExtractor as a function of the source magnitude.
-The resulting completeness curves are shown in Figure~\ref{fig:completeness}.
+The resulting completeness curves are shown in Fig.~\ref{fig:completeness}.
 As in the sensitivity analysis above, we find that the detection completeness curves are nearly identical for the deep \HAWKI\ fields, with 50\% (90\%) source completeness at $K_s\sim25.9$ (25.7) AB.
@@
 The center and right panels show a dramatic example of a galaxy with a strong break between the \textit{HST} and IRAC bands.
+Its photometric redshift and $\Delta(U-V)$ color are indicated with the large red star in the left panel.
 The addition of the $K_s$ band pinpoints the location of the break and shrinks the photometric redshift uncertainties by a factor of two and suggesting $z_\mathrm{phot} > 3$ (right panel).
@@
 
-A comparison of the $K_s$-band and space-based Frontier Fields imaging is shown in Figure~\ref{fig:cutout}, showing just a small cutout of the MACS-0416 ``parallel'' field.
+A comparison of the $K_s$-band and space-based Frontier Fields imaging is shown in Fig.~\ref{fig:cutout}, showing just a small cutout of the MACS-0416 ``parallel'' field.
 The deep optical and near-infrared \textit{Hubble} imaging provides spectacular multiwavelength spatially-resolved information on physical scales of just $\sim$1~kpc (Fig.~\ref{fig:cutout}a).
@@
 We then compute photometric redshifts with EAZY \citep{brammer:08} for the catalogs with and without the $K_s$-band photometry included.
-The left panel of Figure~\ref{fig:delta_UV} shows the difference in the rest-frame $U-V$ color derived from the photometric redshift fit, which probes the strength of the Balmer/4000~\AA\ break and is a proxy for the age and mass-to-light ratio of the underlying stellar population.
+The left panel of Fig.~\ref{fig:delta_UV} shows the difference in the rest-frame $U-V$ color derived from the photometric redshift fit, which probes the strength of the Balmer/4000~\AA\ break and is a proxy for the age and mass-to-light ratio of the underlying stellar population.
 
@@
 
-The right two panels of Figure~\ref{fig:delta_UV} show a single galaxy that illustrates how these systematic effects are not trivial and will likely result in biases in the interpretation of the galaxy population properties derived from the \textit{HST} and IRAC observations alone.
+The right two panels of Fig.~\ref{fig:delta_UV} show a single galaxy that illustrates how these systematic effects are not trivial and will likely result in biases in the interpretation of the galaxy population properties derived from the \textit{HST} and IRAC observations alone.
 The spectral energy distribution (SED) shown is steadily rising through the reddest WFC3/IR bands, and then shows a sharp break with bright detections in the IRAC bands.
@@
 Even though the \textit{measured} $H_{160}-K_s$ color is redder than that inferred from the \textit{HST} + IRAC photometry alone, the final rest-frame $U-V$ color is actually \textit{bluer} as a result of the higher preferred redshift.
-Evolved galaxies at $z>3$ such as the one shown are an intriguing population deserving of detailed study in their own right, and the combined Frontier Fields \textit{Hubble}+$K_s$+IRAC dataset is ideally suited for this purpose.
+Evolved galaxies at $z>3$ such as the one shown in Fig.~\ref{fig:delta_UV}bare an intriguing population deserving of detailed study in their own right, and the combined Frontier Fields \textit{Hubble}+$K_s$+IRAC dataset is ideally suited for this purpose.
 
@@
 \noindent We gratefully acknowledge funding support from the STScI Director's Discretionary Research Fund.
-D.M.
-acknowledges the support of the Research Corporation for Science Advancement's Cottrell Scholarship, and of the National Science Foundation under Grant No.
+DM acknowledges the support of the Research Corporation for Science Advancement's Cottrell Scholarship, and of the National Science Foundation under Grant No.
 1513473.
-D.M.
-and D.L.-V.
-acknowledge support from program HST-AR-14302, provided by NASA through a grant from the Space Telescope Science Institute, which is operated by the Association of Universities for Research in Astronomy, Incorporated, under NASA contract NAS5-26555.
-B.V.
-acknowledges the support from an Australian Re- search Council Discovery Early Career Researcher Award (PD0028506).
+DM and DL-V acknowledge support from program HST-AR-14302, and KEW gratefully acknowledges support by NASA through Hubble Fellowship grant \#HF2-51368, both of which are provided by NASA through a grant from the Space Telescope Science Institute, which is operated by the Association of Universities for Research in Astronomy, Incorporated, under NASA contract NAS5-26555.
+BV acknowledges the support from an Australian Research Council Discovery Early Career Researcher Award (PD0028506).
 This work was supported by NASA Keck PI Data Awards, administered by the NASA Exoplanet Science Institute.
@@
 
+\label{appendix:sip}
 \section{SIP keywords for \HAWKI\ geometric distortion}
@@
 
-\section{Description of provided data products}
+%\section{Description of provided data products}
 
-\clearpage
+%\clearpage
 

### 2016-05-10 Paper0/ms.tex
Fix Fig. 7 caption

@@
 The $K_s$-band at 2.2\,\mum\ crucially fills the gap between the reddest HST filter (1.6\,\mum\,$\sim H$~band) and the IRAC 3.6\,\mum\ passband.
-While reaching the full depths of the space-based imaging is not currently feasible from the ground, the deep $K_s$-band images provide important constraints on both the redshifts and the stellar population properties of galaxies extending well below the characteristic stellar mass across most of the age of the universe, down to, and including, the redshifts of the targeted galaxy clusters ($z \lesssim 0.5$).
+While reaching the full depths of the space-based imaging is not currently feasible from the ground, the deep $K_s$-band images provide import constraints on both the redshifts and the stellar population properties of galaxies extending well below the characteristic stellar mass across most of the age of the universe, down to, and including, the redshifts of the targeted galaxy clusters ($z \lesssim 0.5$).
 Reduced, aligned mosaics of all six survey fields are provided accompanying this manuscript.
@@
 \plotone{aper_depth.pdf} % Figures/exposure_map.pdf
-\caption{\textit{Top panel:} The product of the inverse curves-of-growth (Fig.~\ref{fig:cog}) and the predicted variance within a photometric aperture (Eq.~\ref{eq:aper_var}) gives the point-source sensitivity as a function of aperture size.
+\caption{The product of the inverse curves-of-growth (Fig.~\ref{fig:cog}) and the predicted variance within a photometric aperture (Eq.~\ref{eq:aper_var}) gives the point-source sensitivity as a function of aperture size.
 The S/N is maximized with an aperture somewhat larger than the FWHM; depths evaluated at $D=0\farcs6$ near the maxima are indicated.)

### 2016-06-22 Paper0/ms.tex
Resubmitted

--- 2016-04-29 Paper0/ms.tex
@@
 Elizbeth A. Barker\altaffilmark{1},
-Adriano Fontana\altaffilmark{6},
-Audrey Galametz\altaffilmark{7},
-Anna Ferr\'e-Mateu\altaffilmark{8},
-Tadayuki Kodama\altaffilmark{9},
+Masayuki Tanaka\altaffilmark{6},
+Adriano Fontana\altaffilmark{7},
+Audrey Galametz\altaffilmark{8},
+Anna Ferr\'e-Mateu\altaffilmark{9},
+Tadayuki Kodama\altaffilmark{6},
 Britt Lundgren\altaffilmark{10},
@@
 Mauro Stefanon\altaffilmark{3},
-Masayuki Tanaka\altaffilmark{8},
 Sune Toft\altaffilmark{12},
@@
 \altaffiltext{5}{Australian Astronomical Observatories, PO Box 915 North Ryde NSW 1670, Australia}
-\altaffiltext{6}{INAF, Osservatorio Astronomico di Roma, Via Frascati 33, I-00040, Monteporzio, Italy}
-\altaffiltext{7}{Max-Planck-Institut fur extraterrestrische Physik (MPE), Giessenbachstr. 1, D-85748 Garching, Germany}
-\altaffiltext{8}{Subaru Telescope, 650 N. Aohoku Pl, 96720 Hilo, HI, USA}
-\altaffiltext{9}{National Astronomical Observatory of Japan
-2-21-1 Osawa, Mitaka, Tokyo 181-8588, Japan}
+\altaffiltext{6}{National Astronomical Observatory of Japan 2-21-1 Osawa, Mitaka, Tokyo 181-8588, Japan}
+\altaffiltext{7}{INAF, Osservatorio Astronomico di Roma, Via Frascati 33, I-00040, Monteporzio, Italy}
+\altaffiltext{8}{Max-Planck-Institut fur extraterrestrische Physik (MPE), Giessenbachstr. 1, D-85748 Garching, Germany}
+\altaffiltext{9}{Subaru Telescope, 650 N. Aohoku Pl, 96720 Hilo, HI, USA}
 \altaffiltext{10}{Department of Astronomy, University of Wisconsin, Madison, WI 53706, USA}
@@
 
-The latest effort to further our knowledge of galaxy formation and evolution is represented by the {\it HST} Frontier Fields (HFF) program (Lotz et al., in prep).
-The HFF program is a multi-cycle program consisting of $\sim$560 orbits of Director's Discretionary (DD) time that is imaging six deep fields centered on strong lensing galaxy clusters in parallel with six deep blank fields.
+The latest effort to further our knowledge of galaxy formation and evolution is represented by the {\it HST} Frontier Fields (HFF) program \citep{lotz:16}.
+The HFF program is a multi-cycle \textit{Hubble} program consisting of 840 orbits of Director's Discretionary (DD) time that is imaging six deep fields centered on strong lensing galaxy clusters in parallel with six deep blank fields.
 The primary science goals of the twelve HFF fields are to 1) reveal the population of galaxies at $z=5-10$ that are 10--50 times fainter intrinsically than any presently known, 2) solidify our understanding of the stellar masses and star formation histories of faint galaxies, 3) provide the first statistically meaningful morphological characterization of star-forming galaxies at $z>5$, and 4) find $z>8$ galaxies magnified by the cluster lensing, with some bright enough to make them accessible to spectroscopic follow-up.
@@
 
-The Frontier Fields initiative has inspired a number of separate supporting general observer programs, such as deep \textit{HST} ultraviolet imaging (Siana et al., in prep) and grism spectroscopy \citep[GLASS, ][]{treu:glass}, and deep far-infrared imaging with \textit{Herschel} \citep{rawle:16}.
+The Frontier Fields initiative is complemented by a number of separate supporting general observer programs, such as deep \textit{HST} ultraviolet imaging (Siana et al., in prep) and grism spectroscopy \citep[GLASS, ][]{treu:glass}, and deep far-infrared imaging with \textit{Herschel} \citep{rawle:16}.
 Whereas the main goal of the HFF is to explore the galaxy population in the first billion years of cosmic history, this dataset is also unique for its combination of surveyed area, multiwavelength coverage and depth for studies of galaxy evolution across most of the age of the universe, down to, and including, the redshifts of the targeted galaxy clusters ($z\approx$~0.3--0.5).
@@
 The \HAWKI\ field of view is perfectly suited (Abell 2744, MACS-0416, Abell S1063 and Abell 370) for simultaneous imaging of the cluster+parallel field pairs, which require two separate pointings with MOSFIRE (MACS-0717 and MACS-1149).
-The footprint of additional archival \HAWKI\ coverage of the MACS-1149 field is indicated by the gray square in the lower right panel.\label{fig:layout}}
+The footprints of the AO-assisted GSAOI $K_s$-band imaging of the Abell 2744 and MACS-0416 fields \citep{schirmer:15} are shown in green and the footprint of additional archival \HAWKI\ coverage of the MACS-1149 field is indicated by the gray square in the lower right panel.\label{fig:layout}}
 \end{figure*}
@@
 After dividing by the flat, the background of each exposure is determined in a first pass from the simple median of the four exposures that came both immediately before and after it.
-The background-subtracted exposures are combined into a mosaic from which objects are detected.
-The final refined background of each exposure is determined in a second ``mask pass'' from a median again from the four exposures before and after it but now aggressively masking all pixels that contain flux from the detected objects.
+The first-pass background-subtracted exposures are combined into a mosaic, and objects are identified as pixels with values greater than five times the robust standard deviation \citep{beers:90} of the combined image.
+A buffer with radius 3 pixels is grown around each ``object'' pixel that satisfies this 5$\sigma$ criterion.
+The final refined background of each exposure is determined in a second ``mask pass'' from a median again from the four exposures before and after it but now masking all pixels that contain flux from the detected objects.
 
@@
 The uncertainties shown are taken from the 2MASS catalog; these stars are measured at very high S/N in the deep $K_s$-band mosaics.
+The solid lines and shaded regions show the weighted average and standard deviation of the photometric offsets for each field, with the quantitative values indicated in the labels at the top of the figure.
 \label{fig:2mass}}
@@
 
-Final mosaics of each field with $0\farcs1$ pixels are drizzled from the input exposures with weights
+Final mosaics of each field with $0\farcs1$ pixels are drizzled from the input exposures with weights \citep[following][]{whitaker:nmbs}
 \begin{equation}
@@
 We also compute the robust NMAD scatter \citep{brammer:08} of empty sky pixels within the same cells to empirically calibrate the inverse variance maps that are shown in Fig.~\ref{fig:ivar_mosaics} (see also \S\ref{s:depth}).
-The final image and inverse variance mosaics are provided as an ESO Phase 3 data product\footnote{\url{http://www.eso.org/sci/observing/phase3.html}} and are also currently made available at \url{http://gbrammer.github.io/HAWKI-FF/}\XXX{Note to editor/referee, the final products aren't posted yet but the images are available now and the web releases will be complete by the time of publication.}.
+The final science and inverse variance mosaics of all six fields are provided as ESO Phase 3 data products\footnote{\url{http://www.eso.org/sci/observing/phase3/news.html\#kiff}}.
 The released images are all scaled to a common zeropoint of 26.0 (AB mag), which gives pixel values of order unity faint galaxies near the detection threshold.
@@
 FWHM$\sim$$1\farcs7$--$2^{\prime\prime}$).
+
+The image quality obtained here with HAWK-I demonstrates the excellent natural seeing conditions of the \textit{VLT} site at Cerro Paranal.
+We note here that \HAWKI\ will soon be upgraded with a ground layer adaptive optics module \citep[GRAAL;][]{paufique:10}, that will improve the image quality by factors of 1.5--2 over the natural seeing over the wide \HAWKI\ field of view.
+\cite{schirmer:15} recently presented deep AO-corrected $K_s$-band imaging of the MACS-0416 cluster field obtained with the Gemini South Adaptive Optics Imager \citep[GSAOI;][]{carrasco:12}.
+Thanks to the availability of a bright guide star near the center of the cluster field, GSAOI provided exquisite image quality (FWHM$\sim0\farcs09$) better than even that of \textit{HST} WFC3/IR.
+Though somewhat shallower ($K_s\sim$25.6, 5$\sigma$) and covering a smaller field of view ($1\farcm7\times1\farcm8$, Fig.~\ref{fig:layout}) than the \HAWKI\ mosaics, the GSAOI images provide an auspicious demonstration of the power of wide-field AO-corrected imaging for deep extragalactic science.
 
@@
 \caption{\textit{Left:} Difference in the derived rest-frame $U-V$ colors as a function of photometric redshift for galaxies in the MACS-0416 parallel field, before and after including $K_s$ photometry along with the \textit{HST} and \textit{Spitzer} measurements.
-The scatter in the derived $U-V$ colors at $z>2.5$ is large ($\sim$0.2 mag) where the Balmer/4000~\AA\ break is in the gap between \textit{HST} $H_{160}$ and the \textit{Spitzer} IRAC 3.6~\mum\ bands, which is then well constrained by including $K_s$.
+The scatter in the derived $U-V$ colors (solid black lines) is large at $z>2.5$ ($\sim$0.2 mag) where the Balmer/4000~\AA\ break is in the gap between \textit{HST} $H_{160}$ and the \textit{Spitzer} IRAC 3.6~\mum\ bands, which is then well constrained by including $K_s$.
 This large scatter translates directly into large scatter on other quantities of interest, such as the stellar mass to light ratio and the age of the stellar populations.
@@
 Its photometric redshift and $\Delta(U-V)$ color are indicated with the large red star in the left panel.
+The SED fits (center panel) and derived photometric redshift probability densities (right panel) before and after including the $K_s$-band photometry are shown with gray and blue curves, respectively.
 The addition of the $K_s$ band pinpoints the location of the break and shrinks the photometric redshift uncertainties by a factor of two and suggesting $z_\mathrm{phot} > 3$ (right panel).
@@
 BV acknowledges the support from an Australian Research Council Discovery Early Career Researcher Award (PD0028506).
+MT acknowledges the support by JSPS KAKENHI Grant Number JP15K17617.
 This work was supported by NASA Keck PI Data Awards, administered by the NASA Exoplanet Science Institute.
@@
 This research made use of Astropy, a community-developed core Python package for Astronomy \citep{astropy}, open-source Python modules Numpy, Scipy, Matplotlib, and scikit-image, and NASA's Astrophysics Data System (ADS).
+
+Based on observations collected at the European Organisation for Astronomical Research in the Southern Hemisphere 
+under ESO programmes 090.A-0458, 092.A-0472, and 095.A-0533.
 \\
